### 2018-11-05 paper.tex
abstract

\documentclass[aps,prb,twocolumn,
	groupedaddress,superscriptaddress,
	amsfonts,amssymb,amsmath,floatfix,
	citeautoscript]{revtex4-1}

\usepackage{graphicx}
\usepackage[centering,hmargin=20mm,tmargin=30mm,bmargin=25mm]{geometry}
\usepackage{multirow}
\usepackage{newtxtext}
\usepackage[cmintegrals]{newtxmath}

%----- References -----
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{colorlinks,
	linkcolor={blue!75!black!80!yellow},
	citecolor={blue!75!black!80!yellow},
	urlcolor={blue!75!black!80!yellow}
}

%----- Captions in sans font -----
\makeatletter
\renewcommand\@make@capt@title[2]{%
	\@ifx@empty\float@link{\@firstofone}{\expandafter\href\expandafter{\float@link}}%
	\sffamily{\textbf{#1}}\@caption@fignum@sep#2
}%
\renewcommand\figurename{Figure}
\makeatother

\thickmuskip=5mu plus 2mu minus 1mu  %binary relations (default, 5mu plus 5mu)
\medmuskip=4mu plus 2mu minus 2mu    %binary operations (default, 4mu plus 2mu minus 4mu)

\frenchspacing %Ensure that revTeX does not do "double spaces" after punctuation

\renewcommand{\Im}{\operatorname{Im}}
\renewcommand{\Re}{\operatorname{Re}}
\newcommand{\sgn}{\operatorname{sgn}}

\newcommand{\iu}{\mathrm{i}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\unitvec}[1]{\ensuremath{\hat{\mathbf{#1}}}}
\newcommand{\sub}[1]{\ensuremath{_{\textrm{#1}}}} %Upright multi-character subscript
\newcommand{\super}[1]{\ensuremath{^{\textrm{#1}}}} %Upright multi-character superscript

\newcommand{\HarvardSEAS}{John A. Paulson School of Engineering and Applied Sciences, Harvard University, Cambridge, MA, USA}
\newcommand{\MITPhy}{Department of Physics, Massachusetts Institute of Technology, Cambridge, MA, USA}

%\usepackage[usenames]{color}
%\newcommand{\edited}[1]{{\color{red} #1}}
\newcommand{\comment}[2]{\textcolor{blue!70!black}{\small\textsf{\textsuperscript{\textsc{\MakeLowercase{#1}}}[#2]}}}
\usepackage[normalem]{ulem}
\newcommand{\swap}[2]{\textcolor{red!70!black}{\sout{#1}}\textrightarrow\textcolor{green!50!black}{#2}}
\newcommand{\remove}[1]{\textcolor{red!70!black}{\sout{#1}}}
\newcommand{\inset}[1]{\textcolor{green!50!black}{#1}}
\begin{document}

\title{Theory of interactions between matter and optical phonons in two dimensions}

\author{Nicholas Rivera}\email{nrivera@seas.harvard.edu}\affiliation{\HarvardSEAS}\affiliation{\MITPhy}
\author{Thomas Christensen}\affiliation{\MITPhy}
\author{Prineha Narang}\email{prineha@seas.harvard.edu}\affiliation{\HarvardSEAS}

\date{\today}

\begin{abstract}
Extreme confinement of electromagnetic energy by phonon polaritons promises extremely strong and novel forms of control over the dynamics of matter. To bring such control to its ultimate limit, it is important to consider phonon polaritons in two dimensional systems. However, recent studies have pointed out that in two-dimensional systems, splitting between longitudinal and transverse (LO and TO) optical phonons, which is necessary for the existence of phonon polaritons in three dimensions, is absent.  The question then arises as to whether the ability to exploit optical phonons disappears in lower dimensions. Here, we settle this question, finding that the phonon-polariton of the bulk is essentially replaced by the LO phonon. We present the confinement and propagation losses of LO phonons. We then calculate various measures of strong light-matter interaction such as single and two-photon spontaneous emission enhancement, and discuss methods such as EELS to probe these excitations.
\end{abstract}

\maketitle

\section{Electrodynamics of optical phonons in two-dimensions}

In this section, we develop the theory of electromagnetic waves associated with optical phonons in two dimensions. A theory of electromagnetic waves in polar 2D materials involves first calculating the dielectric permittivity associated with the monolayer.

\section{Strong light-matter interactions enabled by 2D optical phonons}

\subsection{Spontaneous emission}

\subsection{Spontaneous and stimulated electron energy loss}

%%Phonon polaritons, hybrid quasiparticles of photons and optical phonons, offer great promise for deeply sub-diffractional control of electromagnetic fields at mid-IR and THz frequencies. Phonon polaritons share many features in common with plasmon polaritons in conductors. In recent years, it has been shown that phonon polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra}. Due to this remarkable confinement and their relatively high lifetimes of around picoseconds, phonon polaritons open new opportunities for vibrational spectroscopy, radiative heat transfer \cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters \cite{kumar2015tunable,rivera2017making,kurman2018control}. The core features of phonon polaritons, such as mode shape, confinement, and propagation characteristics, are understood from simple Lorentz oscillator models of the dielectric function, enabling successful theoretical accounts of experimental observations.
%
%%we develop a first-principles theory of phonon polaritons in a 2D polar material, specifying the dielectric function in terms of \emph{ab initio} parameters, and finding a universal form for the dispersion relation of 2D phonon polaritons when the wavevector is much smaller than the size of the unit cell.
%
%\
%%\section{Linear response of phonon polaritons in two dimensions}
%%We now move to generalize the derivation of the previous section to describe the optical phonon contribution to Maxwell's equations coming from a two-dimensional material. 
%%Unlike the previous section, we will not demonstrate the results of the calculation on a particular material.
%Here, we are only interested in finding an expression for the optical response of the surface in terms of \emph{ab initio}-derivable parameters, as well as the general form of the dispersion relation of phonon-polaritons in 2D.
%Because of the change of symmetry in the system, we will not be able to perform the analysis by Fourier transforming Maxwell's equations in three dimensions.
%However, we will apply a technique which is often used to analyze electromagnetic modes in 2D systems such as plasmons in 2D conductors.
%An example of the technique is shown in detail in Ref.~\citenum{jablan2009plasmonics}, but we summarize the essential details here. 
%
%%Consider a two-dimensional ionic lattice located in the $xy$ plane ($z=0$) surrounded by a homogeneous environment of dielectric function $\epsilon_{\mathrm{env}}$.
%We consider the simplified case where the lattice is free-standing, to show the essential concept.
%We search for a mode whose magnetic field $\mathbf{H}$ is of the form $\mathbf{H}(z)\e^{\iu\mathbf{q}\cdot\mathbf{r}_{\parallel} - \iu\omega t}$, where $\mathbf{r}_{\parallel}=(x,y,0)$ is the position in the plane, $\mathbf{q}$ is the wavevector and $\omega_{\mathbf{q}}$ is the frequency.
%To simplify the discussion even further and furnish analytical forms for the phonon-polariton modes in 2D, we assume this material has mirror symmetry along the axis of $\mathbf{q}$.
%In this case, the electromagnetic modes can be decomposed into TE and TM-polarized modes.
%As is typical in the study of highly-confined polaritons, whether they are plasmon or phonon polaritons, the TM mode is the highly confined one, and occurs when $\epsilon(\omega) < 0$.
%In our convention, the TM mode is such that $\mathbf{H}$ is in the plane of the lattice and perpendicular to $\mathbf{q}$. 
%Without loss of generality, define the $\mathbf{q}$ direction to be the $x$ direction, making the magnetic field in the $y$ direction.
%	
%
%For $z \neq 0$, the Maxwell equation for the magnetic field is $(\nabla\times\nabla\times - \epsilon_{\mathrm{env}}\frac{\omega^2}{c^2})\mathbf{H}(z)\e^{\iu\mathbf{q}\cdot\mathbf{r}_{\parallel}} = 0$.
%For the TM polarized mode, $\mathbf{H}(z) = H(z)\unitvec{y}$; the vectorial wave equation consequently simplifies to a scalar one
%\begin{equation}\label{eq:2dmaxwell}
%\left(-\frac{\dd^2}{\dd{}z^2}+q^2-\epsilon_{\mathrm{env}}\frac{\omega^2}{c^2} \right)H(z) = 0.
%\end{equation}
%Seeking a surface-bound mode, $H(z)$ must assume the functional form
%	\comment{tc}{Nick, please check (I reverted an earlier edit of mine of this to the stuff below, but I cannot make sense of it): how can we specify $H(z)$ by equality sign here; and then in the following paragraphs we say ``To join these two solutions consistently, we apply \ldots$\unitvec{z}\times(\mathbf{H}^{(+)}-\mathbf{H}^{(-)}) = \mathbf{K} = -\iu\omega\mathbf{P}_{\mathrm{s}}$''?. I cannot make sense of this given that we use equality signs here and that $\mathbf{H}(z) = H(z)\unitvec{y}$.}
%\begin{align}
%	H(z) = \sgn(z)\e^{-\kappa |z|},
%\end{align}
%with $\kappa \equiv \sqrt{q^2-\epsilon_{\mathrm{env}}\frac{\omega^2}{c^2}}$.
%
%To join these two solutions consistently, we apply the interface conditions for the magnetic field. They are $\unitvec{z}\times(\mathbf{H}^{(+)}-\mathbf{H}^{(-)}) = \mathbf{K} = -\iu\omega\mathbf{P}_{\mathrm{s}}$ where $\mathbf{K}$ is the surface current density, expressed through the surface polarization density $\mathbf{P}_{\mathrm{s}}$. 
%Now we use the linear response relation $\mathbf{P}(\mathbf{q},\omega) = \boldsymbol{\Pi}(\mathbf{q},\omega)\mathbf{E}(\mathbf{q},\omega)$ in addition to Ampere's law $\nabla\times\mathbf{H} = -\iu\omega\epsilon_{\mathrm{env}}\mathbf{E}$.
%	\comment{tc}{There's a slight disconnect in that we never define the relation between $\mathbf{P}_{\mathrm{s}}$ and $\mathbf{P}$. Also, there's a bigger issue, I think: if we allow $\boldsymbol{\Pi}$ to be a $3\times 3$ matrix, then we ought to adopt more generic boundary conditions. Specifically, if there's out-of-plane ``current'', then the boundary conditions for normal parts of $\mathbf{B}$ and parallel parts of $\mathbf{E}$ would change.}
%Plugging these relations in yields an implicit equation connecting the polarization response function, the frequency of the 2D phonon polariton, and its frequency. Taking the case where $x$ coincides with a principal axis of the system, we have that 
%\begin{equation}\label{eq:2ddispersion}
%q\Pi_{xx}(q,\omega)  = -2\epsilon_{\mathrm{env}}.
%\end{equation}
%
%To connect this relation to the discussion of the 3D case, we write a microscopic form for the polarization response function of the ionic lattice. It is simply
%\begin{equation}\label{eq:2dsusceptibility}
%\boldsymbol{\Pi}(\mathbf{q},\omega) =  \frac{1}{\epsilon_0 A}\sum\limits_{m,n}\frac{\mathbf{P}_{mn}(\mathbf{q})\otimes\mathbf{P}_{nm}(\mathbf{q})}{\hbar\omega + E_{nm}+\iu 0^+}\left(\e^{\beta E_m}-\e^{\beta E_n} \right),
%\end{equation}
%is the 2D Fourier transform of the polarization density. As the form of the displacement of the ions is still given by Equation~\eqref{eq:phononfield}, $\mathbf{P}_{mn}(\mathbf{q})$ is still given by Equations~(\ref{eq:fourierpolarization}--\ref{eq:oscillatorstrength}), but with the Born charges and eigendisplacements appropriate to the 2D system of interest. Note that although the lattice is two dimensional, in general, the ions may be displaced in any three directions and thus $\boldsymbol{\Pi}$ is still a $3\times3$ matrix.
%Following the same kind of reasoning that lead to Equation~\eqref{eq:lorentzoscillator}, if we parameterize the effect of the different masses and Born charges in the unit cell by $M_{\mathrm{eff}}$ and $Q_{\mathrm{eff}}$, we will find a form for the polarization response function which is given by the second term of Equation~\eqref{eq:lorentzoscillator} but with $V$ replaced by $A$. Defining $N/A \equiv n_{\mathrm{s}}$, Equation~\eqref{eq:2dmaxwell} for the dispersion relation of 2D phonon polaritons tells us that
%\begin{equation}\label{eq:2dphononpolaritondispersion}
%\omega = \sqrt{\omega_{\mathrm{TO}}^2+\frac{n_{\mathrm{s}}Q_{\mathrm{eff}}^{2}}{2M_{\mathrm{eff}}\epsilon_0\epsilon_{\mathrm{env}}}q}.
%\end{equation}
%Note that this dispersion has a strong resemblance to the plasmon dispersion relation of two-dimensional free electron gases, which is  $\omega = \sqrt{n_{\mathrm{s}}e^2q/2m\epsilon_0\epsilon_{\mathrm{env}}}$  in the local limit~\cite{stern1967polarizability}.
%The only difference is the replacement of electron parameters (density, charge, mass) by the corresponding lattice parameters in the phononic case, and a shift in the minimum frequency by the transverse optical phonon frequency. 
%
%We note that the dispersion relation of Equation~\eqref{eq:2dphononpolaritondispersion} can in fact be obtained by taking the infinitely thin-limit of a bulk polar dielectric, and finding the dispersion of the resulting phonon polaritons in the quasi-electrostatic limit.
%In particular, for an isotropic polar dielectric of thickness $2d$ with boundaries at $z=\pm d$, solving Laplace's equations yields as the dispersion of the even parity mode:
%\begin{equation}\label{eq:thinfilmphononpolaritondispersion}
%\tanh(qd) = -\frac{\epsilon_{\mathrm{env}}}{\epsilon(\omega)} \rightarrow qd\epsilon(\omega) = -\epsilon_{\mathrm{env}},
%\end{equation}
%with $\epsilon_{\mathrm{env}}$ the dielectric constant of the surrounding environment and $\epsilon(\omega)$ that of the polar dielectric.
%The right-hand side is obtained in the limit of $d\rightarrow 0$.
%Taking Equation~\eqref{eq:lorentzoscillator} for the dielectric function in bulk, we have that (in the absence of losses) $\epsilon_{\mathrm{env}} + qd\epsilon_{\infty} + q\frac{n_sQ_{\mathrm{eff}}^{2}}{2\epsilon_0 M_{\mathrm{eff}}}\frac{1}{\omega_{\mathrm{TO}}^2-\omega^2} = 0$, where we have identified $V = 2Ad$.
%In the limit of small $d$, i.e., $qd \ll 1$, we recover Equations~\eqref{eq:2ddispersion} and \eqref{eq:2dphononpolaritondispersion} with $\Pi = \frac{n_sQ_{\mathrm{eff}}^{2}}{\epsilon_0 M}\frac{1}{\omega^2_{\mathrm{TO}}-\omega^2}$. It is interesting to see that in the 2D limit, the effect of the screening inside the material altogether vanishes. A similar phenomenon is known in the electrodynamics of plasmons in 2D materials, where the dielectric screening is known to be entirely provided by the environment and is equal to $\epsilon_{\mathrm{env}}$. A useful result which follows is that we may also derive another important figure of merit, the $p$-polarized reflectivity of a 2D polar dielectric, which is related to the local density of states of the polar dielectric, which quantifies the strength of coupling to external probes such as a nano-antenna or a quantum emitter.
%By considering the infinitely thin limit of a polar dielectric with dielectric function $\epsilon(\omega)$ and defining $\epsilon(\omega) = \Pi/2d$, we immediately find that the $p$-polarized reflectivity is given by
%\begin{equation}\label{eq:2dtmreflectivity}
%r_p = \frac{1}{1+2\epsilon_{\mathrm{env}}/q\Pi}.
%\end{equation}
%
%We note that, in general, when considering bulk and 2D versions of a material, at the very least, the effective parameters $Q_{\mathrm{eff}}$ and $M_{\mathrm{eff}}$ will change as a result of changes to the phonon band structure. In other words, although the forms of the expressions for the dielectric function 
%
%
% It is also quite possible that they altogether vanish in the 2D case as opposed to the 3D case. As a recently discovered example of such a phenomenon, it was found that the LO-TO splitting of some 2D polar dielectrics, such as hBN \cite{sohier2017breakdown} altogether vanishes, and only returns to its bulk value after a small but finite wavevector. This would then suggest that a description of phonon polaritons necessitates consideration of the spatially nonlocal dielectric function as in Equation (12), and that these polaritons moreover may only exist when their wavelengths are sufficiently short. Another important change, beyond the theory presented in this work is that the losses may also change drastically in the reduction of dimensions. For example, it was found in several experiments that the damping rate of the TO phonon in hBN nearly doubles upon going from bulk to 2D\cite{gorbachev2011hunting,tran2016quantum}, which may strongly suppress the ability to image the polariton with near-field probes.
%%This is due to the small thickness where the dispersion of the phonon polariton is quite flat, and very high wavevector modes, corresponding to the maximum wavevector detectable by a near-field tip, occur at frequencies very close to the TO phonon frequency.
%%Near the TO phonon-frequency however, the ratio $\Im\Pi/\Re\Pi$ is the highest, corresponding to very low quality-factor modes.
%%In addition to these considerations, the values of the parameters may depend on the background material, which may change the phonon band structure. 
%
%
%We provided this thin-film derivation as a simple way to understand the content of Equations~\eqref{eq:2ddispersion} and \eqref{eq:2dphononpolaritondispersion}.
%In general Equation~\eqref{eq:2dsusceptibility} should be used to calculate $\Pi$ and the corresponding dispersion.
%We note that a similar situation appears in the treatment of 2D plasmons: namely that a relation like Equation~\eqref{eq:2ddispersion} appears, except with $\Pi$ the polarization-polarization response of a 2D electron gas.
%And furthermore that one can also obtain the same equation by considering the ultra-thin limit of a 3D plasmonic material \cite{jablan2009plasmonics,jablan2013plasmons}.
%
%
%\section{Summary and Outlook}
%
%In summary, we have provided a theoretical framework based on linear response theory to calculate the phonon contribution to the dielectric function from first principles.
%Our framework is rather versatile, allowing us to: use first principles calculations to get the dielectric function, predict how nonlocality enters the dielectric function, and predict the effect of reduced dimensionality.
%A particularly interesting case to consider in future work would be to find a system where the optical phonons are drastically different from 3D to 2D.
%Perhaps it is possible that there are some materials in which the 2D optical phonons experience lower losses due to a reduced scattering phase space.
%The formalism we provide here may also be extended to understand phonon-polaritons in other more atypical reduced-dimensional settings, such as zero-dimensional settings in single emitters, i.e., `molecular phonon polaritons', in analogy to recent work on `molecular plasmons' \cite{manjavacas2013tunable,lauchner2015molecular}.
%Notably, we go beyond oversimplifications of the Lorentz oscillator model in which the Born charges are treated as a single scalar quantity and can treat the influence and interplay of many phonon modes that contribute to the dielectric function.
%We corroborated our approach through density functional theoretic calculations.
%
%In future work, besides furnishing an \emph{ab initio} description of nonlocality and reduced-dimension phonon polaritons, there are a number of interesting questions that can be addressed by the framework provided here.
%For example, we may consider situations such as the phonon polaritons associated with optical phonons at the interface of two materials.
%Another system of interest would be to consider two nearby layers of material in which the optical phonons of each material strongly couple to each other. 
%In that case, it would be relevant to evaluate how this strong coupling manifests itself in the infrared dielectric function, and ultimately the confinement and propagation of the phonon polaritons.

\section{Acknowledgements}
%This research used resources of the National Energy Research Scientific Computing Center, a DOE Office of Science User Facility supported by the Office of Science of the U.S. Department of Energy under Contract No. DE-AC02-05CH11231. 
N. R. recognizes the support of the DOE Computational Science Graduate Fellowship (CSGF) fellowship no. DE-FG02-97ER25308. P. N. acknowledges start-up funding from the Harvard John A. Paulson School of Engineering and Applied Sciences. T. C. acknowledges support from the Danish Council for Independent Research (Grant No.\ DFF--6108-00667). The authors thank Joshua Caldwell (Vanderbilt), Dmitri Basov (Columbia), Ido Kaminer (Technion), Siyuan Dai (UT Austin), and Samuel Moore (Columbia) for helpful discussions. This work was supported by the DOE Photonics at Thermodynamic Limits Energy Frontier Research Center under grant no. DE-SC0019140.

\bibliographystyle{apsrev4-1}
\bibliography{references}

\end{document}

### 2018-11-05 paper2.tex
abstract

\documentclass[aps,prb,twocolumn,
	groupedaddress,superscriptaddress,
	amsfonts,amssymb,amsmath,floatfix,
	citeautoscript]{revtex4-1}

\usepackage{graphicx}
\usepackage[centering,hmargin=20mm,tmargin=30mm,bmargin=25mm]{geometry}
\usepackage{multirow}
\usepackage{newtxtext}
\usepackage[cmintegrals]{newtxmath}

%----- References -----
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{colorlinks,
	linkcolor={blue!75!black!80!yellow},
	citecolor={blue!75!black!80!yellow},
	urlcolor={blue!75!black!80!yellow}
}

%----- Captions in sans font -----
\makeatletter
\renewcommand\@make@capt@title[2]{%
	\@ifx@empty\float@link{\@firstofone}{\expandafter\href\expandafter{\float@link}}%
	\sffamily{\textbf{#1}}\@caption@fignum@sep#2
}%
\renewcommand\figurename{Figure}
\makeatother

\thickmuskip=5mu plus 2mu minus 1mu  %binary relations (default, 5mu plus 5mu)
\medmuskip=4mu plus 2mu minus 2mu    %binary operations (default, 4mu plus 2mu minus 4mu)

\frenchspacing %Ensure that revTeX does not do "double spaces" after punctuation

\renewcommand{\Im}{\operatorname{Im}}
\renewcommand{\Re}{\operatorname{Re}}
\newcommand{\sgn}{\operatorname{sgn}}

\newcommand{\iu}{\mathrm{i}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\unitvec}[1]{\ensuremath{\hat{\mathbf{#1}}}}
\newcommand{\sub}[1]{\ensuremath{_{\textrm{#1}}}} %Upright multi-character subscript
\newcommand{\super}[1]{\ensuremath{^{\textrm{#1}}}} %Upright multi-character superscript

\newcommand{\HarvardSEAS}{John A. Paulson School of Engineering and Applied Sciences, Harvard University, Cambridge, MA, USA}
\newcommand{\MITPhy}{Department of Physics, Massachusetts Institute of Technology, Cambridge, MA, USA}

%\usepackage[usenames]{color}
%\newcommand{\edited}[1]{{\color{red} #1}}
\newcommand{\comment}[2]{\textcolor{blue!70!black}{\small\textsf{\textsuperscript{\textsc{\MakeLowercase{#1}}}[#2]}}}
\usepackage[normalem]{ulem}
\newcommand{\swap}[2]{\textcolor{red!70!black}{\sout{#1}}\textrightarrow\textcolor{green!50!black}{#2}}
\newcommand{\remove}[1]{\textcolor{red!70!black}{\sout{#1}}}
\newcommand{\inset}[1]{\textcolor{green!50!black}{#1}}
\begin{document}

\title{Electrodynamics of 2D phonon polaritons from first principles}

\author{Nicholas Rivera$^\perp$}\email{nrivera@seas.harvard.edu}\affiliation{\HarvardSEAS}\affiliation{\MITPhy}
\author{Jennifer Coulter$^\perp$}\affiliation{\HarvardSEAS}
\author{Thomas Christensen}\affiliation{\MITPhy}
\author{Prineha Narang}\email{prineha@seas.harvard.edu}\affiliation{\HarvardSEAS}

\date{\today}

\begin{abstract}
The ability to use photonic quasiparticles to control electromagnetic energy far below the diffraction limit is a defining paradigm in nanophotonics. An important recent development in this field is the measurement and manipulation of extremely confined phonon-polariton modes in polar dielectrics such as silicon carbide and hexagonal boron nitride, which pave the way for nanophotonics and extreme light-matter interactions in the mid-IR to THz frequency range. To further advance this promising field, it is of great interest to predict the optical response of 2D polaritonic materials. Here we develop a unified framework based on quantum linear response theory to calculate the spatially non-local dielectric function of a polar lattice in arbitrary dimensions. Our results also allow us, for the first time, to derive the dielectric function of phonon polaritons in a monolayer, which may also be specified in terms of the same \emph{ab initio}-derived parameters of the underlying lattice. Finally, to show the generality of the framework developed here, we derive the electrodynamics of two-dimensional phonon polaritons, finding a universal form for their electromagnetic modes and dispersion relation, which can be specified in terms of \emph{ab initio} parameters.
\end{abstract}

\maketitle

Phonon polaritons, hybrid quasiparticles of photons and optical phonons, offer great promise for deeply sub-diffractional control of electromagnetic fields at mid-IR and THz frequencies. Phonon polaritons share many features in common with plasmon polaritons in conductors. In recent years, it has been shown that phonon polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra}. Due to this remarkable confinement and their relatively high lifetimes of around picoseconds, phonon polaritons open new opportunities for vibrational spectroscopy, radiative heat transfer \cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters \cite{kumar2015tunable,rivera2017making,kurman2018control}. The core features of phonon polaritons, such as mode shape, confinement, and propagation characteristics, are understood from simple Lorentz oscillator models of the dielectric function, enabling successful theoretical accounts of experimental observations.

we develop a first-principles theory of phonon polaritons in a 2D polar material, specifying the dielectric function in terms of \emph{ab initio} parameters, and finding a universal form for the dispersion relation of 2D phonon polaritons when the wavevector is much smaller than the size of the unit cell.

\
\section{Linear response of phonon polaritons in two dimensions}
We now move to generalize the derivation of the previous section to describe the optical phonon contribution to Maxwell's equations coming from a two-dimensional material. 
Unlike the previous section, we will not demonstrate the results of the calculation on a particular material.
Here, we are only interested in finding an expression for the optical response of the surface in terms of \emph{ab initio}-derivable parameters, as well as the general form of the dispersion relation of phonon-polaritons in 2D.
Because of the change of symmetry in the system, we will not be able to perform the analysis by Fourier transforming Maxwell's equations in three dimensions.
However, we will apply a technique which is often used to analyze electromagnetic modes in 2D systems such as plasmons in 2D conductors.
An example of the technique is shown in detail in Ref.~\citenum{jablan2009plasmonics}, but we summarize the essential details here.

Consider a two-dimensional ionic lattice located in the $xy$ plane ($z=0$) surrounded by a homogeneous environment of dielectric function $\epsilon_{\mathrm{env}}$.
We consider the simplified case where the lattice is free-standing, to show the essential concept.
We search for a mode whose magnetic field $\mathbf{H}$ is of the form $\mathbf{H}(z)\e^{\iu\mathbf{q}\cdot\mathbf{r}_{\parallel} - \iu\omega t}$, where $\mathbf{r}_{\parallel}=(x,y,0)$ is the position in the plane, $\mathbf{q}$ is the wavevector and $\omega_{\mathbf{q}}$ is the frequency.
To simplify the discussion even further and furnish analytical forms for the phonon-polariton modes in 2D, we assume this material has mirror symmetry along the axis of $\mathbf{q}$.
In this case, the electromagnetic modes can be decomposed into TE and TM-polarized modes.
As is typical in the study of highly-confined polaritons, whether they are plasmon or phonon polaritons, the TM mode is the highly confined one, and occurs when $\epsilon(\omega) < 0$.
In our convention, the TM mode is such that $\mathbf{H}$ is in the plane of the lattice and perpendicular to $\mathbf{q}$. 
Without loss of generality, define the $\mathbf{q}$ direction to be the $x$ direction, making the magnetic field in the $y$ direction.

For $z \neq 0$, the Maxwell equation for the magnetic field is $(\nabla\times\nabla\times - \epsilon_{\mathrm{env}}\frac{\omega^2}{c^2})\mathbf{H}(z)\e^{\iu\mathbf{q}\cdot\mathbf{r}_{\parallel}} = 0$.
For the TM polarized mode, $\mathbf{H}(z) = H(z)\unitvec{y}$; the vectorial wave equation consequently simplifies to a scalar one
\begin{equation}\label{eq:2dmaxwell}
\left(-\frac{\dd^2}{\dd{}z^2}+q^2-\epsilon_{\mathrm{env}}\frac{\omega^2}{c^2} \right)H(z) = 0.
\end{equation}
Seeking a surface-bound mode, $H(z)$ must assume the functional form
	\comment{tc}{Nick, please check (I reverted an earlier edit of mine of this to the stuff below, but I cannot make sense of it): how can we specify $H(z)$ by equality sign here; and then in the following paragraphs we say ``To join these two solutions consistently, we apply \ldots$\unitvec{z}\times(\mathbf{H}^{(+)}-\mathbf{H}^{(-)}) = \mathbf{K} = -\iu\omega\mathbf{P}_{\mathrm{s}}$''?. I cannot make sense of this given that we use equality signs here and that $\mathbf{H}(z) = H(z)\unitvec{y}$.}
\begin{align}
	H(z) = \sgn(z)\e^{-\kappa |z|},
\end{align}
with $\kappa \equiv \sqrt{q^2-\epsilon_{\mathrm{env}}\frac{\omega^2}{c^2}}$.

To join these two solutions consistently, we apply the interface conditions for the magnetic field. They are $\unitvec{z}\times(\mathbf{H}^{(+)}-\mathbf{H}^{(-)}) = \mathbf{K} = -\iu\omega\mathbf{P}_{\mathrm{s}}$ where $\mathbf{K}$ is the surface current density, expressed through the surface polarization density $\mathbf{P}_{\mathrm{s}}$. 
Now we use the linear response relation $\mathbf{P}(\mathbf{q},\omega) = \boldsymbol{\Pi}(\mathbf{q},\omega)\mathbf{E}(\mathbf{q},\omega)$ in addition to Ampere's law $\nabla\times\mathbf{H} = -\iu\omega\epsilon_{\mathrm{env}}\mathbf{E}$.
	\comment{tc}{There's a slight disconnect in that we never define the relation between $\mathbf{P}_{\mathrm{s}}$ and $\mathbf{P}$. Also, there's a bigger issue, I think: if we allow $\boldsymbol{\Pi}$ to be a $3\times 3$ matrix, then we ought to adopt more generic boundary conditions. Specifically, if there's out-of-plane ``current'', then the boundary conditions for normal parts of $\mathbf{B}$ and parallel parts of $\mathbf{E}$ would change.}
Plugging these relations in yields an implicit equation connecting the polarization response function, the frequency of the 2D phonon polariton, and its frequency. Taking the case where $x$ coincides with a principal axis of the system, we have that 
\begin{equation}\label{eq:2ddispersion}
q\Pi_{xx}(q,\omega)  = -2\epsilon_{\mathrm{env}}.
\end{equation}

To connect this relation to the discussion of the 3D case, we write a microscopic form for the polarization response function of the ionic lattice. It is simply
\begin{equation}\label{eq:2dsusceptibility}
\boldsymbol{\Pi}(\mathbf{q},\omega) =  \frac{1}{\epsilon_0 A}\sum\limits_{m,n}\frac{\mathbf{P}_{mn}(\mathbf{q})\otimes\mathbf{P}_{nm}(\mathbf{q})}{\hbar\omega + E_{nm}+\iu 0^+}\left(\e^{\beta E_m}-\e^{\beta E_n} \right),
\end{equation}
is the 2D Fourier transform of the polarization density. As the form of the displacement of the ions is still given by Equation~\eqref{eq:phononfield}, $\mathbf{P}_{mn}(\mathbf{q})$ is still given by Equations~(\ref{eq:fourierpolarization}--\ref{eq:oscillatorstrength}), but with the Born charges and eigendisplacements appropriate to the 2D system of interest. Note that although the lattice is two dimensional, in general, the ions may be displaced in any three directions and thus $\boldsymbol{\Pi}$ is still a $3\times3$ matrix.
Following the same kind of reasoning that lead to Equation~\eqref{eq:lorentzoscillator}, if we parameterize the effect of the different masses and Born charges in the unit cell by $M_{\mathrm{eff}}$ and $Q_{\mathrm{eff}}$, we will find a form for the polarization response function which is given by the second term of Equation~\eqref{eq:lorentzoscillator} but with $V$ replaced by $A$. Defining $N/A \equiv n_{\mathrm{s}}$, Equation~\eqref{eq:2dmaxwell} for the dispersion relation of 2D phonon polaritons tells us that
\begin{equation}\label{eq:2dphononpolaritondispersion}
\omega = \sqrt{\omega_{\mathrm{TO}}^2+\frac{n_{\mathrm{s}}Q_{\mathrm{eff}}^{2}}{2M_{\mathrm{eff}}\epsilon_0\epsilon_{\mathrm{env}}}q}.
\end{equation}
Note that this dispersion has a strong resemblance to the plasmon dispersion relation of two-dimensional free electron gases, which is  $\omega = \sqrt{n_{\mathrm{s}}e^2q/2m\epsilon_0\epsilon_{\mathrm{env}}}$  in the local limit~\cite{stern1967polarizability}.
The only difference is the replacement of electron parameters (density, charge, mass) by the corresponding lattice parameters in the phononic case, and a shift in the minimum frequency by the transverse optical phonon frequency.

We note that the dispersion relation of Equation~\eqref{eq:2dphononpolaritondispersion} can in fact be obtained by taking the infinitely thin-limit of a bulk polar dielectric, and finding the dispersion of the resulting phonon polaritons in the quasi-electrostatic limit.
In particular, for an isotropic polar dielectric of thickness $2d$ with boundaries at $z=\pm d$, solving Laplace's equations yields as the dispersion of the even parity mode:
\begin{equation}\label{eq:thinfilmphononpolaritondispersion}
\tanh(qd) = -\frac{\epsilon_{\mathrm{env}}}{\epsilon(\omega)} \rightarrow qd\epsilon(\omega) = -\epsilon_{\mathrm{env}},
\end{equation}
with $\epsilon_{\mathrm{env}}$ the dielectric constant of the surrounding environment and $\epsilon(\omega)$ that of the polar dielectric.
The right-hand side is obtained in the limit of $d\rightarrow 0$.
Taking Equation~\eqref{eq:lorentzoscillator} for the dielectric function in bulk, we have that (in the absence of losses) $\epsilon_{\mathrm{env}} + qd\epsilon_{\infty} + q\frac{n_sQ_{\mathrm{eff}}^{2}}{2\epsilon_0 M_{\mathrm{eff}}}\frac{1}{\omega_{\mathrm{TO}}^2-\omega^2} = 0$, where we have identified $V = 2Ad$.
In the limit of small $d$, i.e., $qd \ll 1$, we recover Equations~\eqref{eq:2ddispersion} and \eqref{eq:2dphononpolaritondispersion} with $\Pi = \frac{n_sQ_{\mathrm{eff}}^{2}}{\epsilon_0 M}\frac{1}{\omega^2_{\mathrm{TO}}-\omega^2}$. It is interesting to see that in the 2D limit, the effect of the screening inside the material altogether vanishes. A similar phenomenon is known in the electrodynamics of plasmons in 2D materials, where the dielectric screening is known to be entirely provided by the environment and is equal to $\epsilon_{\mathrm{env}}$. A useful result which follows is that we may also derive another important figure of merit, the $p$-polarized reflectivity of a 2D polar dielectric, which is related to the local density of states of the polar dielectric, which quantifies the strength of coupling to external probes such as a nano-antenna or a quantum emitter.
By considering the infinitely thin limit of a polar dielectric with dielectric function $\epsilon(\omega)$ and defining $\epsilon(\omega) = \Pi/2d$, we immediately find that the $p$-polarized reflectivity is given by
\begin{equation}\label{eq:2dtmreflectivity}
r_p = \frac{1}{1+2\epsilon_{\mathrm{env}}/q\Pi}.
\end{equation}

We note that, in general, when considering bulk and 2D versions of a material, at the very least, the effective parameters $Q_{\mathrm{eff}}$ and $M_{\mathrm{eff}}$ will change as a result of changes to the phonon band structure. In other words, although the forms of the expressions for the dielectric function

 It is also quite possible that they altogether vanish in the 2D case as opposed to the 3D case. As a recently discovered example of such a phenomenon, it was found that the LO-TO splitting of some 2D polar dielectrics, such as hBN \cite{sohier2017breakdown} altogether vanishes, and only returns to its bulk value after a small but finite wavevector. This would then suggest that a description of phonon polaritons necessitates consideration of the spatially nonlocal dielectric function as in Equation (12), and that these polaritons moreover may only exist when their wavelengths are sufficiently short. Another important change, beyond the theory presented in this work is that the losses may also change drastically in the reduction of dimensions. For example, it was found in several experiments that the damping rate of the TO phonon in hBN nearly doubles upon going from bulk to 2D\cite{gorbachev2011hunting,tran2016quantum}, which may strongly suppress the ability to image the polariton with near-field probes.
%This is due to the small thickness where the dispersion of the phonon polariton is quite flat, and very high wavevector modes, corresponding to the maximum wavevector detectable by a near-field tip, occur at frequencies very close to the TO phonon frequency.
%Near the TO phonon-frequency however, the ratio $\Im\Pi/\Re\Pi$ is the highest, corresponding to very low quality-factor modes.
%In addition to these considerations, the values of the parameters may depend on the background material, which may change the phonon band structure.

We provided this thin-film derivation as a simple way to understand the content of Equations~\eqref{eq:2ddispersion} and \eqref{eq:2dphononpolaritondispersion}.
In general Equation~\eqref{eq:2dsusceptibility} should be used to calculate $\Pi$ and the corresponding dispersion.
We note that a similar situation appears in the treatment of 2D plasmons: namely that a relation like Equation~\eqref{eq:2ddispersion} appears, except with $\Pi$ the polarization-polarization response of a 2D electron gas.
And furthermore that one can also obtain the same equation by considering the ultra-thin limit of a 3D plasmonic material \cite{jablan2009plasmonics,jablan2013plasmons}.

\section{Summary and Outlook}

In summary, we have provided a theoretical framework based on linear response theory to calculate the phonon contribution to the dielectric function from first principles.
Our framework is rather versatile, allowing us to: use first principles calculations to get the dielectric function, predict how nonlocality enters the dielectric function, and predict the effect of reduced dimensionality.
A particularly interesting case to consider in future work would be to find a system where the optical phonons are drastically different from 3D to 2D.
Perhaps it is possible that there are some materials in which the 2D optical phonons experience lower losses due to a reduced scattering phase space.
The formalism we provide here may also be extended to understand phonon-polaritons in other more atypical reduced-dimensional settings, such as zero-dimensional settings in single emitters, i.e., `molecular phonon polaritons', in analogy to recent work on `molecular plasmons' \cite{manjavacas2013tunable,lauchner2015molecular}.
Notably, we go beyond oversimplifications of the Lorentz oscillator model in which the Born charges are treated as a single scalar quantity and can treat the influence and interplay of many phonon modes that contribute to the dielectric function.
We corroborated our approach through density functional theoretic calculations.

In future work, besides furnishing an \emph{ab initio} description of nonlocality and reduced-dimension phonon polaritons, there are a number of interesting questions that can be addressed by the framework provided here.
For example, we may consider situations such as the phonon polaritons associated with optical phonons at the interface of two materials.
Another system of interest would be to consider two nearby layers of material in which the optical phonons of each material strongly couple to each other. 
In that case, it would be relevant to evaluate how this strong coupling manifests itself in the infrared dielectric function, and ultimately the confinement and propagation of the phonon polaritons.

\section{Acknowledgements}
This research used resources of the National Energy Research Scientific Computing Center, a DOE Office of Science User Facility supported by the Office of Science of the U.S. Department of Energy under Contract No. DE-AC02-05CH11231. 
N. R. and J. C. recognize the support of the DOE Computational Science Graduate Fellowship (CSGF) fellowship no. DE-FG02-97ER25308. 
P. N. acknowledges start-up funding from the Harvard John A. Paulson School of Engineering and Applied Sciences. 
T. C. acknowledges support from the Danish Council for Independent Research (Grant No.\ DFF--6108-00667).
%The authors thank XXXX for helpful discussions. Tentatively Josh Caldwell, Jeremy Baumberg, Dirk England, etc.

\bibliographystyle{apsrev4-1}
\bibliography{references}

\end{document}

### 2018-11-07 paper.tex
most of theory

@@
 
-\title{Theory of interactions between matter and optical phonons in two dimensions}
+\title{Nanophotonics with optical phonons in two dimensions}
 
@@
 In this section, we develop the theory of electromagnetic waves associated with optical phonons in two dimensions.
-A theory of electromagnetic waves in polar 2D materials involves first calculating the dielectric permittivity associated with the monolayer.
+The crux of the theory isdetermining the dielectric permittivity associated with the monolayer.
+It is obtained by considering the response of the ions in the monolayer to a longitudinal electromagnetic wave whose electric field is the gradient of a scalar potential $\phi$.
+The coupling Hamiltonian is then given by
+\begin{equation}
+H_{\mathrm{int}} = \int d^2x \rho \phi = -\int d^2x (\nabla\cdot\mathbf{P})\phi,
+\end{equation} 
+with $\rho$ the charge density and $\mathbf{P}$ the polarization density associated with the ionic motion.
+The polarization can be expanded in powers of the displacement of the ions.
+In particular, we can consider the polarization $P_i$ generated in the monolayer when atom $\kappa$ of the unit cell is displaced along the $j$ direction by amount $u_{\kappa,j}$.
+To lowest order:
+\begin{equation}
+P_i - P_i^{0} = \Omega\frac{\partial P_i}{\partial u_{\kappa,j}}u_{\kappa,j} \equiv Z_{kappa,ij}u_{j},
+\end{equation} 
+where $Z_{\kappa,ij}$ is the Born effective charge tensor of ion $\kappa$ and $\Omega$ is the unit cell area.
+With this relation between polarization and ionic displacements, we have expressed the Hamiltonian as a coupling between the scalar potential and ionic displacements.
+We consider the response of the monolayer to a scalar potential of the form $\phi = \phi(\mathbf{q},\omega)e^{i\mathbf{q}\cdot\mathbf{r}-i\omega t}$, where $\mathbf{q}$ is a two-dimensional wavevector in the plane of the monolayer.
+Such a potential corresponds to a longitudinal electric field of the form $\mathbf{E} = i\mathbf{q}\phi(\mathbf{q},\omega)e^{i\mathbf{q}\cdot\mathbf{r}-i\omega t}$.
+It follows that the induced polarization takes the form $P_i  = \epsilon_0\Pi_{ij}(\mathbf{q},\omega)E_{j,\mathrm{tot}}(\mathbf{q},\omega)$, where $\Pi_{ij}(\mathbf{q},\omega)$ is the polarization-polarization response function of the monolayer and  $E_{j,mathrm{tot}}(\mathbf{q},\omega)$ is taken as the \textit{total} field which is a sum of the applied field and the field created by induced polarization.
+The polarization-polarization response function is given by
+\begin{equation}\label{eq:2dsusceptibility}
+\boldsymbol{\Pi}(\mathbf{q},\omega) =  \frac{1}{\epsilon_0 \Omega}\sum\limits_{m,n}\frac{\mathbf{P}_{mn}(\mathbf{q})\otimes\mathbf{P}_{nm}(\mathbf{q})}{\hbar\omega + E_{nm}+\iu 0^+}\left(\e^{\beta E_m}-\e^{\beta E_n} \right),
+\end{equation}
+where $m,n$ are states of the phononic Fock space of the monolayer, $P_{i,mn} = \sum\limits_{\kappa,ij}\langle m | u_{\kappa,j} | n \rangle$ are matrix elements of the polarization associated with phonon modes, $E_{m (n)}$ is the energy of state $m (n)$, and $\beta$ is the inverse temperature.
+Evaluating the contribution of optical phonons in the harmonic approximation to the polarization-polarization response, one finds that the response function is given by:
+\begin{equation}
+\Pi_{qq}(\mathbf{q},\omega) = \frac{1}{\epsilon_0\hbar\Omega} \frac{2\omega_{\mathbf{q}L}}{\omega^2_{\mathbf{q}L}-\omega^2}|\hat{q}\cdot\langle 1_{\mathbf{q}L}|\mathbf{P}(\mathbf{q})|0_{\mathbf{q}\sigma}\rangle|^2,
+\end{equation}
+where 'L' denotes longitudinal polarization, $|0_{\mathbf{q}L}\rangle$ ($|1_{\mathbf{q}L}\rangle$) denotes a state with no (one) longitudinal phonon of wavevector $\mathbf{q}$.
+The frequency $\omega_{\mathbf{q}L}$ in the denominator, as in the case of bulk phonons, is the frequency of the longitudinal phonon of wavevector $\mathbf{q}$ prior to considering LO-TO splitting [Needs reference!].
+This is the consistent with the fact that LO-TO splitting is a collective effect arising from Coulomb interactions and the fact that the equation above represents a single-particle susceptibility.
+Coulomb interactions are accounted for in the random phase approximation, and to include them in the single-particle response would be double-counting.
+The component $qq$ in the response tensor denotes a pair of directions parallel to the wavevector.
+To relate the polarization-polarization response function to the electromagnetic modes supported by a polar monolayer, we solve Maxwell's equations for an evanescent electromagnetic mode supported by a surface with polarization-polarization response tensor $\mathbf{\Pi}$.
+We consider the monolayer to be sandwiched by a superstrate of permittivity $\epsilon_1$ and a substrate of permittivity $\epsilon_2$.
+
+To strip the analysis to its bare essentials, we consider optical phonon response with in-plane isotropy in the long-wavelength limit arising from in-plane LO oscillations.
+A relevant example of a system where these conditions are satisfied is in a hexagonal boron nitride monolayer, considering optical phonons in the so-called upper Reststrahlen band which in bulk spans the frequency range of 1360 to 1610 cm$^{-1}$.
+In a monolayer geometry with translation invariance and in-plane isotropy, the electromagnetic modes of Maxwell's equations can be composed into transverse magnetic (TM) and transverse electric (TE), where the magnetic or electric field respectively is transverse to the in-plane wavevector of the mode.
+In practice, it is the TM mode which is associated with highly confined electromagnetic waves that are of use in nanophotonics.
+TE waves are not supported at the same frequency as transverse magnetic waves, and require exotic conditions to be realized in two-dimensional materials.
+Given the isotropy, we may suppress indices from the response tensor.
+We may also consider without further loss of generality a TM mode with wavenumber $q$ along the $x$-direction in the monolayer and magnetic field $H(z)e^{iqx-i\omega t}$ along the $y$-direction of the monolayer.
+The direction transverse to the monolayer is denoted as $z$.
+With these definitions in place, the Maxwell equation satisfied by the magnetic field is \begin{equation}\label{eq:2dmaxwell}
+\left(-\frac{\dd^2}{\dd{}z^2}+q^2-\epsilon_{\mathrm{env}}\frac{\omega^2}{c^2} \right)H(z) = 0.
+\end{equation}
+We consider a solution of the form $H(z) = e^{-\kappa_1 z}$ for $z > 0$ with $\kappa_1 = \sqrt{q^2-\epsilon_1\frac{\omega^2}{c^2}}$ and a solution of the form $H(z) = -ce^{\kappa_2 z}$ for $z < 0$ with $\kappa_2 = \sqrt{q^2-\epsilon_2\frac{\omega^2}{c^2}}$.
+Continuity of the displacement field across the interface forces $c = \frac{\epsilon_1}{\epsilon_2}$.
+Meanwhile, the boundary condition on the magnetic field is $\unitvec{z}\times(\mathbf{H}^{(+)}-\mathbf{H}^{(-)}) = \mathbf{K} = -\iu\omega\mathbf{P}_{\mathrm{s}}$ where $\mathbf{K}$ is the surface current density, expressed through the surface polarization density $\mathbf{P}_{\mathrm{s}}$.
+
+The condition for an evanescent electromagnetic wave can be written as $\epsilon_{\mathrm{RPA}}(\omega) = 0$, where $\epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}+\frac{1}{2}q\Pi(\mathbf{q},\omega)$ coincides with the phonon contribution to the longitudinal dielectric function of the monolayer in the random-phase approximation.
 

### 2018-11-07 paper.tex
more theory section

@@
 \begin{equation}
-P_i - P_i^{0} = \Omega\frac{\partial P_i}{\partial u_{\kappa,j}}u_{\kappa,j} \equiv Z_{kappa,ij}u_{j},
+P_i - P_i^{0} = \Omega\frac{\partial P_i}{\partial u_{\kappa,j}}u_{\kappa,j} \equiv Z_{\kappa,ij}u_{j},
 \end{equation} 
@@
 Such a potential corresponds to a longitudinal electric field of the form $\mathbf{E} = i\mathbf{q}\phi(\mathbf{q},\omega)e^{i\mathbf{q}\cdot\mathbf{r}-i\omega t}$.
-It follows that the induced polarization takes the form $P_i  = \epsilon_0\Pi_{ij}(\mathbf{q},\omega)E_{j,\mathrm{tot}}(\mathbf{q},\omega)$, where $\Pi_{ij}(\mathbf{q},\omega)$ is the polarization-polarization response function of the monolayer and  $E_{j,mathrm{tot}}(\mathbf{q},\omega)$ is taken as the \textit{total} field which is a sum of the applied field and the field created by induced polarization.
+It follows that the induced polarization takes the form $P_i  = \epsilon_0\Pi_{ij}(\mathbf{q},\omega)E_{j,\mathrm{tot}}(\mathbf{q},\omega)$, where $\Pi_{ij}(\mathbf{q},\omega)$ is the polarization-polarization response function of the monolayer and  $E_{j,\mathrm{tot}}(\mathbf{q},\omega)$ is taken as the \textit{total} field which is a sum of the applied field and the field created by induced polarization.
 The polarization-polarization response function is given by
@@
 \end{equation}
-where $m,n$ are states of the phononic Fock space of the monolayer, $P_{i,mn} = \sum\limits_{\kappa,ij}\langle m | u_{\kappa,j} | n \rangle$ are matrix elements of the polarization associated with phonon modes, $E_{m (n)}$ is the energy of state $m (n)$, and $\beta$ is the inverse temperature.
+where $m,n$ are states of the phononic Fock space of the monolayer, $P_{i,mn} = \sum\limits_{\kappa,j}\langle m | u_{\kappa,j} | n \rangle$ are matrix elements of the polarization associated with phonon modes, $E_{m (n)}$ is the energy of state $m$ $(n)$, and $\beta$ is the inverse temperature.
 Evaluating the contribution of optical phonons in the harmonic approximation to the polarization-polarization response, one finds that the response function is given by:
@@
 \end{equation}
-We consider a solution of the form $H(z) = e^{-\kappa_1 z}$ for $z > 0$ with $\kappa_1 = \sqrt{q^2-\epsilon_1\frac{\omega^2}{c^2}}$ and a solution of the form $H(z) = -ce^{\kappa_2 z}$ for $z < 0$ with $\kappa_2 = \sqrt{q^2-\epsilon_2\frac{\omega^2}{c^2}}$.
-Continuity of the displacement field across the interface forces $c = \frac{\epsilon_1}{\epsilon_2}$.
-Meanwhile, the boundary condition on the magnetic field is $\unitvec{z}\times(\mathbf{H}^{(+)}-\mathbf{H}^{(-)}) = \mathbf{K} = -\iu\omega\mathbf{P}_{\mathrm{s}}$ where $\mathbf{K}$ is the surface current density, expressed through the surface polarization density $\mathbf{P}_{\mathrm{s}}$.
+We consider a solution of the form $H(z) = e^{-\kappa_1 z}$ for $z > 0$ with $\kappa_1 = \sqrt{q^2-\epsilon_1\frac{\omega^2}{c^2}}$ and a solution of the form $H(z) = ce^{\kappa_2 z}$ for $z < 0$ with $\kappa_2 = \sqrt{q^2-\epsilon_2\frac{\omega^2}{c^2}}$.
+The boundary condition on the magnetic field is $H_y^{(+)}-H_y^{(-)} = -K_x = \iu\omega\mathbf{P}_x$ where $\mathbf{K}$ is the surface current density, expressed through the polarization density $\mathbf{P}$ of the monolayer.
+This condition enforces $c =(1+\frac{\epsilon_0}{\epsilon_2} \kappa_2\Pi)^{-1}$.
+Continuity of the electric field in the $x$ direction enforces $\frac{\epsilon_2}{\epsilon_1} = -\frac{\kappa_2}{\kappa_1}c$.
+These two conditions fully specify the TM mode.
+To simplify the discussion, we consider the so-called electrostatic limit, it in which the wavenumber $q$ of the TM mode is much larger than the free-space wavenumber $\frac{\omega}{c}$.
+This is well-satisfied in monolayer optical phonons in 2D.
+With this approximation, the condition for a TM mode corresponding to an optical phonon is
+\begin{equation}
+\epsilon_{\mathrm{env}} + \frac{1}{2}\epsilon_0 q \Pi(q,\omega) \equiv \epsilon_{\mathrm{RPA}}(\omega) = 0.
+\end{equation}
+The quantity $\epsilon_{\mathrm{RPA}}(\omega) = 0$ coincides with the phonon contribution to the longitudinal dielectric function of the monolayer in the random-phase approximation.
 
-The condition for an evanescent electromagnetic wave can be written as $\epsilon_{\mathrm{RPA}}(\omega) = 0$, where $\epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}+\frac{1}{2}q\Pi(\mathbf{q},\omega)$ coincides with the phonon contribution to the longitudinal dielectric function of the monolayer in the random-phase approximation.
+To proceed, we express the polarization matrix element in Equation (4) in terms of parameters of the monolayer such as the Born effective charges.
+Considering the longitudinal phonon contribution to the second-quantized ionic displacement, as in [Needs reference!], one finds that the RPA dielectric function is given by:
+\begin{equation}
+\Pi_{qq}(\mathbf{q},\omega) = \frac{1}{\epsilon_0\hbar\Omega} \frac{2\omega_{\mathbf{q}L}}{\omega^2_{\mathbf{q}TO}-\omega^2}|\hat{q}\cdot\langle 1_{\mathbf{q}L}|\mathbf{P}(\mathbf{q})|0_{\mathbf{q}\sigma}\rangle|^2.
+\end{equation}
+In this expression, $\omega_L$ and been re-named as $\omega_{TO}$, because in the absence of LO-TO splitting, they are degenerate.
+Additionally, we have defined eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}\equiv \frac{e_{\kappa\mathbf{q},L}}{\sqrt{M_{\kappa}}}$, where $e_{\kappa\mathbf{q},L}$ is the unit-normalized polarization vector of atom $\kappa$ in the unit cell oscillating according to a longitudinal phonon of wavevector $\mathbf{q}$ and $M_{\kappa}$ is the mass of atom $\kappa$.
+
+As is well known, the zeros of the RPA dielectric function correspond to longitudinal electromagnetic waves. Thus, the condition for the zeros of the dielectric function correspond to
+\begin{equation}
+\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},T} = \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa,j}Z_{\kappa,ij}\eta_{\kappa,j}  \Big|^2
+\end{equation}
 

### 2018-11-07 paper.tex
need to fill out one more section

@@
 
+In this Letter, we ...
+
 \section{Electrodynamics of optical phonons in two-dimensions}
@@
 \begin{equation}
-\Pi_{qq}(\mathbf{q},\omega) = \frac{1}{\epsilon_0\hbar\Omega} \frac{2\omega_{\mathbf{q}L}}{\omega^2_{\mathbf{q}TO}-\omega^2}|\hat{q}\cdot\langle 1_{\mathbf{q}L}|\mathbf{P}(\mathbf{q})|0_{\mathbf{q}\sigma}\rangle|^2.
+\epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}} - \frac{q}{2\epsilon_0}\frac{\Big|\hat{q}\cdot\sum\limits_{\kappa,j} Z_{\kappa,ij}\eta_{\kappa,j}  \Big|^2}{\omega^2_{\mathrm{TO}}-\omega^2}
 \end{equation}
-In this expression, $\omega_L$ and been re-named as $\omega_{TO}$, because in the absence of LO-TO splitting, they are degenerate.
+In this expression, $\omega_L$ and been re-named as $\omega_{\mathrm{TO}}$, because in the absence of LO-TO splitting, they are degenerate.
 Additionally, we have defined eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}\equiv \frac{e_{\kappa\mathbf{q},L}}{\sqrt{M_{\kappa}}}$, where $e_{\kappa\mathbf{q},L}$ is the unit-normalized polarization vector of atom $\kappa$ in the unit cell oscillating according to a longitudinal phonon of wavevector $\mathbf{q}$ and $M_{\kappa}$ is the mass of atom $\kappa$.
 
-As is well known, the zeros of the RPA dielectric function correspond to longitudinal electromagnetic waves. Thus, the condition for the zeros of the dielectric function correspond to
+As is well known, the zeros of the RPA dielectric function, denoted $\omega_{\mathbf{q}}$, correspond to longitudinal electromagnetic waves.
+Thus, the condition for the zeros of the dielectric function correspond to
+\begin{align}
+\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} &= \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa,j}Z_{\kappa,ij}\eta_{\kappa,j}  \Big|^2 \nonumber \\
+&= V(q)\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa,j}Z_{\kappa,ij}\eta_{\kappa,j}  \Big|^2,
+\end{align}
+where $V(q)$ is the Coulomb interaction in two dimensions, screened by the dielectric environment of the sub- and superstrates.
+This result, as we now show, reveals that the TM mode corresponds simply to the two-dimensional optical phonon.
+In particular, it was shown in [Sohier2017] that in two-dimensional polar materials, the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO-TO splitting given by Equation (8) of the current manuscript.
+\textit{Thus, one of the main results of our manuscript is that the wave solutions Maxwell's equations associated with phonon-polaritons in bulk materials, correspond to pure LO phonons in the monolayer limit.}
+
+We now re-write $\epsilon_{\mathrm{RPA}}$ in a way that is expressed purely in terms of the optical phonon dispersion, derive the conductivity of the monolayer, and derive a universal form for the dispersion of 2D LO phonons in the long-wavelength limit.
+From Equation (8), we may write Equation (7) as
 \begin{equation}
-\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},T} = \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa,j}Z_{\kappa,ij}\eta_{\kappa,j}  \Big|^2
+\epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}\left( 1 + \frac{\omega^2_{\mathbf{q}\mathrm{LO}}-\omega^2_{\mathbf{q}\mathrm{TO}}}{\omega^2_{\mathbf{q}\mathrm{TO}}-\omega^2}\right).
 \end{equation}
+Similarly, the conductivity is prescribed by $\sigma(\mathbf{q},\omega) = -\iu \omega \Pi(\mathbf{q},\omega)$.
+We are now in a position to derive a universal form for the RPA dielectric function of a 2D polar slab in terms of three phenomenological parameters.
+In many cases [ref], the Born charges are independent of $q$ for small $q$, and thus the dispersion can be written as
+\begin{equation}
+\omega_{\mathbf{q},LO} = \sqrt{\omega^2_{\Gamma \mathrm{TO}}+\mathcal{S}q} \approx \omega_{\Gamma \mathrm{TO}} + v_g q,
+\end{equation}
+where $\mathcal{S}$ is a microscopic constant determined by the Born charges of the monolayer, $v_g$ is the group velocity of the LO mode, and $\omega_{\Gamma \mathrm{TO}}$ is the TO phonon frequency at the $\Gamma$ point.
+Then, the universal form of the dispersion relation
+\begin{equation}
+\epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}\left( 1 + \frac{2v_g\omega_{\mathbf{q}\mathrm{TO}}}{\omega^2_{\mathbf{q}\mathrm{TO}}-\omega^2+\iu \omega \tau^{-1}}\right).
+\end{equation}
+where we have also phenomenologically included the phonon dissipation rate $\tau^{-1}$, consistently with  a relaxation-time prescription.
+Thus, all of the optical properties associated with LO phonons is captured by the TO phonon frequency, the slope of the LO dispersion at $\Gamma$, and the dissipation $\tau^{-1}$.
 
 \section{Strong light-matter interactions enabled by 2D optical phonons}
+
+We now move to analyze key figures of merit of 2D LO phonons.
 
@@
 %\begin{equation}\label{eq:2dphononpolaritondispersion}
-%\omega = \sqrt{\omega_{\mathrm{TO}}^2+\frac{n_{\mathrm{s}}Q_{\mathrm{eff}}^{2}}{2M_{\mathrm{eff}}\epsilon_0\epsilon_{\mathrm{env}}}q}.
+%\omega = \sqrt{\omega_{\mathrm{\mathrm{TO}}}^2+\frac{n_{\mathrm{s}}Q_{\mathrm{eff}}^{2}}{2M_{\mathrm{eff}}\epsilon_0\epsilon_{\mathrm{env}}}q}.
 %\end{equation}
@@
 %The right-hand side is obtained in the limit of $d\rightarrow 0$.
-%Taking Equation~\eqref{eq:lorentzoscillator} for the dielectric function in bulk, we have that (in the absence of losses) $\epsilon_{\mathrm{env}} + qd\epsilon_{\infty} + q\frac{n_sQ_{\mathrm{eff}}^{2}}{2\epsilon_0 M_{\mathrm{eff}}}\frac{1}{\omega_{\mathrm{TO}}^2-\omega^2} = 0$, where we have identified $V = 2Ad$.
-%In the limit of small $d$, i.e., $qd \ll 1$, we recover Equations~\eqref{eq:2ddispersion} and \eqref{eq:2dphononpolaritondispersion} with $\Pi = \frac{n_sQ_{\mathrm{eff}}^{2}}{\epsilon_0 M}\frac{1}{\omega^2_{\mathrm{TO}}-\omega^2}$.
+%Taking Equation~\eqref{eq:lorentzoscillator} for the dielectric function in bulk, we have that (in the absence of losses) $\epsilon_{\mathrm{env}} + qd\epsilon_{\infty} + q\frac{n_sQ_{\mathrm{eff}}^{2}}{2\epsilon_0 M_{\mathrm{eff}}}\frac{1}{\omega_{\mathrm{\mathrm{TO}}}^2-\omega^2} = 0$, where we have identified $V = 2Ad$.
+%In the limit of small $d$, i.e., $qd \ll 1$, we recover Equations~\eqref{eq:2ddispersion} and \eqref{eq:2dphononpolaritondispersion} with $\Pi = \frac{n_sQ_{\mathrm{eff}}^{2}}{\epsilon_0 M}\frac{1}{\omega^2_{\mathrm{\mathrm{TO}}}-\omega^2}$.
 It is interesting to see that in the 2D limit, the effect of the screening inside the material altogether vanishes.
@@
 % It is also quite possible that they altogether vanish in the 2D case as opposed to the 3D case.
-As a recently discovered example of such a phenomenon, it was found that the LO-TO splitting of some 2D polar dielectrics, such as hBN \cite{sohier2017breakdown} altogether vanishes, and only returns to its bulk value after a small but finite wavevector.
+As a recently discovered example of such a phenomenon, it was found that the LO-\mathrm{TO} splitting of some 2D polar dielectrics, such as hBN \cite{sohier2017breakdown} altogether vanishes, and only returns to its bulk value after a small but finite wavevector.
 This would then suggest that a description of phonon polaritons necessitates consideration of the spatially nonlocal dielectric function as in Equation (12), and that these polaritons moreover may only exist when their wavelengths are sufficiently short.
 Another important change, beyond the theory presented in this work is that the losses may also change drastically in the reduction of dimensions.
-For example, it was found in several experiments that the damping rate of the TO phonon in hBN nearly doubles upon going from bulk to 2D\cite{gorbachev2011hunting,tran2016quantum}, which may strongly suppress the ability to image the polariton with near-field probes.
-%%This is due to the small thickness where the dispersion of the phonon polariton is quite flat, and very high wavevector modes, corresponding to the maximum wavevector detectable by a near-field tip, occur at frequencies very close to the TO phonon frequency.
-%%Near the TO phonon-frequency however, the ratio $\Im\Pi/\Re\Pi$ is the highest, corresponding to very low quality-factor modes.
+For example, it was found in several experiments that the damping rate of the \mathrm{TO} phonon in hBN nearly doubles upon going from bulk to 2D\cite{gorbachev2011hunting,tran2016quantum}, which may strongly suppress the ability to image the polariton with near-field probes.
+%%This is due to the small thickness where the dispersion of the phonon polariton is quite flat, and very high wavevector modes, corresponding to the maximum wavevector detectable by a near-field tip, occur at frequencies very close to the \mathrm{TO} phonon frequency.
+%%Near the \mathrm{TO} phonon-frequency however, the ratio $\Im\Pi/\Re\Pi$ is the highest, corresponding to very low quality-factor modes.
 %%In addition to these considerations, the values of the parameters may depend on the background material, which may change the phonon band structure. 

### 2018-11-08 paper.tex
theory section done... intro populated

@@
 \begin{abstract}
-Extreme confinement of electromagnetic energy by phonon polaritons promises extremely strong and novel forms of control over the dynamics of matter.
-To bring such control to its ultimate limit, it is important to consider phonon polaritons in two dimensional systems.
-However, recent studies have pointed out that in two-dimensional systems, splitting between longitudinal and transverse (LO and TO) optical phonons, which is necessary for the existence of phonon polaritons in three dimensions, is absent.
-The question then arises as to whether the ability to exploit optical phonons disappears in lower dimensions.
-Here, we settle this question, finding that the phonon-polariton of the bulk is essentially replaced by the LO phonon.
-We present the confinement and propagation losses of LO phonons.
-We then calculate various measures of strong light-matter interaction such as single and two-photon spontaneous emission enhancement, and discuss methods such as EELS to probe these excitations.
+Extreme confinement of electromagnetic energy by phonon-polaritons promises extremely strong and novel forms of control over the dynamics of matter.
+To bring such control to its ultimate limit, it is important to consider phonon-polaritons in two dimensional systems.
+Recent studies have pointed out that in two-dimensional systems, splitting between longitudinal and transverse (LO and TO) optical phonons is absent at the $\Gamma$ point.
+Does this lack of LO-TO splitting imply the lack of a strongly confined electromagnetic (phonon-polariton) mode in polar monolayers?
+Here, we derive a universal form for the spatially-nonlocal dielectric function of a polar monolayer specified by: the TO phonon frequency, the non-zero group velocity of the LO phonon, and the phonon lifetime.
+Our analysis reveals that the phonon-polariton of the bulk is ``replaced" by the LO phonon.
+We present the confinement and propagation losses of LO phonons, finding that high LO phonon confinement and reasonable propagation quality factors coincide in regions which are difficult to detect with near-field optical microscopy, but may be more amenable to electron energy-loss spectroscopy.
+Finally, we show that 2D LO phonons in hBN present extremely strong light-matter interactions with external emitters, providing possibilities for 10,000,000-fold enhanced spontaneous emission and strong coupling with vibrational emitters.
 \end{abstract}
@@
 
-In this Letter, we ...
+Phonon polaritons, hybrid quasiparticles of photons and optical phonons supported in polar materials, offer great promise for deeply sub-diffractional control of electromagnetic fields at mid-IR and THz frequencies.
+Phonon polaritons share many features in common with plasmon polaritons in conductors.
+In recent years, it has been shown that phonon-polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra}.
+Due to this remarkable confinement and their relatively high lifetimes of around picoseconds, phonon-polaritons open new opportunities for vibrational spectroscopy, radiative heat transfer \cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters \cite{kumar2015tunable,rivera2017making,kurman2018control}.
+In many cases, extreme confinement of phonon polaritons is achieved by the use of thin-films, which shrink the in-plane and out-of-plane wavelength of polaritons monotonically with decreasing film thickness.
+In many cases, a monolayer is the ultimate limit of this effect.
+Thus, to bring these classical and quantum nanophotonics applications to their ultimate limits, it is necessary to consider the properties of these modes in two-dimensions.
+However, this picture of monolayers supporting maximally confined phonon-polariton modes is ostensibly complicated by the fact that in two-dimensional polar materials, the LO-TO splitting that gives rise to phonon-polaritons in three-dimensions is absent at the $\Gamma$ point.
+
+In this Letter, we develop core results in the nascent field of optical phononics in two-dimensional materials by deriving a universal form for the dielectric function of a polar monolayer, incorporating spatially nonlocal effects, which is specified in fully in terms of the dispersion relation of LO and TO phonons in two dimensions.
+Using these parameters for important monolayers such as hexagonal boron nitride, we present the confinement and propagation losses of these modes, identifying the frequency region where they should be most easily detected.
+We...
+[insert nicest results here].
 
@@
 In this section, we develop the theory of electromagnetic waves associated with optical phonons in two dimensions.
-The crux of the theory isdetermining the dielectric permittivity associated with the monolayer.
+The crux of the theory is determining the dielectric permittivity associated with the monolayer.
 It is obtained by considering the response of the ions in the monolayer to a longitudinal electromagnetic wave whose electric field is the gradient of a scalar potential $\phi$.
@@
 As is well known, the zeros of the RPA dielectric function, denoted $\omega_{\mathbf{q}}$, correspond to longitudinal electromagnetic waves.
-Thus, the condition for the zeros of the dielectric function correspond to
+Thus, the condition for the zeros of the dielectric function can be expressed as
 \begin{align}
-\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} &= \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa,j}Z_{\kappa,ij}\eta_{\kappa,j}  \Big|^2 \nonumber \\
-&= V(q)\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa,j}Z_{\kappa,ij}\eta_{\kappa,j}  \Big|^2,
+\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} &= \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2 \nonumber \\
+&= V(q)\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2,
 \end{align}
-where $V(q)$ is the Coulomb interaction in two dimensions, screened by the dielectric environment of the sub- and superstrates.
+where $eZ^r = Z$, with $e$ the electron charge, and $V(q)$ is the Coulomb interaction in two dimensions, screened by the dielectric environment of the sub- and superstrates.
 This result, as we now show, reveals that the TM mode corresponds simply to the two-dimensional optical phonon.
@@
 \begin{equation}
-\omega_{\mathbf{q},LO} = \sqrt{\omega^2_{\Gamma \mathrm{TO}}+\mathcal{S}q} \approx \omega_{\Gamma \mathrm{TO}} + v_g q,
+\omega_{\mathbf{q},\mathrm{LO}} = \sqrt{\omega^2_{\Gamma \mathrm{TO}}+\mathcal{S}q} \approx \omega_{\Gamma \mathrm{TO}} + v_g q,
 \end{equation}
 where $\mathcal{S}$ is a microscopic constant determined by the Born charges of the monolayer, $v_g$ is the group velocity of the LO mode, and $\omega_{\Gamma \mathrm{TO}}$ is the TO phonon frequency at the $\Gamma$ point.
-Then, the universal form of the dispersion relation
+The group velocity is specified in terms of microscopic parameters via
 \begin{equation}
-\epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}\left( 1 + \frac{2v_g\omega_{\mathbf{q}\mathrm{TO}}}{\omega^2_{\mathbf{q}\mathrm{TO}}-\omega^2+\iu \omega \tau^{-1}}\right).
+v_g = \frac{e^2 }{4\epsilon_0 \epsilon_{\mathrm{env}}\omega_{\Gamma TO}\Omega}\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2. 
+\end{equation}
+Thus, in the long wavelength limit, we have the following universal parameterization of the RPA dielectric function of a polar monolayer:
+\begin{equation}
+\epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}\left( 1 + \frac{2v_g\omega_{\mathbf{q}\mathrm{TO}}}{\omega^2_{\mathbf{q}\mathrm{TO}}-\omega^2+\iu \omega \tau^{-1}}\right),
 \end{equation}
@@
 Thus, all of the optical properties associated with LO phonons is captured by the TO phonon frequency, the slope of the LO dispersion at $\Gamma$, and the dissipation $\tau^{-1}$.
+
+It has been well-appreciated in the literature on phonon polaritons in thin slabs.
+Here, we expand on the analogy in the two-dimensional case.
+To do so, we take Equation (8) in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $e^2\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2 \equiv  \left(\frac{eZ^r_*}{\sqrt{M_*}}\right)^2 \equiv \left(\frac{Q_*}{\sqrt{M_*}}\right)^2$.
+Then, the LO-TO splitting can be written as $\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} = \frac{Q_*^2}{2\epsilon_0\epsilon_{\mathrm{env}} M_*}q$.
+Now we note that the RHS is exactly the squared-frequency $\omega^2_{\mathbf{q}p}$ for a plasma oscillation in a 2D gas of charged particles with charge $Q_*$ and mass $M_*$.
+This squared frequency can be thought of as the "LP-TP" splitting between longitudinal and transverse plasma oscillations.
+Of course, there are no transverse plasma oscillations, and so ``$\omega_{\mathrm{TP}} = 0$".
+In the three-dimensional plasmon case, `$\omega_{\mathrm{TP}} = 0$", but the difference between the squared longitudinal and transverse plasma oscillation frequencies at zero-wavevector is non-zero and given by $\omega_p^2$.
+This analogy between phononic and plasmonic behavior as a function of dimension is illustrated schematically in Figure 1.
 
@@
 Phonon polaritons share many features in common with plasmon polaritons in conductors.
-In recent years, it has been shown that phonon polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra}.
-Due to this remarkable confinement and their relatively high lifetimes of around picoseconds, phonon polaritons open new opportunities for vibrational spectroscopy, radiative heat transfer \cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters \cite{kumar2015tunable,rivera2017making,kurman2018control}.
-The core features of phonon polaritons, such as mode shape, confinement, and propagation characteristics, are understood from simple Lorentz oscillator models of the dielectric function, enabling successful theoretical accounts of experimental observations.
+In recent years, it has been shown that phonon-polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra}.
+Due to this remarkable confinement and their relatively high lifetimes of around picoseconds, phonon-polaritons open new opportunities for vibrational spectroscopy, radiative heat transfer \cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters \cite{kumar2015tunable,rivera2017making,kurman2018control}.
+The core features of phonon-polaritons, such as mode shape, confinement, and propagation characteristics, are understood from simple Lorentz oscillator models of the dielectric function, enabling successful theoretical accounts of experimental observations.
 %
-%%we develop a first-principles theory of phonon polaritons in a 2D polar material, specifying the dielectric function in terms of \emph{ab initio} parameters, and finding a universal form for the dispersion relation of 2D phonon polaritons when the wavevector is much smaller than the size of the unit cell.
+%%we develop a first-principles theory of phonon-polaritons in a 2D polar material, specifying the dielectric function in terms of \emph{ab initio} parameters, and finding a universal form for the dispersion relation of 2D phonon-polaritons when the wavevector is much smaller than the size of the unit cell.
 %
 %\
-%%\section{Linear response of phonon polaritons in two dimensions}
+%%\section{Linear response of phonon-polaritons in two dimensions}
 %%We now move to generalize the derivation of the previous section to describe the optical phonon contribution to Maxwell's equations coming from a two-dimensional material. 
@@
 %In this case, the electromagnetic modes can be decomposed into TE and TM-polarized modes.
-%As is typical in the study of highly-confined polaritons, whether they are plasmon or phonon polaritons, the TM mode is the highly confined one, and occurs when $\epsilon(\omega) < 0$.
+%As is typical in the study of highly-confined polaritons, whether they are plasmon or phonon-polaritons, the TM mode is the highly confined one, and occurs when $\epsilon(\omega) < 0$.
 %In our convention, the TM mode is such that $\mathbf{H}$ is in the plane of the lattice and perpendicular to $\mathbf{q}$. 
@@
 Specifically, if there's out-of-plane ``current'', then the boundary conditions for normal parts of $\mathbf{B}$ and parallel parts of $\mathbf{E}$ would change.}
-%Plugging these relations in yields an implicit equation connecting the polarization response function, the frequency of the 2D phonon polariton, and its frequency.
+%Plugging these relations in yields an implicit equation connecting the polarization response function, the frequency of the 2D phonon-polariton, and its frequency.
 Taking the case where $x$ coincides with a principal axis of the system, we have that
@@
 %Following the same kind of reasoning that lead to Equation~\eqref{eq:lorentzoscillator}, if we parameterize the effect of the different masses and Born charges in the unit cell by $M_{\mathrm{eff}}$ and $Q_{\mathrm{eff}}$, we will find a form for the polarization response function which is given by the second term of Equation~\eqref{eq:lorentzoscillator} but with $V$ replaced by $A$.
-Defining $N/A \equiv n_{\mathrm{s}}$, Equation~\eqref{eq:2dmaxwell} for the dispersion relation of 2D phonon polaritons tells us that
+Defining $N/A \equiv n_{\mathrm{s}}$, Equation~\eqref{eq:2dmaxwell} for the dispersion relation of 2D phonon-polaritons tells us that
 %\begin{equation}\label{eq:2dphononpolaritondispersion}
@@
 %
-%We note that the dispersion relation of Equation~\eqref{eq:2dphononpolaritondispersion} can in fact be obtained by taking the infinitely thin-limit of a bulk polar dielectric, and finding the dispersion of the resulting phonon polaritons in the quasi-electrostatic limit.
+%We note that the dispersion relation of Equation~\eqref{eq:2dphononpolaritondispersion} can in fact be obtained by taking the infinitely thin-limit of a bulk polar dielectric, and finding the dispersion of the resulting phonon-polaritons in the quasi-electrostatic limit.
 %In particular, for an isotropic polar dielectric of thickness $2d$ with boundaries at $z=\pm d$, solving Laplace's equations yields as the dispersion of the even parity mode:
@@
 As a recently discovered example of such a phenomenon, it was found that the LO-\mathrm{TO} splitting of some 2D polar dielectrics, such as hBN \cite{sohier2017breakdown} altogether vanishes, and only returns to its bulk value after a small but finite wavevector.
-This would then suggest that a description of phonon polaritons necessitates consideration of the spatially nonlocal dielectric function as in Equation (12), and that these polaritons moreover may only exist when their wavelengths are sufficiently short.
+This would then suggest that a description of phonon-polaritons necessitates consideration of the spatially nonlocal dielectric function as in Equation (12), and that these polaritons moreover may only exist when their wavelengths are sufficiently short.
 Another important change, beyond the theory presented in this work is that the losses may also change drastically in the reduction of dimensions.
 For example, it was found in several experiments that the damping rate of the \mathrm{TO} phonon in hBN nearly doubles upon going from bulk to 2D\cite{gorbachev2011hunting,tran2016quantum}, which may strongly suppress the ability to image the polariton with near-field probes.
-%%This is due to the small thickness where the dispersion of the phonon polariton is quite flat, and very high wavevector modes, corresponding to the maximum wavevector detectable by a near-field tip, occur at frequencies very close to the \mathrm{TO} phonon frequency.
+%%This is due to the small thickness where the dispersion of the phonon-polariton is quite flat, and very high wavevector modes, corresponding to the maximum wavevector detectable by a near-field tip, occur at frequencies very close to the \mathrm{TO} phonon frequency.
 %%Near the \mathrm{TO} phonon-frequency however, the ratio $\Im\Pi/\Re\Pi$ is the highest, corresponding to very low quality-factor modes.
@@
 %Perhaps it is possible that there are some materials in which the 2D optical phonons experience lower losses due to a reduced scattering phase space.
-%The formalism we provide here may also be extended to understand phonon-polaritons in other more atypical reduced-dimensional settings, such as zero-dimensional settings in single emitters, i.e., `molecular phonon polaritons', in analogy to recent work on `molecular plasmons' \cite{manjavacas2013tunable,lauchner2015molecular}.
+%The formalism we provide here may also be extended to understand phonon-polaritons in other more atypical reduced-dimensional settings, such as zero-dimensional settings in single emitters, i.e., `molecular phonon-polaritons', in analogy to recent work on `molecular plasmons' \cite{manjavacas2013tunable,lauchner2015molecular}.
 %Notably, we go beyond oversimplifications of the Lorentz oscillator model in which the Born charges are treated as a single scalar quantity and can treat the influence and interplay of many phonon modes that contribute to the dielectric function.
@@
 %
-%In future work, besides furnishing an \emph{ab initio} description of nonlocality and reduced-dimension phonon polaritons, there are a number of interesting questions that can be addressed by the framework provided here.
-%For example, we may consider situations such as the phonon polaritons associated with optical phonons at the interface of two materials.
+%In future work, besides furnishing an \emph{ab initio} description of nonlocality and reduced-dimension phonon-polaritons, there are a number of interesting questions that can be addressed by the framework provided here.
+%For example, we may consider situations such as the phonon-polaritons associated with optical phonons at the interface of two materials.
 %Another system of interest would be to consider two nearby layers of material in which the optical phonons of each material strongly couple to each other. 
-%In that case, it would be relevant to evaluate how this strong coupling manifests itself in the infrared dielectric function, and ultimately the confinement and propagation of the phonon polaritons.
+%In that case, it would be relevant to evaluate how this strong coupling manifests itself in the infrared dielectric function, and ultimately the confinement and propagation of the phonon-polaritons.
 

### 2018-11-09 paper.tex
discussion on anomalies in monolayer vs bulk

@@
 \end{equation}
-where $m,n$ are states of the phononic Fock space of the monolayer, $P_{i,mn} = \sum\limits_{\kappa,j}\langle m | u_{\kappa,j} | n \rangle$ are matrix elements of the polarization associated with phonon modes, $E_{m (n)}$ is the energy of state $m$ $(n)$, and $\beta$ is the inverse temperature.
+where $m,n$ are states of the phononic Fock space of the monolayer, $P_{i,mn} = \sum\limits_{\kappa,j}Z_{\kappa,ij}\langle m | u_{\kappa,j} | n \rangle$ are matrix elements of the polarization associated with phonon modes, $E_{m (n)}$ is the energy of state $m$ $(n)$, and $\beta$ is the inverse temperature.
 Evaluating the contribution of optical phonons in the harmonic approximation to the polarization-polarization response, one finds that the response function is given by:
@@
 Thus, all of the optical properties associated with LO phonons is captured by the TO phonon frequency, the slope of the LO dispersion at $\Gamma$, and the dissipation $\tau^{-1}$.
+In Table 1, we tabulate these parameters for a few polar monolayers whose phonon band-structures have been calculated in the literature: hBN, MoS$_2$, and ...
 
@@
 
+\section{Properties of 2D LO phonon modes in monolayer hBN}
+
+In Fig.
+2a, we present the dispersion relation, operationally defined through the poles of the Fresnel reflection coefficient for $p$-polarized waves, which for a monolayer of conductivity $\sigma(\mathbf{q},\omega) = -\frac{2\iu \omega}{\epsilon q}\left(\epsilon_{\mathrm{RPA}} - \epsilon_{\mathrm{env}} \right)$ below vacuum and on top of a substrate of permittivity $\epsilon_{\mathrm{s}}$, is given by
+\begin{equation}
+r_p(\mathbf{q},\omega) = \frac{(\epsilon_{\mathrm{s}}-1)-\frac{\sigma(\mathbf{q},\omega) q}{\omega\epsilon_0}}{(\epsilon_{\mathrm{s}}+1)-\frac{\sigma(\mathbf{q},\omega) q}{\omega \epsilon_0}}.
+\end{equation}
+In Figs.
+2b, we show the confinement factor (or effective mode index) of the 2D LO phonon mode, defined by $n_{\mathrm{eff}} = \frac{qc}{\omega}$.
+The confinement factor is a key figure of merit in many nanophotonic applications involving quantum emitters, because as was shown in [Rivera], all interactions involving emission of electromagnetic energy by emitters, such as electric and magnetic dipole emission, multipole emission, phosphorescence, and multiphoton emission, scale as some power of this confinement factor.
+Interestingly, in the case of a monolayer, the confinement grows very rapidly in frequency, due to the extremely low group velocity of 2D LO phonons, which is a remarkable four orders of magnitude slower than the speed of light.
+In particular, at frequencies of 1450 $cm^{-1}$, the LO phonon already has a confinement factor in excess of conf, corresponding to a phonon wavelength of 6896.55/conf nm, which is much shorter wavelength than any phonon polariton thus measured so far, as well as shorter wavelength than any plasmon that has been measured in the mid-IR, such as in graphene.
+In fact, this short wavelength is mostly out of the range of measurability in scattering near-field microscope (SNOM) measurements, which with current tip radii, are limited to measuring wavelengths in excess of 60 nm.
+This 2D LO phonon can be measured with SNOM at frequencies close to the TO phonon frequency, where the confinement is less, but this comes at the price of very high dissipation and correspondingly low propagation quality factor, as illustrated in Fig.
+2c.
+This is simply because frequencies near the TO phonon, which is strongly absorbing, correspond to the highest dissipation (measured by $\text{Re }\sigma$ or $\text{Im }\epsilon_{\mathrm{RPA}}$).
+These considerations imply that access to the lower-loss and higher-confined portions of the phononic waves, in the absence of a sharp tip, requires either a free electron probe, as in EELS, or a quantum emitter probe, such as a molecule or quantum well [cite people].
+
+In what remains of this section, we discuss briefly ``anomalies", by which we mean features of the 2D LO phonon dispersion that cannot be predicted by considering the electromagnetic modes of a very thin slab of material whose dielectric function is that of the bulk polar material (with finite LO-TO splitting).
+For example, the long-wavelength linear dispersion of the 2D LO phonon mode given by Equation (10) is not an anomaly, as this would be predicted by taking a sub-nm thick slab of a polar dielectric material.
+However, the precise slope cannot be predicted, as it would depend on the thickness used for the slab, which is ill-defined, as one will find that it does not correspond to the thickness of the boron or nitrogen atoms.
+One immediate anomaly is that in the monolayer, there is only one LO phonon, and thus only one branch of TM modes.
+This is contrary to a thin slab of hexagonal boron nitride, which is hyperbolic $\epsilon_{zz} > 0, \epsilon_{xx} = \epsilon_{yy} < 0$, and thus has many branches of TM modes spaced in wavevector by an amount proportional to $d^{-1}$, with $d$ the film thickness.
+Such additional TM modes would be incorrectly predicted by extrapolating the bulk to a film.
+Another anomaly comes from the fact that the dispersion relation of the 2D LO phonon is non-monotonic, and the frequency starts to decrease for sufficiently large wavevector.
+This can never happen by extrapolating to a monolayer a dielectric function which has finite LO-TO splitting.
+
 \section{Strong light-matter interactions enabled by 2D optical phonons}
 
-We now move to analyze key figures of merit of 2D LO phonons.
+We now move to analyze t
 

### 2018-11-11 paper.tex
text basically done. need to add figures and an equation now.

@@
 Using these parameters for important monolayers such as hexagonal boron nitride, we present the confinement and propagation losses of these modes, identifying the frequency region where they should be most easily detected.
-We...
-[insert nicest results here].
+We then show that these modes essentially provide the ultimate limit of light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced seven orders of magnitude by the emitter-LO phonon coupling.
+For free-electrons, we show a substantial probability of multiple LO phonon excitation can occur.
 
@@
 This is the consistent with the fact that LO-TO splitting is a collective effect arising from Coulomb interactions and the fact that the equation above represents a single-particle susceptibility.
-Coulomb interactions are accounted for in the random phase approximation, and to include them in the single-particle response would be double-counting.
+Coulomb interactions are accounted for in the random phase approximation, and to include them in the single-particle response is an uncontrolled double-counting.
 The component $qq$ in the response tensor denotes a pair of directions parallel to the wavevector.
@@
 
-We now move to analyze t
+The extreme confinement of electromagnetic fields offered by the 2D LO phonon presents an opportunity for quantum optical applications in which one seeks to couple an external emitter such as an atom, molecule, defect, or artificial atomic system to electromagnetic fields.
+Applications of these couplings can be to realize ultra-bright single- or two-photon sources, or realize strong-coupling and the associated phenomenology of Rabi oscillations and polaritons, or resolve spectroscopically "forbidden" transitions.
+In Figure 3, we show key figures of emitter-phonon coupling in polar monolayers.
 
-\subsection{Spontaneous emission}
+In Fig.
+3a, we show the enhancement (or Purcell factor) $F_p$ of spontaneous emission of 2D LO phonons at frequency $\omega_0$ by an external emitter at different distances $d$ away from the surface.
+The rate of spontaneous emission $\Gamma$ into 2D LO phonons obeys
+\begin{equation}
+\Gamma = \frac{3c^3}{2\omega_0^3}\int dq~q^2e^{-2qz}\text{Im }r_p(q,\omega_0) \times \Gamma_0 \equiv F_p(\omega_0)\Gamma_0,
+\end{equation}
+where $\Gamma_0$ is the free-space spontaneous emission rate.
+As can be seen from Fig.
+3a, for an emitter position 5 nm away from the surface, the Purcell factor can be very high, corresponding to over a million-fold increase in the rate of 2D LO phonon emission compared to photon emission into free-space.
+While this rate-enhancement is already quite high, it is a natural question what the highest possible spontaneous-emission rate enhancement is as a function of frequency.
+Recently, material-geometry-independent limits for spontaneous emission enhancement have been derived.
+In particular, the maximum rate of spontaneous emission enhancement is bounded by
+\begin{equation}
+...
+\end{equation}
+This is plotted in Fig. 3b for the case of monolayer hBN.
 
-\subsection{Spontaneous and stimulated electron energy loss}
+Finally, we address one manner of externally probing 2D LO phonons without the need for SNOM or emitters: with electron energy losses in free electrons.
+As the energy losses under consideration will be on the order of 200 meV for single phonon excitation in monolayer hBN, the electron energy spread which determines the width of EELS peak (including the zero-loss peak) must be substantially smaller than 200 meV.
+Fortunately, this is possible in high-resolution EELS, which is particularly useful for probing surface phenomena, and has been applied to measure the dispersion relation of plasmons in 2D or quasi-2D systems.
+Through advances in monochromation, electrons have also been recently used to probe optical phonons in bulk.
+In Fig.
+4, we show the EELS probability for electrons with energies on the order of 100 keV traversing a polar monolayer .
+Because of the strong coupling between matter and LO phonons shown in Fig.
+3, one expects multiple LO phonon excitation to be possible.
+Thus, this effect must be taken into account in the EELS.
+\begin{equation}
+...
+\end{equation}
+
+The ability to probe low-loss and highly confined electromagnetic modes associated with 2D LO phonons in polar materials provides a new platform for nanophotonics in the mid- and far-infrared spectral range.
+The identification of the phonon polariton of bulk and thin-film geometries with the 2D LO phonon made in this manuscript may one to allow extend the rich phenomenology of optical phonons to nanophotonic applications.
+It also may lead to useful new ways to study LO phonons, arising from the fact that 2D LO phonons, unlike their 3D counterparts, have their electromagnetic energy extend a considerable distance from the material boundary.
+Due to the strong electromagnetic interactions between emitters and 2D LO phonons shown in this manuscript, it may be possible to design interesting new hybrid states of matter and phonons based on quantum electrodynamical strong coupling.
+The highly confined optical phonon waves in polar monolayers may also provide interesting new opportunities in near-field radiative heat transfer, in which it has been long known that thin-film surface phonon polaritons play a critical role.
+An important avenue of future study would be the \textit{ab initio} calculation of lifetimes of 2D LO phonons associated with three-phonon processes and electron-phonon interactions.
+It would be of great interest to study the prospects of techniques such as isotopic purification and cryogenic temperatures for reducing the decay of these 2D LO phonons.
 

### 2018-11-29 paper.tex
wrote the last section on emitter strong coupling to phonons, defined scope of the rest of paper

@@
 Our analysis reveals that the phonon-polariton of the bulk is ``replaced" by the LO phonon.
-We present the confinement and propagation losses of LO phonons, finding that high LO phonon confinement and reasonable propagation quality factors coincide in regions which are difficult to detect with near-field optical microscopy, but may be more amenable to electron energy-loss spectroscopy.
-Finally, we show that 2D LO phonons in hBN present extremely strong light-matter interactions with external emitters, providing possibilities for 10,000,000-fold enhanced spontaneous emission and strong coupling with vibrational emitters.
+We present the confinement and propagation losses of LO phonons, finding that high LO phonon confinement and reasonable propagation quality factors coincide in regions which are difficult to detect with current near-field optical microscope techniques.
+We conclude by studying the interaction of external emitters with two-dimensional hBN nanostructures, finding extreme enhancement of spontaneous emission due to 2D phonon channels, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN phonons.
 \end{abstract}
@@
 Using these parameters for important monolayers such as hexagonal boron nitride, we present the confinement and propagation losses of these modes, identifying the frequency region where they should be most easily detected.
-We then show that these modes essentially provide the ultimate limit of light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced seven orders of magnitude by the emitter-LO phonon coupling.
-For free-electrons, we show a substantial probability of multiple LO phonon excitation can occur.
+We then show that these modes essentially provide the ultimate limit of light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced seven orders of magnitude by the emitter-LO phonon coupling, which due to the narrow bandwidth of phonon-polariton resonances, can lead to the possibility of realizing multi-mode strong coupling between an external emitter and optical phonons in hexagonal boron nitride.
 
@@
 In Fig.
-3a, we show the enhancement (or Purcell factor) $F_p$ of spontaneous emission of 2D LO phonons at frequency $\omega_0$ by an external emitter at different distances $d$ away from the surface.
+3a, we show the enhancement (or Purcell factor) $F_p$ of spontaneous emission of 2D LO phonons at frequency $\omega_0$ by an external emitter at different distances $d$ away from an infinite sheet of two-dimensional boron nitride.
 The rate of spontaneous emission $\Gamma$ into 2D LO phonons obeys
@@
 3a, for an emitter position 5 nm away from the surface, the Purcell factor can be very high, corresponding to over a million-fold increase in the rate of 2D LO phonon emission compared to photon emission into free-space.
-While this rate-enhancement is already quite high, it is a natural question what the highest possible spontaneous-emission rate enhancement is as a function of frequency.
-Recently, material-geometry-independent limits for spontaneous emission enhancement have been derived.
-In particular, the maximum rate of spontaneous emission enhancement is bounded by
-\begin{equation}
-...
-\end{equation}
-This is plotted in Fig. 3b for the case of monolayer hBN.
-
-Finally, we address one manner of externally probing 2D LO phonons without the need for SNOM or emitters: with electron energy losses in free electrons.
-As the energy losses under consideration will be on the order of 200 meV for single phonon excitation in monolayer hBN, the electron energy spread which determines the width of EELS peak (including the zero-loss peak) must be substantially smaller than 200 meV.
-Fortunately, this is possible in high-resolution EELS, which is particularly useful for probing surface phenomena, and has been applied to measure the dispersion relation of plasmons in 2D or quasi-2D systems.
-Through advances in monochromation, electrons have also been recently used to probe optical phonons in bulk.
+While this rate-enhancement is already quite high, it is a natural question how this coupling can be improved by nanostructuring, which will take the local density of electromagnetic states of Fig.
+3a and concentrate it around frequencies associated with resonances of the nanostructure.For simplicity, we consider the effect of nanostructuring the hBN into a disk, which leads to the formation of modes as in Fig.
+3b, in which we plot the surface charge density associated with each resonant mode.
 In Fig.
-4, we show the EELS probability for electrons with energies on the order of 100 keV traversing a polar monolayer .
-Because of the strong coupling between matter and LO phonons shown in Fig.
-3, one expects multiple LO phonon excitation to be possible.
-Thus, this effect must be taken into account in the EELS.
-\begin{equation}
-...
-\end{equation}
+3c, we plot the enhancement of the spontaneous emission rate for an external emitter polarized perpendicularly to the plane of the disk and placed 5 nm away from the surface of the disk.
+We find that the rate of spontaneous emission of 2D optical phonons is approximately 8 orders of magnitude larger than the rate of spontaneous emission of photons in the far field.
+For an infrared emitter at a transition wavelength of 7 $\mu$m with a free-space radiative lifetime of 1 $\mu$s, 5 nm away from an hBN disk, the coupling rate to 2D optical phonons would be on the same scale as the optical phonon frequency itself (about $2 \times 10^{14}$ $ s^{-1}$).
+This coupling rate thus implies coupling between an emitter and the field in the regime of ultra-strong coupling.
+Thus, the extreme confinement of electromagnetic energy associated with LO phonons in two dimensions enables the possibility of realizing ultrastrong coupling of an atom or molecule with optical phonons in a polar material, allowing the potential realization of new coupled states of quantum emitters and phonons.
 

### 2018-12-21 paper.tex
Nick's work from Overleaf

@@
 Phonon polaritons share many features in common with plasmon polaritons in conductors.
-In recent years, it has been shown that phonon-polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra}.
+In recent years, it has been shown that phonon-polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra,li2018infrared,ma2018plane}.
 Due to this remarkable confinement and their relatively high lifetimes of around picoseconds, phonon-polaritons open new opportunities for vibrational spectroscopy, radiative heat transfer \cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters \cite{kumar2015tunable,rivera2017making,kurman2018control}.
-In many cases, extreme confinement of phonon polaritons is achieved by the use of thin-films, which shrink the in-plane and out-of-plane wavelength of polaritons monotonically with decreasing film thickness.
+In many cases, extreme confinement of phonon polaritons is achieved by the use of thin-films, which shrink the in-plane and out-of-plane wavelength of polaritons monotonically with decreasing film thickness \cite{dai2014tunable,dubrovkin2018ultra}.
 In many cases, a monolayer is the ultimate limit of this effect.
 Thus, to bring these classical and quantum nanophotonics applications to their ultimate limits, it is necessary to consider the properties of these modes in two-dimensions.
-However, this picture of monolayers supporting maximally confined phonon-polariton modes is ostensibly complicated by the fact that in two-dimensional polar materials, the LO-TO splitting that gives rise to phonon-polaritons in three-dimensions is absent at the $\Gamma$ point.
+However, this picture of monolayers supporting maximally confined phonon-polariton modes is ostensibly complicated by the fact that in two-dimensional polar materials, the LO-TO splitting that gives rise to phonon-polaritons in three-dimensions is absent at the $\Gamma$ point \cite{sanchez2002vibrational,mele2002electric,serrano2007vibrational,sohier2017breakdown}.
 
 	In this Letter, we develop core results in the nascent field of optical phononics in two-dimensional materials by deriving a universal form for the dielectric function of a polar monolayer, incorporating spatially nonlocal effects, which is specified in fully in terms of the dispersion relation of LO and TO phonons in two dimensions.
-Using these parameters for important monolayers such as hexagonal boron nitride, we present the confinement and propagation losses of these modes, identifying the frequency region where they should be most easily detected.
-We then show that these modes essentially provide the ultimate limit of light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced seven orders of magnitude by the emitter-LO phonon coupling, which due to the narrow bandwidth of phonon-polariton resonances, can lead to the possibility of realizing multi-mode strong coupling between an external emitter and optical phonons in hexagonal boron nitride.
+Using parameters from Ref.
+\cite{sohier2017breakdown} for important monolayers such as hexagonal boron nitride, we present the confinement and propagation losses of these modes, identifying the frequency region where they should be most easily detected.
+We then show that these modes essentially provide the ultimate limit of light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced by up to eight orders of magnitude by the emitter-LO phonon coupling, which due to the narrow bandwidth of phonon-polariton resonances, can lead to the possibility of realizing the multi-mode strong coupling and ultra-strong coupling regimes of quantum electrodynamics for an external emitter coupled with optical phonons in hexagonal boron nitride.
 
@@
 	\begin{equation}
-H_{\mathrm{int}} = \int d^2x \rho \phi = -\int d^2x (\nabla\cdot\mathbf{P})\phi,
+	H_{\mathrm{int}} = \int d^2x ~\rho \phi = -\int d^2x~ (\nabla\cdot\mathbf{P})\phi,
 	\end{equation} 
@@
 	\begin{equation}\label{eq:2dsusceptibility}
-\boldsymbol{\Pi}(\mathbf{q},\omega) =  \frac{1}{\epsilon_0 \Omega}\sum\limits_{m,n}\frac{\mathbf{P}_{mn}(\mathbf{q})\otimes\mathbf{P}_{nm}(\mathbf{q})}{\hbar\omega + E_{nm}+\iu 0^+}\left(\e^{\beta E_m}-\e^{\beta E_n} \right),
+	\boldsymbol{\Pi}(\mathbf{q},\omega) =  \frac{1}{\epsilon_0 \Omega\mathcal{Z}}\sum\limits_{m,n}\frac{\mathbf{P}_{mn}(\mathbf{q})\otimes\mathbf{P}_{nm}(\mathbf{q})}{\hbar\omega + E_{nm}+\iu 0^+}\left(\e^{-\beta E_m}-\e^{-\beta E_n} \right),
 	\end{equation}
-where $m,n$ are states of the phononic Fock space of the monolayer, $P_{i,mn} = \sum\limits_{\kappa,j}Z_{\kappa,ij}\langle m | u_{\kappa,j} | n \rangle$ are matrix elements of the polarization associated with phonon modes, $E_{m (n)}$ is the energy of state $m$ $(n)$, and $\beta$ is the inverse temperature.
+	where $m,n$ are states of the phononic Fock space of the monolayer, $P_{i,mn} = \sum\limits_{\kappa,j}Z_{\kappa,ij}\langle m | u_{\kappa,j} | n \rangle$ are matrix elements of the polarization associated with phonon modes, $E_{m (n)}$ is the energy of state $m$ $(n)$, $\beta$ is the inverse temperature, and $\mathcal{Z}$ is the grand partition function.
 Evaluating the contribution of optical phonons in the harmonic approximation to the polarization-polarization response, one finds that the response function is given by:
@@
 	To proceed, we express the polarization matrix element in Equation (4) in terms of parameters of the monolayer such as the Born effective charges.
-Considering the longitudinal phonon contribution to the second-quantized ionic displacement, as in [Needs reference!], one finds that the RPA dielectric function is given by:
+Considering the longitudinal phonon contribution to the second-quantized ionic displacement, as in \cite{srivastava1990physics}, one finds that the RPA dielectric function is given by:
 	\begin{equation}
@@
 	\end{equation}
-In this expression, $\omega_L$ and been re-named as $\omega_{\mathrm{TO}}$, because in the absence of LO-TO splitting, they are degenerate.
+	In this expression, $\omega_L$ has been re-named as $\omega_{\mathrm{TO}}$, because in the absence of LO-TO splitting, they are degenerate.
 Additionally, we have defined eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}\equiv \frac{e_{\kappa\mathbf{q},L}}{\sqrt{M_{\kappa}}}$, where $e_{\kappa\mathbf{q},L}$ is the unit-normalized polarization vector of atom $\kappa$ in the unit cell oscillating according to a longitudinal phonon of wavevector $\mathbf{q}$ and $M_{\kappa}$ is the mass of atom $\kappa$.
-
+	\begin{figure}[t]
+		\includegraphics[width=85mm]{fig1.pdf}
+		\caption{\textbf{LO phonons as the basic electromagnetic waves of a polar monolayer.} (a) Schematic structure of a polar monolayer such as hexagonal boron nitride.
+(b) Properties of LO and TO phonons in 3D and 2D.
+In 3D, there is a finite LO-TO splitting at zero wavevector, while in 2D there is none.
+Despite this, the 2D LO phonon plays the role of the phonon polariton in 3D and thin films.
+(c) Analogous physics appears in electron gases in 3D and 2D, where the 3D plasma frequency is analogous to the 3D LO-TO splitting.
+In 2D, the plasma frequency at zero wavevector is zero, but the electromagnetic physics is determined by the dispersion of 2D plasmons, which replace the plasmon polariton of bulk and thin films.}
+	\end{figure}
 	As is well known, the zeros of the RPA dielectric function, denoted $\omega_{\mathbf{q}}$, correspond to longitudinal electromagnetic waves.
@@
 This result, as we now show, reveals that the TM mode corresponds simply to the two-dimensional optical phonon.
-In particular, it was shown in [Sohier2017] that in two-dimensional polar materials, the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO-TO splitting given by Equation (8) of the current manuscript.
+As shown recently in \cite{sohier2017breakdown}, in two-dimensional polar materials the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO-TO splitting.
+The expression for the LO-TO splitting is none other than Equation (8) of the current manuscript.
 \textit{Thus, one of the main results of our manuscript is that the wave solutions Maxwell's equations associated with phonon-polaritons in bulk materials, correspond to pure LO phonons in the monolayer limit.}
-
+	\begin{figure*}[t]
+		\includegraphics[width=135mm]{fig2.pdf}
+		\caption{\textbf{Propagation characteristics of 2D LO phonons in an un-structured monolayer of boron nitride.} (a) Imaginary part of the Fresnel reflection coefficient for p-polarized waves in free-standing monolayer hBN, which peaks at the dispersion relation of the 2D LO phonon.
+(b) Confinement factor of 2D LO phonon modes, which measures how much smaller the 2D LO phonon wavelength is compared to that of a photon at the same frequency.
+(c) Propagation quality factor, which measures the number of wavelengths of propagation of the 2D LO phonon mode, for different values of the damping rate of phonon.}
+	\end{figure*}
 	We now re-write $\epsilon_{\mathrm{RPA}}$ in a way that is expressed purely in terms of the optical phonon dispersion, derive the conductivity of the monolayer, and derive a universal form for the dispersion of 2D LO phonons in the long-wavelength limit.
@@
 Thus, all of the optical properties associated with LO phonons is captured by the TO phonon frequency, the slope of the LO dispersion at $\Gamma$, and the dissipation $\tau^{-1}$.
-In Table 1, we tabulate these parameters for a few polar monolayers whose phonon band-structures have been calculated in the literature: hBN, MoS$_2$, and ...
+%In Table 1, we tabulate these parameters for a few polar monolayers whose phonon band-structures have been calculated in the literature: hBN, MoS$_2$, and ...
 
@@
 In the three-dimensional plasmon case, `$\omega_{\mathrm{TP}} = 0$", but the difference between the squared longitudinal and transverse plasma oscillation frequencies at zero-wavevector is non-zero and given by $\omega_p^2$.
-This analogy between phononic and plasmonic behavior as a function of dimension is illustrated schematically in Figure 1.
+This analogy between phononic and plasmonic behavior as a function of dimension is illustrated schematically in Figures 1(b,c).
 
@@
 	\end{equation}
-In Figs.
-2b, we show the confinement factor (or effective mode index) of the 2D LO phonon mode, defined by $n_{\mathrm{eff}} = \frac{qc}{\omega}$.
-The confinement factor is a key figure of merit in many nanophotonic applications involving quantum emitters, because as was shown in [Rivera], all interactions involving emission of electromagnetic energy by emitters, such as electric and magnetic dipole emission, multipole emission, phosphorescence, and multiphoton emission, scale as some power of this confinement factor.
+	In Figure 2b, we show the confinement factor (or effective mode index) of the 2D LO phonon mode, defined by $n_{\mathrm{eff}} = \frac{qc}{\omega}$.
+The confinement factor is a key figure of merit in many nanophotonic applications involving quantum emitters, because as was shown in \cite{archambault2010quantum,koppens2011graphene,rivera2016shrinking,rivera2017making}, all interactions involving emission of electromagnetic energy by emitters, such as electric and magnetic dipole emission, multipole emission, phosphorescence, and multiphoton emission, scale as some power of this confinement factor.
 Interestingly, in the case of a monolayer, the confinement grows very rapidly in frequency, due to the extremely low group velocity of 2D LO phonons, which is a remarkable four orders of magnitude slower than the speed of light.
-In particular, at frequencies of 1450 $cm^{-1}$, the LO phonon already has a confinement factor in excess of conf, corresponding to a phonon wavelength of 6896.55/conf nm, which is much shorter wavelength than any phonon polariton thus measured so far, as well as shorter wavelength than any plasmon that has been measured in the mid-IR, such as in graphene.
-In fact, this short wavelength is mostly out of the range of measurability in scattering near-field microscope (SNOM) measurements, which with current tip radii, are limited to measuring wavelengths in excess of 60 nm.
+In particular, at frequencies of 1450 $cm^{-1}$, the LO phonon already has a confinement factor in excess of 400, corresponding to a phonon wavelength below 20 nm, which is much shorter wavelength than any phonon polariton thus measured so far, as well as shorter wavelength than any plasmon that has been measured in the mid-IR, such as in graphene.
+In fact, this short wavelength is mostly out of the range of measurability in current scattering near-field microscope (SNOM) measurements, which with current tip radii, are limited to measuring wavelengths no shorter than 60 nm.
 This 2D LO phonon can be measured with SNOM at frequencies close to the TO phonon frequency, where the confinement is less, but this comes at the price of very high dissipation and correspondingly low propagation quality factor, as illustrated in Fig.
-2c.
-This is simply because frequencies near the TO phonon, which is strongly absorbing, correspond to the highest dissipation (measured by $\text{Re }\sigma$ or $\text{Im }\epsilon_{\mathrm{RPA}}$).
-These considerations imply that access to the lower-loss and higher-confined portions of the phononic waves, in the absence of a sharp tip, requires either a free electron probe, as in EELS, or a quantum emitter probe, such as a molecule or quantum well [cite people].
+2c, where Re $q$/Im $q$ < 5.
+This is simply because frequencies near the TO phonon resonance correspond to very high dissipation (measured by $\text{Re }\sigma$ or $\text{Im }\epsilon_{\mathrm{RPA}}$).
+These considerations imply that access to the lower-loss and higher-confined portions of the phononic waves, in the absence of a sharp tip, requires either a free electron probe, as in EELS, where slow electrons can be used to probe plasmon wavelengths of just a few nanometers in monolayer metals, as well as the nonlocal bulk plasmon dispersion in metals \cite{nagao2001dispersion,diaconescu2007acoustic,de2010optical}.
+EELS has been recently employed to measure phonon polaritons in ultrathin films of hBN \cite{govyadinov2017probing}.
+Yet another interesting type of probe, with relevance to fundamental physics and quantum optics applications, is a quantum emitter such as an atom, molecule, or quantum well.
+A quantum emitter, which one can consider as a sub-nanoscale tip, can see very high wavevectors of the 2D LO phonon if sufficiently close to the polar monolayer.
+Recently, it was demonstrated using nanostructures of bulk hBN that the interaction of vibrational emitters with phonon polaritons is on the border of the strong coupling regime \cite{autore2018boron}.
 
@@
 	\section{Strong light-matter interactions enabled by 2D optical phonons}
-
+	\begin{figure}[t]
+		\includegraphics[width=80mm]{fig3.pdf}
+		\caption{\textbf{Extreme spontaneous emission enhancement due to 2D LO phonons in nanostructured geometries.} Enhancement of the spontaneous emission rate for an emitter concentric with the disk at a distance $z=5$ nm away from the disk.
+For a disk with a diameter of 20 nm, the rate of emission enhancement can be 100 million-fold.
+For an emitter with a free-space decay rate of $10^6$ cm$^{-1}$ at 7 $\mu$m, the emitter would experience a decay rate comparable to the frequency of the disk mode, leading to ultra-strong coupling of an external emitter with 2D LO phonons.}
+	\end{figure}
 	The extreme confinement of electromagnetic fields offered by the 2D LO phonon presents an opportunity for quantum optical applications in which one seeks to couple an external emitter such as an atom, molecule, defect, or artificial atomic system to electromagnetic fields.
 Applications of these couplings can be to realize ultra-bright single- or two-photon sources, or realize strong-coupling and the associated phenomenology of Rabi oscillations and polaritons, or resolve spectroscopically "forbidden" transitions.
-In Figure 3, we show key figures of emitter-phonon coupling in polar monolayers.
 
 	In Fig.
-3a, we show the enhancement (or Purcell factor) $F_p$ of spontaneous emission of 2D LO phonons at frequency $\omega_0$ by an external emitter at different distances $d$ away from an infinite sheet of two-dimensional boron nitride.
-The rate of spontaneous emission $\Gamma$ into 2D LO phonons obeys
-\begin{equation}
-\Gamma = \frac{3c^3}{2\omega_0^3}\int dq~q^2e^{-2qz}\text{Im }r_p(q,\omega_0) \times \Gamma_0 \equiv F_p(\omega_0)\Gamma_0,
-\end{equation}
-where $\Gamma_0$ is the free-space spontaneous emission rate.
-As can be seen from Fig.
-3a, for an emitter position 5 nm away from the surface, the Purcell factor can be very high, corresponding to over a million-fold increase in the rate of 2D LO phonon emission compared to photon emission into free-space.
-While this rate-enhancement is already quite high, it is a natural question how this coupling can be improved by nanostructuring, which will take the local density of electromagnetic states of Fig.
-3a and concentrate it around frequencies associated with resonances of the nanostructure.For simplicity, we consider the effect of nanostructuring the hBN into a disk, which leads to the formation of modes as in Fig.
-3b, in which we plot the surface charge density associated with each resonant mode.
+3, we consider the quantum coupling of a dipole emitter to 2D LO phonon modes in nanostructured monolayer hBN.
+For simplicity, we consider hBN nanostructured as a disk, which leads to the formation of sharp resonances.
 In Fig.
-3c, we plot the enhancement of the spontaneous emission rate for an external emitter polarized perpendicularly to the plane of the disk and placed 5 nm away from the surface of the disk.
+3, we plot the enhancement of the spontaneous emission rate for an external emitter polarized perpendicularly to the plane of the disk and placed 5 nm away from the surface of the disk.
 We find that the rate of spontaneous emission of 2D optical phonons is approximately 8 orders of magnitude larger than the rate of spontaneous emission of photons in the far field.
@@
 acknowledges support from the Danish Council for Independent Research (Grant No.\ DFF--6108-00667).
-The authors thank Joshua Caldwell (Vanderbilt), Dmitri Basov (Columbia), Ido Kaminer (Technion), Siyuan Dai (UT Austin), and Samuel Moore (Columbia) for helpful discussions.
+The authors thank Joshua Caldwell, Dmitri Basov, Ido Kaminer, Siyuan Dai, Samuel Moore), Jennifer Coulter, and Christopher Ciccarino for helpful discussions.
 This work was supported by the DOE Photonics at Thermodynamic Limits Energy Frontier Research Center under grant no.

### 2018-12-21 paper.tex
Edits and comments from Thomas

--- 2018-11-29 paper.tex
@@
 
+% --- PACKAGES ---
 \usepackage{graphicx}
@@
 
-%----- References -----
+%-----SI UNITS-----
+\usepackage{siunitx}
+\sisetup{range-phrase =\text{\,--\,},
+	list-units   =single,
+	range-units  =single,
+	list-pair-separator = {\ \text{and}\ },
+	list-separator = {,\ \linebreak[0]},
+	list-final-separator = {,\ \linebreak[0]\text{and}\ }}
+
+%--- REFERENCING ---
 \usepackage{xcolor}
@@
 
-%----- Captions in sans font -----
+%--- CAPTIONS (sans serif) ---
 \makeatletter
@@
 
+% --- SPACING ---
+\usepackage{microtype} % Kerning and symbol-stretching
+\usepackage{xspace}
+
 \thickmuskip=5mu plus 2mu minus 1mu  %binary relations (default, 5mu plus 5mu)
@@
 
+% --- COMMANDS ---
 \renewcommand{\Im}{\operatorname{Im}}
@@
 
+%abbreviations
+\newcommand{\ie}{i.e.\@\xspace} %Gobble-spaces of the "small" type (small is ensured by adding \@)
+\newcommand{\cf}{cf.\@\xspace}
+\newcommand{\eg}{e.g.\@\xspace}
+\newcommand{\etc}{etc.\@\xspace}
+
 \newcommand{\HarvardSEAS}{John A. Paulson School of Engineering and Applied Sciences, Harvard University, Cambridge, MA, USA}
@@
 
-%\usepackage[usenames]{color}
-%\newcommand{\edited}[1]{{\color{red} #1}}
-\newcommand{\comment}[2]{\textcolor{blue!70!black}{\small\textsf{\textsuperscript{\textsc{\MakeLowercase{#1}}}[#2]}}}
-\usepackage[normalem]{ulem}
-\newcommand{\swap}[2]{\textcolor{red!70!black}{\sout{#1}}\textrightarrow\textcolor{green!50!black}{#2}}
-\newcommand{\remove}[1]{\textcolor{red!70!black}{\sout{#1}}}
-\newcommand{\inset}[1]{\textcolor{green!50!black}{#1}}
+% ----- COMMENTS AND META-ANNOTATIONS -----
+\usepackage{textcomp} % for \textrightarrow
+\usepackage{xifthen}
+\usepackage{etoolbox}
+\newboolean{togglecomments}
+\newboolean{togglechanges}
+
+% toggle to true to see comments (otherwise hidden)
+\setboolean{togglecomments}{true}  
+% toggle to false to see mixed versions (otherwise edits are shown exclusively)
+\setboolean{togglechanges}{false}
+
+\newcommand{\comment}[2]{%
+    \ifbool{togglecomments}%
+    {\textcolor{blue!70!black}{\small\textsf{%
+    \textsuperscript{\textsc{\textsf{\MakeLowercase{#1}}}}%
+    [#2]}}} % if true, show comments
+    {}}     % if false, do nothing
+\newcommand{\swap}[2]{\ifbool{togglechanges}
+    {#2}  % TC-only version
+    {\textcolor{red!70!black}{[#1]}\text{\textrightarrow{}}\textcolor{green!50!black}{[#2]}}}
+\newcommand{\remove}[1]{\ifbool{togglechanges}
+    {}    % TC-only version
+    {\textcolor{red!70!black}{#1}}}
+\newcommand{\inset}[1]{\ifbool{togglechanges}
+    {{}#1}  % TC-only version
+    {\textcolor{green!50!black}{{}#1}}}
+\makeatletter
+\newcommand{\citeremind}[1]{%
+	\unskip%
+    \textcolor{blue!75!black!80!yellow}{${}^\blacksquare$%
+	\ifthenelse{\isempty{#1}}{}{\textsuperscript{\tiny\textsf{#1}}}%
+	}\xspace}
+\makeatother
+\newcommand{\FIXME}[1]{\textcolor{red!70!black}{[#1]}}
+\newcommand{\optional}[1]{\textcolor{yellow!70!gray}{#1}}
+
+% --- DOCUMENT ---
 \begin{document}
 
+% --- TITLE & AUTHORSHIP ---
 \title{Nanophotonics with optical phonons in two dimensions}
 
-\author{Nicholas Rivera}\email{nrivera@seas.harvard.edu}\affiliation{\HarvardSEAS}\affiliation{\MITPhy}
-\author{Thomas Christensen}\affiliation{\MITPhy}
-\author{Prineha Narang}\email{prineha@seas.harvard.edu}\affiliation{\HarvardSEAS}
+\author{Nicholas Rivera}
+\email{nrivera@seas.harvard.edu}
+\affiliation{\HarvardSEAS}\affiliation{\MITPhy}
+%
+\author{Thomas Christensen}
+\affiliation{\MITPhy}
+%
+\author{Prineha Narang}
+\email{prineha@seas.harvard.edu}
+\affiliation{\HarvardSEAS}
 
@@
 
+% --- ABSTRACT ---
 \begin{abstract}
-Extreme confinement of electromagnetic energy by phonon-polaritons promises extremely strong and novel forms of control over the dynamics of matter.
+    Extreme confinement of electromagnetic energy by phonon-polaritons \swap{promises}{holds the promise of} extremely strong and novel forms of control over the dynamics of matter. 
     To bring such control to its ultimate limit, it is important to consider phonon-polaritons in two dimensional systems. 
-Recent studies have pointed out that in two-dimensional systems, splitting between longitudinal and transverse (LO and TO) optical phonons is absent at the $\Gamma$ point.
-Does this lack of LO-TO splitting imply the lack of a strongly confined electromagnetic (phonon-polariton) mode in polar monolayers?
+    \comment{tc}{It is not necessarily clear to me why 2D phonon-polaritonics would be an obvious component in bringing phonon-polaritons to their ``ultimate limit''?}
+    Recent studies have pointed out that in two-dimensional systems, splitting between longitudinal and transverse optical (LO and TO) phonons is absent at the $\Gamma$ point.  
+    Does this lack of LO--TO splitting imply the \swap{lack}{absence} of a \swap{strongly confined electromagnetic (phonon-polariton) mode}{phonon-polariton} in polar monolayers? 
     Here, we derive a universal form for the spatially-nonlocal dielectric function of a polar monolayer specified by: the TO phonon frequency, the non-zero group velocity of the LO phonon, and the phonon lifetime.
-Our analysis reveals that the phonon-polariton of the bulk is ``replaced" by the LO phonon.
-We present the confinement and propagation losses of LO phonons, finding that high LO phonon confinement and reasonable propagation quality factors coincide in regions which are difficult to detect with current near-field optical microscope techniques.
-We conclude by studying the interaction of external emitters with two-dimensional hBN nanostructures, finding extreme enhancement of spontaneous emission due to 2D phonon channels, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN phonons.
+    \comment{tc}{I'm a bit hesitant to highlight the ``spatially-nonlocal'' dielectric function: the real reason this (simplified) form is nonlocal is just dimensional---\eg graphene's dielectric function is also nonlocal in a similar fashion (even in the low-momentum, low-frequency limit).
+The associated surface conductivity is local; so the in-plane polarization response is also local.}
+    Our analysis reveals that the phonon-polariton of the bulk is \swap{``replaced''}{essentially supplanted/replaced} by the LO phonon\inset{ of the two-dimensional system}.
+    We \swap{present}{discuss} the confinement and propagation losses of the LO \swap{phonons}{phonon-polaritons}, finding that high \remove{LO phonon} confinement and reasonable propagation quality factors coincide in regions which are difficult to detect with current near-field optical microscope techniques.
+    We conclude by studying the interaction of external emitters with two-dimensional hBN nanostructures, finding extreme enhancement of spontaneous emission due to \swap{2D phonon channels}{coupling with localized 2D phonon-polaritons}, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN \swap{phonons}{phonon-polaritons}.
 \end{abstract}
@@
 
-Phonon polaritons, hybrid quasiparticles of photons and optical phonons supported in polar materials, offer great promise for deeply sub-diffractional control of electromagnetic fields at mid-IR and THz frequencies.
-Phonon polaritons share many features in common with plasmon polaritons in conductors.
-In recent years, it has been shown that phonon-polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra}.
-Due to this remarkable confinement and their relatively high lifetimes of around picoseconds, phonon-polaritons open new opportunities for vibrational spectroscopy, radiative heat transfer \cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters \cite{kumar2015tunable,rivera2017making,kurman2018control}.
-In many cases, extreme confinement of phonon polaritons is achieved by the use of thin-films, which shrink the in-plane and out-of-plane wavelength of polaritons monotonically with decreasing film thickness.
+\comment{tc}{Both `phonon-polaritons' and `phonon polaritons' are used; let's settle on one.}
+
+\comment{tc}{For title, how about `Phonon polaritons in two-dimensional hexagonal boron nitride'?}
+
+% --- INTRODUCTION ---
+Phonon polaritons, hybrid quasiparticles of photons and optical phonons supported in polar materials, \swap{offer}{hold} great promise for deeply sub-diffractional control of electromagnetic fields at mid-\swap{IR}{infrared} and \swap{THz}{terahertz} frequencies.
+\swap{P}{Qualitatively, p}honon polaritons share many features in common with plasmon polaritons in conductors.
+\swap{In}{E.g., in} recent years, it has been shown that phonon-polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra,li2018infrared,ma2018plane}.
+Due to this remarkable confinement and their relatively high lifetimes of around picoseconds, phonon-polaritons open new opportunities for vibrational spectroscopy,\citeremind{} radiative heat transfer~\cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters~\cite{kumar2015tunable,rivera2017making,kurman2018control}.
+In many cases, extreme confinement of phonon polaritons is achieved by the use of thin-films, which \swap{shrink}{reduce} the in-\remove{plane} and out-of-plane wavelength of polaritons monotonically with decreasing film thickness~\cite{dai2014tunable,dubrovkin2018ultra}.
+\comment{tc}{The previous and following sentences both start with `many cases' here; best to rewrite one (probably the latter).}
 In many cases, a monolayer is the ultimate limit of this effect.
 Thus, to bring these classical and quantum nanophotonics applications to their ultimate limits, it is necessary to consider the properties of these modes in two-dimensions.
-However, this picture of monolayers supporting maximally confined phonon-polariton modes is ostensibly complicated by the fact that in two-dimensional polar materials, the LO-TO splitting that gives rise to phonon-polaritons in three-dimensions is absent at the $\Gamma$ point.
+\comment{tc}{I'm not sure that starting with `however' in this sentence is optimal: we want to convey an additional motivation for studying them (specifically, the transition from 3D to 2D is nontrivial)---maybe this sentence could be rewritten with this perspective in mind.}
+However, this picture of monolayers supporting maximally confined phonon-polariton modes is ostensibly complicated by the fact that in two-dimensional polar materials, the LO--TO splitting that gives rise to phonon-polaritons in three-dimensions is absent at the $\Gamma$ point~\cite{sanchez2002vibrational,mele2002electric,serrano2007vibrational,sohier2017breakdown}.
 
-In this Letter, we develop core results in the nascent field of optical phononics in two-dimensional materials by deriving a universal form for the dielectric function of a polar monolayer, incorporating spatially nonlocal effects, which is specified in fully in terms of the dispersion relation of LO and TO phonons in two dimensions.
-Using these parameters for important monolayers such as hexagonal boron nitride, we present the confinement and propagation losses of these modes, identifying the frequency region where they should be most easily detected.
-We then show that these modes essentially provide the ultimate limit of light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced seven orders of magnitude by the emitter-LO phonon coupling, which due to the narrow bandwidth of phonon-polariton resonances, can lead to the possibility of realizing multi-mode strong coupling between an external emitter and optical phonons in hexagonal boron nitride.
+In this Letter, we develop \swap{core results in}{a framework and a several concrete examples for understanding the} the nascent field of optical phononics in two-dimensional materials\swap{ by deriving}{.
+We do so by deriving} a universal form for the dielectric function of a polar monolayer, \swap{incorporating spatially nonlocal effects}{including spatial nonlocality}, which \swap{is specified fully in terms of the dispersion relation of LO and TO phonons in two dimensions}{which depends solely on the LO and TO phonon frequencies---and their dispersion with momentum---in the two-dimensional system}.
+Using parameters from Ref.~\citenum{sohier2017breakdown} for \swap{ important monolayers such as }{the canonical two-dimensional polar monolayer---}hexagonal boron nitride\swap{, }{---}we present the confinement and propagation losses of \swap{these modes}{the supported phonon-polaritons}, identifying the frequency region where they should be most easily detected.
+\swap{We then}{Finally, we} show that these modes \swap{essentially provide the ultimate limit of}{enable extreme} light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced by up to eight orders of magnitude by the emitter--LO phonon coupling\swap{, which due to}{.
+Given} the narrow bandwidth of phonon-polariton resonances, \swap{can lead to }{this suggests} the possibility of realizing the multi-mode strong coupling and ultra-strong coupling regimes of quantum electrodynamics \swap{for an external emitter coupled with optical phonons in hexagonal boron nitride}{in a two-dimensional hexagonal boron nitride platform}.
 
+% -----------------
 \section{Electrodynamics of optical phonons in two-dimensions}
+\comment{tc}{Structurally, this section would benefit from being split into two sections: one solely tasked with deriving/establishing the forms of the (non-interacting) polarization-polarization response function---and possibly its equivalent conductivity; and another one solely tasked with deriving and discussing the dispersion relation for the associated longitudinal optical modes.
+Right now, this section goes back and forth between the two, which (I think) reduces its focus and gives way for unnecessary repetition.}
 
-In this section, we develop the theory of electromagnetic waves associated with optical phonons in two dimensions.
-The crux of the theory is determining the dielectric permittivity associated with the monolayer.
-It is obtained by considering the response of the ions in the monolayer to a longitudinal electromagnetic wave whose electric field is the gradient of a scalar potential $\phi$.
-The coupling Hamiltonian is then given by
+In this section, we develop \swap{the}{a} theory of electromagnetic \swap{waves associated with}{response due to} optical phonons in two\swap{ dimensions}{-dimensional systems}.
+The crux of the theory \swap{is determining}{involves} the dielectric \swap{permittivity}{function} associated with the monolayer.
+\swap{It is obtained by considering the response of the ions in the monolayer to a longitudinal electromagnetic wave whose electric field is the gradient of a scalar potential $\phi$.}{To obtain it, we consider the response of the ions of the monolayer due(/when subjected) to a longitudinal electromagnetic potential $\phi$}.
+The \swap{coupling}{interaction} Hamiltonian is \swap{then given by}{in this case simply}
 \begin{equation}
-H_{\mathrm{int}} = \int d^2x \rho \phi = -\int d^2x (\nabla\cdot\mathbf{P})\phi,
+    H_{\mathrm{int}} = \int \dd^2x ~\rho \phi = -\int \dd^2x~ (\nabla\cdot\mathbf{P})\phi,
+    \label{eq:coupling}
 \end{equation} 
-with $\rho$ the charge density and $\mathbf{P}$ the polarization density associated with the ionic motion.
-The polarization can be expanded in powers of the displacement of the ions.
+with $\rho$ the \inset{induced} charge density and $\mathbf{P}$ the \inset{total} polarization density associated with the ionic motion. 
+\swap{The polarization \inset{density} can be expanded in powers of the displacement of the ions.
 In particular, we can consider the polarization $P_i$ generated in the monolayer when atom $\kappa$ of the unit cell is displaced along the $j$ direction by amount $u_{\kappa,j}$.
-To lowest order:
+To lowest order:}{Within linear response, the induced polarization density can be straightforwardly evaluated from the displacement amplitude $\mathbf{u}_{\kappa}$ of every atom $\kappa$ within the unit cell.
+Specifically, to first-order, the induced polarization is:}
+\comment{tc}{I don't see how the second equality can be consistent: since, by Taylor expansion, $P_i = P_i^0 + u_{\kappa,j}\partial_{u_{\kappa,j}}P_i + \mathcal{O}(u_{\kappa,j}^2)$---so where does the unit cell area $\Omega$ come from?
+The issue seems to arise from alternatingly interpreting the LHS/RHS $P_i$ as total/density quantities.
+The right-most equation also seems to be missing a $\kappa$-subscript in $u_{j}$.}
 \begin{equation}
@@
 where $Z_{\kappa,ij}$ is the Born effective charge tensor of ion $\kappa$ and $\Omega$ is the unit cell area.
-With this relation between polarization and ionic displacements, we have expressed the Hamiltonian as a coupling between the scalar potential and ionic displacements.
-We consider the response of the monolayer to a scalar potential of the form $\phi = \phi(\mathbf{q},\omega)e^{i\mathbf{q}\cdot\mathbf{r}-i\omega t}$, where $\mathbf{q}$ is a two-dimensional wavevector in the plane of the monolayer.
-Such a potential corresponds to a longitudinal electric field of the form $\mathbf{E} = i\mathbf{q}\phi(\mathbf{q},\omega)e^{i\mathbf{q}\cdot\mathbf{r}-i\omega t}$.
-It follows that the induced polarization takes the form $P_i  = \epsilon_0\Pi_{ij}(\mathbf{q},\omega)E_{j,\mathrm{tot}}(\mathbf{q},\omega)$, where $\Pi_{ij}(\mathbf{q},\omega)$ is the polarization-polarization response function of the monolayer and  $E_{j,\mathrm{tot}}(\mathbf{q},\omega)$ is taken as the \textit{total} field which is a sum of the applied field and the field created by induced polarization.
+\comment{tc}{We do not give a definition of $P_i^0$: we also do not comment on what the implications are if it is nonzero (it could be..!)---as far as I can see, we are assuming it is zero for the remaining analysis.}
+With this relation between polarization and ionic displacements, \swap{we have expressed the Hamiltonian as a coupling between the scalar potential and ionic displacements}{the interaction Hamiltonian in Equation~\eqref{eq:coupling} couples the scalar potential and the ionic displacements}.
+We consider the response of the monolayer to a scalar potential of the form $\phi\inset{(\mathbf{r})} = \phi(\mathbf{q},\omega)\e^{\iu\mathbf{q}\cdot\mathbf{r}-\iu\omega t}$, where $\mathbf{q}$ is a two-dimensional wavevector in the plane of the monolayer.
+Such a potential corresponds to a longitudinal electric field \remove{of the form} $\mathbf{E}\inset{(\mathbf{r})} = \iu\mathbf{q}\swap{\phi(\mathbf{q},\omega)\e^{\iu\mathbf{q}\cdot\mathbf{r}-\iu\omega t}}{\phi(\mathbf{r})}$.
+\comment{tc}{I don't think it really ``follows'' from the preceding that $P_i = ...
+E_{j,\text{tot}}(\mathbf{q},\omega)$.
+First, this specializes to the RPA via a mean-field approximation, and second, it introduces a response function (so, really, it is more of a definition).
+It is probably better to phrase these next few sentences rather carefully and with specificity, \eg with references to the usual Kubo formula/approach etc.
+It may also make sense to split the $P_i = ...
+E_{j,\text{tot}}(\mathbf{q},\omega)$ off onto its own line, since this is the equation that lends practical utility to the $\boldsymbol{\Pi}$ response function.
+Finally, I don't think we actually need to mention the RPA or anything like it at this point: so far, this is just the non-interacting polarization-polarization response---we effectively arrive at the RPA later on, when we start doing electromagnetic calculations (\ie around Eq.~\eqref{eq:2dmaxwell}); only then do we actually make use of $\mathbf{P} \propto \boldsymbol{\Pi}\mathbf{E}_{\mathrm{tot}}$.}
+It follows that the induced polarization takes the form $P_i\inset{(\mathbf{q})}  = \epsilon_0\Pi_{ij}(\mathbf{q},\omega)E_{j,\mathrm{tot}}(\mathbf{q},\omega)$, where $\Pi_{ij}(\mathbf{q},\omega)$ is the polarization-polarization response function of the monolayer and  $E_{j,\mathrm{tot}}(\mathbf{q},\omega)$ is taken as the \textit{total} field which is a sum of the applied field and the field created by induced polarization.
 The polarization-polarization response function is given by
 \begin{equation}\label{eq:2dsusceptibility}
-\boldsymbol{\Pi}(\mathbf{q},\omega) =  \frac{1}{\epsilon_0 \Omega}\sum\limits_{m,n}\frac{\mathbf{P}_{mn}(\mathbf{q})\otimes\mathbf{P}_{nm}(\mathbf{q})}{\hbar\omega + E_{nm}+\iu 0^+}\left(\e^{\beta E_m}-\e^{\beta E_n} \right),
+    \boldsymbol{\Pi}(\mathbf{q},\omega) =  \frac{1}{\epsilon_0 \Omega\mathcal{Z}}\sum\limits_{m,n}\frac{\mathbf{P}_{mn}(\mathbf{q})\otimes\mathbf{P}_{nm}(\mathbf{q})}{\hbar\omega + E_{nm}+\iu 0^+}\Big(\e^{-\beta E_m}-\e^{-\beta E_n} \Big),
 \end{equation}
-where $m,n$ are states of the phononic Fock space of the monolayer, $P_{i,mn} = \sum\limits_{\kappa,j}Z_{\kappa,ij}\langle m | u_{\kappa,j} | n \rangle$ are matrix elements of the polarization associated with phonon modes, $E_{m (n)}$ is the energy of state $m$ $(n)$, and $\beta$ is the inverse temperature.
+where $m,n$ are states of the phononic Fock space of the monolayer, $P_{mn,i} = \sum_{\kappa,j}Z_{\kappa,ij}\langle m | u_{\kappa,j} | n \rangle$ are matrix elements of the polarization associated with phonon modes,
+\comment{tc}{Probably best to include $\mathbf{q}$ dependence explicitly in both LHS and RHS of the definition of $P_{mn,i}\textcolor{red!50!black}{(\mathbf{q})}$.}
+$E_{m}$ ($E_n$) is the energy of state $m$ ($n$), $\beta\inset{\equiv 1/k_{\text{\textsc{b}}}T}$ is the inverse temperature, and $\mathcal{Z}$ is the grand partition function.
 Evaluating the contribution of optical phonons in the harmonic approximation to the polarization-polarization response, one finds that the response function is given by: 
+\comment{tc}{It's not clear how we obtain Eq.~\eqref{eq:polpolresponse} from Eq.~\eqref{eq:2dsusceptibility}: \eg, I'm not sure what you mean by the harmonic approximation here?
+There's also a low-temperature approximation at play here.
+Finally, every phonon band except the LO one is dropped.
+Finally, we have not previously defined the $\mathbf{P}(\mathbf{q})$ polarization operator.
+I guess I'm just asking for us to be more explicit and detailed here.}
 \begin{equation}
-\Pi_{qq}(\mathbf{q},\omega) = \frac{1}{\epsilon_0\hbar\Omega} \frac{2\omega_{\mathbf{q}L}}{\omega^2_{\mathbf{q}L}-\omega^2}|\hat{q}\cdot\langle 1_{\mathbf{q}L}|\mathbf{P}(\mathbf{q})|0_{\mathbf{q}\sigma}\rangle|^2,
+    \Pi_{qq}(\mathbf{q},\omega) = \frac{1}{\epsilon_0\hbar\Omega} \frac{2\omega_{\mathbf{q},\mathrm{L}}}{\omega^2_{\mathbf{q},\mathrm{L}}-\omega^2}|\hat{\mathbf{q}}\cdot\langle 1_{\mathbf{q},\mathrm{L}}|\mathbf{P}(\mathbf{q})|0_{\mathbf{q}\sigma}\rangle|^2,
+    \label{eq:polpolresponse}
 \end{equation}
-where 'L' denotes longitudinal polarization, $|0_{\mathbf{q}L}\rangle$ ($|1_{\mathbf{q}L}\rangle$) denotes a state with no (one) longitudinal phonon of wavevector $\mathbf{q}$.
-The frequency $\omega_{\mathbf{q}L}$ in the denominator, as in the case of bulk phonons, is the frequency of the longitudinal phonon of wavevector $\mathbf{q}$ prior to considering LO-TO splitting [Needs reference!].
-This is the consistent with the fact that LO-TO splitting is a collective effect arising from Coulomb interactions and the fact that the equation above represents a single-particle susceptibility.
-Coulomb interactions are accounted for in the random phase approximation, and to include them in the single-particle response is an uncontrolled double-counting.
+where L-subscripts denote longitudinal polarization, $|0_{\mathbf{q},\mathrm{L}}\rangle$ ($|1_{\mathbf{q},\mathrm{L}}\rangle$) denotes a state with no (one) longitudinal phonon of wavevector $\mathbf{q}$.
+The frequency $\omega_{\mathbf{q},\mathrm{L}}$ in the denominator, as in the case of bulk phonons, is the frequency of the longitudinal phonon of wavevector $\mathbf{q}$ prior to considering LO--TO splitting.\citeremind{}
+This is consistent with the fact that LO--TO splitting is a collective effect arising from Coulomb interactions and the fact that the equation above represents a single-particle susceptibility.
+Coulomb interactions are accounted for in the random phase approximation,
+\comment{tc}{Another reason for us to be more clear about the implications of the mean-field assumption before Eq.~\eqref{eq:2dsusceptibility}; might as well establish equivalence with RPA clearly and only once.}
+and to include them in the single-particle response \swap{is an}{would amount to an} uncontrolled double-counting. 
+\comment{tc}{Could it be better to interleave this $qq$-comment into the lead-up sentence for Eq.~\eqref{eq:polpolresponse}?
+Basically, we want to say that we specialize to the diagonal components.
+Finally; is it really necessary for us to make that scope reduction here?
+Wouldn't the formula nearly be simpler without it?}
 The component $qq$ in the response tensor denotes a pair of directions parallel to the wavevector.  
+\comment{tc}{It might be better to break this off into a new paragraph here (and possibly merge (and maybe shorten) with the following paragraph) since this is sort of orthogonal to the previous stuff.}
 To relate the polarization-polarization response function to the electromagnetic modes supported by a polar monolayer, we solve Maxwell's equations for an evanescent electromagnetic mode supported by a surface with polarization-polarization response tensor $\mathbf{\Pi}$.
@@
 
+\comment{tc}{I don't think this needs to be as long as it currently is---also, some of the details, \eg position of Reststrahlen band, assumption of long wave-length etc., would probably fit in more meaningfully when those approximations are actually applied.
+For Eq.~\eqref{eq:2dmaxwell}, only the homogeneity assumption is necessary (corresponding to in-plane translation invariance)}
 To strip the analysis to its bare essentials, we consider optical phonon response with in-plane isotropy in the long-wavelength limit arising from in-plane LO oscillations.
-A relevant example of a system where these conditions are satisfied is in a hexagonal boron nitride monolayer, considering optical phonons in the so-called upper Reststrahlen band which in bulk spans the frequency range of 1360 to 1610 cm$^{-1}$.
+A relevant example of a system where these conditions are satisfied is in a hexagonal boron nitride monolayer, considering optical phonons in the so-called upper Reststrahlen band which in bulk spans the frequency range of \SIrange[range-phrase=\text{ to }]{1360}{1610}{\per\cm}.
 In a monolayer geometry with translation invariance and in-plane isotropy, the electromagnetic modes of Maxwell's equations can be composed into transverse magnetic (TM) and transverse electric (TE), where the magnetic or electric field respectively is transverse to the in-plane wavevector of the mode.
@@
 Given the isotropy, we may suppress indices from the response tensor.
-We may also consider without further loss of generality a TM mode with wavenumber $q$ along the $x$-direction in the monolayer and magnetic field $H(z)e^{iqx-i\omega t}$ along the $y$-direction of the monolayer.
+We may also consider without \remove{further} loss of generality a TM mode with wavenumber $q$ along the $x$-direction in the monolayer and magnetic field $H(z)\e^{\iu qx-\iu\omega t}$ along the $y$-direction of the monolayer.
 The direction transverse to the monolayer is denoted as $z$.
-With these definitions in place, the Maxwell equation satisfied by the magnetic field is \begin{equation}\label{eq:2dmaxwell}
-\left(-\frac{\dd^2}{\dd{}z^2}+q^2-\epsilon_{\mathrm{env}}\frac{\omega^2}{c^2} \right)H(z) = 0.
+With these definitions in place, the Maxwell equation satisfied by the magnetic field is
+\begin{equation}\label{eq:2dmaxwell}
+    \bigg(-\frac{\dd^2}{\dd{}z^2}+q^2-\epsilon_{\mathrm{env}}\frac{\omega^2}{c^2} \bigg)H(z) = 0.
 \end{equation}
-We consider a solution of the form $H(z) = e^{-\kappa_1 z}$ for $z > 0$ with $\kappa_1 = \sqrt{q^2-\epsilon_1\frac{\omega^2}{c^2}}$ and a solution of the form $H(z) = ce^{\kappa_2 z}$ for $z < 0$ with $\kappa_2 = \sqrt{q^2-\epsilon_2\frac{\omega^2}{c^2}}$.
+\comment{tc}{I'm not sure this needs to be its own equation.
+Even if it does, I don't think that the term $\epsilon_{\mathrm{env}}\omega^2/c^2$ is correct; shouldn't this really be $\epsilon_{1,2}\omega^2/c^2$, with choice of $\{1,2\}$ depending on sign of $z$?}
+We consider a solution of the form $H(z) = \e^{-\kappa_1 z}$ for $z > 0$ with $\kappa_1 = \sqrt{q^2-\epsilon_1\frac{\omega^2}{c^2}}$ and a solution of the form $H(z) = c\e^{\kappa_2 z}$ for $z < 0$ with $\kappa_2 = \sqrt{q^2-\epsilon_2\frac{\omega^2}{c^2}}$.
+\comment{tc}{Should probably choose a different name for the constant `$c$' to avoid clashing with the speed of light $c$.
+Maybe $c_1$ or $a$.
+Or, alternatively, let $H(z) = h_\pm\e^{\mp\kappa_\pm z}$ with $\kappa_\pm = \sqrt{q^2-\epsilon_\pm k_0^2}$ with upper/lower signs applicable for $\pm z>0$ (instead of using $\{1,2\}$ subscripts)---and then specify the boundary condition as $h_-/h_+ = \ldots (=c)$.
+Personally, I favor the latter option but it requires a few more fixes.
+If you go with that option, you can probably drop the parentheses around $H_y^{(\pm)}$ in the following.}
 The boundary condition on the magnetic field is $H_y^{(+)}-H_y^{(-)} = -K_x = \iu\omega\mathbf{P}_x$ where $\mathbf{K}$ is the surface current density, expressed through the polarization density $\mathbf{P}$ of the monolayer.
@@
 Continuity of the electric field in the $x$ direction enforces $\frac{\epsilon_2}{\epsilon_1} = -\frac{\kappa_2}{\kappa_1}c$. 
-These two conditions fully specify the TM mode.
-To simplify the discussion, we consider the so-called electrostatic limit, it in which the wavenumber $q$ of the TM mode is much larger than the free-space wavenumber $\frac{\omega}{c}$.
+\swap{These two conditions fully specify the TM mode}{Combining the two conditions, we obtain the usual dispersion equation for the TM mode of a polarizable 2D monolayer, namely $\epsilon_1/\kappa_1 + \epsilon_2/\kappa_2 +\epsilon_0\Pi=0$.}.
+%
+\swap{To simplify the discussion, we consider the so-called electrostatic limit, it in which the wavenumber $q$ of the TM mode is much larger than the free-space wavenumber $\frac{\omega}{c}$.
 This is well-satisfied in monolayer optical phonons in 2D.
-With this approximation, the condition for a TM mode corresponding to an optical phonon is
+With this approximation, the condition for a TM mode corresponding to an optical phonon is}
+{Given the deeply subwavelength nature of 2D phonon-polaritons, \ie since $q\gg\omega/c$ such that $\kappa_{1,2}\simeq q$, the dispersion equation can be reduced to its quasistatic limit without consequential loss of accuracy:}
 \begin{equation}
     \epsilon_{\mathrm{env}} + \frac{1}{2}\epsilon_0 q \Pi(q,\omega) \equiv \epsilon_{\mathrm{RPA}}(\omega) = 0.
+    \label{eq:dispeq}
 \end{equation}
+\comment{tc}{Need to define $\epsilon_{\mathrm{env}}\equiv (\epsilon_1+\epsilon_2)/2$.}
 The quantity $\epsilon_{\mathrm{RPA}}(\omega) = 0$ coincides with the phonon contribution to the longitudinal dielectric function of the monolayer in the random-phase approximation. 
+\comment{tc}{I do not completely understand the intention of this last sentence: are we trying to say that this coincides with the zeros of the dielectric function?
+If so, maybe we should be a bit more explicit in our thinking here (e.g., we can say that we identify the LHS with the RPA dielectric function---and maybe comment on that the two quantities naturally have the same zeros)}
 
-To proceed, we express the polarization matrix element in Equation (4) in terms of parameters of the monolayer such as the Born effective charges.
-Considering the longitudinal phonon contribution to the second-quantized ionic displacement, as in [Needs reference!], one finds that the RPA dielectric function is given by:
+\comment{tc}{Wouldn't it be more meaningful to get these kind of formal things all handled before we start discussing the 2D excitations?
+Specifically, we might as well introduce the eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}$ when we discuss $\boldsymbol{\Pi}$?
+That way, the discussion of modes could be more focused and wouldn't need this detour.
+}
+To proceed, we express the polarization matrix element in Equation~\eqref{eq:polpolresponse} in terms of parameters of the monolayer such as the Born effective charges. 
+Considering the longitudinal phonon contribution to the second-quantized ionic displacement, as in Ref.~\citenum{srivastava1990physics}, one finds that the RPA dielectric function is given by:
 \begin{equation}
-\epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}} - \frac{q}{2\epsilon_0}\frac{\Big|\hat{q}\cdot\sum\limits_{\kappa,j} Z_{\kappa,ij}\eta_{\kappa,j}  \Big|^2}{\omega^2_{\mathrm{TO}}-\omega^2}
+    \epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}} - \frac{q}{2\epsilon_0}\frac{\big|\hat{\mathbf{q}}\cdot\sum_{\kappa,j} Z_{\kappa,ij}\eta_{\kappa,j}  \big|^2}{\omega^2_{\mathrm{TO}}-\omega^2}.
+    \label{eq:epsRPA}
 \end{equation}
-In this expression, $\omega_L$ and been re-named as $\omega_{\mathrm{TO}}$, because in the absence of LO-TO splitting, they are degenerate.
-Additionally, we have defined eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}\equiv \frac{e_{\kappa\mathbf{q},L}}{\sqrt{M_{\kappa}}}$, where $e_{\kappa\mathbf{q},L}$ is the unit-normalized polarization vector of atom $\kappa$ in the unit cell oscillating according to a longitudinal phonon of wavevector $\mathbf{q}$ and $M_{\kappa}$ is the mass of atom $\kappa$.
+\comment{tc}{There's an issue with Eq.~\eqref{eq:epsRPA} AFAICT (and with all other occurrences of $\sum_{\kappa,j}Z_{\kappa,ij}\eta_{\kappa,j}$): it has a lingering $i$-index dependence.
+I guess this is the result of a mix between vector- and index-notation.}
+In this expression, $\omega_\mathrm{L}$ has been re-named as $\omega_{\mathrm{TO}}$, because in the absence of LO--TO splitting, they are degenerate. 
+\comment{tc}{There's some unpleasantness in substituting an \emph{longitudinal} mode for a \emph{transverse} mode in the $\mathrm{L}\rightarrow\mathrm{TO}$ swap here.
+I guess it might be good to be reminded here that $\omega_{\mathrm{L}}$ is the longitudinal phonon mode \emph{before} LO--TO splitting (I understand that this sentence is trying to hint that; but a more direct hint may be useful).}
+Additionally, we have defined \inset{scaled} eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}\equiv \frac{\mathbf{e}_{\kappa\mathbf{q},\mathrm{L}}}{\sqrt{M_{\kappa}}}$, where $\mathbf{e}_{\kappa\mathbf{q},\mathrm{L}}$ is the unit-normalized polarization vector
+\comment{tc}{We had hats on $\mathbf{e}_{\kappa\mathbf{q}}$ in the previous paper.}
+of atom $\kappa$ in the unit cell oscillating according to a longitudinal phonon of wavevector $\mathbf{q}$ and $M_{\kappa}$ is the mass of atom $\kappa$.
+	\comment{tc}{A few things: 
+	\begin{enumerate}
+	\item Here, we switch to using vectorial notation.
+My preference is probably that we use it whenever possible, but we have used index-notation elsewhere (also in cases where vector notation is not awkward)---we should probably try to stick to one set of notations.
+I think it's meaningful to go with mainly vector-notation, since that's also what we used in the previous paper.
+	\item I recall something I think I also asked about in the previous paper: shouldn't there be an $\e^{\iu\mathbf{q}\cdot\mathbf{b}_\kappa}$ dependence here?
+Looking at the previous paper's Eq.~(8b) and the text immediately after, that also seems to be an issue there (compare the definition of $\mathbf{S}_{\mathbf{q}\sigma}$ with (8b) and (9); things don't end up agreeing, AFAICT).
+What's the argument for dropping this exponential factor?
+	\item This is a more general comment: the notation we use here deviates quite a bit from the notation we used in the previous paper: wouldn't it make more sense to go with essentially the same notation?
+More generally, reading through this, I occasionally feel that the level of detail is in the ``uncanny valley'': it gives some things with substantial detail, and for several other things, many will feel a need to consult our previous paper.
+One could argue (I'm not necessarily doing so, just raising it for consideration) that the derivation of the 2D polarization-polarization response (specifically, its structure) does not involve sufficient differences between 2D and 3D to merit a new derivation here.
+E.g., we could almost start with Eq.~(9) from our previous paper (with $V\rightarrow \Omega$, instead of bothering to go through the derivation.
+Of course, that makes the content somewhat more sparse, but it would possibly be a more focused paper that way.
+Anyway, something to consider.
+	\item We discuss the dielectric function/permittivity of the 2D sheet quite a bit.
+In my view, though, it would be much more natural (and, I think, more immediately recognizable to researchers in the 2D optical-excitations field) to discuss the conductivity instead.
+This would also make for a clearer discussion of which parts of the response actually are nonlocal (since, in the low-$q$ limit, the 2D permittivity of any material must have the generic form $\epsilon\simeq 1 + \alpha q$ (vacuum cladding)).
+	\end{enumerate}}
 
-As is well known, the zeros of the RPA dielectric function, denoted $\omega_{\mathbf{q}}$, correspond to longitudinal electromagnetic waves.
-Thus, the condition for the zeros of the dielectric function can be expressed as
+% ~~~~~~~~~~~~~~
+\begin{figure}[t]
+    \includegraphics[width=85mm]{fig1.pdf}
+    \caption{%
+        \textbf{LO phonons as the basic electromagnetic waves of a polar monolayer.} (a) Schematic structure of a polar monolayer such as hexagonal boron nitride.
+(b) Properties of LO and TO phonons in 3D and 2D.
+In 3D, there is a finite LO--TO splitting at zero wavevector, while in 2D there is none.
+Despite this, the 2D LO phonon plays the role of the phonon polariton in 3D and thin films.
+(c) Analogous physics appears in electron gases in 3D and 2D, where the 3D plasma frequency is analogous to the 3D LO--TO splitting.
+In 2D, the plasma frequency at zero wavevector is zero, but the electromagnetic physics is determined by the dispersion of 2D plasmons, which replace the plasmon polariton of bulk and thin films.
+        \comment{tc}{We already discussed reasonable changes to the figures in-person, so I won't comment on that this time around.}
+        \label{fig:1}
+        }
+\end{figure}~
+% ~~~~~~~~~~~~~~
+
+\swap{As is well known, t}{T}he zeros of the RPA dielectric function, denoted $\omega_{\mathbf{q}}$, correspond to longitudinal electromagnetic \swap{waves}{excitations}.
+\swap{Thus, the condition for the zeros of the dielectric function can be expressed as}{Thus, from Equation~\eqref{eq:epsRPA}, we obtain the dispersion equation for the 2D phonon polaritons:}
+\comment{tc}{I think we are repeating/cycling a bit here; we already mentioned this previously, and also already derived a form for the phonon-polariton dispersion equation (Eq.~\eqref{eq:dispeq}).}
 \begin{align}
@@
     &= V(q)\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2,
+    \label{eq:dispersion}
 \end{align}
-where $eZ^r = Z$, with $e$ the electron charge, and $V(q)$ is the Coulomb interaction in two dimensions, screened by the dielectric environment of the sub- and superstrates.
+where $eZ^r \equiv Z$, 
+\comment{tc}{I guess I see the motivation for introducing a quantity like $Z^r$---but it's not obvious to me why it should be an $r$-superscript?
+Maybe $\tilde{Z}$?
+Oh, in any event, this definition needs subscripts.}
+with $e$ the electron charge, and $V(q)$ is the Coulomb interaction in two dimensions, \comment{tc}{Probably should define $V(q)$ explicitly (to remind that $V(q)\propto 1/q$)} screened by the dielectric environment of the sub- and superstrates.
 This result, as we now show, reveals that the TM mode corresponds simply to the two-dimensional optical phonon.
-In particular, it was shown in [Sohier2017] that in two-dimensional polar materials, the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO-TO splitting given by Equation (8) of the current manuscript.
-\textit{Thus, one of the main results of our manuscript is that the wave solutions Maxwell's equations associated with phonon-polaritons in bulk materials, correspond to pure LO phonons in the monolayer limit.}
+As shown recently in Ref.~\citenum{sohier2017breakdown}, in two-dimensional polar materials the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO--TO splitting.
+\comment{tc}{This might be a good spot to mention/highlight that this has an analogy in 2D plasmonics: the 2D plasmon is just the conventional ``bulk'' plasmon of the 2D sheet---\ie there is no notion of a surface plasmon there.
+The intrinsic mode becomes the surface mode, and vice versa.}
+The expression for the LO--TO splitting is none other than Equation~\eqref{eq:dispersion} of the current manuscript.
+\comment{tc}{Might be a good spot to point out that the splitting vanishes in the $q\rightarrow 0$ limit.}
+\emph{Thus, one of the main results of our manuscript is that the wave solutions Maxwell's equations associated with phonon-polaritons in bulk materials, correspond to pure LO phonons in the monolayer limit.}
+\comment{tc}{Probably not a good idea to have large block of text in italics, even if we feel it is important.}
 
-We now re-write $\epsilon_{\mathrm{RPA}}$ in a way that is expressed purely in terms of the optical phonon dispersion, derive the conductivity of the monolayer, and derive a universal form for the dispersion of 2D LO phonons in the long-wavelength limit.
-From Equation (8), we may write Equation (7) as
+% ~~~~~~~~~~~~~~
+\begin{figure*}[t]
+    \includegraphics[width=135mm]{fig2.pdf}
+    \caption{%
+        \textbf{Propagation characteristics of 2D LO phonons in an un-structured monolayer of boron nitride.} (a) Imaginary part of the Fresnel reflection coefficient for p-polarized waves in free-standing monolayer hBN, which peaks at the dispersion relation of the 2D LO phonon.
+(b) Confinement factor of 2D LO phonon modes, which measures how much smaller the 2D LO phonon wavelength is compared to that of a photon at the same frequency.
+(c) Propagation quality factor, which measures the number of wavelengths of propagation of the 2D LO phonon mode, for different values of the damping rate of phonon.
+        \label{fig:2}
+        }
+\end{figure*}
+% ~~~~~~~~~~~~~~
+
+We now rewrite $\epsilon_{\mathrm{RPA}}$ in a way that is expressed purely in terms of the optical phonon dispersion, derive the conductivity of the monolayer, and derive a universal form for the dispersion of 2D LO phonons in the long-wavelength limit.
+From Equation~\eqref{eq:dispersion}, we may write Equation~\eqref{eq:epsRPA} as
 \begin{equation}
-\epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}\left( 1 + \frac{\omega^2_{\mathbf{q}\mathrm{LO}}-\omega^2_{\mathbf{q}\mathrm{TO}}}{\omega^2_{\mathbf{q}\mathrm{TO}}-\omega^2}\right).
+    \epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}\Bigg( 1 + \frac{\omega^2_{\mathbf{q},\mathrm{LO}}-\omega^2_{\mathbf{q},\mathrm{TO}}}{\omega^2_{\mathbf{q},\mathrm{TO}}-\omega^2}\Bigg).
 \end{equation}
 Similarly, the conductivity is prescribed by $\sigma(\mathbf{q},\omega) = -\iu \omega \Pi(\mathbf{q},\omega)$.
-We are now in a position to derive a universal form for the RPA dielectric function of a 2D polar slab in terms of three phenomenological parameters.
-In many cases [ref], the Born charges are independent of $q$ for small $q$, and thus the dispersion can be written as
+\comment{tc}{This conductivity-definition should probably come sooner; along with the definition of which quantities $\boldsymbol{\Pi}$ connects.}
+We are now in a position to derive a universal form for the RPA dielectric function of a 2D polar slab in terms of three phenomenological parameters. In many cases,\citeremind{}
+\comment{tc}{It seems likely that this is true in all cases (just from thinking of Taylor series)\ldots If so, let's be more clear.
+(Though, of course, it could depend on the direction of $\mathbf{q}$; but we already assumed isotropy elsewhere.}
+the Born charges are independent of $q$ for small $q$, and thus the dispersion can be written as
 \begin{equation}
-\omega_{\mathbf{q},\mathrm{LO}} = \sqrt{\omega^2_{\Gamma \mathrm{TO}}+\mathcal{S}q} \approx \omega_{\Gamma \mathrm{TO}} + v_g q,
+    \omega_{\mathbf{q},\mathrm{LO}} = \sqrt{\omega^2_{\Gamma, \mathrm{TO}}+\mathcal{S}q} \simeq \omega_{\Gamma,\mathrm{TO}} + v_{\mathrm{g}} q,
+    \label{eq:simpledispersion}
 \end{equation}
-where $\mathcal{S}$ is a microscopic constant determined by the Born charges of the monolayer, $v_g$ is the group velocity of the LO mode, and $\omega_{\Gamma \mathrm{TO}}$ is the TO phonon frequency at the $\Gamma$ point.
+where $\mathcal{S}$ is a microscopic constant determined by the Born charges of the monolayer, $v_{\mathrm{g}}$ is the group velocity of the LO mode, and $\omega_{\Gamma,\mathrm{TO}}$ is the TO phonon frequency at the $\Gamma$ point.
+\comment{tc}{There is probably no reason to have both $\mathcal{S}$ and $v_{\mathrm{g}}$; they are (simply) related.
+In that case, this paragraph could also be contracted substantially and the two sentences merged into one}
 The group velocity is specified in terms of microscopic parameters via 
 \begin{equation}
-v_g = \frac{e^2 }{4\epsilon_0 \epsilon_{\mathrm{env}}\omega_{\Gamma TO}\Omega}\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2. 
+    v_{\mathrm{g}} = \frac{e^2 }{4\epsilon_0 \epsilon_{\mathrm{env}}\omega_{\Gamma, \mathrm{TO}}\Omega}\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2. 
 \end{equation}
@@
 \begin{equation}
-\epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}\left( 1 + \frac{2v_g\omega_{\mathbf{q}\mathrm{TO}}}{\omega^2_{\mathbf{q}\mathrm{TO}}-\omega^2+\iu \omega \tau^{-1}}\right),
+    \epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}\Bigg( 1 + \frac{2v_{\mathrm{g}}\omega_{\mathbf{q},\mathrm{TO}}}{\omega^2_{\mathbf{q},\mathrm{TO}}-\omega^2+\iu \omega \tau^{-1}}\Bigg),
 \end{equation}
 where we have also phenomenologically included the phonon dissipation rate $\tau^{-1}$, consistently with  a relaxation-time prescription. 
+\comment{tc}{I guess it would make sense to introduce $\tau^{-1}$ sooner.}
 Thus, all of the optical properties associated with LO phonons is captured by the TO phonon frequency, the slope of the LO dispersion at $\Gamma$, and the dissipation $\tau^{-1}$.
-In Table 1, we tabulate these parameters for a few polar monolayers whose phonon band-structures have been calculated in the literature: hBN, MoS$_2$, and ...
+%In Table 1, we tabulate these parameters for a few polar monolayers whose phonon band-structures have been calculated in the literature: hBN, MoS$_2$, and ...
 
@@
 Here, we expand on the analogy in the two-dimensional case.
-To do so, we take Equation (8) in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $e^2\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2 \equiv  \left(\frac{eZ^r_*}{\sqrt{M_*}}\right)^2 \equiv \left(\frac{Q_*}{\sqrt{M_*}}\right)^2$.
-Then, the LO-TO splitting can be written as $\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} = \frac{Q_*^2}{2\epsilon_0\epsilon_{\mathrm{env}} M_*}q$.
+To do so, we take Equation~\eqref{eq:dispersion} in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $e^2\big|\sum_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \big|^2 \equiv  \Big(\frac{eZ^r_*}{\sqrt{M_*}}\Big)^2 \equiv \Big(\frac{Q_*}{\sqrt{M_*}}\Big)^2$.
+Then, the LO--TO splitting can be written as $\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} = \frac{Q_*^2}{2\epsilon_0\epsilon_{\mathrm{env}} M_*}q$.
 Now we note that the RHS is exactly the squared-frequency $\omega^2_{\mathbf{q}p}$ for a plasma oscillation in a 2D gas of charged particles with charge $Q_*$ and mass $M_*$.
-This squared frequency can be thought of as the "LP-TP" splitting between longitudinal and transverse plasma oscillations.
-Of course, there are no transverse plasma oscillations, and so ``$\omega_{\mathrm{TP}} = 0$".
-In the three-dimensional plasmon case, `$\omega_{\mathrm{TP}} = 0$", but the difference between the squared longitudinal and transverse plasma oscillation frequencies at zero-wavevector is non-zero and given by $\omega_p^2$.
-This analogy between phononic and plasmonic behavior as a function of dimension is illustrated schematically in Figure 1.
+This squared frequency can be thought of as the ``LP--TP'' splitting between longitudinal and transverse plasma oscillations.
+Of course, there are no transverse plasma oscillations, and so ``$\omega_{\mathrm{TP}} = 0$''.
+In the three-dimensional plasmon case, `$\omega_{\mathrm{TP}} = 0$', but the difference between the squared longitudinal and transverse plasma oscillation frequencies at zero-wavevector is non-zero and given by $\omega_p^2$.
+This analogy between phononic and plasmonic behavior as a function of dimension is illustrated schematically in Figures~\ref{fig:1}(b,c).
+\comment{tc}{We never mention Fig.~\ref{fig:1}(a). We should try to mention/ref' every subfigure at least once in the text.}
 
+% -----------------
 \section{Properties of 2D LO phonon modes in monolayer hBN}
 
-In Fig.
-2a, we present the dispersion relation, operationally defined through the poles of the Fresnel reflection coefficient for $p$-polarized waves, which for a monolayer of conductivity $\sigma(\mathbf{q},\omega) = -\frac{2\iu \omega}{\epsilon q}\left(\epsilon_{\mathrm{RPA}} - \epsilon_{\mathrm{env}} \right)$ below vacuum and on top of a substrate of permittivity $\epsilon_{\mathrm{s}}$, is given by
+In Figure~\ref{fig:2}(a), we present the dispersion relation, operationally defined through the poles of the Fresnel reflection coefficient for $p$-polarized waves, which for a monolayer of conductivity $\sigma(\mathbf{q},\omega) = -\frac{2\iu \omega}{\epsilon q}(\epsilon_{\mathrm{RPA}} - \epsilon_{\mathrm{env}})$
+\comment{tc}{We already have a definition of $\sigma(\mathbf{q},\omega)$ previously ($=-\iu\omega\Pi(\mathbf{q},\omega)$), so I don't think we need this additional definition here.
+If anything, it would be nice to have $\sigma(\mathbf{q},\omega)$ in terms of $\omega_{(\mathrm{TO},\mathrm{LO})}$ (including its local limit).}
+below vacuum and on top of a substrate of permittivity $\epsilon_{\mathrm{s}}$, is given by
+\comment{tc}{We should probably not have two competing but equivalent substrate permittivity definitions (\ie we have both $\epsilon_{\mathrm{s}}$ and $\epsilon_1$ right now; we could just stick with $\epsilon_1$ (or, even better, $\epsilon_-$).
+In that case, we could also just give the general form of $r_p$ with $(\varepsilon_- - \varepsilon_+)+\ldots$ in the numerator and $(\varepsilon_-+\varepsilon_+)+\ldots$ in the denominator.}
+\comment{tc}{The $-\sigma q/\epsilon_0\omega$ terms are missing an imaginary unit. I'm pretty sure it should be $+\iu\sigma q/\epsilon_0\omega$ but you may want to check the sign.}
 \begin{equation}
@@
 \end{equation}
-In Figs.
-2b, we show the confinement factor (or effective mode index) of the 2D LO phonon mode, defined by $n_{\mathrm{eff}} = \frac{qc}{\omega}$.
-The confinement factor is a key figure of merit in many nanophotonic applications involving quantum emitters, because as was shown in [Rivera], all interactions involving emission of electromagnetic energy by emitters, such as electric and magnetic dipole emission, multipole emission, phosphorescence, and multiphoton emission, scale as some power of this confinement factor.
-Interestingly, in the case of a monolayer, the confinement grows very rapidly in frequency, due to the extremely low group velocity of 2D LO phonons, which is a remarkable four orders of magnitude slower than the speed of light.
-In particular, at frequencies of 1450 $cm^{-1}$, the LO phonon already has a confinement factor in excess of conf, corresponding to a phonon wavelength of 6896.55/conf nm, which is much shorter wavelength than any phonon polariton thus measured so far, as well as shorter wavelength than any plasmon that has been measured in the mid-IR, such as in graphene.
-In fact, this short wavelength is mostly out of the range of measurability in scattering near-field microscope (SNOM) measurements, which with current tip radii, are limited to measuring wavelengths in excess of 60 nm.
-This 2D LO phonon can be measured with SNOM at frequencies close to the TO phonon frequency, where the confinement is less, but this comes at the price of very high dissipation and correspondingly low propagation quality factor, as illustrated in Fig.
-2c.
-This is simply because frequencies near the TO phonon, which is strongly absorbing, correspond to the highest dissipation (measured by $\text{Re }\sigma$ or $\text{Im }\epsilon_{\mathrm{RPA}}$).
-These considerations imply that access to the lower-loss and higher-confined portions of the phononic waves, in the absence of a sharp tip, requires either a free electron probe, as in EELS, or a quantum emitter probe, such as a molecule or quantum well [cite people].
+In Figure~\ref{fig:2}(b), we show the confinement factor (or effective mode index) of the 2D LO phonon mode, defined by $n_{\mathrm{eff}} = \frac{qc}{\omega}$. 
+\comment{tc}{I guess you want a real-part sign somewhere in this definition?
+(Of course, it's reasonable to not do this if it's just interpreted as an effective index, but questionable if interpreted as a confinement factor.}
+The confinement factor is a key figure of merit in many nanophotonic applications involving quantum emitters,\swap{ because as was shown in~\cite{archambault2010quantum,koppens2011graphene,rivera2016shrinking,rivera2017making},}{ since} all interactions involving emission of electromagnetic energy by emitters\swap{, such as}{---\eg} electric and magnetic dipole emission, multipole emission, phosphorescence, and multiphoton emission\swap{, }{---}scale \inset{approximately} as some power of this confinement factor\swap{.}{~\cite{archambault2010quantum,koppens2011graphene,rivera2016shrinking,rivera2017making}.}
 
-In what remains of this section, we discuss briefly ``anomalies", by which we mean features of the 2D LO phonon dispersion that cannot be predicted by considering the electromagnetic modes of a very thin slab of material whose dielectric function is that of the bulk polar material (with finite LO-TO splitting).
-For example, the long-wavelength linear dispersion of the 2D LO phonon mode given by Equation (10) is not an anomaly, as this would be predicted by taking a sub-nm thick slab of a polar dielectric material.
+Interestingly, in the case of a \inset{polar} monolayer, the confinement grows very rapidly in frequency, due to the extremely low group velocity of 2D LO phonons, which is a remarkable four orders of magnitude slower than the speed of light.
+In particular, at frequencies of \SI{1450}{\per\cm}, the LO phonon already has a confinement factor in excess of \num{400}, corresponding to a phonon wavelength below \SI{20}{\nm}, which is much shorter wavelength than any phonon polariton \remove{thus} measured so far, \swap{as well as shorter wavelength than any plasmon that has been measured in the mid-\swap{IR}{infrared}, such as in graphene.}{and, similarly shorter than any plasmonic wavelength, even in graphene.}
+In fact, this short\inset{ a} wavelength is \swap{mostly}{essentially} \swap{out of the range of measurability in}{beyond the range of} current scattering near-field microscope (SNOM) measurements, \swap{which with current tip radii, are limited to measuring wavelengths no shorter than \SI{60}{\nm}}{which are limited (by tip radii) to wavelengths greater than $\approx\,\SI{60}{\nm}$}.
+\comment{tc}{How did Koppens measure his nonlocal stuff then? 60 nm is not that high.}
+\swap{This 2D LO phonon can be measured with SNOM at frequencies close to the TO phonon frequency, where the confinement is less, but this comes at the price of very high dissipation and correspondingly low propagation quality factor, as illustrated in Figure~\ref{fig:2}(c), where $\Re q/\Im q < 5$.
+This is simply because frequencies near the TO phonon resonance correspond to very high dissipation (measured by $\Re\sigma$ or $\Im\epsilon_{\mathrm{RPA}}$).}
+{The 2D LO phonon polariton could in principle be measured closer to the TO frequency, where confinement is smaller; unfortunately, as shown in Figure~\ref{fig:2}(c), near the TO frequency, dissipation is far higher (and corresponding propagative quality factors $\Re q/\Im q$ far lower) \optional{due to large $\Re\sigma$ (or, equivalently, large $\Im\epsilon_{\mathrm{RPA}}$)}.}
+\comment{tc}{Might be good to actually plot $\sigma(\omega)$}
+These considerations imply that access to the lower-loss and higher-confined portions of the \swap{phononic waves}{2D phonon polaritons}, in the absence of a sharp tip, requires either a free electron probe, as in EELS, \comment{tc}{Acronym without definition} where slow electrons can be used to probe plasmon wavelengths of just a few nanometers in monolayer metals, as well as the nonlocal bulk plasmon dispersion in metals~\cite{nagao2001dispersion, de2010optical}{}\citeremind{diaconescu2007acoustic}.
+\comment{tc}{This sentence doesn't quite work, because it has an `either' construction but no second option; I guess an ``emitter-part'' is missing?}
+EELS has been recently employed to measure phonon polaritons in ultrathin films of hBN~\cite{govyadinov2017probing}.
+\swap{Yet a}{A}nother interesting \swap{type of probe}{class of probes}, with relevance to fundamental physics and quantum optics applications, is a quantum emitter such as an atom, molecule, or quantum well.
+A quantum emitter\swap{, which one can consider as a sub-nanoscale tip, }{---qualitatively akin to a sub-nanoscale SNOM tip---}can \swap{see}{efficiently excite} very high wavevectors of the 2D LO phonon if sufficiently close to the polar monolayer.
+\comment{tc}{This last sentence is a bit unmotivated and feels disconnected. Is it necessary?}
+Recently, it was demonstrated using nanostructures of bulk hBN that the interaction of vibrational emitters with phonon polaritons is on the border of the strong coupling regime~\cite{autore2018boron}.
+\comment{tc}{It feels strange to may that we have this discussion of quantum emitters here without discussing Fig.~\ref{fig:3}---there's no need for us to pick up the ball twice.}
+
+In what remains of this section, we  briefly discuss ``anomalies'', by which we mean features of the 2D LO phonon dispersion that cannot be predicted by considering the electromagnetic modes of a very thin slab of material whose dielectric function is that of the bulk polar material (with finite LO--TO splitting).
+For example, the long-wavelength linear dispersion of the 2D LO phonon mode given by Equation~\eqref{eq:simpledispersion} is not an anomaly, as this would be predicted by taking a sub-nm thick slab of a polar dielectric material.
 However, the precise slope cannot be predicted, as it would depend on the thickness used for the slab, which is ill-defined, as one will find that it does not correspond to the thickness of the boron or nitrogen atoms.
+\comment{tc}{I'm not sure I agree with that: it seems reasonable to take the interlayer spacing of bulk hBN.
+That would make sense when thinking about surface vs.\ bulk polarization.
+Generally though, I'm surprised to hear that this choice is consequential: that's not the case in plasmonics.}
 One immediate anomaly is that in the monolayer, there is only one LO phonon, and thus only one branch of TM modes.
 This is contrary to a thin slab of hexagonal boron nitride, which is hyperbolic $\epsilon_{zz} > 0, \epsilon_{xx} = \epsilon_{yy} < 0$, and thus has many branches of TM modes spaced in wavevector by an amount proportional to $d^{-1}$, with $d$ the film thickness.
+\comment{tc}{Well, this doesn't seem to be too bad of an anomaly: if I decided to treat the 2D sheet as a thin slab, then all the unphysical modes would be pushed to enormously high frequencies due to the $d^{-1}$-scaling; so it wouldn't affect any response-quantity I might be interested in?}
 Such additional TM modes would be incorrectly predicted by extrapolating the bulk to a film.
 Another anomaly comes from the fact that the dispersion relation of the 2D LO phonon is non-monotonic, and the frequency starts to decrease for sufficiently large wavevector.
-This can never happen by extrapolating to a monolayer a dielectric function which has finite LO-TO splitting.
+This can never happen by extrapolating to a monolayer a dielectric function which has finite LO--TO splitting.
+\comment{tc}{Generally, I think this discussion would be much improved by actually having calculations of a thin slab (like we discussed; similarly to what we did for the Argentene paper) of varying ``thinness'', and seeing how the calculations agree/disagree.}
 
+% ~~~~~~~~~~~~~~
+\begin{figure}[t]
+    \includegraphics[width=80mm]{fig3.pdf}
+    \caption{%
+        \textbf{Extreme spontaneous emission enhancement due to 2D LO phonons in nanostructured geometries.} Enhancement of the spontaneous emission rate for an emitter concentric \comment{tc}{Maybe go with simpler wording?} with the disk at a distance $z=\SI{5}{\nm}$ away from the disk.
+For a disk with a diameter of \SI{20}{\nm}, the rate of emission enhancement can be \inset{enhanced} \num{100} million-fold.
+For an emitter with a free-space decay rate of $\SI{1e6}{\per\cm}$ at \SI{7}{\micro\m}, the emitter would experience a decay rate comparable to the frequency of the disk mode, leading to ultra-strong coupling of an external emitter with 2D LO phonons.
+        \comment{tc}{The actual maximal value of the enhancement here is sensitively dependent on our choice of $\tau^{-1}$.
+I recall we took the value on the basis of one of Koppens' papers.
+If we don't state the value (and that it is realistic), our claim is not very convincing or impressive.
+We should probably state it in both text and caption.}
+        \label{fig:3}
+        }
+\end{figure}
+% ~~~~~~~~~~~~~~
+
+% -----------------
 \section{Strong light-matter interactions enabled by 2D optical phonons}
+\comment{tc}{Disclaimer: I didn't read this section very carefully.}
 
 The extreme confinement of electromagnetic fields offered by the 2D LO phonon presents an opportunity for quantum optical applications in which one seeks to couple an external emitter such as an atom, molecule, defect, or artificial atomic system to electromagnetic fields.
-Applications of these couplings can be to realize ultra-bright single- or two-photon sources, or realize strong-coupling and the associated phenomenology of Rabi oscillations and polaritons, or resolve spectroscopically "forbidden" transitions.
-In Figure 3, we show key figures of emitter-phonon coupling in polar monolayers.
+Applications of these couplings \swap{can be to realize}{are} ultra-bright single- or two-photon sources, \swap{or realize}{realizing the} strong-coupling \inset{regime} and the associated phenomenology of Rabi oscillations and polaritons, or \swap{resolve}{resolving} spectroscopically ``forbidden'' transitions.
+\comment{tc}{I guess this part needs citations?}
 
-In Fig.
-3a, we show the enhancement (or Purcell factor) $F_p$ of spontaneous emission of 2D LO phonons at frequency $\omega_0$ by an external emitter at different distances $d$ away from an infinite sheet of two-dimensional boron nitride.
-The rate of spontaneous emission $\Gamma$ into 2D LO phonons obeys
-\begin{equation}
-\Gamma = \frac{3c^3}{2\omega_0^3}\int dq~q^2e^{-2qz}\text{Im }r_p(q,\omega_0) \times \Gamma_0 \equiv F_p(\omega_0)\Gamma_0,
-\end{equation}
-where $\Gamma_0$ is the free-space spontaneous emission rate.
-As can be seen from Fig.
-3a, for an emitter position 5 nm away from the surface, the Purcell factor can be very high, corresponding to over a million-fold increase in the rate of 2D LO phonon emission compared to photon emission into free-space.
-While this rate-enhancement is already quite high, it is a natural question how this coupling can be improved by nanostructuring, which will take the local density of electromagnetic states of Fig.
-3a and concentrate it around frequencies associated with resonances of the nanostructure.For simplicity, we consider the effect of nanostructuring the hBN into a disk, which leads to the formation of modes as in Fig.
-3b, in which we plot the surface charge density associated with each resonant mode.
-In Fig.
-3c, we plot the enhancement of the spontaneous emission rate for an external emitter polarized perpendicularly to the plane of the disk and placed 5 nm away from the surface of the disk.
+In Figure~\ref{fig:3}, we consider the \remove{quantum }coupling of a dipole emitter to \inset{the in }2D \swap{LO phonon modes}{localized phonon polaritons of} nanostructured monolayer hBN.
+For simplicity, we consider hBN nanostructured as a disk, which leads to the formation of sharp resonances\inset{, quantized along the azimuthal and radial directions}.
+In Figure~\ref{fig:3}, we plot the enhancement of the spontaneous emission rate for an external emitter polarized perpendicularly to the plane of the disk and placed \SI{5}{\nm} away from the surface of the disk.
 We find that the rate of spontaneous emission of 2D optical phonons is approximately 8 orders of magnitude larger than the rate of spontaneous emission of photons in the far field.
-For an infrared emitter at a transition wavelength of 7 $\mu$m with a free-space radiative lifetime of 1 $\mu$s, 5 nm away from an hBN disk, the coupling rate to 2D optical phonons would be on the same scale as the optical phonon frequency itself (about $2 \times 10^{14}$ $ s^{-1}$).
+For an infrared emitter at a transition wavelength of \SI{7}{\micro\m} with a free-space radiative lifetime of \SI{1}{\micro\s}, \SI{5}{\nm} away from an hBN disk, the coupling rate to 2D optical phonons would be on the same scale as the optical phonon frequency itself (about $\SI{2e14}{\per\s}$).
+\comment{tc}{We're using two different sets of unit for frequency (inverse seconds and inverse centimeters)}
 This coupling rate thus implies coupling between an emitter and the field in the regime of ultra-strong coupling.
@@
 In recent years, it has been shown that phonon-polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra}.
-Due to this remarkable confinement and their relatively high lifetimes of around picoseconds, phonon-polaritons open new opportunities for vibrational spectroscopy, radiative heat transfer \cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters \cite{kumar2015tunable,rivera2017making,kurman2018control}.
+Due to this remarkable confinement and their relatively high lifetimes of around picoseconds, phonon-polaritons open new opportunities for vibrational spectroscopy, radiative heat transfer~\cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters~\cite{kumar2015tunable,rivera2017making,kurman2018control}.
 The core features of phonon-polaritons, such as mode shape, confinement, and propagation characteristics, are understood from simple Lorentz oscillator models of the dielectric function, enabling successful theoretical accounts of experimental observations.
@@
 % It is also quite possible that they altogether vanish in the 2D case as opposed to the 3D case.
-As a recently discovered example of such a phenomenon, it was found that the LO-\mathrm{TO} splitting of some 2D polar dielectrics, such as hBN \cite{sohier2017breakdown} altogether vanishes, and only returns to its bulk value after a small but finite wavevector.
+As a recently discovered example of such a phenomenon, it was found that the LO-\mathrm{TO} splitting of some 2D polar dielectrics, such as hBN~\cite{sohier2017breakdown} altogether vanishes, and only returns to its bulk value after a small but finite wavevector.
 This would then suggest that a description of phonon-polaritons necessitates consideration of the spatially nonlocal dielectric function as in Equation (12), and that these polaritons moreover may only exist when their wavelengths are sufficiently short.
@@
 %We note that a similar situation appears in the treatment of 2D plasmons: namely that a relation like Equation~\eqref{eq:2ddispersion} appears, except with $\Pi$ the polarization-polarization response of a 2D electron gas.
-%And furthermore that one can also obtain the same equation by considering the ultra-thin limit of a 3D plasmonic material \cite{jablan2009plasmonics,jablan2013plasmons}.
+%And furthermore that one can also obtain the same equation by considering the ultra-thin limit of a 3D plasmonic material~\cite{jablan2009plasmonics,jablan2013plasmons}.
 %
@@
 %Perhaps it is possible that there are some materials in which the 2D optical phonons experience lower losses due to a reduced scattering phase space.
-%The formalism we provide here may also be extended to understand phonon-polaritons in other more atypical reduced-dimensional settings, such as zero-dimensional settings in single emitters, i.e., `molecular phonon-polaritons', in analogy to recent work on `molecular plasmons' \cite{manjavacas2013tunable,lauchner2015molecular}.
+%The formalism we provide here may also be extended to understand phonon-polaritons in other more atypical reduced-dimensional settings, such as zero-dimensional settings in single emitters, i.e., `molecular phonon-polaritons', in analogy to recent work on `molecular plasmons'~\cite{manjavacas2013tunable,lauchner2015molecular}.
 %Notably, we go beyond oversimplifications of the Lorentz oscillator model in which the Born charges are treated as a single scalar quantity and can treat the influence and interplay of many phonon modes that contribute to the dielectric function.
@@
 
+% --- ACKNOWLEDGEMENTS ---
 \section{Acknowledgements}
@@
 acknowledges support from the Danish Council for Independent Research (Grant No.\ DFF--6108-00667).
-The authors thank Joshua Caldwell (Vanderbilt), Dmitri Basov (Columbia), Ido Kaminer (Technion), Siyuan Dai (UT Austin), and Samuel Moore (Columbia) for helpful discussions.
+The authors thank Joshua Caldwell, Dmitri Basov, Ido Kaminer, Siyuan Dai, Samuel Moore, Jennifer Coulter, and Christopher Ciccarino for helpful discussions.
 This work was supported by the DOE Photonics at Thermodynamic Limits Energy Frontier Research Center under grant no.
@@
 
+% --- BIBLIOGRAPHY ---
 \bibliographystyle{apsrev4-1}

### 2018-12-26 paper.tex
revisions
small handful of revisions left pertaining to the figures

@@
 % --- TITLE & AUTHORSHIP ---
-\title{Nanophotonics with optical phonons in two dimensions}
+\title{Phonon-polaritonic reponse without LO-TO splitting in two-dimensional materials}
 
@@
 \begin{abstract}
-    Extreme confinement of electromagnetic energy by phonon-polaritons \swap{promises}{holds the promise of} extremely strong and novel forms of control over the dynamics of matter. 
+    Extreme confinement of electromagnetic energy by phonon-polaritons holds the promise of extremely strong and novel forms of control over the dynamics of matter. 
     To bring such control to its ultimate limit, it is important to consider phonon-polaritons in two dimensional systems.
-    \comment{tc}{It is not necessarily clear to me why 2D phonon-polaritonics would be an obvious component in bringing phonon-polaritons to their ``ultimate limit''?}
 Recent studies have pointed out that in two-dimensional systems, splitting between longitudinal and transverse optical (LO and TO) phonons is absent at the $\Gamma$ point.
-    Does this lack of LO--TO splitting imply the \swap{lack}{absence} of a \swap{strongly confined electromagnetic (phonon-polariton) mode}{phonon-polariton} in polar monolayers? 
-    Here, we derive a universal form for the spatially-nonlocal dielectric function of a polar monolayer specified by: the TO phonon frequency, the non-zero group velocity of the LO phonon, and the phonon lifetime.
-    \comment{tc}{I'm a bit hesitant to highlight the ``spatially-nonlocal'' dielectric function: the real reason this (simplified) form is nonlocal is just dimensional---\eg graphene's dielectric function is also nonlocal in a similar fashion (even in the low-momentum, low-frequency limit).
-The associated surface conductivity is local; so the in-plane polarization response is also local.}
-    Our analysis reveals that the phonon-polariton of the bulk is \swap{``replaced''}{essentially supplanted/replaced} by the LO phonon\inset{ of the two-dimensional system}.
-    We \swap{present}{discuss} the confinement and propagation losses of the LO \swap{phonons}{phonon-polaritons}, finding that high \remove{LO phonon} confinement and reasonable propagation quality factors coincide in regions which are difficult to detect with current near-field optical microscope techniques.
-    We conclude by studying the interaction of external emitters with two-dimensional hBN nanostructures, finding extreme enhancement of spontaneous emission due to \swap{2D phonon channels}{coupling with localized 2D phonon-polaritons}, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN \swap{phonons}{phonon-polaritons}.
+Does this lack of LO--TO splitting imply the absence of a phonon-polariton in polar monolayers?
+    Here, we derive an expression for  the spatially non-local conductivity of a polar monolayer specified by the wavevector-dependent LO and TO phonon dispersions.
+In the long-wavelength (local) limit, we find a universal form for the conductivity in terms of the TO phonon frequency at the $\Gamma$ point, its lifetime, and the group velocity of the LO phonon.
+Our analysis reveals that the phonon-polariton of the bulk is essentially replaced by the LO phonon of the two-dimensional system.
+We discuss the confinement and propagation losses of the LO phonons, finding that high  confinement and reasonable propagation quality factors coincide in regions which are difficult to detect with current near-field optical microscope techniques.
+    We conclude by studying the interaction of external emitters with two-dimensional hBN nanostructures, finding extreme enhancement of spontaneous emission due to coupling with localized 2D LO phonons, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN phonons.
 \end{abstract}
@@
 
-\comment{tc}{Both `phonon-polaritons' and `phonon polaritons' are used; let's settle on one.}
+% --- INTRODUCTION ---
+phonon-polaritons, hybrid quasiparticles of photons and optical phonons supported in polar materials, hold great promise for deeply sub-diffractional control of electromagnetic fields at mid-infrared and terahertz frequencies.
+Qualitatively, phonon-polaritons share many features in common with plasmon polaritons in conductors.
+Recently, it has been shown that phonon-polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra,li2018infrared,ma2018plane}.
+Due to this remarkable confinement and their relatively high lifetimes of around picoseconds, phonon-polaritons open new opportunities for vibrational spectroscopy,\cite{autore2018boron} radiative heat transfer~\cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters~\cite{kumar2015tunable,rivera2017making,kurman2018control}.
+In many cases, extreme confinement of phonon-polaritons is achieved by the use of thin-films, which shrink the in- and out-of-plane wavelength of polaritons monotonically with decreasing film thickness~\cite{dai2014tunable,dubrovkin2018ultra}.
+A monolayer is the ultimate limit of this effect when the properties of the bulk material can be used in the monolayer limit.
+In polar materials, the transition from 3D to 2D is not so simple, as in a polar monolayer, the LO--TO splitting that gives rise to phonon-polaritons in three-dimensions is absent at the $\Gamma$ point~\cite{sanchez2002vibrational,mele2002electric,serrano2007vibrational,sohier2017breakdown}.
+This raises a fundamental question about whether strongly confined electromagnetic modes exist in polar monolayers at all.
 
-\comment{tc}{For title, how about `Phonon polaritons in two-dimensional hexagonal boron nitride'?}
-
-% --- INTRODUCTION ---
-Phonon polaritons, hybrid quasiparticles of photons and optical phonons supported in polar materials, \swap{offer}{hold} great promise for deeply sub-diffractional control of electromagnetic fields at mid-\swap{IR}{infrared} and \swap{THz}{terahertz} frequencies.
-\swap{P}{Qualitatively, p}honon polaritons share many features in common with plasmon polaritons in conductors.
-\swap{In}{E.g., in} recent years, it has been shown that phonon-polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra,li2018infrared,ma2018plane}.
-Due to this remarkable confinement and their relatively high lifetimes of around picoseconds, phonon-polaritons open new opportunities for vibrational spectroscopy,\citeremind{} radiative heat transfer~\cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters~\cite{kumar2015tunable,rivera2017making,kurman2018control}.
-In many cases, extreme confinement of phonon polaritons is achieved by the use of thin-films, which \swap{shrink}{reduce} the in-\remove{plane} and out-of-plane wavelength of polaritons monotonically with decreasing film thickness~\cite{dai2014tunable,dubrovkin2018ultra}.
-\comment{tc}{The previous and following sentences both start with `many cases' here; best to rewrite one (probably the latter).}
-In many cases, a monolayer is the ultimate limit of this effect.
-Thus, to bring these classical and quantum nanophotonics applications to their ultimate limits, it is necessary to consider the properties of these modes in two-dimensions.
-\comment{tc}{I'm not sure that starting with `however' in this sentence is optimal: we want to convey an additional motivation for studying them (specifically, the transition from 3D to 2D is nontrivial)---maybe this sentence could be rewritten with this perspective in mind.}
-However, this picture of monolayers supporting maximally confined phonon-polariton modes is ostensibly complicated by the fact that in two-dimensional polar materials, the LO--TO splitting that gives rise to phonon-polaritons in three-dimensions is absent at the $\Gamma$ point~\cite{sanchez2002vibrational,mele2002electric,serrano2007vibrational,sohier2017breakdown}.
-
-In this Letter, we develop \swap{core results in}{a framework and a several concrete examples for understanding the} the nascent field of optical phononics in two-dimensional materials\swap{ by deriving}{.
-We do so by deriving} a universal form for the dielectric function of a polar monolayer, \swap{incorporating spatially nonlocal effects}{including spatial nonlocality}, which \swap{is specified fully in terms of the dispersion relation of LO and TO phonons in two dimensions}{which depends solely on the LO and TO phonon frequencies---and their dispersion with momentum---in the two-dimensional system}.
-Using parameters from Ref.~\citenum{sohier2017breakdown} for \swap{ important monolayers such as }{the canonical two-dimensional polar monolayer---}hexagonal boron nitride\swap{, }{---}we present the confinement and propagation losses of \swap{these modes}{the supported phonon-polaritons}, identifying the frequency region where they should be most easily detected.
-\swap{We then}{Finally, we} show that these modes \swap{essentially provide the ultimate limit of}{enable extreme} light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced by up to eight orders of magnitude by the emitter--LO phonon coupling\swap{, which due to}{.
-Given} the narrow bandwidth of phonon-polariton resonances, \swap{can lead to }{this suggests} the possibility of realizing the multi-mode strong coupling and ultra-strong coupling regimes of quantum electrodynamics \swap{for an external emitter coupled with optical phonons in hexagonal boron nitride}{in a two-dimensional hexagonal boron nitride platform}.
+In this Letter, we develop a framework and a several concrete examples for understanding the the nascent field of optical phononics in two-dimensional materials.
+We do so by deriving a universal form for the conductivity of a polar monolayer, including spatial nonlocality, which depends solely on the LO and TO phonon frequencies---and their dispersion with momentum---in the two-dimensional system.
+Using parameters from Ref.~\citenum{sohier2017breakdown} for the canonical two-dimensional polar monolayer --- hexagonal boron nitride --- we present the confinement and propagation losses of the 2D phononic modes, identifying the frequency region where they should be most easily detected.
+Finally, we show that these modes enable extreme light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced by up to eight orders of magnitude by the emitter--LO phonon coupling.
+Given the narrow bandwidth of phonon-polariton resonances, this suggests the possibility of realizing the multi-mode strong coupling and ultra-strong coupling regimes of quantum electrodynamics in a two-dimensional hexagonal boron nitride platform.
 
@@
 \section{Electrodynamics of optical phonons in two-dimensions}
-\comment{tc}{Structurally, this section would benefit from being split into two sections: one solely tasked with deriving/establishing the forms of the (non-interacting) polarization-polarization response function---and possibly its equivalent conductivity; and another one solely tasked with deriving and discussing the dispersion relation for the associated longitudinal optical modes.
+
+%\comment{tc}{Structurally, this section would benefit from being split into two sections: one solely tasked with deriving/establishing the forms of the (non-interacting) polarization-polarization response function---and possibly its equivalent conductivity; and another one solely tasked with deriving and discussing the dispersion relation for the associated longitudinal optical modes.
 Right now, this section goes back and forth between the two, which (I think) reduces its focus and gives way for unnecessary repetition.}
 
-In this section, we develop \swap{the}{a} theory of electromagnetic \swap{waves associated with}{response due to} optical phonons in two\swap{ dimensions}{-dimensional systems}.
-The crux of the theory \swap{is determining}{involves} the dielectric \swap{permittivity}{function} associated with the monolayer.
-\swap{It is obtained by considering the response of the ions in the monolayer to a longitudinal electromagnetic wave whose electric field is the gradient of a scalar potential $\phi$.}{To obtain it, we consider the response of the ions of the monolayer due(/when subjected) to a longitudinal electromagnetic potential $\phi$}.
-The \swap{coupling}{interaction} Hamiltonian is \swap{then given by}{in this case simply}
+In this section, we develop a theory of electromagnetic response due to optical phonons in two-dimensional systems.
+The crux of the theory involves the conductivity associated with the monolayer.
+To obtain it, we consider the response of the ions of the monolayer due(/when subjected) to a longitudinal electromagnetic potential $\phi$.
+The interaction Hamiltonian is in this case simply
 \begin{equation}
@@
 \end{equation} 
-with $\rho$ the \inset{induced} charge density and $\mathbf{P}$ the \inset{total} polarization density associated with the ionic motion. 
-\swap{The polarization \inset{density} can be expanded in powers of the displacement of the ions.
-In particular, we can consider the polarization $P_i$ generated in the monolayer when atom $\kappa$ of the unit cell is displaced along the $j$ direction by amount $u_{\kappa,j}$.
-To lowest order:}{Within linear response, the induced polarization density can be straightforwardly evaluated from the displacement amplitude $\mathbf{u}_{\kappa}$ of every atom $\kappa$ within the unit cell.
-Specifically, to first-order, the induced polarization is:}
-\comment{tc}{I don't see how the second equality can be consistent: since, by Taylor expansion, $P_i = P_i^0 + u_{\kappa,j}\partial_{u_{\kappa,j}}P_i + \mathcal{O}(u_{\kappa,j}^2)$---so where does the unit cell area $\Omega$ come from?
-The issue seems to arise from alternatingly interpreting the LHS/RHS $P_i$ as total/density quantities.
-The right-most equation also seems to be missing a $\kappa$-subscript in $u_{j}$.}
+with $\rho$ the induced charge density and $\mathbf{P}$ the total polarization density associated with the ionic motion. 
+Within linear response, the induced polarization density can be straightforwardly evaluated from the displacement amplitude $\mathbf{u}_{\kappa}$ of every atom $\kappa$ within the unit cell.
+Specifically, to first-order, the induced polarization is (in repeated-index notation):
 \begin{equation}
-    P_i - P_i^{0} = \Omega\frac{\partial P_i}{\partial u_{\kappa,j}}u_{\kappa,j} \equiv Z_{\kappa,ij}u_{j},
+    P_i - P_i^{0} = \frac{\partial P_i}{\partial u_{\kappa,j}}u_{\kappa,j} \equiv \frac{1}{\Omega}Z_{\kappa,ij}u_{\kappa,j},
 \end{equation} 
-where $Z_{\kappa,ij}$ is the Born effective charge tensor of ion $\kappa$ and $\Omega$ is the unit cell area.
-\comment{tc}{We do not give a definition of $P_i^0$: we also do not comment on what the implications are if it is nonzero (it could be..!)---as far as I can see, we are assuming it is zero for the remaining analysis.}
-With this relation between polarization and ionic displacements, \swap{we have expressed the Hamiltonian as a coupling between the scalar potential and ionic displacements}{the interaction Hamiltonian in Equation~\eqref{eq:coupling} couples the scalar potential and the ionic displacements}.
-We consider the response of the monolayer to a scalar potential of the form $\phi\inset{(\mathbf{r})} = \phi(\mathbf{q},\omega)\e^{\iu\mathbf{q}\cdot\mathbf{r}-\iu\omega t}$, where $\mathbf{q}$ is a two-dimensional wavevector in the plane of the monolayer.
-Such a potential corresponds to a longitudinal electric field \remove{of the form} $\mathbf{E}\inset{(\mathbf{r})} = \iu\mathbf{q}\swap{\phi(\mathbf{q},\omega)\e^{\iu\mathbf{q}\cdot\mathbf{r}-\iu\omega t}}{\phi(\mathbf{r})}$.
-\comment{tc}{I don't think it really ``follows'' from the preceding that $P_i = ...
-E_{j,\text{tot}}(\mathbf{q},\omega)$.
-First, this specializes to the RPA via a mean-field approximation, and second, it introduces a response function (so, really, it is more of a definition).
-It is probably better to phrase these next few sentences rather carefully and with specificity, \eg with references to the usual Kubo formula/approach etc.
-It may also make sense to split the $P_i = ...
-E_{j,\text{tot}}(\mathbf{q},\omega)$ off onto its own line, since this is the equation that lends practical utility to the $\boldsymbol{\Pi}$ response function.
-Finally, I don't think we actually need to mention the RPA or anything like it at this point: so far, this is just the non-interacting polarization-polarization response---we effectively arrive at the RPA later on, when we start doing electromagnetic calculations (\ie around Eq.~\eqref{eq:2dmaxwell}); only then do we actually make use of $\mathbf{P} \propto \boldsymbol{\Pi}\mathbf{E}_{\mathrm{tot}}$.}
-It follows that the induced polarization takes the form $P_i\inset{(\mathbf{q})}  = \epsilon_0\Pi_{ij}(\mathbf{q},\omega)E_{j,\mathrm{tot}}(\mathbf{q},\omega)$, where $\Pi_{ij}(\mathbf{q},\omega)$ is the polarization-polarization response function of the monolayer and  $E_{j,\mathrm{tot}}(\mathbf{q},\omega)$ is taken as the \textit{total} field which is a sum of the applied field and the field created by induced polarization.
-The polarization-polarization response function is given by
+where $Z_{\kappa,ij} \equiv \Omega\frac{\partial P_i}{\partial u_{\kappa,j}} $ is the Born effective charge tensor of ion $\kappa$ and $\Omega$ is the unit cell area.
+$P_i^{0}$ is the equilibrium polarization in the absence of displacements, which is zero here.
+With this relation between polarization and ionic displacements, the interaction Hamiltonian in Equation~\eqref{eq:coupling} couples the scalar potential and the ionic displacements.
+We consider the response of the monolayer to a scalar potential of the form $\phi(\mathbf{r}) = \phi(\mathbf{q},\omega)\e^{\iu\mathbf{q}\cdot\mathbf{r}-\iu\omega t}$, where $\mathbf{q}$ is a two-dimensional wavevector in the plane of the monolayer.
+Such a potential corresponds to a longitudinal electric field  $\mathbf{E}(\mathbf{r}) = \iu\mathbf{q}\phi(\mathbf{r})$.
+
+In what follows, we assume the validity of the random-phase approximation (RPA) for dealing with Coulombic interactions between ions in the polar lattice.
+Within the RPA, these Coulombic interactions are accounted for by taking the induced polarization $\mathbf{P}(\mathbf{q},\omega)$ to be proportional \textit{total} electric field, $\mathbf{E}_{\mathrm{tot}}(\mathbf{q},\omega)$, defined to be the sum of the externally applied electric field and the electric field created by the induced polarization.
+The polarization and total-field are connected by the polarization-polarization response (tensor) function  $\boldsymbol{\Pi}(\mathbf{q},\omega)$ via
+\begin{equation}
+\mathbf{P}(\mathbf{q},\omega)  = \epsilon_0\boldsymbol{\Pi}(\mathbf{q},\omega)\mathbf{E}_{\mathrm{tot}}(\mathbf{q},\omega),
+\end{equation}
+where boldfaced quantities refer to vectors or tensors as appropriate.
+From the Kubo formula, it follows that
 \begin{equation}\label{eq:2dsusceptibility}
@@
 \end{equation}
-where $m,n$ are states of the phononic Fock space of the monolayer, $P_{mn,i} = \sum_{\kappa,j}Z_{\kappa,ij}\langle m | u_{\kappa,j} | n \rangle$ are matrix elements of the polarization associated with phonon modes,
-\comment{tc}{Probably best to include $\mathbf{q}$ dependence explicitly in both LHS and RHS of the definition of $P_{mn,i}\textcolor{red!50!black}{(\mathbf{q})}$.}
-$E_{m}$ ($E_n$) is the energy of state $m$ ($n$), $\beta\inset{\equiv 1/k_{\text{\textsc{b}}}T}$ is the inverse temperature, and $\mathcal{Z}$ is the grand partition function.
-Evaluating the contribution of optical phonons in the harmonic approximation to the polarization-polarization response, one finds that the response function is given by: 
-\comment{tc}{It's not clear how we obtain Eq.~\eqref{eq:polpolresponse} from Eq.~\eqref{eq:2dsusceptibility}: \eg, I'm not sure what you mean by the harmonic approximation here?
-There's also a low-temperature approximation at play here.
-Finally, every phonon band except the LO one is dropped.
-Finally, we have not previously defined the $\mathbf{P}(\mathbf{q})$ polarization operator.
-I guess I'm just asking for us to be more explicit and detailed here.}
+where $m,n$ are states of the phononic Fock space of the monolayer, $P_{mn,i}(\mathbf{q}) = \sum_{\kappa,j}Z_{\kappa,ij}\langle m | u_{\kappa,j}(\mathbf{q}) | n \rangle$ are matrix elements of the polarization associated with phonon modes (where $u_{\kappa,j}(\mathbf{q})$ is the Fourier transform of the phonon displacement operator), $E_{m}$ ($E_n$) is the energy of state $m$ ($n$), $\beta \equiv 1/k_{\text{\textsc{B}}}T$ is the inverse temperature, and $\mathcal{Z}$ is the grand partition function.
+We now evaluate the contribution of optical phonons to the polarization-polarization response in the low-temperature limit $T << \frac{\hbar\omega_{\mathrm{ph}}}{k_B}$ with $\omega_{\mathrm{ph}}$ a characteristic optical phonon frequency.
+\textbf{Need to include details on $qq$ components.} Taking a material with long-wavelength isotropy, such as hBN,we only have to consider the $qq$-component in the response tensor, where $qq$ denotes a pair of directions parallel to the wavevector.
+We thus find that the response function is given by \citeremind{}:
 \begin{equation}
@@
 This is consistent with the fact that LO--TO splitting is a collective effect arising from Coulomb interactions and the fact that the equation above represents a single-particle susceptibility.
-Coulomb interactions are accounted for in the random phase approximation,
-\comment{tc}{Another reason for us to be more clear about the implications of the mean-field assumption before Eq.~\eqref{eq:2dsusceptibility}; might as well establish equivalence with RPA clearly and only once.}
-and to include them in the single-particle response \swap{is an}{would amount to an} uncontrolled double-counting. 
-\comment{tc}{Could it be better to interleave this $qq$-comment into the lead-up sentence for Eq.~\eqref{eq:polpolresponse}?
-Basically, we want to say that we specialize to the diagonal components.
-Finally; is it really necessary for us to make that scope reduction here?
-Wouldn't the formula nearly be simpler without it?}
-The component $qq$ in the response tensor denotes a pair of directions parallel to the wavevector.  
-\comment{tc}{It might be better to break this off into a new paragraph here (and possibly merge (and maybe shorten) with the following paragraph) since this is sort of orthogonal to the previous stuff.}
+Coulomb interactions are accounted for in the random phase approximation, and to include them in the single-particle response amounts to an uncontrolled double-counting.
+
+To proceed, we express the polarization matrix element in Equation~\eqref{eq:polpolresponse} in terms of parameters of the monolayer such as the Born effective charges. 
+Considering the longitudinal phonon contribution to the second-quantized ionic displacement, as in Ref.~\citenum{srivastava1990physics}, one finds that the RPA dielectric function is given by:
+\begin{equation}
+    \epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}} - \frac{q}{2\epsilon_0}\frac{\big|\hat{q}\cdot\sum_{\kappa} \mathbf{Z}_{\kappa}\eta_{\kappa}  \big|^2}{\omega^2_{\mathrm{TO}}-\omega^2},
+    \label{eq:epsRPA},
+\end{equation}
+  where $\hat{q}$ is a unit vector along $\mathbf{q}$.
+In this expression, $\omega_\mathrm{L}$, the LO phonon frequency prior to LO-T0 splitting, has been re-named as $\omega_{\mathrm{TO}}$, the transverse optical phonon frequency, because in the absence of LO--TO splitting, they are degenerate.
+This is done in order to make the form of the final results more closely resemble their 3D counterparts.
+Additionally, we have defined scaled eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}\equiv \frac{\hat{e}_{\kappa\mathbf{q},\mathrm{L}}}{\sqrt{M_{\kappa}}}$, where $\hat{e}_{\kappa\mathbf{q},\mathrm{L}}$ is the unit-normalized polarization vector of atom $\kappa$ in the unit cell oscillating according to a longitudinal phonon of wavevector $\mathbf{q}$ and $M_{\kappa}$ is the mass of atom $\kappa$.
+
+  %\comment{tc}{I don't think this needs to be as long as it currently is---also, some of the details, \eg position of Reststrahlen band, assumption of long wave-length etc., would probably fit in more meaningfully when those approximations are actually applied.
+For Eq.~\eqref{eq:2dmaxwell}, only the homogeneity assumption is necessary (corresponding to in-plane translation invariance)}
+
 To relate the polarization-polarization response function to the electromagnetic modes supported by a polar monolayer, we solve Maxwell's equations for an evanescent electromagnetic mode supported by a surface with polarization-polarization response tensor $\mathbf{\Pi}$.
 We consider the monolayer to be sandwiched by a superstrate of permittivity $\epsilon_1$ and a substrate of permittivity $\epsilon_2$.
-
-\comment{tc}{I don't think this needs to be as long as it currently is---also, some of the details, \eg position of Reststrahlen band, assumption of long wave-length etc., would probably fit in more meaningfully when those approximations are actually applied.
-For Eq.~\eqref{eq:2dmaxwell}, only the homogeneity assumption is necessary (corresponding to in-plane translation invariance)}
 To strip the analysis to its bare essentials, we consider optical phonon response with in-plane isotropy in the long-wavelength limit arising from in-plane LO oscillations.
-A relevant example of a system where these conditions are satisfied is in a hexagonal boron nitride monolayer, considering optical phonons in the so-called upper Reststrahlen band which in bulk spans the frequency range of \SIrange[range-phrase=\text{ to }]{1360}{1610}{\per\cm}.
+A relevant example of a system where these conditions are satisfied is in a hexagonal boron nitride monolayer (see Fig.
+1(a) for schematic atomic structure).
 In a monolayer geometry with translation invariance and in-plane isotropy, the electromagnetic modes of Maxwell's equations can be composed into transverse magnetic (TM) and transverse electric (TE), where the magnetic or electric field respectively is transverse to the in-plane wavevector of the mode.
 In practice, it is the TM mode which is associated with highly confined electromagnetic waves that are of use in nanophotonics.
-TE waves are not supported at the same frequency as transverse magnetic waves, and require exotic conditions to be realized in two-dimensional materials.
 Given the isotropy, we may suppress indices from the response tensor.
-We may also consider without \remove{further} loss of generality a TM mode with wavenumber $q$ along the $x$-direction in the monolayer and magnetic field $H(z)\e^{\iu qx-\iu\omega t}$ along the $y$-direction of the monolayer.
+We may also consider without loss of generality a TM mode with wavenumber $q$ along the $x$-direction in the monolayer and magnetic field $H(z)\e^{\iu qx-\iu\omega t}$ along the $y$-direction of the monolayer.
 The direction transverse to the monolayer is denoted as $z$.
@@
 \begin{equation}\label{eq:2dmaxwell}
-    \bigg(-\frac{\dd^2}{\dd{}z^2}+q^2-\epsilon_{\mathrm{env}}\frac{\omega^2}{c^2} \bigg)H(z) = 0.
+    \bigg(-\frac{\dd^2}{\dd{}z^2}+q^2-\epsilon_{1,2}\frac{\omega^2}{c^2} \bigg)H(z) = 0,
 \end{equation}
-\comment{tc}{I'm not sure this needs to be its own equation.
-Even if it does, I don't think that the term $\epsilon_{\mathrm{env}}\omega^2/c^2$ is correct; shouldn't this really be $\epsilon_{1,2}\omega^2/c^2$, with choice of $\{1,2\}$ depending on sign of $z$?}
-We consider a solution of the form $H(z) = \e^{-\kappa_1 z}$ for $z > 0$ with $\kappa_1 = \sqrt{q^2-\epsilon_1\frac{\omega^2}{c^2}}$ and a solution of the form $H(z) = c\e^{\kappa_2 z}$ for $z < 0$ with $\kappa_2 = \sqrt{q^2-\epsilon_2\frac{\omega^2}{c^2}}$.
-\comment{tc}{Should probably choose a different name for the constant `$c$' to avoid clashing with the speed of light $c$.
-Maybe $c_1$ or $a$.
-Or, alternatively, let $H(z) = h_\pm\e^{\mp\kappa_\pm z}$ with $\kappa_\pm = \sqrt{q^2-\epsilon_\pm k_0^2}$ with upper/lower signs applicable for $\pm z>0$ (instead of using $\{1,2\}$ subscripts)---and then specify the boundary condition as $h_-/h_+ = \ldots (=c)$.
-Personally, I favor the latter option but it requires a few more fixes.
-If you go with that option, you can probably drop the parentheses around $H_y^{(\pm)}$ in the following.}
-The boundary condition on the magnetic field is $H_y^{(+)}-H_y^{(-)} = -K_x = \iu\omega\mathbf{P}_x$ where $\mathbf{K}$ is the surface current density, expressed through the polarization density $\mathbf{P}$ of the monolayer.
-This condition enforces $c =(1+\frac{\epsilon_0}{\epsilon_2} \kappa_2\Pi)^{-1}$. 
-Continuity of the electric field in the $x$ direction enforces $\frac{\epsilon_2}{\epsilon_1} = -\frac{\kappa_2}{\kappa_1}c$. 
-\swap{These two conditions fully specify the TM mode}{Combining the two conditions, we obtain the usual dispersion equation for the TM mode of a polarizable 2D monolayer, namely $\epsilon_1/\kappa_1 + \epsilon_2/\kappa_2 +\epsilon_0\Pi=0$.}.
-%
-\swap{To simplify the discussion, we consider the so-called electrostatic limit, it in which the wavenumber $q$ of the TM mode is much larger than the free-space wavenumber $\frac{\omega}{c}$.
-This is well-satisfied in monolayer optical phonons in 2D.
-With this approximation, the condition for a TM mode corresponding to an optical phonon is}
-{Given the deeply subwavelength nature of 2D phonon-polaritons, \ie since $q\gg\omega/c$ such that $\kappa_{1,2}\simeq q$, the dispersion equation can be reduced to its quasistatic limit without consequential loss of accuracy:}
+where $\epsilon_1$ applies for $z>0$ and $\epsilon_2$ applies for $z<0$.
+We consider a solution of the form $H(z) = h_\pm\e^{\mp\kappa_\pm z}$ with $\kappa_\pm = \sqrt{q^2-\epsilon_{1,2} k_0^2}$ with $\pm$ and $1,2$ corresponding each to $\pm z > 0$ respectively.
+The boundary condition on the magnetic field is $h^+-h^- = -K_x = \iu\omega\mathbf{P}_x$ where $\mathbf{K}$ is the surface current density, expressed through the polarization density $\mathbf{P}$ of the monolayer.
+This condition enforces $\frac{h^-}{h^+} =(1+\frac{\epsilon_0}{\epsilon_2} \kappa_2\Pi)^{-1}$.
+Continuity of the electric field in the $x$ direction enforces $\frac{\epsilon_2}{\epsilon_1} = -\frac{\kappa_2}{\kappa_1}\frac{h^-}{h^+}$. 
+Combining the two conditions, we obtain the usual dispersion equation for the TM mode of a polarizable 2D monolayer, namely $\epsilon_1/\kappa_1 + \epsilon_2/\kappa_2 +\epsilon_0\Pi=0$.
+Given the deeply subwavelength nature of 2D phonon-polaritons, \ie since $q\gg\omega/c$ such that $\kappa_{1,2}\simeq q$, the dispersion equation can be reduced to its quasistatic limit without consequential loss of accuracy:
 \begin{equation}
-    \epsilon_{\mathrm{env}} + \frac{1}{2}\epsilon_0 q \Pi(q,\omega) \equiv \epsilon_{\mathrm{RPA}}(\omega) = 0.
+    \epsilon_{\mathrm{env}} + \frac{1}{2}\epsilon_0 q \Pi(q,\omega) \equiv \epsilon_{\mathrm{RPA}}(\omega) = 0,
     \label{eq:dispeq}
 \end{equation}
-\comment{tc}{Need to define $\epsilon_{\mathrm{env}}\equiv (\epsilon_1+\epsilon_2)/2$.}
-The quantity $\epsilon_{\mathrm{RPA}}(\omega) = 0$ coincides with the phonon contribution to the longitudinal dielectric function of the monolayer in the random-phase approximation. 
-\comment{tc}{I do not completely understand the intention of this last sentence: are we trying to say that this coincides with the zeros of the dielectric function?
+with $\epsilon_{\mathrm{env}}\equiv (\epsilon_1+\epsilon_2)/2$.
+The quantity $\epsilon_{\mathrm{RPA}}(\omega)$ precisely coincides with the optical phonon contribution to the RPA dielectric function of the monolayer.
+\textbf{Probably want to explain more how I know this is the RPA dielectric function despite getting it from Maxwell.}
+%\comment{tc}{I do not completely understand the intention of this last sentence: are we trying to say that this coincides with the zeros of the dielectric function?
 If so, maybe we should be a bit more explicit in our thinking here (e.g., we can say that we identify the LHS with the RPA dielectric function---and maybe comment on that the two quantities naturally have the same zeros)}
 
-\comment{tc}{Wouldn't it be more meaningful to get these kind of formal things all handled before we start discussing the 2D excitations?
-Specifically, we might as well introduce the eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}$ when we discuss $\boldsymbol{\Pi}$?
-That way, the discussion of modes could be more focused and wouldn't need this detour.
-}
-To proceed, we express the polarization matrix element in Equation~\eqref{eq:polpolresponse} in terms of parameters of the monolayer such as the Born effective charges. 
-Considering the longitudinal phonon contribution to the second-quantized ionic displacement, as in Ref.~\citenum{srivastava1990physics}, one finds that the RPA dielectric function is given by:
-\begin{equation}
-    \epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}} - \frac{q}{2\epsilon_0}\frac{\big|\hat{\mathbf{q}}\cdot\sum_{\kappa,j} Z_{\kappa,ij}\eta_{\kappa,j}  \big|^2}{\omega^2_{\mathrm{TO}}-\omega^2}.
-    \label{eq:epsRPA}
-\end{equation}
-\comment{tc}{There's an issue with Eq.~\eqref{eq:epsRPA} AFAICT (and with all other occurrences of $\sum_{\kappa,j}Z_{\kappa,ij}\eta_{\kappa,j}$): it has a lingering $i$-index dependence.
+%\comment{tc}{There's an issue with Eq.~\eqref{eq:epsRPA} AFAICT (and with all other occurrences of $\sum_{\kappa,j}Z_{\kappa,ij}\eta_{\kappa,j}$): it has a lingering $i$-index dependence.
 I guess this is the result of a mix between vector- and index-notation.}
-In this expression, $\omega_\mathrm{L}$ has been re-named as $\omega_{\mathrm{TO}}$, because in the absence of LO--TO splitting, they are degenerate. 
-\comment{tc}{There's some unpleasantness in substituting an \emph{longitudinal} mode for a \emph{transverse} mode in the $\mathrm{L}\rightarrow\mathrm{TO}$ swap here.
-I guess it might be good to be reminded here that $\omega_{\mathrm{L}}$ is the longitudinal phonon mode \emph{before} LO--TO splitting (I understand that this sentence is trying to hint that; but a more direct hint may be useful).}
-Additionally, we have defined \inset{scaled} eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}\equiv \frac{\mathbf{e}_{\kappa\mathbf{q},\mathrm{L}}}{\sqrt{M_{\kappa}}}$, where $\mathbf{e}_{\kappa\mathbf{q},\mathrm{L}}$ is the unit-normalized polarization vector
-\comment{tc}{We had hats on $\mathbf{e}_{\kappa\mathbf{q}}$ in the previous paper.}
-of atom $\kappa$ in the unit cell oscillating according to a longitudinal phonon of wavevector $\mathbf{q}$ and $M_{\kappa}$ is the mass of atom $\kappa$.
-	\comment{tc}{A few things: 
-	\begin{enumerate}
-	\item Here, we switch to using vectorial notation.
-My preference is probably that we use it whenever possible, but we have used index-notation elsewhere (also in cases where vector notation is not awkward)---we should probably try to stick to one set of notations.
-I think it's meaningful to go with mainly vector-notation, since that's also what we used in the previous paper.
-	\item I recall something I think I also asked about in the previous paper: shouldn't there be an $\e^{\iu\mathbf{q}\cdot\mathbf{b}_\kappa}$ dependence here?
-Looking at the previous paper's Eq.~(8b) and the text immediately after, that also seems to be an issue there (compare the definition of $\mathbf{S}_{\mathbf{q}\sigma}$ with (8b) and (9); things don't end up agreeing, AFAICT).
-What's the argument for dropping this exponential factor?
-	\item This is a more general comment: the notation we use here deviates quite a bit from the notation we used in the previous paper: wouldn't it make more sense to go with essentially the same notation?
-More generally, reading through this, I occasionally feel that the level of detail is in the ``uncanny valley'': it gives some things with substantial detail, and for several other things, many will feel a need to consult our previous paper.
-One could argue (I'm not necessarily doing so, just raising it for consideration) that the derivation of the 2D polarization-polarization response (specifically, its structure) does not involve sufficient differences between 2D and 3D to merit a new derivation here.
-E.g., we could almost start with Eq.~(9) from our previous paper (with $V\rightarrow \Omega$, instead of bothering to go through the derivation.
-Of course, that makes the content somewhat more sparse, but it would possibly be a more focused paper that way.
-Anyway, something to consider.
-	\item We discuss the dielectric function/permittivity of the 2D sheet quite a bit.
-In my view, though, it would be much more natural (and, I think, more immediately recognizable to researchers in the 2D optical-excitations field) to discuss the conductivity instead.
-This would also make for a clearer discussion of which parts of the response actually are nonlocal (since, in the low-$q$ limit, the 2D permittivity of any material must have the generic form $\epsilon\simeq 1 + \alpha q$ (vacuum cladding)).
-	\end{enumerate}}
 
@@
 In 3D, there is a finite LO--TO splitting at zero wavevector, while in 2D there is none.
-Despite this, the 2D LO phonon plays the role of the phonon polariton in 3D and thin films.
+Despite this, the 2D LO phonon plays the role of the phonon-polariton in 3D and thin films.
 (c) Analogous physics appears in electron gases in 3D and 2D, where the 3D plasma frequency is analogous to the 3D LO--TO splitting.
@@
 
-\swap{As is well known, t}{T}he zeros of the RPA dielectric function, denoted $\omega_{\mathbf{q}}$, correspond to longitudinal electromagnetic \swap{waves}{excitations}.
-\swap{Thus, the condition for the zeros of the dielectric function can be expressed as}{Thus, from Equation~\eqref{eq:epsRPA}, we obtain the dispersion equation for the 2D phonon polaritons:}
-\comment{tc}{I think we are repeating/cycling a bit here; we already mentioned this previously, and also already derived a form for the phonon-polariton dispersion equation (Eq.~\eqref{eq:dispeq}).}
+From Equation~\eqref{eq:epsRPA}, we obtain the dispersion equation for the 2D phonon-polaritons:
 \begin{align}
-    \omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} &= \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2 \nonumber \\
-    &= V(q)\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2,
+    \omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} &= \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa}\tilde{\mathbf{Z}}_{\kappa}\boldsymbol{\eta}_{\kappa}  \Big|^2 \nonumber \\
+    &= V(q)\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa}\tilde{\mathbf{Z}}_{\kappa}\boldsymbol{\eta}_{\kappa}  \Big|^2,
     \label{eq:dispersion}
 \end{align}
-where $eZ^r \equiv Z$, 
-\comment{tc}{I guess I see the motivation for introducing a quantity like $Z^r$---but it's not obvious to me why it should be an $r$-superscript?
-Maybe $\tilde{Z}$?
-Oh, in any event, this definition needs subscripts.}
-with $e$ the electron charge, and $V(q)$ is the Coulomb interaction in two dimensions, \comment{tc}{Probably should define $V(q)$ explicitly (to remind that $V(q)\propto 1/q$)} screened by the dielectric environment of the sub- and superstrates.
+where $e\tilde{\mathbf{Z}} \equiv \mathbf{Z}$, with $e$ the electron charge, and $V(q)=\frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}$ is the Coulomb interaction in two dimensions, screened by the dielectric environment of the sub- and superstrates.
 This result, as we now show, reveals that the TM mode corresponds simply to the two-dimensional optical phonon.
 As shown recently in Ref.~\citenum{sohier2017breakdown}, in two-dimensional polar materials the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO--TO splitting.
-\comment{tc}{This might be a good spot to mention/highlight that this has an analogy in 2D plasmonics: the 2D plasmon is just the conventional ``bulk'' plasmon of the 2D sheet---\ie there is no notion of a surface plasmon there.
-The intrinsic mode becomes the surface mode, and vice versa.}
-The expression for the LO--TO splitting is none other than Equation~\eqref{eq:dispersion} of the current manuscript.
-\comment{tc}{Might be a good spot to point out that the splitting vanishes in the $q\rightarrow 0$ limit.}
+The expression for the LO--TO splitting is none other than Equation~\eqref{eq:dispersion} of the current manuscript, which vanishes in the long-wavelength ($q\rightarrow 0$) limit.
 \emph{Thus, one of the main results of our manuscript is that the wave solutions Maxwell's equations associated with phonon-polaritons in bulk materials, correspond to pure LO phonons in the monolayer limit.}
-\comment{tc}{Probably not a good idea to have large block of text in italics, even if we feel it is important.}
+
+It has been well-appreciated in the literature on phonon-polaritons in thin slabs.
+Here, we expand on the analogy in the two-dimensional case.
+To do so, we take Equation~\eqref{eq:dispersion} in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $e^2\big|\sum_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \big|^2 \equiv  \Big(\frac{eZ^r_*}{\sqrt{M_*}}\Big)^2 \equiv \Big(\frac{Q_*}{\sqrt{M_*}}\Big)^2$.
+Then, the LO--TO splitting can be written as $\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} = \frac{Q_*^2}{2\epsilon_0\epsilon_{\mathrm{env}} M_*}q$.
+Now we note that the RHS is exactly the squared-frequency $\omega^2_{\mathbf{q}p}$ for a plasma oscillation in a 2D gas of charged particles with charge $Q_*$ and mass $M_*$.
+This squared frequency can be thought of as the ``LP--TP'' splitting between longitudinal and transverse plasma oscillations.
+Of course, there are no transverse plasma oscillations, and so ``$\omega_{\mathrm{TP}} = 0$''.
+In the three-dimensional plasmon case, `$\omega_{\mathrm{TP}} = 0$', but the difference between the squared longitudinal and transverse plasma oscillation frequencies at zero-wavevector is non-zero and given by $\omega_p^2$.
+This analogy between phononic and plasmonic behavior as a function of dimension is illustrated schematically in Figures~\ref{fig:1}(b,c).
 
@@
 
-It has been well-appreciated in the literature on phonon polaritons in thin slabs.
-Here, we expand on the analogy in the two-dimensional case.
-To do so, we take Equation~\eqref{eq:dispersion} in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $e^2\big|\sum_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \big|^2 \equiv  \Big(\frac{eZ^r_*}{\sqrt{M_*}}\Big)^2 \equiv \Big(\frac{Q_*}{\sqrt{M_*}}\Big)^2$.
-Then, the LO--TO splitting can be written as $\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} = \frac{Q_*^2}{2\epsilon_0\epsilon_{\mathrm{env}} M_*}q$.
-Now we note that the RHS is exactly the squared-frequency $\omega^2_{\mathbf{q}p}$ for a plasma oscillation in a 2D gas of charged particles with charge $Q_*$ and mass $M_*$.
-This squared frequency can be thought of as the ``LP--TP'' splitting between longitudinal and transverse plasma oscillations.
-Of course, there are no transverse plasma oscillations, and so ``$\omega_{\mathrm{TP}} = 0$''.
-In the three-dimensional plasmon case, `$\omega_{\mathrm{TP}} = 0$', but the difference between the squared longitudinal and transverse plasma oscillation frequencies at zero-wavevector is non-zero and given by $\omega_p^2$.
-This analogy between phononic and plasmonic behavior as a function of dimension is illustrated schematically in Figures~\ref{fig:1}(b,c).
-\comment{tc}{We never mention Fig.~\ref{fig:1}(a). We should try to mention/ref' every subfigure at least once in the text.}
-
 % -----------------
@@
 
-In Figure~\ref{fig:2}(a), we present the dispersion relation, operationally defined through the poles of the Fresnel reflection coefficient for $p$-polarized waves, which for a monolayer of conductivity $\sigma(\mathbf{q},\omega) = -\frac{2\iu \omega}{\epsilon q}(\epsilon_{\mathrm{RPA}} - \epsilon_{\mathrm{env}})$
-\comment{tc}{We already have a definition of $\sigma(\mathbf{q},\omega)$ previously ($=-\iu\omega\Pi(\mathbf{q},\omega)$), so I don't think we need this additional definition here.
-If anything, it would be nice to have $\sigma(\mathbf{q},\omega)$ in terms of $\omega_{(\mathrm{TO},\mathrm{LO})}$ (including its local limit).}
-below vacuum and on top of a substrate of permittivity $\epsilon_{\mathrm{s}}$, is given by
-\comment{tc}{We should probably not have two competing but equivalent substrate permittivity definitions (\ie we have both $\epsilon_{\mathrm{s}}$ and $\epsilon_1$ right now; we could just stick with $\epsilon_1$ (or, even better, $\epsilon_-$).
-In that case, we could also just give the general form of $r_p$ with $(\varepsilon_- - \varepsilon_+)+\ldots$ in the numerator and $(\varepsilon_-+\varepsilon_+)+\ldots$ in the denominator.}
-\comment{tc}{The $-\sigma q/\epsilon_0\omega$ terms are missing an imaginary unit. I'm pretty sure it should be $+\iu\sigma q/\epsilon_0\omega$ but you may want to check the sign.}
+In Figure~\ref{fig:2}(a), we present the dispersion relation, operationally defined through the poles of the Fresnel reflection coefficient for $p$-polarized waves, which for a monolayer of conductivity $\sigma(\mathbf{q},\omega)$
+below vacuum and on top of a substrate of permittivity $\epsilon_{2}$, is given by
 \begin{equation}
-    r_p(\mathbf{q},\omega) = \frac{(\epsilon_{\mathrm{s}}-1)-\frac{\sigma(\mathbf{q},\omega) q}{\omega\epsilon_0}}{(\epsilon_{\mathrm{s}}+1)-\frac{\sigma(\mathbf{q},\omega) q}{\omega \epsilon_0}}.
+    r_p(\mathbf{q},\omega) = \frac{(\epsilon_{2}-1)-i\frac{\sigma(\mathbf{q},\omega) q}{\omega\epsilon_0}}{(\epsilon_{2}+1)-i\frac{\sigma(\mathbf{q},\omega) q}{\omega \epsilon_0}}.
 \end{equation}
-In Figure~\ref{fig:2}(b), we show the confinement factor (or effective mode index) of the 2D LO phonon mode, defined by $n_{\mathrm{eff}} = \frac{qc}{\omega}$. 
-\comment{tc}{I guess you want a real-part sign somewhere in this definition?
-(Of course, it's reasonable to not do this if it's just interpreted as an effective index, but questionable if interpreted as a confinement factor.}
-The confinement factor is a key figure of merit in many nanophotonic applications involving quantum emitters,\swap{ because as was shown in~\cite{archambault2010quantum,koppens2011graphene,rivera2016shrinking,rivera2017making},}{ since} all interactions involving emission of electromagnetic energy by emitters\swap{, such as}{---\eg} electric and magnetic dipole emission, multipole emission, phosphorescence, and multiphoton emission\swap{, }{---}scale \inset{approximately} as some power of this confinement factor\swap{.}{~\cite{archambault2010quantum,koppens2011graphene,rivera2016shrinking,rivera2017making}.}
+In Figure~\ref{fig:2}(b), we show the confinement factor (or effective mode index) of the 2D LO phonon mode, defined by $n_{\mathrm{eff}} = \frac{\text{Re } qc}{\omega}$. 
+The confinement factor is a key figure of merit in many nanophotonic applications involving quantum emitters, since all interactions involving emission of electromagnetic energy by emitters---\eg electric and magnetic dipole emission, multipole emission, phosphorescence, and multiphoton emission ---scale approximately as some power of this confinement factor~\cite{archambault2010quantum,koppens2011graphene,rivera2016shrinking,rivera2017making,sloan2018extreme}.
 
-Interestingly, in the case of a \inset{polar} monolayer, the confinement grows very rapidly in frequency, due to the extremely low group velocity of 2D LO phonons, which is a remarkable four orders of magnitude slower than the speed of light.
-In particular, at frequencies of \SI{1450}{\per\cm}, the LO phonon already has a confinement factor in excess of \num{400}, corresponding to a phonon wavelength below \SI{20}{\nm}, which is much shorter wavelength than any phonon polariton \remove{thus} measured so far, \swap{as well as shorter wavelength than any plasmon that has been measured in the mid-\swap{IR}{infrared}, such as in graphene.}{and, similarly shorter than any plasmonic wavelength, even in graphene.}
-In fact, this short\inset{ a} wavelength is \swap{mostly}{essentially} \swap{out of the range of measurability in}{beyond the range of} current scattering near-field microscope (SNOM) measurements, \swap{which with current tip radii, are limited to measuring wavelengths no shorter than \SI{60}{\nm}}{which are limited (by tip radii) to wavelengths greater than $\approx\,\SI{60}{\nm}$}.
-\comment{tc}{How did Koppens measure his nonlocal stuff then? 60 nm is not that high.}
-\swap{This 2D LO phonon can be measured with SNOM at frequencies close to the TO phonon frequency, where the confinement is less, but this comes at the price of very high dissipation and correspondingly low propagation quality factor, as illustrated in Figure~\ref{fig:2}(c), where $\Re q/\Im q < 5$.
-This is simply because frequencies near the TO phonon resonance correspond to very high dissipation (measured by $\Re\sigma$ or $\Im\epsilon_{\mathrm{RPA}}$).}
-{The 2D LO phonon polariton could in principle be measured closer to the TO frequency, where confinement is smaller; unfortunately, as shown in Figure~\ref{fig:2}(c), near the TO frequency, dissipation is far higher (and corresponding propagative quality factors $\Re q/\Im q$ far lower) \optional{due to large $\Re\sigma$ (or, equivalently, large $\Im\epsilon_{\mathrm{RPA}}$)}.}
+Interestingly, in the case of a polar monolayer, the confinement grows very rapidly in frequency, due to the extremely low group velocity of 2D LO phonons, which is a remarkable four orders of magnitude slower than the speed of light.
+In particular, at frequencies of \SI{1450}{\per\cm}, the LO phonon already has a confinement factor in excess of \num{400}, corresponding to a phonon wavelength below \SI{20}{\nm}, which is much shorter wavelength than any phonon-polariton measured so far, and, similarly shorter than any plasmonic wavelength, even in graphene.
+In fact, this short a wavelength is essentially out of the range of measurability beyond the range of current scattering near-field microscope (SNOM) measurements, which are limited (by tip radii) to wavelengths greater than $\approx\,\SI{60}{\nm}$.
+
+The 2D LO phonon-polariton could in principle be measured closer to the TO frequency, where confinement is smaller; unfortunately, as shown in Figure~\ref{fig:2}(c), near the TO frequency, dissipation is far higher (and corresponding propagative quality factors $\Re q/\Im q$ far lower) due to large $\Re\sigma$ (or, equivalently, large $\Im\epsilon_{\mathrm{RPA}}$).
 \comment{tc}{Might be good to actually plot $\sigma(\omega)$}
-These considerations imply that access to the lower-loss and higher-confined portions of the \swap{phononic waves}{2D phonon polaritons}, in the absence of a sharp tip, requires either a free electron probe, as in EELS, \comment{tc}{Acronym without definition} where slow electrons can be used to probe plasmon wavelengths of just a few nanometers in monolayer metals, as well as the nonlocal bulk plasmon dispersion in metals~\cite{nagao2001dispersion, de2010optical}{}\citeremind{diaconescu2007acoustic}.
-\comment{tc}{This sentence doesn't quite work, because it has an `either' construction but no second option; I guess an ``emitter-part'' is missing?}
-EELS has been recently employed to measure phonon polaritons in ultrathin films of hBN~\cite{govyadinov2017probing}.
-\swap{Yet a}{A}nother interesting \swap{type of probe}{class of probes}, with relevance to fundamental physics and quantum optics applications, is a quantum emitter such as an atom, molecule, or quantum well.
-A quantum emitter\swap{, which one can consider as a sub-nanoscale tip, }{---qualitatively akin to a sub-nanoscale SNOM tip---}can \swap{see}{efficiently excite} very high wavevectors of the 2D LO phonon if sufficiently close to the polar monolayer.
-\comment{tc}{This last sentence is a bit unmotivated and feels disconnected. Is it necessary?}
-Recently, it was demonstrated using nanostructures of bulk hBN that the interaction of vibrational emitters with phonon polaritons is on the border of the strong coupling regime~\cite{autore2018boron}.
-\comment{tc}{It feels strange to may that we have this discussion of quantum emitters here without discussing Fig.~\ref{fig:3}---there's no need for us to pick up the ball twice.}
+These considerations imply that access to the lower-loss and higher-confined portions of the dispersion relation of phonon-polaritons, in the absence of a sharp tip, requires a near-field probe such as a free electron probe, as used in electron energy loss spectroscopy (EELS),  where slow electrons can be used to probe plasmon wavelengths of just a few nanometers in monolayer metals, as well as the nonlocal bulk plasmon dispersion in metals~\cite{nagao2001dispersion, de2010optical,diaconescu2007acoustic}.
+EELS has been recently employed to measure phonon-polaritons in ultrathin films of hBN~\cite{govyadinov2017probing}.
+Another interesting class of near-field probes, with relevance to fundamental physics and quantum optics applications, is a quantum emitter such as an atom, molecule, or quantum well.
+Recently, it was demonstrated using nanostructures of bulk hBN that the interaction of vibrational emitters with phonon-polaritons is on the border of the strong coupling regime~\cite{autore2018boron}.
 
@@
 This can never happen by extrapolating to a monolayer a dielectric function which has finite LO--TO splitting.
-\comment{tc}{Generally, I think this discussion would be much improved by actually having calculations of a thin slab (like we discussed; similarly to what we did for the Argentene paper) of varying ``thinness'', and seeing how the calculations agree/disagree.}
 
@@
     \caption{%
-        \textbf{Extreme spontaneous emission enhancement due to 2D LO phonons in nanostructured geometries.} Enhancement of the spontaneous emission rate for an emitter concentric \comment{tc}{Maybe go with simpler wording?} with the disk at a distance $z=\SI{5}{\nm}$ away from the disk.
-For a disk with a diameter of \SI{20}{\nm}, the rate of emission enhancement can be \inset{enhanced} \num{100} million-fold.
+        \textbf{Extreme spontaneous emission enhancement due to 2D LO phonons in nanostructured geometries.} Plotted is the enhancement of the spontaneous emission rate for an emitter.
+The emitter is concentric with the disk and is at a distance $z=\SI{5}{\nm}$ away from the disk.
+For a disk with a diameter of \SI{20}{\nm}, and assuming a relaxation rate $\tau^{-1} = 5$ cm$^{-1}$, the rate of emission enhancement can be enhanced \num{100} million-fold.
 For an emitter with a free-space decay rate of $\SI{1e6}{\per\cm}$ at \SI{7}{\micro\m}, the emitter would experience a decay rate comparable to the frequency of the disk mode, leading to ultra-strong coupling of an external emitter with 2D LO phonons.
-        \comment{tc}{The actual maximal value of the enhancement here is sensitively dependent on our choice of $\tau^{-1}$.
-I recall we took the value on the basis of one of Koppens' papers.
-If we don't state the value (and that it is realistic), our claim is not very convincing or impressive.
-We should probably state it in both text and caption.}
+Note that $\tau^{-1} = 5$ cm$^{-1}$ is of the order of the loss rate in bulk hBN.
         \label{fig:3}
@@
 \section{Strong light-matter interactions enabled by 2D optical phonons}
-\comment{tc}{Disclaimer: I didn't read this section very carefully.}
+%\comment{tc}{Disclaimer: I didn't read this section very carefully.}
 
 The extreme confinement of electromagnetic fields offered by the 2D LO phonon presents an opportunity for quantum optical applications in which one seeks to couple an external emitter such as an atom, molecule, defect, or artificial atomic system to electromagnetic fields.
-Applications of these couplings \swap{can be to realize}{are} ultra-bright single- or two-photon sources, \swap{or realize}{realizing the} strong-coupling \inset{regime} and the associated phenomenology of Rabi oscillations and polaritons, or \swap{resolve}{resolving} spectroscopically ``forbidden'' transitions.
-\comment{tc}{I guess this part needs citations?}
+Applications of these couplings are ultra-bright single- or two-photon sources, realizing the strong-coupling regime and the associated phenomenology of Rabi oscillations and polaritons, or resolving spectroscopically ``forbidden'' transitions \citeremind{}.
 
-In Figure~\ref{fig:3}, we consider the \remove{quantum }coupling of a dipole emitter to \inset{the in }2D \swap{LO phonon modes}{localized phonon polaritons of} nanostructured monolayer hBN.
-For simplicity, we consider hBN nanostructured as a disk, which leads to the formation of sharp resonances\inset{, quantized along the azimuthal and radial directions}.
+In Figure~\ref{fig:3}, we consider the coupling of a dipole emitter to localized phonon-polaritons of nanostructured monolayer hBN.
+For simplicity, we consider hBN nanostructured as a disk, which leads to the formation of sharp resonances quantized along the azimuthal and radial directions.
+The disk is taken to have a radius of \SI{20}{cm} and a loss rate $\tau^{-1} = 5$ cm$^{-1}$, which is of the order of the loss rate in bulk hBN \citeremind{}.
 In Figure~\ref{fig:3}, we plot the enhancement of the spontaneous emission rate for an external emitter polarized perpendicularly to the plane of the disk and placed \SI{5}{\nm} away from the surface of the disk.
 We find that the rate of spontaneous emission of 2D optical phonons is approximately 8 orders of magnitude larger than the rate of spontaneous emission of photons in the far field.
-For an infrared emitter at a transition wavelength of \SI{7}{\micro\m} with a free-space radiative lifetime of \SI{1}{\micro\s}, \SI{5}{\nm} away from an hBN disk, the coupling rate to 2D optical phonons would be on the same scale as the optical phonon frequency itself (about $\SI{2e14}{\per\s}$).
-\comment{tc}{We're using two different sets of unit for frequency (inverse seconds and inverse centimeters)}
+For an infrared emitter at a transition wavelength of \SI{7}{\micro\m} with a free-space radiative lifetime of \SI{1}{\micro\s}, \SI{5}{\nm} away from an hBN disk, the coupling rate to 2D optical phonons would be on the same scale as the optical phonon frequency itself (about $1430$ cm$^{-1}$ or $\SI{2.70e14}{\per\s}$).
 This coupling rate thus implies coupling between an emitter and the field in the regime of ultra-strong coupling.
@@
 The ability to probe low-loss and highly confined electromagnetic modes associated with 2D LO phonons in polar materials provides a new platform for nanophotonics in the mid- and far-infrared spectral range.
-The identification of the phonon polariton of bulk and thin-film geometries with the 2D LO phonon made in this manuscript may one to allow extend the rich phenomenology of optical phonons to nanophotonic applications.
+The identification of the phonon-polariton of bulk and thin-film geometries with the 2D LO phonon made in this manuscript may one to allow extend the rich phenomenology of optical phonons to nanophotonic applications.
 It also may lead to useful new ways to study LO phonons, arising from the fact that 2D LO phonons, unlike their 3D counterparts, have their electromagnetic energy extend a considerable distance from the material boundary.
 Due to the strong electromagnetic interactions between emitters and 2D LO phonons shown in this manuscript, it may be possible to design interesting new hybrid states of matter and phonons based on quantum electrodynamical strong coupling.
-The highly confined optical phonon waves in polar monolayers may also provide interesting new opportunities in near-field radiative heat transfer, in which it has been long known that thin-film surface phonon polaritons play a critical role.
+The highly confined optical phonon waves in polar monolayers may also provide interesting new opportunities in near-field radiative heat transfer, in which it has been long known that thin-film surface phonon-polaritons play a critical role.
 An important avenue of future study would be the \textit{ab initio} calculation of lifetimes of 2D LO phonons associated with three-phonon processes and electron-phonon interactions.
@@
 
-%%Phonon polaritons, hybrid quasiparticles of photons and optical phonons, offer great promise for deeply sub-diffractional control of electromagnetic fields at mid-IR and THz frequencies.
-Phonon polaritons share many features in common with plasmon polaritons in conductors.
+%%phonon-polaritons, hybrid quasiparticles of photons and optical phonons, offer great promise for deeply sub-diffractional control of electromagnetic fields at mid-IR and THz frequencies.
+phonon-polaritons share many features in common with plasmon polaritons in conductors.
 In recent years, it has been shown that phonon-polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra}.

### 2018-12-28 paper.tex
all comments implemented except change to fig. 2
+ substantial re-org/re-arrangement of main text to improve readability
+ shifting focus from polarization-polarization response to conductivity
+ making experimental ramifications more evident

@@
 % --- TITLE & AUTHORSHIP ---
-\title{Phonon-polaritonic reponse without LO-TO splitting in two-dimensional materials}
+\title{Phonon-polaritonic reponse without $\Gamma$-point LO-TO splitting in two-dimensional materials\\ or \\ Nanophotonics with optical phonons in two dimensions}
 
@@
 % -----------------
-\section{Electrodynamics of optical phonons in two-dimensions}
+\section{Optical response of optical phonons in two-dimensions}
 
@@
 where boldfaced quantities refer to vectors or tensors as appropriate.
-From the Kubo formula, it follows that
+The polarization-polarization response function is related to the more conductivity via the relation $\boldsymbol{\sigma}(\mathbf{q},\omega) = -\iu\omega\epsilon_0\boldsymbol{\Pi}(\mathbf{q},\omega)$.
+From the Kubo formula, it follows that the conductivity is
 \begin{equation}\label{eq:2dsusceptibility}
-    \boldsymbol{\Pi}(\mathbf{q},\omega) =  \frac{1}{\epsilon_0 \Omega\mathcal{Z}}\sum\limits_{m,n}\frac{\mathbf{P}_{mn}(\mathbf{q})\otimes\mathbf{P}_{nm}(\mathbf{q})}{\hbar\omega + E_{nm}+\iu 0^+}\Big(\e^{-\beta E_m}-\e^{-\beta E_n} \Big),
+    \boldsymbol{\sigma}(\mathbf{q},\omega) =  \frac{-\iu \omega}{ \Omega\mathcal{Z}}\sum\limits_{m,n}\frac{\mathbf{P}_{mn}(\mathbf{q})\otimes\mathbf{P}_{nm}(\mathbf{q})}{\hbar\omega + E_{nm}+\iu 0^+}\Big(\e^{-\beta E_m}-\e^{-\beta E_n} \Big),
 \end{equation}
@@
 We now evaluate the contribution of optical phonons to the polarization-polarization response in the low-temperature limit $T << \frac{\hbar\omega_{\mathrm{ph}}}{k_B}$ with $\omega_{\mathrm{ph}}$ a characteristic optical phonon frequency.
-\textbf{Need to include details on $qq$ components.} Taking a material with long-wavelength isotropy, such as hBN,we only have to consider the $qq$-component in the response tensor, where $qq$ denotes a pair of directions parallel to the wavevector.
-We thus find that the response function is given by \citeremind{}:
+Taking a material with long-wavelength isotropy, such as hBN,we only have to consider the $qq$-component in the response tensor, where $qq$ denotes a pair of directions parallel to the wavevector.
+Denoting $\sigma_{qq}$ as simply $\sigma$, we find that the conductivity is given by \cite{rivera2018ab}:
 \begin{equation}
-    \Pi_{qq}(\mathbf{q},\omega) = \frac{1}{\epsilon_0\hbar\Omega} \frac{2\omega_{\mathbf{q},\mathrm{L}}}{\omega^2_{\mathbf{q},\mathrm{L}}-\omega^2}|\hat{\mathbf{q}}\cdot\langle 1_{\mathbf{q},\mathrm{L}}|\mathbf{P}(\mathbf{q})|0_{\mathbf{q}\sigma}\rangle|^2,
+    \sigma(\mathbf{q},\omega) = \frac{-\iu\omega}{\hbar\Omega} \frac{2\omega_{\mathbf{q},\mathrm{L}}}{\omega^2_{\mathbf{q},\mathrm{L}}-\omega^2-\iu\omega\tau^{-1}}|\hat{\mathbf{q}}\cdot\langle 1_{\mathbf{q},\mathrm{L}}|\mathbf{P}(\mathbf{q})|0_{\mathbf{q},L}\rangle|^2,
     \label{eq:polpolresponse}
@@
 where L-subscripts denote longitudinal polarization, $|0_{\mathbf{q},\mathrm{L}}\rangle$ ($|1_{\mathbf{q},\mathrm{L}}\rangle$) denotes a state with no (one) longitudinal phonon of wavevector $\mathbf{q}$.
+We have also phenomenologically included the phonon dissipation rate $\tau^{-1}$, consistently with  a relaxation-time prescription.
 The frequency $\omega_{\mathbf{q},\mathrm{L}}$ in the denominator, as in the case of bulk phonons, is the frequency of the longitudinal phonon of wavevector $\mathbf{q}$ prior to considering LO--TO splitting.\citeremind{}
@@
 Coulomb interactions are accounted for in the random phase approximation, and to include them in the single-particle response amounts to an uncontrolled double-counting.
+%\textbf{May want to to include details on $qq$ components.}
 
-To proceed, we express the polarization matrix element in Equation~\eqref{eq:polpolresponse} in terms of parameters of the monolayer such as the Born effective charges. 
-Considering the longitudinal phonon contribution to the second-quantized ionic displacement, as in Ref.~\citenum{srivastava1990physics}, one finds that the RPA dielectric function is given by:
+To proceed, we express the polarization matrix element in Equation~\eqref{eq:polpolresponse} in terms of the Born effective charges of the monolayer and the phonon displacement eigenvectors.
+Considering the longitudinal phonon contribution to the second-quantized ionic displacement, as in Ref.~\citenum{srivastava1990physics}, we find that the conductivity within the RPA is given by
 \begin{equation}
-    \epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}} - \frac{q}{2\epsilon_0}\frac{\big|\hat{q}\cdot\sum_{\kappa} \mathbf{Z}_{\kappa}\eta_{\kappa}  \big|^2}{\omega^2_{\mathrm{TO}}-\omega^2},
-    \label{eq:epsRPA},
+    \sigma(\mathbf{q},\omega) = -\frac{\iu\omega }{\Omega}\frac{\Big|\hat{q}\cdot\sum\limits_{\kappa}\mathbf{Z}_{\kappa}\boldsymbol{\eta}_{\kappa} \Big|^2}{\omega^2_{\mathbf{q},\mathrm{L}}-\omega^2-\iu\omega\tau^{-1}}
 \end{equation}
@@
 
+While the conductivity is the main electromagnetic quantity of interest for electrodynamics applications, we briefly state the form of the (2D) permittivity, as its zeros immediately yield the longitudinal modes of the system, which are the LO phonons.
+The permittivity within the RPA, denoted $\epsilon_{\textrm{RPA}}$ is related to the polarization-polarization response function via \cite{jablan2009plasmonics} $\epsilon_{\textrm{RPA}} = \epsilon_{\mathrm{env}} + \frac{1}{2}q\Pi(q,\omega)$ and the conductivity via $\epsilon_{\textrm{RPA}} = \epsilon_{\mathrm{env}} + \frac{\iu q\sigma(q,\omega)}{2\omega\epsilon_0}$.
+Here, $\epsilon_{\mathrm{env}}$ is the average permittivity of a the bulk above and below the monolayer, and is added to take into account the polarization arising from these bulk materials.
+Based on Equation (6), the zeros $\omega_{\mathbf{q}}$ of the RPA dielectric function satisfy:
+\begin{equation}
+    \omega^2_{\mathbf{q}} - \omega^2_{\mathrm{TO}} = V(q)\frac{1}{\Omega}q^2\Big|\hat{q}\cdot\sum\limits_{\kappa}\tilde{\mathbf{Z}}_{\kappa}\boldsymbol{\eta}_{\kappa}\Big|^2,
+\end{equation}
+  where $V(q)$ is the Coulomb interaction in Fourier space, which in two dimensions, is given by $V(q) = \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}$.
+Given that the zeros of the dielectric function are associated with longitudinal modes, one expects that $\omega_{\mathbf{q}}$ is in fact the frequency of the LO phonon mode.
+This is consistent with the result of Ref.~\citenum{sohier2017breakdown}, in which Sohier \textit{et al} show that in two-dimensional polar materials, the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO--TO splitting and zero LO-TO splitting at the $\Gamma$ point of the Brillouin zone.
+
+  Given these results, we now re-express the conductivity explicitly in terms of the 2D phonon dispersion, and derive a universal form for the conductivity in the local ($q\rightarrow 0$) limit specified in terms of three parameters: the TO phonon frequency, the group velocity of the LO phonon at the $\Gamma$ point, and the damping rate.
+From Equation (7), we can immediately write the conductivity as
+  \begin{equation}
+  \sigma(q,\omega) = \frac{-2\iu\epsilon_0\epsilon_{\mathrm{env}}}{q}\frac{\omega^2_{\mathbf{q},\mathrm{LO}}-\omega^2_{\mathrm{TO}}}{\omega^2_{\mathrm{TO}}-\omega^2-\iu\omega\tau^{-1}}.
+  \end{equation}
+ For small $q$,  the Born charges are (to lowest-order) constant, and so the LO phonon dispersion takes the form $\omega_{\mathbf{q},\mathrm{LO}}=\sqrt{\omega^2_{\mathrm{TO}}+2v_g\omega_{\mathrm{TO}}q} \approx \omega_{\mathrm{TO}} + v_gq$, where $v_g$, the LO phonon group velocity, is defined from microscopic parameters through the relation
+ \begin{equation}
+    v_{\mathrm{g}} = \frac{e^2 }{4\epsilon_0 \epsilon_{\mathrm{env}}\omega_{\Gamma, \mathrm{TO}}\Omega}\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2. 
+\end{equation}
+Thus, in the long wavelength limit, we have the following universal parameterization of the conductivity of a polar monolayer:
+\begin{equation}
+    \sigma(\omega) =  \frac{-4\iu\epsilon_0\epsilon_{\mathrm{env}}\omega_{\mathrm{TO}}v_g}{\omega^2_{\mathrm{TO}}-\omega^2-\iu\omega\tau^{-1}}.
+\end{equation}
+From this relation, it follows that given the properties of the 2D phonons (from experiments or from \emph{ab initio} calculations), one can immediately specify the conductivity.
+Alternatively, from optical measurements (including far-field measurements) which allow one to extract the conductivity, it becomes possible to extract the group velocity of 2D LO phonons and thus the small-wavevector dispersion of those phonons.
+
+Before moving on to analyze the electrodynamics of 2D phonon-polaritons, we make two comments on lack of LO-TO splitting in 2D polar materials.
+The first is that this situation is in stark contrast to the situation of polar materials in 3D, which have a finite LO-TO splitting at the $\Gamma$ point.
+In the absence of such LO-TO splitting in 3D, there would be no frequency compatible with the existence of a phonon polariton.
+On the contrary, we will show that in 2D, despite the absence of LO-TO splitting at the $\Gamma$ point, there persists a strongly confined evanescent electromagnetic mode with a high local density of states which in all respects is similar to a phonon polariton of a thin film, but is in fact the 2D LO phonon of the polar monolayer.
+
+The second comment is that much of what has been discussed here has a strong analogy with the theory of optical response in electron gases in 2D, and particularly the relation between plasmons in 2D and 3D.
+To elaborate on this analogy, To do so, we take Equation (7) in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $e^2\big|\sum_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \big|^2 \equiv  \Big(\frac{eZ^r_*}{\sqrt{M_*}}\Big)^2 \equiv \Big(\frac{Q_*}{\sqrt{M_*}}\Big)^2$.
+Then, the LO--TO splitting can be written as $\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} = \frac{Q_*^2}{2\epsilon_0\epsilon_{\mathrm{env}} M_*}q$.
+Now we note that the RHS is exactly the squared-frequency $\omega^2_{\mathbf{q}p}$ for a plasma oscillation in a 2D gas of charged particles with charge $Q_*$ and mass $M_*$.
+This squared frequency can be thought of as the ``LP--TP'' splitting between longitudinal and transverse plasma oscillations.
+Of course, there are no transverse plasma oscillations due to the structureless nature of the electron, and so ``$\omega_{\mathrm{TP}}"$  should be considered as 0.
+In the three-dimensional plasmon case, `$\omega_{\mathrm{TP}} = 0$', but the difference between the squared longitudinal and transverse plasma oscillation frequencies at zero-wavevector is non-zero and given by $\omega_p^2$.
+In other words, the plasma frequency in electron gases is analogous to the LO-TO splitting in polar materials.
+This analogy between phononic and plasmonic behavior as a function of dimension is illustrated schematically in Figures~\ref{fig:1}(b,c).
+
   %\comment{tc}{I don't think this needs to be as long as it currently is---also, some of the details, \eg position of Reststrahlen band, assumption of long wave-length etc., would probably fit in more meaningfully when those approximations are actually applied.
 For Eq.~\eqref{eq:2dmaxwell}, only the homogeneity assumption is necessary (corresponding to in-plane translation invariance)}
-
-To relate the polarization-polarization response function to the electromagnetic modes supported by a polar monolayer, we solve Maxwell's equations for an evanescent electromagnetic mode supported by a surface with polarization-polarization response tensor $\mathbf{\Pi}$.
+\section{Electrodynamics of optical phonons in two-dimensions}
+To relate for example the polarization-polarization response function  to the electromagnetic modes supported by a polar monolayer, we solve Maxwell's equations for an evanescent electromagnetic mode supported by a surface with polarization-polarization response tensor $\mathbf{\Pi}$.
 We consider the monolayer to be sandwiched by a superstrate of permittivity $\epsilon_1$ and a substrate of permittivity $\epsilon_2$.
@@
 with $\epsilon_{\mathrm{env}}\equiv (\epsilon_1+\epsilon_2)/2$.
-The quantity $\epsilon_{\mathrm{RPA}}(\omega)$ precisely coincides with the optical phonon contribution to the RPA dielectric function of the monolayer.
-\textbf{Probably want to explain more how I know this is the RPA dielectric function despite getting it from Maxwell.}
+Thus, the condition for ``phonon-polaritons" coincides precisely with the condition for longitudinal optical phonons.
+Thus, one of the main results of our manuscript is that despite the lack of LO-TO splitting at the $\Gamma$ point, there nevertheless exists a strongly confined evanescent mode in many respects similar to the phonon-polaritons of thin films.
+We now analyze the dispersion relation of phonon polaritons in a specific material, hexagonal boron nitride, in more detail, showing the possibility of highly confined electromagnetic modes with a large local density of states.
+%The quantity $\epsilon_{\mathrm{RPA}}(\omega)$ precisely coincides with the optical phonon contribution to the RPA dielectric function of the monolayer.
+%\textbf{Probably want to explain more how I know this is the RPA dielectric function despite getting it from Maxwell.}
 %\comment{tc}{I do not completely understand the intention of this last sentence: are we trying to say that this coincides with the zeros of the dielectric function?
@@
 
-From Equation~\eqref{eq:epsRPA}, we obtain the dispersion equation for the 2D phonon-polaritons:
-\begin{align}
-    \omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} &= \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa}\tilde{\mathbf{Z}}_{\kappa}\boldsymbol{\eta}_{\kappa}  \Big|^2 \nonumber \\
-    &= V(q)\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa}\tilde{\mathbf{Z}}_{\kappa}\boldsymbol{\eta}_{\kappa}  \Big|^2,
-    \label{eq:dispersion}
-\end{align}
-where $e\tilde{\mathbf{Z}} \equiv \mathbf{Z}$, with $e$ the electron charge, and $V(q)=\frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}$ is the Coulomb interaction in two dimensions, screened by the dielectric environment of the sub- and superstrates.
+%From Equation~\eqref{eq:epsRPA}, we obtain the dispersion equation for the 2D phonon-polaritons:
+%\begin{align}
+%    \omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} &= \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa}\tilde{\mathbf{Z}}_{\kappa}\boldsymbol{\eta}_{\kappa}  \Big|^2 \nonumber \\
+%    &= V(q)\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa}\tilde{\mathbf{Z}}_{\kappa}\boldsymbol{\eta}_{\kappa}  \Big|^2,
+%    \label{eq:dispersion}
+%\end{align}
+%where $e\tilde{\mathbf{Z}} \equiv \mathbf{Z}$, with $e$ the electron charge, and $V(q)=\frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}$ is the Coulomb interaction in two dimensions, screened by the dielectric environment of the sub- and superstrates.
 This result, as we now show, reveals that the TM mode corresponds simply to the two-dimensional optical phonon.
-As shown recently in Ref.~\citenum{sohier2017breakdown}, in two-dimensional polar materials the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO--TO splitting.
-The expression for the LO--TO splitting is none other than Equation~\eqref{eq:dispersion} of the current manuscript, which vanishes in the long-wavelength ($q\rightarrow 0$) limit.
-\emph{Thus, one of the main results of our manuscript is that the wave solutions Maxwell's equations associated with phonon-polaritons in bulk materials, correspond to pure LO phonons in the monolayer limit.}
 
-It has been well-appreciated in the literature on phonon-polaritons in thin slabs.
-Here, we expand on the analogy in the two-dimensional case.
-To do so, we take Equation~\eqref{eq:dispersion} in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $e^2\big|\sum_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \big|^2 \equiv  \Big(\frac{eZ^r_*}{\sqrt{M_*}}\Big)^2 \equiv \Big(\frac{Q_*}{\sqrt{M_*}}\Big)^2$.
-Then, the LO--TO splitting can be written as $\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} = \frac{Q_*^2}{2\epsilon_0\epsilon_{\mathrm{env}} M_*}q$.
-Now we note that the RHS is exactly the squared-frequency $\omega^2_{\mathbf{q}p}$ for a plasma oscillation in a 2D gas of charged particles with charge $Q_*$ and mass $M_*$.
-This squared frequency can be thought of as the ``LP--TP'' splitting between longitudinal and transverse plasma oscillations.
-Of course, there are no transverse plasma oscillations, and so ``$\omega_{\mathrm{TP}} = 0$''.
-In the three-dimensional plasmon case, `$\omega_{\mathrm{TP}} = 0$', but the difference between the squared longitudinal and transverse plasma oscillation frequencies at zero-wavevector is non-zero and given by $\omega_p^2$.
-This analogy between phononic and plasmonic behavior as a function of dimension is illustrated schematically in Figures~\ref{fig:1}(b,c).
+%It has been well-appreciated in the literature on phonon-polaritons in thin slabs. Here, we expand on the analogy in the two-dimensional case.
 
@@
 
-We now rewrite $\epsilon_{\mathrm{RPA}}$ in a way that is expressed purely in terms of the optical phonon dispersion, derive the conductivity of the monolayer, and derive a universal form for the dispersion of 2D LO phonons in the long-wavelength limit.
+%We now rewrite $\epsilon_{\mathrm{RPA}}$ in a way that is expressed purely in terms of the optical phonon dispersion, derive the conductivity of the monolayer, and derive a universal form for the dispersion of 2D LO phonons in the long-wavelength limit.
 From Equation~\eqref{eq:dispersion}, we may write Equation~\eqref{eq:epsRPA} as
-\begin{equation}
-    \epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}\Bigg( 1 + \frac{\omega^2_{\mathbf{q},\mathrm{LO}}-\omega^2_{\mathbf{q},\mathrm{TO}}}{\omega^2_{\mathbf{q},\mathrm{TO}}-\omega^2}\Bigg).
-\end{equation}
-Similarly, the conductivity is prescribed by $\sigma(\mathbf{q},\omega) = -\iu \omega \Pi(\mathbf{q},\omega)$.
-\comment{tc}{This conductivity-definition should probably come sooner; along with the definition of which quantities $\boldsymbol{\Pi}$ connects.}
-We are now in a position to derive a universal form for the RPA dielectric function of a 2D polar slab in terms of three phenomenological parameters. In many cases,\citeremind{}
-\comment{tc}{It seems likely that this is true in all cases (just from thinking of Taylor series)\ldots If so, let's be more clear.
+%\begin{equation}
+%    \epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}\Bigg( 1 + \frac{\omega^2_{\mathbf{q},\mathrm{LO}}-\omega^2_{\mathbf{q},\mathrm{TO}}}{\omega^2_{\mathbf{q},\mathrm{TO}}-\omega^2}\Bigg).
+%\end{equation}
+%Similarly, the conductivity is prescribed by $\sigma(\mathbf{q},\omega) = -\iu \omega \Pi(\mathbf{q},\omega)$.
+%\comment{tc}{This conductivity-definition should probably come sooner; along with the definition of which quantities $\boldsymbol{\Pi}$ connects.}
+%We are now in a position to derive a universal form for the RPA dielectric function of a 2D polar slab in terms of three phenomenological parameters. In many cases,\citeremind{}
+%\comment{tc}{It seems likely that this is true in all cases (just from thinking of Taylor series)\ldots If so, let's be more clear.
 (Though, of course, it could depend on the direction of $\mathbf{q}$; but we already assumed isotropy elsewhere.}
-the Born charges are independent of $q$ for small $q$, and thus the dispersion can be written as
-\begin{equation}
-    \omega_{\mathbf{q},\mathrm{LO}} = \sqrt{\omega^2_{\Gamma, \mathrm{TO}}+\mathcal{S}q} \simeq \omega_{\Gamma,\mathrm{TO}} + v_{\mathrm{g}} q,
-    \label{eq:simpledispersion}
-\end{equation}
-where $\mathcal{S}$ is a microscopic constant determined by the Born charges of the monolayer, $v_{\mathrm{g}}$ is the group velocity of the LO mode, and $\omega_{\Gamma,\mathrm{TO}}$ is the TO phonon frequency at the $\Gamma$ point.
-\comment{tc}{There is probably no reason to have both $\mathcal{S}$ and $v_{\mathrm{g}}$; they are (simply) related.
+%the Born charges are independent of $q$ for small $q$, and thus the dispersion can be written as
+%\begin{equation}
+%    \omega_{\mathbf{q},\mathrm{LO}} = \sqrt{\omega^2_{\Gamma, \mathrm{TO}}+\mathcal{S}q} \simeq \omega_{\Gamma,\mathrm{TO}} + v_{\mathrm{g}} q,
+%    \label{eq:simpledispersion}
+%\end{equation}
+%where $\mathcal{S}$ is a microscopic constant determined by the Born charges of the monolayer, $v_{\mathrm{g}}$ is the group velocity of the LO mode, and $\omega_{\Gamma,\mathrm{TO}}$ is the TO phonon frequency at the $\Gamma$ point.
+%\comment{tc}{There is probably no reason to have both $\mathcal{S}$ and $v_{\mathrm{g}}$; they are (simply) related.
 In that case, this paragraph could also be contracted substantially and the two sentences merged into one}
-The group velocity is specified in terms of microscopic parameters via 
-\begin{equation}
-    v_{\mathrm{g}} = \frac{e^2 }{4\epsilon_0 \epsilon_{\mathrm{env}}\omega_{\Gamma, \mathrm{TO}}\Omega}\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2. 
-\end{equation}
-Thus, in the long wavelength limit, we have the following universal parameterization of the RPA dielectric function of a polar monolayer:
-\begin{equation}
-    \epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}\Bigg( 1 + \frac{2v_{\mathrm{g}}\omega_{\mathbf{q},\mathrm{TO}}}{\omega^2_{\mathbf{q},\mathrm{TO}}-\omega^2+\iu \omega \tau^{-1}}\Bigg),
-\end{equation}
-where we have also phenomenologically included the phonon dissipation rate $\tau^{-1}$, consistently with  a relaxation-time prescription. 
-\comment{tc}{I guess it would make sense to introduce $\tau^{-1}$ sooner.}
-Thus, all of the optical properties associated with LO phonons is captured by the TO phonon frequency, the slope of the LO dispersion at $\Gamma$, and the dissipation $\tau^{-1}$.
+%The group velocity is specified in terms of microscopic parameters via 
  %In Table 1, we tabulate these parameters for a few polar monolayers whose phonon band-structures have been calculated in the literature: hBN, MoS$_2$, and ...
@@
 % -----------------
-\section{Properties of 2D LO phonon modes in monolayer hBN}
+%\section{Properties of 2D LO phonon modes in monolayer hBN}
 
@@
 For simplicity, we consider hBN nanostructured as a disk, which leads to the formation of sharp resonances quantized along the azimuthal and radial directions.
-The disk is taken to have a radius of \SI{20}{cm} and a loss rate $\tau^{-1} = 5$ cm$^{-1}$, which is of the order of the loss rate in bulk hBN \citeremind{}.
+The disk is taken to have a radius of \SI{20}{cm} and a loss rate $\tau^{-1} = 5$ cm$^{-1}$, which is of the order of the loss rate in bulk hBN.
 In Figure~\ref{fig:3}, we plot the enhancement of the spontaneous emission rate for an external emitter polarized perpendicularly to the plane of the disk and placed \SI{5}{\nm} away from the surface of the disk.

### 2019-01-03 paper.tex
minor comments and edits

@@
 Within linear response, the induced polarization density can be straightforwardly evaluated from the displacement amplitude $\mathbf{u}_{\kappa}$ of every atom $\kappa$ within the unit cell.
-Specifically, to first-order, the induced polarization is (in repeated-index notation):
+Specifically, to first order, the induced polarization is (in repeated-index notation):
 \begin{equation}
@@
 $P_i^{0}$ is the equilibrium polarization in the absence of displacements, which is zero here.
+	\comment{tc}{I'm a bit unsure, but I think there's some subtlety to disregarding this term without further comment: bulk, static polarization \emph{changes} can be defined (giving rise to e.g.
+piezo-electricity, etc.), as can bulk, static unit-cell polarization, calculable from King-Smith \& Vanderbilt's formula but with unit-cell indeterminacy (\eg discussed in this \href{https://doi.org/10.1103/RevModPhys.82.1959}{RMP}).
+	There might be a sense in which $P_i^0$ is not really zero but just ill-defined for bulk problems.
+	For a finite system though, $P_i^0$ could definitely be nonzero. }
 With this relation between polarization and ionic displacements, the interaction Hamiltonian in Equation~\eqref{eq:coupling} couples the scalar potential and the ionic displacements.
@@
 In what follows, we assume the validity of the random-phase approximation (RPA) for dealing with Coulombic interactions between ions in the polar lattice.
-Within the RPA, these Coulombic interactions are accounted for by taking the induced polarization $\mathbf{P}(\mathbf{q},\omega)$ to be proportional \textit{total} electric field, $\mathbf{E}_{\mathrm{tot}}(\mathbf{q},\omega)$, defined to be the sum of the externally applied electric field and the electric field created by the induced polarization.
+Within the RPA, these Coulombic interactions are accounted for by taking the induced polarization $\mathbf{P}(\mathbf{q},\omega)$ to be proportional to the \emph{total} electric field, $\mathbf{E}_{\mathrm{tot}}(\mathbf{q},\omega)$, defined to be the sum of the externally applied electric field and the electric field created by the induced polarization.
 The polarization and total-field are connected by the polarization-polarization response (tensor) function  $\boldsymbol{\Pi}(\mathbf{q},\omega)$ via
@@
 where boldfaced quantities refer to vectors or tensors as appropriate.
-The polarization-polarization response function is related to the more conductivity via the relation $\boldsymbol{\sigma}(\mathbf{q},\omega) = -\iu\omega\epsilon_0\boldsymbol{\Pi}(\mathbf{q},\omega)$.
+The polarization-polarization response function is related to the conductivity via the relation $\boldsymbol{\sigma}(\mathbf{q},\omega) = -\iu\omega\epsilon_0\boldsymbol{\Pi}(\mathbf{q},\omega)$.
 From the Kubo formula, it follows that the conductivity is
@@
     \sigma(\mathbf{q},\omega) = -\frac{\iu\omega }{\Omega}\frac{\Big|\hat{q}\cdot\sum\limits_{\kappa}\mathbf{Z}_{\kappa}\boldsymbol{\eta}_{\kappa} \Big|^2}{\omega^2_{\mathbf{q},\mathrm{L}}-\omega^2-\iu\omega\tau^{-1}}
+    \label{eq:conductivity_simplified}
 \end{equation}
 where $\hat{q}$ is a unit vector along $\mathbf{q}$.
+  \comment{tc}{$\hat{q}$ was already used in \eqref{eq:polpolresponse}, so probably should be introduced there.
+Also, it should probably be $\hat{\mathbf{q}}$ rather than $\hat{q}$ here and in \eqref{eq:zerosofresponse}.
+  Same for $\hat{e}_{\kappa\mathbf{q},\mathrm{L}} \rightarrow \hat{\mathbf{e}}_{\kappa\mathbf{q},\mathrm{L}}$.}
 In this expression, $\omega_\mathrm{L}$, the LO phonon frequency prior to LO-T0 splitting, has been re-named as $\omega_{\mathrm{TO}}$, the transverse optical phonon frequency, because in the absence of LO--TO splitting, they are degenerate.
@@
 Here, $\epsilon_{\mathrm{env}}$ is the average permittivity of a the bulk above and below the monolayer, and is added to take into account the polarization arising from these bulk materials.
-Based on Equation (6), the zeros $\omega_{\mathbf{q}}$ of the RPA dielectric function satisfy:
+Based on Equation~\eqref{eq:conductivity_simplified}, the zeros $\omega_{\mathbf{q}}$ of the RPA dielectric function satisfy:
 \begin{equation}
     \omega^2_{\mathbf{q}} - \omega^2_{\mathrm{TO}} = V(q)\frac{1}{\Omega}q^2\Big|\hat{q}\cdot\sum\limits_{\kappa}\tilde{\mathbf{Z}}_{\kappa}\boldsymbol{\eta}_{\kappa}\Big|^2,
+    \label{eq:zerosofresponse}
 \end{equation}
@@
 Given that the zeros of the dielectric function are associated with longitudinal modes, one expects that $\omega_{\mathbf{q}}$ is in fact the frequency of the LO phonon mode.
-This is consistent with the result of Ref.~\citenum{sohier2017breakdown}, in which Sohier \textit{et al} show that in two-dimensional polar materials, the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO--TO splitting and zero LO-TO splitting at the $\Gamma$ point of the Brillouin zone.
+This is consistent with the result of Ref.~\citenum{sohier2017breakdown}, in which\relax{{}\ } \citeauthor{sohier2017breakdown} show that in two-dimensional polar materials, the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO--TO splitting and zero LO-TO splitting at the $\Gamma$ point of the Brillouin zone.
 
@@
 To strip the analysis to its bare essentials, we consider optical phonon response with in-plane isotropy in the long-wavelength limit arising from in-plane LO oscillations.
-A relevant example of a system where these conditions are satisfied is in a hexagonal boron nitride monolayer (see Fig.
-1(a) for schematic atomic structure).
+A relevant example of a system where these conditions are satisfied is in a hexagonal boron nitride monolayer (see Figure~\ref{fig:1}(a) for schematic atomic structure).
 In a monolayer geometry with translation invariance and in-plane isotropy, the electromagnetic modes of Maxwell's equations can be composed into transverse magnetic (TM) and transverse electric (TE), where the magnetic or electric field respectively is transverse to the in-plane wavevector of the mode.
@@
 where $\epsilon_1$ applies for $z>0$ and $\epsilon_2$ applies for $z<0$.
+\comment{tc}{Probably makes sense to use $\varepsilon_+$ instead of $\varepsilon_1$ and $\varepsilon_-$ for $\varepsilon_2$, to have just a single class of subscripts for this notion.}
 We consider a solution of the form $H(z) = h_\pm\e^{\mp\kappa_\pm z}$ with $\kappa_\pm = \sqrt{q^2-\epsilon_{1,2} k_0^2}$ with $\pm$ and $1,2$ corresponding each to $\pm z > 0$ respectively.
@@
 \begin{equation}
-    r_p(\mathbf{q},\omega) = \frac{(\epsilon_{2}-1)-i\frac{\sigma(\mathbf{q},\omega) q}{\omega\epsilon_0}}{(\epsilon_{2}+1)-i\frac{\sigma(\mathbf{q},\omega) q}{\omega \epsilon_0}}.
+    r_p(\mathbf{q},\omega) = \frac{(\epsilon_{2}-1)-\iu\frac{\sigma(\mathbf{q},\omega) q}{\omega\epsilon_0}}{(\epsilon_{2}+1)-\iu\frac{\sigma(\mathbf{q},\omega) q}{\omega \epsilon_0}}.
 \end{equation}

### 2019-01-03 paper.tex
starting on the edits..

--- 2018-12-28 paper.tex
@@
 % --- TITLE & AUTHORSHIP ---
-\title{Phonon-polaritonic reponse without $\Gamma$-point LO-TO splitting in two-dimensional materials\\ or \\ Nanophotonics with optical phonons in two dimensions}
+\title{Nanophotonics with optical phonons: Phonon-polaritonic response without $\Gamma$-point LO-TO splitting in two-dimensional materials}
 
@@
 Extreme confinement of electromagnetic energy by phonon-polaritons holds the promise of extremely strong and novel forms of control over the dynamics of matter.
-    To bring such control to its ultimate limit, it is important to consider phonon-polaritons in two dimensional systems.
+To bring such control to the atomic-scale limit, it is important to consider phonon-polaritons in two dimensional systems.
 Recent studies have pointed out that in two-dimensional systems, splitting between longitudinal and transverse optical (LO and TO) phonons is absent at the $\Gamma$ point.

### 2019-01-03 paper.tex
ready for more edits

--- 2019-01-03 paper.tex
@@
 % --- TITLE & AUTHORSHIP ---
-\title{Nanophotonics with optical phonons: Phonon-polaritonic response without $\Gamma$-point LO-TO splitting in two-dimensional materials}
+\title{Phonon polaritonics in two-dimensional materials \\ OR \\ Nanophotonics with optical phonons: Phonon-polaritonic response without $\Gamma$-point LO-TO splitting in two-dimensional materials}
 
@@
     Extreme confinement of electromagnetic energy by phonon-polaritons holds the promise of extremely strong and novel forms of control over the dynamics of matter. 
-To bring such control to the atomic-scale limit, it is important to consider phonon-polaritons in two dimensional systems.
+    To bring such control to its ultimate limit, it is important to consider phonon-polaritons in two dimensional systems.
 Recent studies have pointed out that in two-dimensional systems, splitting between longitudinal and transverse optical (LO and TO) phonons is absent at the $\Gamma$ point.
@@
 % --- INTRODUCTION ---
-phonon-polaritons, hybrid quasiparticles of photons and optical phonons supported in polar materials, hold great promise for deeply sub-diffractional control of electromagnetic fields at mid-infrared and terahertz frequencies.
+Phonon-polaritons, hybrid quasiparticles of photons and optical phonons supported in polar materials, hold great promise for deeply sub-diffractional control of electromagnetic fields at mid-infrared and terahertz frequencies.
 Qualitatively, phonon-polaritons share many features in common with plasmon polaritons in conductors.
@@
 
-In this Letter, we develop a framework and a several concrete examples for understanding the the nascent field of optical phononics in two-dimensional materials.
-We do so by deriving a universal form for the conductivity of a polar monolayer, including spatial nonlocality, which depends solely on the LO and TO phonon frequencies---and their dispersion with momentum---in the two-dimensional system.
+In this Letter, we develop a framework and concrete examples for understanding the the nascent field of optical phononics in two-dimensional materials.
+We do so by deriving a universal form for the conductivity of a polar monolayer,  which depends solely on the LO and TO phonon frequencies---and their dispersion with momentum---in the two-dimensional system.
 Using parameters from Ref.~\citenum{sohier2017breakdown} for the canonical two-dimensional polar monolayer --- hexagonal boron nitride --- we present the confinement and propagation losses of the 2D phononic modes, identifying the frequency region where they should be most easily detected.
 Finally, we show that these modes enable extreme light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced by up to eight orders of magnitude by the emitter--LO phonon coupling.
-Given the narrow bandwidth of phonon-polariton resonances, this suggests the possibility of realizing the multi-mode strong coupling and ultra-strong coupling regimes of quantum electrodynamics in a two-dimensional hexagonal boron nitride platform.
+For an infrared emitter with a sufficiently high free-space radiative decay rate ($> 10^6 s$ $^{-1}$), the enhanced linewidth of the emitter becomes comparable to the spacing between different phonon polaritonic resonances of an hBN nanostructure.
+This suggests the possibility of realizing the multi-mode strong coupling and ultra-strong coupling regimes of quantum electrodynamics in a two-dimensional hexagonal boron nitride platform.
+Our results for hBN are particularly relevant due its widespread use in 2D van der Waals heterostructures.
+In addition to providing functionality as a layer which improves the electrical and optical properties of other 2D materials, such as graphene, our results suggest that in these heterostructures, the hBN layers could provide a platform for nanophotonics and quantum optics which operates in the mid-infrared spectral range.
 
@@
 In what follows, we assume the validity of the random-phase approximation (RPA) for dealing with Coulombic interactions between ions in the polar lattice.
-Within the RPA, these Coulombic interactions are accounted for by taking the induced polarization $\mathbf{P}(\mathbf{q},\omega)$ to be proportional \textit{total} electric field, $\mathbf{E}_{\mathrm{tot}}(\mathbf{q},\omega)$, defined to be the sum of the externally applied electric field and the electric field created by the induced polarization.
+Within the RPA, these Coulombic interactions are accounted for by taking the induced polarization $\mathbf{P}(\mathbf{q},\omega)$ to be proportional \emph{total} electric field, $\mathbf{E}_{\mathrm{tot}}(\mathbf{q},\omega)$, defined to be the sum of the externally applied electric field and the electric field created by the induced polarization.
 The polarization and total-field are connected by the polarization-polarization response (tensor) function  $\boldsymbol{\Pi}(\mathbf{q},\omega)$ via
@@
 We now evaluate the contribution of optical phonons to the polarization-polarization response in the low-temperature limit $T << \frac{\hbar\omega_{\mathrm{ph}}}{k_B}$ with $\omega_{\mathrm{ph}}$ a characteristic optical phonon frequency.
-Taking a material with long-wavelength isotropy, such as hBN,we only have to consider the $qq$-component in the response tensor, where $qq$ denotes a pair of directions parallel to the wavevector.
+Taking a material with long-wavelength isotropy, such as hBN, we only have to consider the $qq$-component in the response tensor, where $qq$ denotes a pair of directions parallel to the wavevector.
 Denoting $\sigma_{qq}$ as simply $\sigma$, we find that the conductivity is given by \cite{rivera2018ab}:
@@
 \end{equation}
-where L-subscripts denote longitudinal polarization, $|0_{\mathbf{q},\mathrm{L}}\rangle$ ($|1_{\mathbf{q},\mathrm{L}}\rangle$) denotes a state with no (one) longitudinal phonon of wavevector $\mathbf{q}$.
+where L-subscripts denote longitudinal polarization, $|0_{\mathbf{q},\mathrm{L}}\rangle$ ($|1_{\mathbf{q},\mathrm{L}}\rangle$) denotes a state with no (one) longitudinal phonon of wavevector $\mathbf{q}$, and $\hat{\mathbf{q}}$ denotes a unit vector in the direction of $\mathbf{q}$.
 We have also phenomenologically included the phonon dissipation rate $\tau^{-1}$, consistently with  a relaxation-time prescription.
-The frequency $\omega_{\mathbf{q},\mathrm{L}}$ in the denominator, as in the case of bulk phonons, is the frequency of the longitudinal phonon of wavevector $\mathbf{q}$ prior to considering LO--TO splitting.\citeremind{}
+The frequency $\omega_{\mathbf{q},\mathrm{L}}$ in the denominator, as in the case of bulk phonons, is the frequency of the longitudinal phonon of wavevector $\mathbf{q}$ prior to considering LO--TO splitting \cite{BornHuang:1954}.
 This is consistent with the fact that LO--TO splitting is a collective effect arising from Coulomb interactions and the fact that the equation above represents a single-particle susceptibility.
@@
 \begin{equation}
-    \sigma(\mathbf{q},\omega) = -\frac{\iu\omega }{\Omega}\frac{\Big|\hat{q}\cdot\sum\limits_{\kappa}\mathbf{Z}_{\kappa}\boldsymbol{\eta}_{\kappa} \Big|^2}{\omega^2_{\mathbf{q},\mathrm{L}}-\omega^2-\iu\omega\tau^{-1}}
+    \sigma(\mathbf{q},\omega) = -\frac{\iu\omega }{\Omega}\frac{\Big|\hat{\mathbf{q}}\cdot\sum\limits_{\kappa}\mathbf{Z}_{\kappa}\boldsymbol{\eta}_{\kappa} \Big|^2}{\omega^2_{\mathbf{q},\mathrm{L}}-\omega^2-\iu\omega\tau^{-1}}
+    \label{eq:conductivity_simplified}
 \end{equation}
-  where $\hat{q}$ is a unit vector along $\mathbf{q}$.
-In this expression, $\omega_\mathrm{L}$, the LO phonon frequency prior to LO-T0 splitting, has been re-named as $\omega_{\mathrm{TO}}$, the transverse optical phonon frequency, because in the absence of LO--TO splitting, they are degenerate.
+  In this expression, $\omega_\mathrm{L}$, the LO phonon frequency prior to LO-TO splitting, has been re-named as $\omega_{\mathrm{TO}}$, the transverse optical phonon frequency, because in the absence of LO--TO splitting, they are degenerate.
 This is done in order to make the form of the final results more closely resemble their 3D counterparts.
-Additionally, we have defined scaled eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}\equiv \frac{\hat{e}_{\kappa\mathbf{q},\mathrm{L}}}{\sqrt{M_{\kappa}}}$, where $\hat{e}_{\kappa\mathbf{q},\mathrm{L}}$ is the unit-normalized polarization vector of atom $\kappa$ in the unit cell oscillating according to a longitudinal phonon of wavevector $\mathbf{q}$ and $M_{\kappa}$ is the mass of atom $\kappa$.
+Additionally, we have defined scaled eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}\equiv \frac{\hat{\mathbf{e}}_{\kappa\mathbf{q},\mathrm{L}}}{\sqrt{M_{\kappa}}}$, where $\hat{\mathbf{e}}_{\kappa\mathbf{q},\mathrm{L}}$ is the unit-normalized polarization vector of atom $\kappa$ in the unit cell oscillating according to a longitudinal phonon of wavevector $\mathbf{q}$ and $M_{\kappa}$ is the mass of atom $\kappa$.
 
@@
 The permittivity within the RPA, denoted $\epsilon_{\textrm{RPA}}$ is related to the polarization-polarization response function via \cite{jablan2009plasmonics} $\epsilon_{\textrm{RPA}} = \epsilon_{\mathrm{env}} + \frac{1}{2}q\Pi(q,\omega)$ and the conductivity via $\epsilon_{\textrm{RPA}} = \epsilon_{\mathrm{env}} + \frac{\iu q\sigma(q,\omega)}{2\omega\epsilon_0}$.
-Here, $\epsilon_{\mathrm{env}}$ is the average permittivity of a the bulk above and below the monolayer, and is added to take into account the polarization arising from these bulk materials.
-Based on Equation (6), the zeros $\omega_{\mathbf{q}}$ of the RPA dielectric function satisfy:
+Here, $\epsilon_{\mathrm{env}}$ is the average permittivity of the bulk above and below the monolayer, and is added to take into account the polarization arising from these bulk materials.
+Based on Equation~\eqref{eq:conductivity_simplified}, the zeros $\omega_{\mathbf{q}}$ of the RPA dielectric function satisfy:
 \begin{equation}
-    \omega^2_{\mathbf{q}} - \omega^2_{\mathrm{TO}} = V(q)\frac{1}{\Omega}q^2\Big|\hat{q}\cdot\sum\limits_{\kappa}\tilde{\mathbf{Z}}_{\kappa}\boldsymbol{\eta}_{\kappa}\Big|^2,
+    \omega^2_{\mathbf{q}} - \omega^2_{\mathrm{TO}} = V(q)\frac{1}{\Omega}q^2\Big|\hat{\mathbf{q}}\cdot\sum\limits_{\kappa}\tilde{\mathbf{Z}}_{\kappa}\boldsymbol{\eta}_{\kappa}\Big|^2,
+    \label{eq:zerosofresponse}
 \end{equation}
@@
 Given that the zeros of the dielectric function are associated with longitudinal modes, one expects that $\omega_{\mathbf{q}}$ is in fact the frequency of the LO phonon mode.
-This is consistent with the result of Ref.~\citenum{sohier2017breakdown}, in which Sohier \textit{et al} show that in two-dimensional polar materials, the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO--TO splitting and zero LO-TO splitting at the $\Gamma$ point of the Brillouin zone.
+This is consistent with the result of Ref.~\citenum{sohier2017breakdown}, in which \relax{{}\ } \citeauthor{sohier2017breakdown} show that in two-dimensional polar materials, the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO--TO splitting and zero LO--TO splitting at the $\Gamma$ point of the Brillouin zone.
 
@@
     \sigma(\omega) =  \frac{-4\iu\epsilon_0\epsilon_{\mathrm{env}}\omega_{\mathrm{TO}}v_g}{\omega^2_{\mathrm{TO}}-\omega^2-\iu\omega\tau^{-1}}.
+    \label{eq:localsigma}
 \end{equation}
@@
 
-Before moving on to analyze the electrodynamics of 2D phonon-polaritons, we make two comments on lack of LO-TO splitting in 2D polar materials.
-The first is that this situation is in stark contrast to the situation of polar materials in 3D, which have a finite LO-TO splitting at the $\Gamma$ point.
-In the absence of such LO-TO splitting in 3D, there would be no frequency compatible with the existence of a phonon polariton.
-On the contrary, we will show that in 2D, despite the absence of LO-TO splitting at the $\Gamma$ point, there persists a strongly confined evanescent electromagnetic mode with a high local density of states which in all respects is similar to a phonon polariton of a thin film, but is in fact the 2D LO phonon of the polar monolayer.
+Before moving on to analyze the electrodynamics of 2D phonon-polaritons, we make two comments on lack of LO--TO splitting in 2D polar materials.
+The first is that this situation is in stark contrast to the situation of polar materials in 3D, which have a finite LO--TO splitting at the $\Gamma$ point.
+In the absence of such LO--TO splitting in 3D, there would be no frequency compatible with the existence of a phonon  polariton.
+On the contrary, we will show that in 2D, despite the absence of LO--TO splitting at the $\Gamma$ point, there persists a strongly confined evanescent electromagnetic mode with a high local density of states which in all respects is similar to a phonon polariton of a thin film, but is in fact the 2D LO phonon of the polar monolayer.
 
@@
 In the three-dimensional plasmon case, `$\omega_{\mathrm{TP}} = 0$', but the difference between the squared longitudinal and transverse plasma oscillation frequencies at zero-wavevector is non-zero and given by $\omega_p^2$.
-In other words, the plasma frequency in electron gases is analogous to the LO-TO splitting in polar materials.
+In other words, the plasma frequency in electron gases is analogous to the LO--TO splitting in polar materials.
+The change in the dispersion of 3D versus 2D plasmons, like 3D versus 2D LO phonons, arises from the change in dimensionality of the Coulomb interaction from 3D to 2D.
 This analogy between phononic and plasmonic behavior as a function of dimension is illustrated schematically in Figures~\ref{fig:1}(b,c).
@@
 To relate for example the polarization-polarization response function  to the electromagnetic modes supported by a polar monolayer, we solve Maxwell's equations for an evanescent electromagnetic mode supported by a surface with polarization-polarization response tensor $\mathbf{\Pi}$.
-We consider the monolayer to be sandwiched by a superstrate of permittivity $\epsilon_1$ and a substrate of permittivity $\epsilon_2$.
+We consider the monolayer to be sandwiched by a superstrate of permittivity $\epsilon_+$ and a substrate of permittivity $\epsilon_-$.
 To strip the analysis to its bare essentials, we consider optical phonon response with in-plane isotropy in the long-wavelength limit arising from in-plane LO oscillations.
-A relevant example of a system where these conditions are satisfied is in a hexagonal boron nitride monolayer (see Fig.
-1(a) for schematic atomic structure).
-In a monolayer geometry with translation invariance and in-plane isotropy, the electromagnetic modes of Maxwell's equations can be composed into transverse magnetic (TM) and transverse electric (TE), where the magnetic or electric field respectively is transverse to the in-plane wavevector of the mode.
+A relevant example of a system where these conditions are satisfied is in a hexagonal boron nitride monolayer (see Figure~\ref{fig:1}(a) for schematic atomic structure).
+In a monolayer geometry with translation invariance and in-plane isotropy, the solutions of Maxwell's equations can be decomposed into transverse magnetic (TM) and transverse electric (TE) parts, where the magnetic or electric field respectively is transverse to the in-plane wavevector of the mode.
 In practice, it is the TM mode which is associated with highly confined electromagnetic waves that are of use in nanophotonics.
@@
 \begin{equation}\label{eq:2dmaxwell}
-    \bigg(-\frac{\dd^2}{\dd{}z^2}+q^2-\epsilon_{1,2}\frac{\omega^2}{c^2} \bigg)H(z) = 0,
+    \bigg(-\frac{\dd^2}{\dd{}z^2}+q^2-\epsilon_{\pm}\frac{\omega^2}{c^2} \bigg)H(z) = 0,
 \end{equation}
 where $\epsilon_1$ applies for $z>0$ and $\epsilon_2$ applies for $z<0$.
-We consider a solution of the form $H(z) = h_\pm\e^{\mp\kappa_\pm z}$ with $\kappa_\pm = \sqrt{q^2-\epsilon_{1,2} k_0^2}$ with $\pm$ and $1,2$ corresponding each to $\pm z > 0$ respectively.
+%\comment{tc}{Probably makes sense to use $\varepsilon_+$ instead of $\varepsilon_1$ and $\varepsilon_-$ for $\varepsilon_2$, to have just a single class of subscripts for this notion.}
+We consider a solution of the form $H(z) = h_\pm\e^{\mp\kappa_\pm z}$ with $\kappa_\pm = \sqrt{q^2-\epsilon_{\pm} k_0^2}$ with $\pm$ corresponding to $\pm z > 0$ respectively.
 The boundary condition on the magnetic field is $h^+-h^- = -K_x = \iu\omega\mathbf{P}_x$ where $\mathbf{K}$ is the surface current density, expressed through the polarization density $\mathbf{P}$ of the monolayer.
-This condition enforces $\frac{h^-}{h^+} =(1+\frac{\epsilon_0}{\epsilon_2} \kappa_2\Pi)^{-1}$.
-Continuity of the electric field in the $x$ direction enforces $\frac{\epsilon_2}{\epsilon_1} = -\frac{\kappa_2}{\kappa_1}\frac{h^-}{h^+}$. 
-Combining the two conditions, we obtain the usual dispersion equation for the TM mode of a polarizable 2D monolayer, namely $\epsilon_1/\kappa_1 + \epsilon_2/\kappa_2 +\epsilon_0\Pi=0$.
-Given the deeply subwavelength nature of 2D phonon-polaritons, \ie since $q\gg\omega/c$ such that $\kappa_{1,2}\simeq q$, the dispersion equation can be reduced to its quasistatic limit without consequential loss of accuracy:
+This condition enforces $\frac{h^-}{h^+} =(1+\frac{\epsilon_0}{\epsilon_-} \kappa_-\Pi)^{-1}$.
+Continuity of the electric field in the $x$ direction enforces $\frac{\epsilon_-}{\epsilon_+} = -\frac{\kappa_-}{\kappa_+}\frac{h^-}{h^+}$. 
+Combining the two conditions, we obtain the usual dispersion equation for the TM mode of a polarizable 2D monolayer, namely $\epsilon_+/\kappa_+ + \epsilon_-/\kappa_- +\epsilon_0\Pi=0$.
+Given the deeply subwavelength nature of 2D phonon-polaritons, \ie since $q\gg\omega/c$ such that $\kappa_{\pm}\simeq q$, the dispersion equation can be reduced to its quasistatic limit without consequential loss of accuracy:
 \begin{equation}
@@
 \end{equation}
-with $\epsilon_{\mathrm{env}}\equiv (\epsilon_1+\epsilon_2)/2$.
+with $\epsilon_{\mathrm{env}}\equiv (\epsilon_++\epsilon_-)/2$.
 Thus, the condition for ``phonon-polaritons" coincides precisely with the condition for longitudinal optical phonons.
-Thus, one of the main results of our manuscript is that despite the lack of LO-TO splitting at the $\Gamma$ point, there nevertheless exists a strongly confined evanescent mode in many respects similar to the phonon-polaritons of thin films.
+Thus, one of the main results of our manuscript is that despite the lack of LO--TO splitting at the $\Gamma$ point, there nevertheless exists a strongly confined evanescent mode in many respects similar to the phonon-polaritons of thin films.
 We now analyze the dispersion relation of phonon polaritons in a specific material, hexagonal boron nitride, in more detail, showing the possibility of highly confined electromagnetic modes with a large local density of states.
@@
 In 2D, the plasma frequency at zero wavevector is zero, but the electromagnetic physics is determined by the dispersion of 2D plasmons, which replace the plasmon polariton of bulk and thin films.
-        \comment{tc}{We already discussed reasonable changes to the figures in-person, so I won't comment on that this time around.}
+        %\comment{tc}{We already discussed reasonable changes to the figures in-person, so I won't comment on that this time around.}
         \label{fig:1}
@@
 \begin{figure*}[t]
-    \includegraphics[width=135mm]{fig2.pdf}
+    \includegraphics[width=135mm]{figure2hBN.pdf}
     \caption{%
-        \textbf{Propagation characteristics of 2D LO phonons in an un-structured monolayer of boron nitride.} (a) Imaginary part of the Fresnel reflection coefficient for p-polarized waves in free-standing monolayer hBN, which peaks at the dispersion relation of the 2D LO phonon.
-(b) Confinement factor of 2D LO phonon modes, which measures how much smaller the 2D LO phonon wavelength is compared to that of a photon at the same frequency.
-(c) Propagation quality factor, which measures the number of wavelengths of propagation of the 2D LO phonon mode, for different values of the damping rate of phonon.
+        \textbf{Properties of phonon polaritons in an monolayer of boron nitride compared to bulk.} (a) Real and imaginary parts of the conductivity of 2D hexagonal boron nitride in units of $e^2/\hbar$ for different values of the loss-rate.
+(b) Propagation quality factor, which measures the number of wavelengths of propagation of the 2D phonon polariton.
+(c) Dispersion relation of phonon polaritons in the monolayer (black) and thin films whose thicknesses are taken to be 1,2,4, and 8 interlayer spacings in hBN.
+For the thin-film, the fundamental mode and the first higher-order mode are plotted.
         \label{fig:2}
@@
 % -----------------
-%\section{Properties of 2D LO phonon modes in monolayer hBN}
+\section{Properties of 2D LO phonon modes in monolayer hBN}
 
-In Figure~\ref{fig:2}(a), we present the dispersion relation, operationally defined through the poles of the Fresnel reflection coefficient for $p$-polarized waves, which for a monolayer of conductivity $\sigma(\mathbf{q},\omega)$
-below vacuum and on top of a substrate of permittivity $\epsilon_{2}$, is given by
-\begin{equation}
-    r_p(\mathbf{q},\omega) = \frac{(\epsilon_{2}-1)-i\frac{\sigma(\mathbf{q},\omega) q}{\omega\epsilon_0}}{(\epsilon_{2}+1)-i\frac{\sigma(\mathbf{q},\omega) q}{\omega \epsilon_0}}.
-\end{equation}
-In Figure~\ref{fig:2}(b), we show the confinement factor (or effective mode index) of the 2D LO phonon mode, defined by $n_{\mathrm{eff}} = \frac{\text{Re } qc}{\omega}$. 
-The confinement factor is a key figure of merit in many nanophotonic applications involving quantum emitters, since all interactions involving emission of electromagnetic energy by emitters---\eg electric and magnetic dipole emission, multipole emission, phosphorescence, and multiphoton emission ---scale approximately as some power of this confinement factor~\cite{archambault2010quantum,koppens2011graphene,rivera2016shrinking,rivera2017making,sloan2018extreme}.
+In Figure ~\ref{fig:2}(a), we present the conductivity of 2D hBN of Eq.~\ref{eq:localsigma}, using parameters from Ref.
+$\cite{sohier2017breakdown}$ calculated from density functional theory within the local density approximation.
+From the conductivity, the dispersion relation of phonon polaritons on an infinite sheet is given by $q=\frac{2\iu\omega\epsilon_0\epsilon_{\mathrm{env}}}{\sigma(\omega)}$.
+The dispersion, assuming $\tau = \infty$, and $\epsilon_{\mathrm{env}}=1$ is shown in \ref{fig:2}(c) (black line).
+A key figure of merit for applications involving the propagation of phonon polaritons, is the propagation quality factor, defined by $\frac{\text{Re }q}{\text{Im }q} = \frac{\text{Im }\sigma(\omega)}{\text{Re }\sigma(\omega)}$, which is shown in \ref{fig:2}(b).
 
-Interestingly, in the case of a polar monolayer, the confinement grows very rapidly in frequency, due to the extremely low group velocity of 2D LO phonons, which is a remarkable four orders of magnitude slower than the speed of light.
-In particular, at frequencies of \SI{1450}{\per\cm}, the LO phonon already has a confinement factor in excess of \num{400}, corresponding to a phonon wavelength below \SI{20}{\nm}, which is much shorter wavelength than any phonon-polariton measured so far, and, similarly shorter than any plasmonic wavelength, even in graphene.
+For monolayer hBN, the wavevector grows very rapidly with frequency, due to the extremely low group velocity of 2D LO phonons, which is a remarkable four orders of magnitude slower than the speed of light.
+In particular, at frequencies of \SI{1450}{\per\cm}, the phonon-polariton has a wavelength of about 15 nm, which is much shorter wavelength than any phonon-polariton measured so far, and, similarly shorter than any plasmonic wavelength, even in graphene.
 In fact, this short a wavelength is essentially out of the range of measurability beyond the range of current scattering near-field microscope (SNOM) measurements, which are limited (by tip radii) to wavelengths greater than $\approx\,\SI{60}{\nm}$.
-
-The 2D LO phonon-polariton could in principle be measured closer to the TO frequency, where confinement is smaller; unfortunately, as shown in Figure~\ref{fig:2}(c), near the TO frequency, dissipation is far higher (and corresponding propagative quality factors $\Re q/\Im q$ far lower) due to large $\Re\sigma$ (or, equivalently, large $\Im\epsilon_{\mathrm{RPA}}$).
-\comment{tc}{Might be good to actually plot $\sigma(\omega)$}
-These considerations imply that access to the lower-loss and higher-confined portions of the dispersion relation of phonon-polaritons, in the absence of a sharp tip, requires a near-field probe such as a free electron probe, as used in electron energy loss spectroscopy (EELS),  where slow electrons can be used to probe plasmon wavelengths of just a few nanometers in monolayer metals, as well as the nonlocal bulk plasmon dispersion in metals~\cite{nagao2001dispersion, de2010optical,diaconescu2007acoustic}.
+The 2D LO phonon-polariton could in principle be measured by SNOM closer to the TO frequency, where confinement is smaller; unfortunately, as shown in Figure~\ref{fig:2}(b), near the TO frequency, dissipation is far higher (and corresponding propagative quality factors $\Re q/\Im q$ far lower) due to large $\Re\sigma$ (or, equivalently, large $\Im\epsilon_{\mathrm{RPA}}$).
+%\comment{tc}{Might be good to actually plot $\sigma(\omega)$}
+These considerations imply that access to the lower-loss and higher-confined portions of the dispersion relation of phonon-polaritons, in the absence of a sharp tip, requires a near-field probe such as a free electron probe, as used in electron energy loss spectroscopy (EELS),  where slow electrons can be used to probe plasmon wavelengths of just a few nanometers in monolayer metals, as well as the nonlocal bulk plasmon dispersion in metals~\cite{nagao2001dispersion,de2010optical,diaconescu2007low}.
 EELS has been recently employed to measure phonon-polaritons in ultrathin films of hBN~\cite{govyadinov2017probing}.
@@
 
-In what remains of this section, we  briefly discuss ``anomalies'', by which we mean features of the 2D LO phonon dispersion that cannot be predicted by considering the electromagnetic modes of a very thin slab of material whose dielectric function is that of the bulk polar material (with finite LO--TO splitting).
-For example, the long-wavelength linear dispersion of the 2D LO phonon mode given by Equation~\eqref{eq:simpledispersion} is not an anomaly, as this would be predicted by taking a sub-nm thick slab of a polar dielectric material.
+%In Figure~\ref{fig:2}(a), we present the dispersion relation, operationally defined through the poles of the Fresnel reflection coefficient for $p$-polarized waves, which for a monolayer of conductivity $\sigma(\mathbf{q},\omega)$
+%below vacuum and on top of a substrate of permittivity $\epsilon_{2}$, is given by
+%\begin{equation}
+%    r_p(\mathbf{q},\omega) = \frac{(\epsilon_{2}-1)-i\frac{\sigma(\mathbf{q},\omega) q}{\omega\epsilon_0}}{(\epsilon_{2}+1)-i\frac{\sigma(\mathbf{q},\omega) q}{\omega \epsilon_0}}.
+%\end{equation}
+%In Figure~\ref{fig:2}(b), we show the confinement factor (or effective mode index) of the 2D LO phonon mode, defined by $n_{\mathrm{eff}} = \frac{\text{Re } qc}{\omega}$. 
+%The confinement factor is a key figure of merit in many nanophotonic applications involving quantum emitters, since all interactions involving emission of electromagnetic energy by emitters---\eg electric and magnetic dipole emission, multipole emission, phosphorescence, and multiphoton emission ---scale approximately as some power of this confinement factor~\cite{archambault2010quantum,koppens2011graphene,rivera2016shrinking,rivera2017making,sloan2018extreme}.
+
+%Interestingly, in the case of a polar monolayer, the confinement grows very rapidly in frequency, due to the extremely low group velocity of 2D LO phonons, which is a remarkable four orders of magnitude slower than the speed of light.
+In particular, at frequencies of \SI{1450}{\per\cm}, the LO phonon already has a confinement factor in excess of \num{400}, corresponding to a phonon wavelength below \SI{20}{\nm}, which is much shorter wavelength than any phonon-polariton measured so far, and, similarly shorter than any plasmonic wavelength, even in graphene.
+In fact, this short a wavelength is essentially out of the range of measurability beyond the range of current scattering near-field microscope (SNOM) measurements, which are limited (by tip radii) to wavelengths greater than $\approx\,\SI{60}{\nm}$.
+
+In what remains of this section, we  briefly discuss the relation between the dispersion of an hBN monolayer versus the atomically-thick limit of a thin film of a material with hBN's bulk dielectric function.
+To aid this discussion, in \ref{fig:2}(c), we show the dispersion relation of thin films of "bulk hBN" with film thicknesses of 1, 2, 4, and 8 times the interlayer spacing of bulk hBN, which is roughly 0.33 nm.
+For these plots, we take hBN to be cleaved such that the optic-axis is perpendicular to the plane of the film.
+In such a case, the components of the bulk permittivity perpendicular and parallel to the c-axis ($\epsilon_{\perp}$ and $\epsilon_{||}$ respectively) are given by
+\begin{equation}
+\epsilon_{\perp,||}(\omega) = \epsilon_{\infty\perp,||}\left(1 + \frac{\omega^2_{\mathrm{LO}\perp,||}-\omega^2_{\mathrm{TO}\perp,||}}{\omega^2-\omega^2_{\mathrm{TO}\perp,||}} \right),
+\end{equation}
+with $\epsilon_{\infty\perp,||} = 2.95, 4.87$, $\omega_{\mathrm{TO}\perp,||} = 760,1360$ cm$^{-1}$, and $\omega_{\mathrm{LO}\perp,||} = 830,1614$ cm$^{-1}$ \cite{caldwell2014sub,dai2014tunable}.
+Losses are ignored in this discussion altogether.
+In the range between $\omega_{\mathrm{TO}}$ and $\omega_{\mathrm{LO}}$, the corresponding component of the permittivity is negative, while the other component is positive.
+This "hyperbolicity" leads to a dispersion for hBN thin films that have multiple branches at a given frequency, as can be seen in \ref{fig:2}(c).
+This trend persists even when the thickness of the bulk is taken down to a single layer, albeit being pushed to high wavevectors.
+This is in contrast with the true monolayer, in which there is only one LO phonon mode.
+Given that the "phonon polariton" of 2D is the LO phonon, there can be only one branch of the dispersion.
+Another "anomaly" between 2D and 3D is that the dispersion relation of the 2D LO phonon is non-monotonic, and the frequency starts to decrease for sufficiently large wavevector.
+This can never happen by extrapolating to a monolayer a dielectric function which has finite LO--TO splitting.
+This "turn-back" point corresponds to a frequency of around $1500$ cm$^{-1}$ and a phonon wavelength of less than 1 nm.
+Thus, this deviation is unlikely to be probed in typical optical experiments.
+However it means that for the monolayer, the range of accessible phonon polaritonic frequencies for optical applications is reduced compared to thick films.
+Finally, while we have focused on anomalies between the atomically-thin limit of bulk and a true monolayer, a comparison of the monolayer with the fundamental phonon polariton mode of the "one-atom-thick thin film" suggests that one has a reasonable qualitative understanding of the monolayer dispersion from this limit.
+
+%For example, the long-wavelength linear dispersion of the 2D LO phonon mode given by Equation~\eqref{eq:simpledispersion} is not an anomaly, as this would be predicted by taking a sub-nm thick slab of a polar dielectric material.
 However, the precise slope cannot be predicted, as it would depend on the thickness used for the slab, which is ill-defined, as one will find that it does not correspond to the thickness of the boron or nitrogen atoms.
-\comment{tc}{I'm not sure I agree with that: it seems reasonable to take the interlayer spacing of bulk hBN.
+%\comment{tc}{I'm not sure I agree with that: it seems reasonable to take the interlayer spacing of bulk hBN.
 That would make sense when thinking about surface vs.\ bulk polarization.
 Generally though, I'm surprised to hear that this choice is consequential: that's not the case in plasmonics.}
-One immediate anomaly is that in the monolayer, there is only one LO phonon, and thus only one branch of TM modes.
+%One immediate anomaly is that in the monolayer, there is only one LO phonon, and thus only one branch of TM modes.
 This is contrary to a thin slab of hexagonal boron nitride, which is hyperbolic $\epsilon_{zz} > 0, \epsilon_{xx} = \epsilon_{yy} < 0$, and thus has many branches of TM modes spaced in wavevector by an amount proportional to $d^{-1}$, with $d$ the film thickness.
-\comment{tc}{Well, this doesn't seem to be too bad of an anomaly: if I decided to treat the 2D sheet as a thin slab, then all the unphysical modes would be pushed to enormously high frequencies due to the $d^{-1}$-scaling; so it wouldn't affect any response-quantity I might be interested in?}
-Such additional TM modes would be incorrectly predicted by extrapolating the bulk to a film.
-Another anomaly comes from the fact that the dispersion relation of the 2D LO phonon is non-monotonic, and the frequency starts to decrease for sufficiently large wavevector.
-This can never happen by extrapolating to a monolayer a dielectric function which has finite LO--TO splitting.
 
@@
 The extreme confinement of electromagnetic fields offered by the 2D LO phonon presents an opportunity for quantum optical applications in which one seeks to couple an external emitter such as an atom, molecule, defect, or artificial atomic system to electromagnetic fields.
-Applications of these couplings are ultra-bright single- or two-photon sources, realizing the strong-coupling regime and the associated phenomenology of Rabi oscillations and polaritons, or resolving spectroscopically ``forbidden'' transitions \citeremind{}.
+Applications of these couplings are ultra-bright single- or two-photon sources, realizing the strong-coupling regime and the associated phenomenology of Rabi oscillations and polaritons, or resolving spectroscopically ``forbidden'' transitions \cite{koppens2011graphene,autore2018boron,rivera2016shrinking,rivera2017making,kurman2018control}.
 
@@
 We find that the rate of spontaneous emission of 2D optical phonons is approximately 8 orders of magnitude larger than the rate of spontaneous emission of photons in the far field.
-For an infrared emitter at a transition wavelength of \SI{7}{\micro\m} with a free-space radiative lifetime of \SI{1}{\micro\s}, \SI{5}{\nm} away from an hBN disk, the coupling rate to 2D optical phonons would be on the same scale as the optical phonon frequency itself (about $1430$ cm$^{-1}$ or $\SI{2.70e14}{\per\s}$).
+For an infrared emitter at a transition wavelength of \SI{7}{\micro\m} with a free-space radiative lifetime of \SI{1}{\micro\s}, \SI{5}{\nm} away from an hBN disk, the coupling rate to 2D optical phonons (about $\SI{1e14}{\per\s}$ or 0.065 eV) would be on the same scale as the optical phonon frequency itself (about $1430$ cm$^{-1}$ or $\SI{2.70e14}{\per\s}$ or 0.18 eV).
 This coupling rate thus implies coupling between an emitter and the field in the regime of ultra-strong coupling.
-Thus, the extreme confinement of electromagnetic energy associated with LO phonons in two dimensions enables the possibility of realizing ultrastrong coupling of an atom or molecule with optical phonons in a polar material, allowing the potential realization of new coupled states of quantum emitters and phonons.
+Moreover, the coupling rate for the 20 nm disk (orange), for an emitter with a far field decay rate of $> \SI{3e14}{\per\s}$ would have a sufficient coupling strength to the resonances in orange for its linewidth to span multiple resonances and thus be in a multi-mode ultra-strong coupling regime.
+Thus, the extreme confinement of electromagnetic energy associated with LO phonons in two dimensions enables the possibility of realizing ultrastrong coupling of an atom or molecule with optical phonons in a polar material, allowing the potential realization of new coupled states of quantum emitters and phonons such as atom-phonon polariton bound states.
 

### 2019-01-04 paper.tex
edits.

@@
 \usepackage{graphicx}
-\usepackage[centering,hmargin=20mm,tmargin=30mm,bmargin=25mm]{geometry}
+\usepackage[centering,hmargin=20mm,tmargin=28mm,bmargin=25mm]{geometry}
 \usepackage{multirow}
@@
 % --- TITLE & AUTHORSHIP ---
-\title{Phonon polaritonics in two-dimensional materials \\ OR \\ Nanophotonics with optical phonons: Phonon-polaritonic response without $\Gamma$-point LO-TO splitting in two-dimensional materials}
+\title{Phonon polaritonics in two-dimensional materials}
 
@@
 \author{Thomas Christensen}
-\affiliation{\MITPhy}
+\affiliation{\HarvardSEAS}\affiliation{\MITPhy}
 %
@@
     Extreme confinement of electromagnetic energy by phonon-polaritons holds the promise of extremely strong and novel forms of control over the dynamics of matter. 
-    To bring such control to its ultimate limit, it is important to consider phonon-polaritons in two dimensional systems.
+    To bring such control to the atomic-scale limit, it is important to consider phonon-polaritons in two dimensional systems.
 Recent studies have pointed out that in two-dimensional systems, splitting between longitudinal and transverse optical (LO and TO) phonons is absent at the $\Gamma$ point.
 Does this lack of LO--TO splitting imply the absence of a phonon-polariton in polar monolayers?
-    Here, we derive an expression for  the spatially non-local conductivity of a polar monolayer specified by the wavevector-dependent LO and TO phonon dispersions.
+    Here, we derive a first-principles expression for the spatially non-local conductivity of a polar monolayer specified by the wavevector-dependent LO and TO phonon dispersions.
 In the long-wavelength (local) limit, we find a universal form for the conductivity in terms of the TO phonon frequency at the $\Gamma$ point, its lifetime, and the group velocity of the LO phonon.
 Our analysis reveals that the phonon-polariton of the bulk is essentially replaced by the LO phonon of the two-dimensional system.
-We discuss the confinement and propagation losses of the LO phonons, finding that high  confinement and reasonable propagation quality factors coincide in regions which are difficult to detect with current near-field optical microscope techniques.
-    We conclude by studying the interaction of external emitters with two-dimensional hBN nanostructures, finding extreme enhancement of spontaneous emission due to coupling with localized 2D LO phonons, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN phonons.
+We calculate the confinement and propagation losses of the LO phonons, finding that high confinement and reasonable propagation quality factors coincide in regions which are difficult to detect with current near-field optical microscopy techniques.
+    Finally, we study the interaction of external emitters with two-dimensional hBN nanostructures, finding extreme enhancement of spontaneous emission due to coupling with localized 2D LO phonons, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN phonons.
+Using the strong electromagnetic interactions between emitters and 2D LO phonons calculated in this work, it is now possible to design new hybrid states of light, electrons and phonons based on quantum electrodynamical strong coupling.
 \end{abstract}
@@
 % --- INTRODUCTION ---
-Phonon-polaritons, hybrid quasiparticles of photons and optical phonons supported in polar materials, hold great promise for deeply sub-diffractional control of electromagnetic fields at mid-infrared and terahertz frequencies.
-Qualitatively, phonon-polaritons share many features in common with plasmon polaritons in conductors.
+Phonon-polaritons, hybrid quasiparticles of photons and optical phonons supported in polar materials, hold tremendous promise for deeply sub-diffraction control of electromagnetic fields at mid-infrared and terahertz frequencies.
+Qualitatively, phonon-polaritons share many features with plasmon-polaritons in conductors.
 Recently, it has been shown that phonon-polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra,li2018infrared,ma2018plane}.
@@
 In many cases, extreme confinement of phonon-polaritons is achieved by the use of thin-films, which shrink the in- and out-of-plane wavelength of polaritons monotonically with decreasing film thickness~\cite{dai2014tunable,dubrovkin2018ultra}.
-A monolayer is the ultimate limit of this effect when the properties of the bulk material can be used in the monolayer limit.
+A monolayer is the ultimate, atomic-scale limit of this effect when the properties of the bulk material can be used in the monolayer limit.
 In polar materials, the transition from 3D to 2D is not so simple, as in a polar monolayer, the LO--TO splitting that gives rise to phonon-polaritons in three-dimensions is absent at the $\Gamma$ point~\cite{sanchez2002vibrational,mele2002electric,serrano2007vibrational,sohier2017breakdown}.
@@
 
-In this Letter, we develop a framework and concrete examples for understanding the the nascent field of optical phononics in two-dimensional materials.
+In this Letter, we develop a theoretical and computational framework with quantitative, concrete examples to enable a new understanding of optical phononics in two-dimensional materials.
 We do so by deriving a universal form for the conductivity of a polar monolayer,  which depends solely on the LO and TO phonon frequencies---and their dispersion with momentum---in the two-dimensional system.
-Using parameters from Ref.~\citenum{sohier2017breakdown} for the canonical two-dimensional polar monolayer --- hexagonal boron nitride --- we present the confinement and propagation losses of the 2D phononic modes, identifying the frequency region where they should be most easily detected.
+Using parameters from Ref.~\citenum{sohier2017breakdown} for the canonical two-dimensional polar monolayer --- hexagonal boron nitride (hBN) --- we present the confinement and propagation losses of the 2D phononic modes, identifying the frequency region where they should be most easily detected.
 Finally, we show that these modes enable extreme light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced by up to eight orders of magnitude by the emitter--LO phonon coupling.
 For an infrared emitter with a sufficiently high free-space radiative decay rate ($> 10^6 s$ $^{-1}$), the enhanced linewidth of the emitter becomes comparable to the spacing between different phonon polaritonic resonances of an hBN nanostructure.
-This suggests the possibility of realizing the multi-mode strong coupling and ultra-strong coupling regimes of quantum electrodynamics in a two-dimensional hexagonal boron nitride platform.
+This suggests the possibility of realizing the multi-mode strong coupling and ultra-strong coupling regimes of quantum electrodynamics in a two-dimensional hBN platform.
 Our results for hBN are particularly relevant due its widespread use in 2D van der Waals heterostructures.
@@
 
-% -----------------
 \section{Optical response of optical phonons in two-dimensions}
-
-%\comment{tc}{Structurally, this section would benefit from being split into two sections: one solely tasked with deriving/establishing the forms of the (non-interacting) polarization-polarization response function---and possibly its equivalent conductivity; and another one solely tasked with deriving and discussing the dispersion relation for the associated longitudinal optical modes.
-Right now, this section goes back and forth between the two, which (I think) reduces its focus and gives way for unnecessary repetition.}
-
 In this section, we develop a theory of electromagnetic response due to optical phonons in two-dimensional systems.
-The crux of the theory involves the conductivity associated with the monolayer.
-To obtain it, we consider the response of the ions of the monolayer due(/when subjected) to a longitudinal electromagnetic potential $\phi$.
-The interaction Hamiltonian is in this case simply
+The key prediction is the conductivity of the monolayer.
+To obtain it, we consider the response of the ions of the monolayer due to a longitudinal electromagnetic potential $\phi$.
+The interaction Hamiltonian in this case is
 \begin{equation}
@@
 with $\rho$ the induced charge density and $\mathbf{P}$ the total polarization density associated with the ionic motion. 
-Within linear response, the induced polarization density can be straightforwardly evaluated from the displacement amplitude $\mathbf{u}_{\kappa}$ of every atom $\kappa$ within the unit cell.
+Within linear response theory, the induced polarization density can be straightforwardly evaluated from the displacement amplitude $\mathbf{u}_{\kappa}$ of every atom $\kappa$ within the unit cell.
 Specifically, to first order, the induced polarization is (in repeated-index notation):
@@
 
-In what follows, we assume the validity of the random-phase approximation (RPA) for dealing with Coulombic interactions between ions in the polar lattice.
+In what follows, we assume the validity of the random-phase approximation (RPA) in calculating Coulombic interactions between ions in the polar lattice.
 Within the RPA, these Coulombic interactions are accounted for by taking the induced polarization $\mathbf{P}(\mathbf{q},\omega)$ to be proportional \emph{total} electric field, $\mathbf{E}_{\mathrm{tot}}(\mathbf{q},\omega)$, defined to be the sum of the externally applied electric field and the electric field created by the induced polarization.
@@
 Coulomb interactions are accounted for in the random phase approximation, and to include them in the single-particle response amounts to an uncontrolled double-counting.
-%\textbf{May want to to include details on $qq$ components.}
 
-To proceed, we express the polarization matrix element in Equation~\eqref{eq:polpolresponse} in terms of the Born effective charges of the monolayer and the phonon displacement eigenvectors.
+Next, we express the polarization matrix element in Equation~\eqref{eq:polpolresponse} in terms of the Born effective charges of the monolayer and the phonon displacement eigenvectors.
 Considering the longitudinal phonon contribution to the second-quantized ionic displacement, as in Ref.~\citenum{srivastava1990physics}, we find that the conductivity within the RPA is given by
@@
 This squared frequency can be thought of as the ``LP--TP'' splitting between longitudinal and transverse plasma oscillations.
-Of course, there are no transverse plasma oscillations due to the structureless nature of the electron, and so ``$\omega_{\mathrm{TP}}"$  should be considered as 0.
+Of course, there are no transverse plasma oscillations due to the structure-less nature of the electron, and so ``$\omega_{\mathrm{TP}}"$  should be considered as 0.
 In the three-dimensional plasmon case, `$\omega_{\mathrm{TP}} = 0$', but the difference between the squared longitudinal and transverse plasma oscillation frequencies at zero-wavevector is non-zero and given by $\omega_p^2$.
@@
 
-  %\comment{tc}{I don't think this needs to be as long as it currently is---also, some of the details, \eg position of Reststrahlen band, assumption of long wave-length etc., would probably fit in more meaningfully when those approximations are actually applied.
-For Eq.~\eqref{eq:2dmaxwell}, only the homogeneity assumption is necessary (corresponding to in-plane translation invariance)}
 \section{Electrodynamics of optical phonons in two-dimensions}
@@
 with $\epsilon_{\mathrm{env}}\equiv (\epsilon_++\epsilon_-)/2$.
-Thus, the condition for ``phonon-polaritons" coincides precisely with the condition for longitudinal optical phonons.
+Thus, the condition for `phonon-polaritons' coincides precisely with the condition for longitudinal optical phonons.
 Thus, one of the main results of our manuscript is that despite the lack of LO--TO splitting at the $\Gamma$ point, there nevertheless exists a strongly confined evanescent mode in many respects similar to the phonon-polaritons of thin films.
 We now analyze the dispersion relation of phonon polaritons in a specific material, hexagonal boron nitride, in more detail, showing the possibility of highly confined electromagnetic modes with a large local density of states.
-%The quantity $\epsilon_{\mathrm{RPA}}(\omega)$ precisely coincides with the optical phonon contribution to the RPA dielectric function of the monolayer.
-%\textbf{Probably want to explain more how I know this is the RPA dielectric function despite getting it from Maxwell.}
-%\comment{tc}{I do not completely understand the intention of this last sentence: are we trying to say that this coincides with the zeros of the dielectric function?
-If so, maybe we should be a bit more explicit in our thinking here (e.g., we can say that we identify the LHS with the RPA dielectric function---and maybe comment on that the two quantities naturally have the same zeros)}
 
-%\comment{tc}{There's an issue with Eq.~\eqref{eq:epsRPA} AFAICT (and with all other occurrences of $\sum_{\kappa,j}Z_{\kappa,ij}\eta_{\kappa,j}$): it has a lingering $i$-index dependence.
-I guess this is the result of a mix between vector- and index-notation.}
-
-% ~~~~~~~~~~~~~~
 \begin{figure}[t]
@@
 Despite this, the 2D LO phonon plays the role of the phonon-polariton in 3D and thin films.
-(c) Analogous physics appears in electron gases in 3D and 2D, where the 3D plasma frequency is analogous to the 3D LO--TO splitting.
+(c) Analogous physics appears in electron gases in 3D and 2D, where the 3D plasma frequency is similar to the 3D LO--TO splitting.
 In 2D, the plasma frequency at zero wavevector is zero, but the electromagnetic physics is determined by the dispersion of 2D plasmons, which replace the plasmon polariton of bulk and thin films.
@@
 \end{figure}~
-% ~~~~~~~~~~~~~~
 
-%From Equation~\eqref{eq:epsRPA}, we obtain the dispersion equation for the 2D phonon-polaritons:
-%\begin{align}
-%    \omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} &= \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa}\tilde{\mathbf{Z}}_{\kappa}\boldsymbol{\eta}_{\kappa}  \Big|^2 \nonumber \\
-%    &= V(q)\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa}\tilde{\mathbf{Z}}_{\kappa}\boldsymbol{\eta}_{\kappa}  \Big|^2,
-%    \label{eq:dispersion}
-%\end{align}
-%where $e\tilde{\mathbf{Z}} \equiv \mathbf{Z}$, with $e$ the electron charge, and $V(q)=\frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}$ is the Coulomb interaction in two dimensions, screened by the dielectric environment of the sub- and superstrates.
-This result, as we now show, reveals that the TM mode corresponds simply to the two-dimensional optical phonon.
-
-%It has been well-appreciated in the literature on phonon-polaritons in thin slabs. Here, we expand on the analogy in the two-dimensional case.
-
-% ~~~~~~~~~~~~~~
 \begin{figure*}[t]
@@
 \end{figure*}
-% ~~~~~~~~~~~~~~
 
-%We now rewrite $\epsilon_{\mathrm{RPA}}$ in a way that is expressed purely in terms of the optical phonon dispersion, derive the conductivity of the monolayer, and derive a universal form for the dispersion of 2D LO phonons in the long-wavelength limit.
-From Equation~\eqref{eq:dispersion}, we may write Equation~\eqref{eq:epsRPA} as
-%\begin{equation}
-%    \epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}\Bigg( 1 + \frac{\omega^2_{\mathbf{q},\mathrm{LO}}-\omega^2_{\mathbf{q},\mathrm{TO}}}{\omega^2_{\mathbf{q},\mathrm{TO}}-\omega^2}\Bigg).
-%\end{equation}
-%Similarly, the conductivity is prescribed by $\sigma(\mathbf{q},\omega) = -\iu \omega \Pi(\mathbf{q},\omega)$.
-%\comment{tc}{This conductivity-definition should probably come sooner; along with the definition of which quantities $\boldsymbol{\Pi}$ connects.}
-%We are now in a position to derive a universal form for the RPA dielectric function of a 2D polar slab in terms of three phenomenological parameters. In many cases,\citeremind{}
-%\comment{tc}{It seems likely that this is true in all cases (just from thinking of Taylor series)\ldots If so, let's be more clear.
-(Though, of course, it could depend on the direction of $\mathbf{q}$; but we already assumed isotropy elsewhere.}
-%the Born charges are independent of $q$ for small $q$, and thus the dispersion can be written as
-%\begin{equation}
-%    \omega_{\mathbf{q},\mathrm{LO}} = \sqrt{\omega^2_{\Gamma, \mathrm{TO}}+\mathcal{S}q} \simeq \omega_{\Gamma,\mathrm{TO}} + v_{\mathrm{g}} q,
-%    \label{eq:simpledispersion}
-%\end{equation}
-%where $\mathcal{S}$ is a microscopic constant determined by the Born charges of the monolayer, $v_{\mathrm{g}}$ is the group velocity of the LO mode, and $\omega_{\Gamma,\mathrm{TO}}$ is the TO phonon frequency at the $\Gamma$ point.
-%\comment{tc}{There is probably no reason to have both $\mathcal{S}$ and $v_{\mathrm{g}}$; they are (simply) related.
-In that case, this paragraph could also be contracted substantially and the two sentences merged into one}
-%The group velocity is specified in terms of microscopic parameters via 
- %In Table 1, we tabulate these parameters for a few polar monolayers whose phonon band-structures have been calculated in the literature: hBN, MoS$_2$, and ...
-
-% -----------------
 \section{Properties of 2D LO phonon modes in monolayer hBN}
@@
 The 2D LO phonon-polariton could in principle be measured by SNOM closer to the TO frequency, where confinement is smaller; unfortunately, as shown in Figure~\ref{fig:2}(b), near the TO frequency, dissipation is far higher (and corresponding propagative quality factors $\Re q/\Im q$ far lower) due to large $\Re\sigma$ (or, equivalently, large $\Im\epsilon_{\mathrm{RPA}}$).
-%\comment{tc}{Might be good to actually plot $\sigma(\omega)$}
+
 These considerations imply that access to the lower-loss and higher-confined portions of the dispersion relation of phonon-polaritons, in the absence of a sharp tip, requires a near-field probe such as a free electron probe, as used in electron energy loss spectroscopy (EELS), where slow electrons can be used to probe plasmon wavelengths of just a few nanometers in monolayer metals, as well as the nonlocal bulk plasmon dispersion in metals~\cite{nagao2001dispersion,de2010optical,diaconescu2007low}.
@@
 
-%In Figure~\ref{fig:2}(a), we present the dispersion relation, operationally defined through the poles of the Fresnel reflection coefficient for $p$-polarized waves, which for a monolayer of conductivity $\sigma(\mathbf{q},\omega)$
-%below vacuum and on top of a substrate of permittivity $\epsilon_{2}$, is given by
-%\begin{equation}
-%    r_p(\mathbf{q},\omega) = \frac{(\epsilon_{2}-1)-i\frac{\sigma(\mathbf{q},\omega) q}{\omega\epsilon_0}}{(\epsilon_{2}+1)-i\frac{\sigma(\mathbf{q},\omega) q}{\omega \epsilon_0}}.
-%\end{equation}
-%In Figure~\ref{fig:2}(b), we show the confinement factor (or effective mode index) of the 2D LO phonon mode, defined by $n_{\mathrm{eff}} = \frac{\text{Re } qc}{\omega}$. 
-%The confinement factor is a key figure of merit in many nanophotonic applications involving quantum emitters, since all interactions involving emission of electromagnetic energy by emitters---\eg electric and magnetic dipole emission, multipole emission, phosphorescence, and multiphoton emission ---scale approximately as some power of this confinement factor~\cite{archambault2010quantum,koppens2011graphene,rivera2016shrinking,rivera2017making,sloan2018extreme}.
-
-%Interestingly, in the case of a polar monolayer, the confinement grows very rapidly in frequency, due to the extremely low group velocity of 2D LO phonons, which is a remarkable four orders of magnitude slower than the speed of light.
-In particular, at frequencies of \SI{1450}{\per\cm}, the LO phonon already has a confinement factor in excess of \num{400}, corresponding to a phonon wavelength below \SI{20}{\nm}, which is much shorter wavelength than any phonon-polariton measured so far, and, similarly shorter than any plasmonic wavelength, even in graphene.
-In fact, this short a wavelength is essentially out of the range of measurability beyond the range of current scattering near-field microscope (SNOM) measurements, which are limited (by tip radii) to wavelengths greater than $\approx\,\SI{60}{\nm}$.
-
-In what remains of this section, we  briefly discuss the relation between the dispersion of an hBN monolayer versus the atomically-thick limit of a thin film of a material with hBN's bulk dielectric function.
-To aid this discussion, in \ref{fig:2}(c), we show the dispersion relation of thin films of "bulk hBN" with film thicknesses of 1, 2, 4, and 8 times the interlayer spacing of bulk hBN, which is roughly 0.33 nm.
+In the rest of this section, we discuss the relation between the dispersion of an hBN monolayer versus the atomically-thick limit of a thin film of a material with hBN's bulk dielectric function.
+To aid this discussion, in \ref{fig:2}(c), we show the dispersion relation of thin films of `bulk' hBN with film thicknesses of 1, 2, 4, and 8 times the interlayer spacing of bulk hBN, which is roughly 0.33 nm.
 For these plots, we take hBN to be cleaved such that the optic-axis is perpendicular to the plane of the film.
@@
 In the range between $\omega_{\mathrm{TO}}$ and $\omega_{\mathrm{LO}}$, the corresponding component of the permittivity is negative, while the other component is positive.
-This "hyperbolicity" leads to a dispersion for hBN thin films that have multiple branches at a given frequency, as can be seen in \ref{fig:2}(c).
+This hyperbolicity leads to a dispersion for hBN thin films that have multiple branches at a given frequency, as can be seen in \ref{fig:2}(c).
 This trend persists even when the thickness of the bulk is taken down to a single layer, albeit being pushed to high wavevectors.
 This is in contrast with the true monolayer, in which there is only one LO phonon mode.
-Given that the "phonon polariton" of 2D is the LO phonon, there can be only one branch of the dispersion.
-Another "anomaly" between 2D and 3D is that the dispersion relation of the 2D LO phonon is non-monotonic, and the frequency starts to decrease for sufficiently large wavevector.
+Given that the phonon polariton of 2D is the LO phonon, there can be only one branch of the dispersion.
+Another anomaly between 2D and 3D is that the dispersion relation of the 2D LO phonon is non-monotonic, and the frequency starts to decrease for sufficiently large wavevector.
 This can never happen by extrapolating to a monolayer a dielectric function which has finite LO--TO splitting.
-This "turn-back" point corresponds to a frequency of around $1500$ cm$^{-1}$ and a phonon wavelength of less than 1 nm.
+This `turn-back' point corresponds to a frequency of around $1500$ cm$^{-1}$ and a phonon wavelength of less than 1 nm.
 Thus, this deviation is unlikely to be probed in typical optical experiments.
 However it means that for the monolayer, the range of accessible phonon polaritonic frequencies for optical applications is reduced compared to thick films.
-Finally, while we have focused on anomalies between the atomically-thin limit of bulk and a true monolayer, a comparison of the monolayer with the fundamental phonon polariton mode of the "one-atom-thick thin film" suggests that one has a reasonable qualitative understanding of the monolayer dispersion from this limit.
+Finally, while we have focused on anomalies between the atomically-thin limit of bulk and a true monolayer, a comparison of the monolayer with the fundamental phonon polariton mode of the `one-atom-thick thin film' suggests that we have a reasonable qualitative understanding of the monolayer dispersion from this limit.
 
-%For example, the long-wavelength linear dispersion of the 2D LO phonon mode given by Equation~\eqref{eq:simpledispersion} is not an anomaly, as this would be predicted by taking a sub-nm thick slab of a polar dielectric material.
-However, the precise slope cannot be predicted, as it would depend on the thickness used for the slab, which is ill-defined, as one will find that it does not correspond to the thickness of the boron or nitrogen atoms.
-%\comment{tc}{I'm not sure I agree with that: it seems reasonable to take the interlayer spacing of bulk hBN.
-That would make sense when thinking about surface vs.\ bulk polarization.
-Generally though, I'm surprised to hear that this choice is consequential: that's not the case in plasmonics.}
-%One immediate anomaly is that in the monolayer, there is only one LO phonon, and thus only one branch of TM modes.
-This is contrary to a thin slab of hexagonal boron nitride, which is hyperbolic $\epsilon_{zz} > 0, \epsilon_{xx} = \epsilon_{yy} < 0$, and thus has many branches of TM modes spaced in wavevector by an amount proportional to $d^{-1}$, with $d$ the film thickness.
-
-% ~~~~~~~~~~~~~~
 \begin{figure}[t]
@@
 \end{figure}
-% ~~~~~~~~~~~~~~
 
-% -----------------
 \section{Strong light-matter interactions enabled by 2D optical phonons}
-%\comment{tc}{Disclaimer: I didn't read this section very carefully.}
 
 The extreme confinement of electromagnetic fields offered by the 2D LO phonon presents an opportunity for quantum optical applications in which one seeks to couple an external emitter such as an atom, molecule, defect, or artificial atomic system to electromagnetic fields.
-Applications of these couplings are ultra-bright single- or two-photon sources, realizing the strong-coupling regime and the associated phenomenology of Rabi oscillations and polaritons, or resolving spectroscopically ``forbidden'' transitions \cite{koppens2011graphene,autore2018boron,rivera2016shrinking,rivera2017making,kurman2018control}.
+Applications of these couplings are ultra-bright single- or two-photon sources, realizing the strong-coupling regime and the associated phenomenology of Rabi oscillations and polaritons, or resolving spectroscopically `forbidden'' transitions \cite{koppens2011graphene,autore2018boron,rivera2016shrinking,rivera2017making,kurman2018control}.
 
@@
 The ability to probe low-loss and highly confined electromagnetic modes associated with 2D LO phonons in polar materials provides a new platform for nanophotonics in the mid- and far-infrared spectral range.
-The identification of the phonon-polariton of bulk and thin-film geometries with the 2D LO phonon made in this manuscript may one to allow extend the rich phenomenology of optical phonons to nanophotonic applications.
-It also may lead to useful new ways to study LO phonons, arising from the fact that 2D LO phonons, unlike their 3D counterparts, have their electromagnetic energy extend a considerable distance from the material boundary.
-Due to the strong electromagnetic interactions between emitters and 2D LO phonons shown in this manuscript, it may be possible to design interesting new hybrid states of matter and phonons based on quantum electrodynamical strong coupling.
+The identification of the phonon-polariton of bulk and thin-film geometries with the 2D LO phonon made in this Letter would extend the rich phenomenology of optical phonons to nanophotonic applications.
+This work also points the way to useful new approaches to study LO phonons, arising from the fact that 2D LO phonons, unlike their 3D counterparts, have their electromagnetic energy extend a considerable distance from the material boundary.
+Due to the strong electromagnetic interactions between emitters and 2D LO phonons shown in this Letter, it is now possible to design interesting new hybrid states of matter and phonons based on quantum electrodynamical strong coupling.
 The highly confined optical phonon waves in polar monolayers may also provide interesting new opportunities in near-field radiative heat transfer, in which it has been long known that thin-film surface phonon-polaritons play a critical role.
 An important avenue of future study would be the \textit{ab initio} calculation of lifetimes of 2D LO phonons associated with three-phonon processes and electron-phonon interactions.
-It would be of great interest to study the prospects of techniques such as isotopic purification and cryogenic temperatures for reducing the decay of these 2D LO phonons.
+In further work, it would be of great interest to study the effects of isotopic purification and cryogenic temperatures on reducing the decay rate of these 2D LO phonons.
 
-%%phonon-polaritons, hybrid quasiparticles of photons and optical phonons, offer great promise for deeply sub-diffractional control of electromagnetic fields at mid-IR and THz frequencies.
-phonon-polaritons share many features in common with plasmon polaritons in conductors.
-In recent years, it has been shown that phonon-polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra}.
-Due to this remarkable confinement and their relatively high lifetimes of around picoseconds, phonon-polaritons open new opportunities for vibrational spectroscopy, radiative heat transfer~\cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters~\cite{kumar2015tunable,rivera2017making,kurman2018control}.
-The core features of phonon-polaritons, such as mode shape, confinement, and propagation characteristics, are understood from simple Lorentz oscillator models of the dielectric function, enabling successful theoretical accounts of experimental observations.
-%
-%%we develop a first-principles theory of phonon-polaritons in a 2D polar material, specifying the dielectric function in terms of \emph{ab initio} parameters, and finding a universal form for the dispersion relation of 2D phonon-polaritons when the wavevector is much smaller than the size of the unit cell.
-%
-%\
-%%\section{Linear response of phonon-polaritons in two dimensions}
-%%We now move to generalize the derivation of the previous section to describe the optical phonon contribution to Maxwell's equations coming from a two-dimensional material. 
-%%Unlike the previous section, we will not demonstrate the results of the calculation on a particular material.
-%Here, we are only interested in finding an expression for the optical response of the surface in terms of \emph{ab initio}-derivable parameters, as well as the general form of the dispersion relation of phonon-polaritons in 2D.
-%Because of the change of symmetry in the system, we will not be able to perform the analysis by Fourier transforming Maxwell's equations in three dimensions.
-%However, we will apply a technique which is often used to analyze electromagnetic modes in 2D systems such as plasmons in 2D conductors.
-%An example of the technique is shown in detail in Ref.~\citenum{jablan2009plasmonics}, but we summarize the essential details here. 
-%
-%%Consider a two-dimensional ionic lattice located in the $xy$ plane ($z=0$) surrounded by a homogeneous environment of dielectric function $\epsilon_{\mathrm{env}}$.
-%We consider the simplified case where the lattice is free-standing, to show the essential concept.
-%We search for a mode whose magnetic field $\mathbf{H}$ is of the form $\mathbf{H}(z)\e^{\iu\mathbf{q}\cdot\mathbf{r}_{\parallel} - \iu\omega t}$, where $\mathbf{r}_{\parallel}=(x,y,0)$ is the position in the plane, $\mathbf{q}$ is the wavevector and $\omega_{\mathbf{q}}$ is the frequency.
-%To simplify the discussion even further and furnish analytical forms for the phonon-polariton modes in 2D, we assume this material has mirror symmetry along the axis of $\mathbf{q}$.
-%In this case, the electromagnetic modes can be decomposed into TE and TM-polarized modes.
-%As is typical in the study of highly-confined polaritons, whether they are plasmon or phonon-polaritons, the TM mode is the highly confined one, and occurs when $\epsilon(\omega) < 0$.
-%In our convention, the TM mode is such that $\mathbf{H}$ is in the plane of the lattice and perpendicular to $\mathbf{q}$. 
-%Without loss of generality, define the $\mathbf{q}$ direction to be the $x$ direction, making the magnetic field in the $y$ direction.
-%	
-%
-%For $z \neq 0$, the Maxwell equation for the magnetic field is $(\nabla\times\nabla\times - \epsilon_{\mathrm{env}}\frac{\omega^2}{c^2})\mathbf{H}(z)\e^{\iu\mathbf{q}\cdot\mathbf{r}_{\parallel}} = 0$.
-%For the TM polarized mode, $\mathbf{H}(z) = H(z)\unitvec{y}$; the vectorial wave equation consequently simplifies to a scalar one
-%\begin{equation}\label{eq:2dmaxwell}
-%\left(-\frac{\dd^2}{\dd{}z^2}+q^2-\epsilon_{\mathrm{env}}\frac{\omega^2}{c^2} \right)H(z) = 0.
-%\end{equation}
-%Seeking a surface-bound mode, $H(z)$ must assume the functional form
-%	\comment{tc}{Nick, please check (I reverted an earlier edit of mine of this to the stuff below, but I cannot make sense of it): how can we specify $H(z)$ by equality sign here; and then in the following paragraphs we say ``To join these two solutions consistently, we apply \ldots$\unitvec{z}\times(\mathbf{H}^{(+)}-\mathbf{H}^{(-)}) = \mathbf{K} = -\iu\omega\mathbf{P}_{\mathrm{s}}$''?.
-I cannot make sense of this given that we use equality signs here and that $\mathbf{H}(z) = H(z)\unitvec{y}$.}
-%\begin{align}
-%	H(z) = \sgn(z)\e^{-\kappa |z|},
-%\end{align}
-%with $\kappa \equiv \sqrt{q^2-\epsilon_{\mathrm{env}}\frac{\omega^2}{c^2}}$.
-%
-%To join these two solutions consistently, we apply the interface conditions for the magnetic field.
-They are $\unitvec{z}\times(\mathbf{H}^{(+)}-\mathbf{H}^{(-)}) = \mathbf{K} = -\iu\omega\mathbf{P}_{\mathrm{s}}$ where $\mathbf{K}$ is the surface current density, expressed through the surface polarization density $\mathbf{P}_{\mathrm{s}}$.
-%Now we use the linear response relation $\mathbf{P}(\mathbf{q},\omega) = \boldsymbol{\Pi}(\mathbf{q},\omega)\mathbf{E}(\mathbf{q},\omega)$ in addition to Ampere's law $\nabla\times\mathbf{H} = -\iu\omega\epsilon_{\mathrm{env}}\mathbf{E}$.
-%	\comment{tc}{There's a slight disconnect in that we never define the relation between $\mathbf{P}_{\mathrm{s}}$ and $\mathbf{P}$.
-Also, there's a bigger issue, I think: if we allow $\boldsymbol{\Pi}$ to be a $3\times 3$ matrix, then we ought to adopt more generic boundary conditions.
-Specifically, if there's out-of-plane ``current'', then the boundary conditions for normal parts of $\mathbf{B}$ and parallel parts of $\mathbf{E}$ would change.}
-%Plugging these relations in yields an implicit equation connecting the polarization response function, the frequency of the 2D phonon-polariton, and its frequency.
-Taking the case where $x$ coincides with a principal axis of the system, we have that
-%\begin{equation}\label{eq:2ddispersion}
-%q\Pi_{xx}(q,\omega)  = -2\epsilon_{\mathrm{env}}.
-%\end{equation}
-%
-%To connect this relation to the discussion of the 3D case, we write a microscopic form for the polarization response function of the ionic lattice. It is simply
-%\begin{equation}\label{eq:2dsusceptibility}
-%\boldsymbol{\Pi}(\mathbf{q},\omega) =  \frac{1}{\epsilon_0 A}\sum\limits_{m,n}\frac{\mathbf{P}_{mn}(\mathbf{q})\otimes\mathbf{P}_{nm}(\mathbf{q})}{\hbar\omega + E_{nm}+\iu 0^+}\left(\e^{\beta E_m}-\e^{\beta E_n} \right),
-%\end{equation}
-%is the 2D Fourier transform of the polarization density.
-As the form of the displacement of the ions is still given by Equation~\eqref{eq:phononfield}, $\mathbf{P}_{mn}(\mathbf{q})$ is still given by Equations~(\ref{eq:fourierpolarization}--\ref{eq:oscillatorstrength}), but with the Born charges and eigendisplacements appropriate to the 2D system of interest.
-Note that although the lattice is two dimensional, in general, the ions may be displaced in any three directions and thus $\boldsymbol{\Pi}$ is still a $3\times3$ matrix.
-%Following the same kind of reasoning that lead to Equation~\eqref{eq:lorentzoscillator}, if we parameterize the effect of the different masses and Born charges in the unit cell by $M_{\mathrm{eff}}$ and $Q_{\mathrm{eff}}$, we will find a form for the polarization response function which is given by the second term of Equation~\eqref{eq:lorentzoscillator} but with $V$ replaced by $A$.
-Defining $N/A \equiv n_{\mathrm{s}}$, Equation~\eqref{eq:2dmaxwell} for the dispersion relation of 2D phonon-polaritons tells us that
-%\begin{equation}\label{eq:2dphononpolaritondispersion}
-%\omega = \sqrt{\omega_{\mathrm{\mathrm{TO}}}^2+\frac{n_{\mathrm{s}}Q_{\mathrm{eff}}^{2}}{2M_{\mathrm{eff}}\epsilon_0\epsilon_{\mathrm{env}}}q}.
-%\end{equation}
-%Note that this dispersion has a strong resemblance to the plasmon dispersion relation of two-dimensional free electron gases, which is  $\omega = \sqrt{n_{\mathrm{s}}e^2q/2m\epsilon_0\epsilon_{\mathrm{env}}}$  in the local limit~\cite{stern1967polarizability}.
-%The only difference is the replacement of electron parameters (density, charge, mass) by the corresponding lattice parameters in the phononic case, and a shift in the minimum frequency by the transverse optical phonon frequency.
-%
-%We note that the dispersion relation of Equation~\eqref{eq:2dphononpolaritondispersion} can in fact be obtained by taking the infinitely thin-limit of a bulk polar dielectric, and finding the dispersion of the resulting phonon-polaritons in the quasi-electrostatic limit.
-%In particular, for an isotropic polar dielectric of thickness $2d$ with boundaries at $z=\pm d$, solving Laplace's equations yields as the dispersion of the even parity mode:
-%\begin{equation}\label{eq:thinfilmphononpolaritondispersion}
-%\tanh(qd) = -\frac{\epsilon_{\mathrm{env}}}{\epsilon(\omega)} \rightarrow qd\epsilon(\omega) = -\epsilon_{\mathrm{env}},
-%\end{equation}
-%with $\epsilon_{\mathrm{env}}$ the dielectric constant of the surrounding environment and $\epsilon(\omega)$ that of the polar dielectric.
-%The right-hand side is obtained in the limit of $d\rightarrow 0$.
-%Taking Equation~\eqref{eq:lorentzoscillator} for the dielectric function in bulk, we have that (in the absence of losses) $\epsilon_{\mathrm{env}} + qd\epsilon_{\infty} + q\frac{n_sQ_{\mathrm{eff}}^{2}}{2\epsilon_0 M_{\mathrm{eff}}}\frac{1}{\omega_{\mathrm{\mathrm{TO}}}^2-\omega^2} = 0$, where we have identified $V = 2Ad$.
-%In the limit of small $d$, i.e., $qd \ll 1$, we recover Equations~\eqref{eq:2ddispersion} and \eqref{eq:2dphononpolaritondispersion} with $\Pi = \frac{n_sQ_{\mathrm{eff}}^{2}}{\epsilon_0 M}\frac{1}{\omega^2_{\mathrm{\mathrm{TO}}}-\omega^2}$.
-It is interesting to see that in the 2D limit, the effect of the screening inside the material altogether vanishes.
-A similar phenomenon is known in the electrodynamics of plasmons in 2D materials, where the dielectric screening is known to be entirely provided by the environment and is equal to $\epsilon_{\mathrm{env}}$.
-A useful result which follows is that we may also derive another important figure of merit, the $p$-polarized reflectivity of a 2D polar dielectric, which is related to the local density of states of the polar dielectric, which quantifies the strength of coupling to external probes such as a nano-antenna or a quantum emitter.
-%By considering the infinitely thin limit of a polar dielectric with dielectric function $\epsilon(\omega)$ and defining $\epsilon(\omega) = \Pi/2d$, we immediately find that the $p$-polarized reflectivity is given by
-%\begin{equation}\label{eq:2dtmreflectivity}
-%r_p = \frac{1}{1+2\epsilon_{\mathrm{env}}/q\Pi}.
-%\end{equation}
-%
-%We note that, in general, when considering bulk and 2D versions of a material, at the very least, the effective parameters $Q_{\mathrm{eff}}$ and $M_{\mathrm{eff}}$ will change as a result of changes to the phonon band structure.
-In other words, although the forms of the expressions for the dielectric function
-%
-%
-% It is also quite possible that they altogether vanish in the 2D case as opposed to the 3D case.
-As a recently discovered example of such a phenomenon, it was found that the LO-\mathrm{TO} splitting of some 2D polar dielectrics, such as hBN~\cite{sohier2017breakdown} altogether vanishes, and only returns to its bulk value after a small but finite wavevector.
-This would then suggest that a description of phonon-polaritons necessitates consideration of the spatially nonlocal dielectric function as in Equation (12), and that these polaritons moreover may only exist when their wavelengths are sufficiently short.
-Another important change, beyond the theory presented in this work is that the losses may also change drastically in the reduction of dimensions.
-For example, it was found in several experiments that the damping rate of the \mathrm{TO} phonon in hBN nearly doubles upon going from bulk to 2D\cite{gorbachev2011hunting,tran2016quantum}, which may strongly suppress the ability to image the polariton with near-field probes.
-%%This is due to the small thickness where the dispersion of the phonon-polariton is quite flat, and very high wavevector modes, corresponding to the maximum wavevector detectable by a near-field tip, occur at frequencies very close to the \mathrm{TO} phonon frequency.
-%%Near the \mathrm{TO} phonon-frequency however, the ratio $\Im\Pi/\Re\Pi$ is the highest, corresponding to very low quality-factor modes.
-%%In addition to these considerations, the values of the parameters may depend on the background material, which may change the phonon band structure. 
-%
-%
-%We provided this thin-film derivation as a simple way to understand the content of Equations~\eqref{eq:2ddispersion} and \eqref{eq:2dphononpolaritondispersion}.
-%In general Equation~\eqref{eq:2dsusceptibility} should be used to calculate $\Pi$ and the corresponding dispersion.
-%We note that a similar situation appears in the treatment of 2D plasmons: namely that a relation like Equation~\eqref{eq:2ddispersion} appears, except with $\Pi$ the polarization-polarization response of a 2D electron gas.
-%And furthermore that one can also obtain the same equation by considering the ultra-thin limit of a 3D plasmonic material~\cite{jablan2009plasmonics,jablan2013plasmons}.
-%
-%
-%\section{Summary and Outlook}
-%
-%In summary, we have provided a theoretical framework based on linear response theory to calculate the phonon contribution to the dielectric function from first principles.
-%Our framework is rather versatile, allowing us to: use first principles calculations to get the dielectric function, predict how nonlocality enters the dielectric function, and predict the effect of reduced dimensionality.
-%A particularly interesting case to consider in future work would be to find a system where the optical phonons are drastically different from 3D to 2D.
-%Perhaps it is possible that there are some materials in which the 2D optical phonons experience lower losses due to a reduced scattering phase space.
-%The formalism we provide here may also be extended to understand phonon-polaritons in other more atypical reduced-dimensional settings, such as zero-dimensional settings in single emitters, i.e., `molecular phonon-polaritons', in analogy to recent work on `molecular plasmons'~\cite{manjavacas2013tunable,lauchner2015molecular}.
-%Notably, we go beyond oversimplifications of the Lorentz oscillator model in which the Born charges are treated as a single scalar quantity and can treat the influence and interplay of many phonon modes that contribute to the dielectric function.
-%We corroborated our approach through density functional theoretic calculations.
-%
-%In future work, besides furnishing an \emph{ab initio} description of nonlocality and reduced-dimension phonon-polaritons, there are a number of interesting questions that can be addressed by the framework provided here.
-%For example, we may consider situations such as the phonon-polaritons associated with optical phonons at the interface of two materials.
-%Another system of interest would be to consider two nearby layers of material in which the optical phonons of each material strongly couple to each other. 
-%In that case, it would be relevant to evaluate how this strong coupling manifests itself in the infrared dielectric function, and ultimately the confinement and propagation of the phonon-polaritons.
-
-% --- ACKNOWLEDGEMENTS ---
-\section{Acknowledgements}
-%This research used resources of the National Energy Research Scientific Computing Center, a DOE Office of Science User Facility supported by the Office of Science of the U.S.
-Department of Energy under Contract No.
-DE-AC02-05CH11231.
+\section{Acknowledgments}
 N.
@@
 DE-FG02-97ER25308.
-P.
-N.
-acknowledges start-up funding from the Harvard John A.
-Paulson School of Engineering and Applied Sciences.
 T.
@@
 The authors thank Joshua Caldwell, Dmitri Basov, Ido Kaminer, Siyuan Dai, Samuel Moore, Jennifer Coulter, and Christopher Ciccarino for helpful discussions.
-This work was supported by the DOE Photonics at Thermodynamic Limits Energy Frontier Research Center under grant no.
-DE-SC0019140.
+This work was supported by the DOE Photonics at Thermodynamic Limits Energy Frontier Research Center under grant number DE-SC0019140.
 
-% --- BIBLIOGRAPHY ---
 \bibliographystyle{apsrev4-1}

### 2019-01-06 paper.tex
edits.

@@
 \begin{abstract}
-    Extreme confinement of electromagnetic energy by phonon-polaritons holds the promise of extremely strong and novel forms of control over the dynamics of matter. 
+    Extreme confinement of electromagnetic energy by phonon-polaritons holds the promise of strong and new forms of control over the dynamics of matter. 
     To bring such control to the atomic-scale limit, it is important to consider phonon-polaritons in two dimensional systems.
@@
 acknowledges support from the Danish Council for Independent Research (Grant No.\ DFF--6108-00667).
-The authors thank Joshua Caldwell, Dmitri Basov, Ido Kaminer, Siyuan Dai, Samuel Moore, Jennifer Coulter, and Christopher Ciccarino for helpful discussions.
+The authors thank Prof.
+Ido Kaminer (Technion), Siyuan Dai (Columbia), Samuel Moore (Columbia), Jennifer Coulter (Harvard), and Christopher Ciccarino (Harvard) for helpful discussions.
 This work was supported by the DOE Photonics at Thermodynamic Limits Energy Frontier Research Center under grant number DE-SC0019140.

### 2019-01-06 paper.tex
minor edits

@@
 In addition to providing functionality as a layer which improves the electrical and optical properties of other 2D materials, such as graphene, our results suggest that in these heterostructures, the hBN layers could provide a platform for nanophotonics and quantum optics which operates in the mid-infrared spectral range.
+While we focus on hBN in this manuscript, many of our findings apply to other polar monolayers.
 
@@
 The disk is taken to have a radius of \SI{20}{cm} and a loss rate $\tau^{-1} = 5$ cm$^{-1}$, which is of the order of the loss rate in bulk hBN.
+We parameterize the coupling between the dipole and phonon-polaritons through the rate of spontaneous emission $\Gamma$ of phonon polaritons by the dipole.
+It is related to the Green's function of the Maxwell equations for the nanostructure via the expression \cite{novotny2012principles}:
+\begin{equation}
+\frac{\Gamma}{\Gamma_0} = 1 + \frac{6\pi c}{\omega_0}\hat{n}\cdot\text{Im }G(\mathbf{r}_0,\mathbf{r}_0,\omega_0)\cdot\hat{n},
+\end{equation}
+where $\hat{n}$ is the orientation of the dipole, $\mathbf{r}_0$ is its position, and $\omega_0$ is its frequency.
+The Green's function is computed through a boundary element method (as in Ref.
+\cite{jin2017infrared}) assuming the validity of the electrostatic approximation, which is well-respected due to the extreme sub-wavelength confinement of the electromagnetic field of the phonon polariton.
 In Figure~\ref{fig:3}, we plot the enhancement of the spontaneous emission rate for an external emitter polarized perpendicularly to the plane of the disk and placed \SI{5}{\nm} away from the surface of the disk.
 We find that the rate of spontaneous emission of 2D optical phonons is approximately 8 orders of magnitude larger than the rate of spontaneous emission of photons in the far field.
-For an infrared emitter at a transition wavelength of \SI{7}{\micro\m} with a free-space radiative lifetime of \SI{1}{\micro\s}, \SI{5}{\nm} away from an hBN disk, the coupling rate to 2D optical phonons (about $\SI{1e14}{\per\s}$ or 0.065 eV) would be on the same scale as the optical phonon frequency itself (about $1430$ cm$^{-1}$ or $\SI{2.70e14}{\per\s}$ or 0.18 eV).
+We note that in this approach, the coupling of the dipole to phonon polaritons is manifested through the phonon contribution to the dielectric function of the disk.
+This should be equivalent to an approach that considers the coupling of a bound electron in an emitter to LO phonons in the disk through a (two-dimensional) Frohlich coupling -- i.e., a coupling of the atomic electron to the electric potential resulting from the polarization associated with an LO phonon mode \cite{sohier2016two}.
+
+In \ref{fig:3}, we show that for an infrared emitter at a transition wavelength of \SI{7}{\micro\m} with a free-space radiative lifetime of \SI{1}{\micro\s}, \SI{5}{\nm} away from an hBN disk, the coupling rate to 2D optical phonons (about $\SI{1e14}{\per\s}$ or 0.065 eV) would be on the same scale as the optical phonon frequency itself (about $1430$ cm$^{-1}$ or $\SI{2.70e14}{\per\s}$ or 0.18 eV).
 This coupling rate thus implies coupling between an emitter and the field in the regime of ultra-strong coupling.

### 2019-01-07 paper.tex
minor things from spot-reading; did not read through properly yet.

--- 2019-01-06 paper.tex
@@
 	list-separator = {,\ \linebreak[0]},
-	list-final-separator = {,\ \linebreak[0]\text{and}\ }}
+	list-final-separator = {,\ \linebreak[0]\text{and}\ },
+	retain-unity-mantissa = false,
+}
 
@@
 
+\comment{tc}{@NR: There's a somewhat inconsistent use of quotations: \ie `X' vs.\ ``X'' throughout the paper.
+In most cases, double quotations are probably correct.
+Generally though, I would favor removing them in most cases, as they often read mostly as implied uncertainty or vagueness.}
+
 % --- INTRODUCTION ---
@@
 We do so by deriving a universal form for the conductivity of a polar monolayer,  which depends solely on the LO and TO phonon frequencies---and their dispersion with momentum---in the two-dimensional system.
-Using parameters from Ref.~\citenum{sohier2017breakdown} for the canonical two-dimensional polar monolayer --- hexagonal boron nitride (hBN) --- we present the confinement and propagation losses of the 2D phononic modes, identifying the frequency region where they should be most easily detected.
+Using parameters from Ref.~\citenum{sohier2017breakdown} for the canonical two-dimensional polar monolayer---hexagonal boron nitride (hBN)---we present the confinement and propagation losses of the 2D phononic modes, identifying the frequency region where they should be most easily detected.
 Finally, we show that these modes enable extreme light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced by up to eight orders of magnitude by the emitter--LO phonon coupling.
-For an infrared emitter with a sufficiently high free-space radiative decay rate ($> 10^6 s$ $^{-1}$), the enhanced linewidth of the emitter becomes comparable to the spacing between different phonon polaritonic resonances of an hBN nanostructure.
+For an infrared emitter with a sufficiently high free-space radiative decay rate (${>}\, \SI{1e6}{\per\s}$), the enhanced linewidth of the emitter becomes comparable to the spacing between different phonon polaritonic resonances of an hBN nanostructure.
 This suggests the possibility of realizing the multi-mode strong coupling and ultra-strong coupling regimes of quantum electrodynamics in a two-dimensional hBN platform.
@@
 where $m,n$ are states of the phononic Fock space of the monolayer, $P_{mn,i}(\mathbf{q}) = \sum_{\kappa,j}Z_{\kappa,ij}\langle m | u_{\kappa,j}(\mathbf{q}) | n \rangle$ are matrix elements of the polarization associated with phonon modes (where $u_{\kappa,j}(\mathbf{q})$ is the Fourier transform of the phonon displacement operator), $E_{m}$ ($E_n$) is the energy of state $m$ ($n$), $\beta \equiv 1/k_{\text{\textsc{B}}}T$ is the inverse temperature, and $\mathcal{Z}$ is the grand partition function.
-We now evaluate the contribution of optical phonons to the polarization-polarization response in the low-temperature limit $T << \frac{\hbar\omega_{\mathrm{ph}}}{k_B}$ with $\omega_{\mathrm{ph}}$ a characteristic optical phonon frequency.
+We now evaluate the contribution of optical phonons to the polarization-polarization response in the low-temperature limit $T \ll \hbar\omega_{\mathrm{ph}}/k_{\mathrm{B}}$ with $\omega_{\mathrm{ph}}$ a characteristic optical phonon frequency.
 Taking a material with long-wavelength isotropy, such as hBN, we only have to consider the $qq$-component in the response tensor, where $qq$ denotes a pair of directions parallel to the wavevector.
@@
 \end{equation}
-  In this expression, $\omega_\mathrm{L}$, the LO phonon frequency prior to LO-TO splitting, has been re-named as $\omega_{\mathrm{TO}}$, the transverse optical phonon frequency, because in the absence of LO--TO splitting, they are degenerate.
+  In this expression, $\omega_\mathrm{L}$, the LO phonon frequency prior to LO--TO splitting, has been re-named as $\omega_{\mathrm{TO}}$, the transverse optical phonon frequency, because in the absence of LO--TO splitting, they are degenerate.
 This is done in order to make the form of the final results more closely resemble their 3D counterparts.
@@
 Given that the zeros of the dielectric function are associated with longitudinal modes, one expects that $\omega_{\mathbf{q}}$ is in fact the frequency of the LO phonon mode.
-This is consistent with the result of Ref.~\citenum{sohier2017breakdown}, in which \relax{{}\ } \citeauthor{sohier2017breakdown} show that in two-dimensional polar materials, the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO--TO splitting and zero LO--TO splitting at the $\Gamma$ point of the Brillouin zone.
+This is consistent with the result of Ref.~\citenum{sohier2017breakdown}, in which \relax{{}\ }\citeauthor{sohier2017breakdown} show that in two-dimensional polar materials, the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO--TO splitting and zero LO--TO splitting at the $\Gamma$ point of the Brillouin zone.
 
 Given these results, we now re-express the conductivity explicitly in terms of the 2D phonon dispersion, and derive a universal form for the conductivity in the local ($q\rightarrow 0$) limit specified in terms of three parameters: the TO phonon frequency, the group velocity of the LO phonon at the $\Gamma$ point, and the damping rate.
-From Equation (7), we can immediately write the conductivity as
+From Equation~\eqref{eq:zerosofresponse}, we can immediately write the conductivity as
   \begin{equation}
@@
   \end{equation}
- For small $q$,  the Born charges are (to lowest-order) constant, and so the LO phonon dispersion takes the form $\omega_{\mathbf{q},\mathrm{LO}}=\sqrt{\omega^2_{\mathrm{TO}}+2v_g\omega_{\mathrm{TO}}q} \approx \omega_{\mathrm{TO}} + v_gq$, where $v_g$, the LO phonon group velocity, is defined from microscopic parameters through the relation
+ For small $q$,  the Born charges are (to lowest-order) constant, and so the LO phonon dispersion takes the form $\omega_{\mathbf{q},\mathrm{LO}}=\sqrt{\omega^2_{\mathrm{TO}}+2v_{\mathrm{g}}\omega_{\mathrm{TO}}q} \simeq \omega_{\mathrm{TO}} + v_{\mathrm{g}}q$, where $v_{\mathrm{g}}$, the LO phonon group velocity, is defined from microscopic parameters through the relation
  \begin{equation}
@@
 \begin{equation}
-    \sigma(\omega) =  \frac{-4\iu\epsilon_0\epsilon_{\mathrm{env}}\omega_{\mathrm{TO}}v_g}{\omega^2_{\mathrm{TO}}-\omega^2-\iu\omega\tau^{-1}}.
+    \sigma(\omega) =  \frac{-4\iu\epsilon_0\epsilon_{\mathrm{env}}\omega_{\mathrm{TO}}v_{\mathrm{g}}}{\omega^2_{\mathrm{TO}}-\omega^2-\iu\omega\tau^{-1}}.
     \label{eq:localsigma}
@@
 The second comment is that much of what has been discussed here has a strong analogy with the theory of optical response in electron gases in 2D, and particularly the relation between plasmons in 2D and 3D.
-To elaborate on this analogy, To do so, we take Equation (7) in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $e^2\big|\sum_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \big|^2 \equiv  \Big(\frac{eZ^r_*}{\sqrt{M_*}}\Big)^2 \equiv \Big(\frac{Q_*}{\sqrt{M_*}}\Big)^2$.
+To elaborate on this analogy, To do so, we take Equation~\eqref{eq:zerosofresponse} in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $e^2\big|\sum_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \big|^2 \equiv  \Big(\frac{eZ^r_*}{\sqrt{M_*}}\Big)^2 \equiv \Big(\frac{Q_*}{\sqrt{M_*}}\Big)^2$.
 Then, the LO--TO splitting can be written as $\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} = \frac{Q_*^2}{2\epsilon_0\epsilon_{\mathrm{env}} M_*}q$.
@@
 Of course, there are no transverse plasma oscillations due to the structure-less nature of the electron, and so ``$\omega_{\mathrm{TP}}"$  should be considered as 0.
-In the three-dimensional plasmon case, `$\omega_{\mathrm{TP}} = 0$', but the difference between the squared longitudinal and transverse plasma oscillation frequencies at zero-wavevector is non-zero and given by $\omega_p^2$.
+In the three-dimensional plasmon case, ``$\omega_{\mathrm{TP}} = 0$'', but the difference between the squared longitudinal and transverse plasma oscillation frequencies at zero-wavevector is non-zero and given by $\omega_{\mathrm{p}}^2$.
 In other words, the plasma frequency in electron gases is analogous to the LO--TO splitting in polar materials.
@@
 \begin{figure}[t]
-    \includegraphics[width=85mm]{fig1.pdf}
+    \includegraphics[width=0.95\linewidth]{fig1.pdf}
     \caption{%
@@
 \begin{figure*}[t]
-    \includegraphics[width=135mm]{figure2hBN.pdf}
+    \includegraphics[width=.7\linewidth]{figure2hBN.pdf}
     \caption{%
-        \textbf{Properties of phonon polaritons in an monolayer of boron nitride compared to bulk.} (a) Real and imaginary parts of the conductivity of 2D hexagonal boron nitride in units of $e^2/\hbar$ for different values of the loss-rate.
-(b) Propagation quality factor, which measures the number of wavelengths of propagation of the 2D phonon polariton.
-(c) Dispersion relation of phonon polaritons in the monolayer (black) and thin films whose thicknesses are taken to be 1,2,4, and 8 interlayer spacings in hBN.
+        \textbf{Properties of phonon polaritons in a monolayer of boron nitride compared to bulk.} (a)~Real and imaginary parts of the conductivity of 2D hexagonal boron nitride in units of $e^2/\hbar$ for different values of the loss-rate.
+(b)~Propagation quality factor, which measures the number of wavelengths of propagation of the 2D phonon polariton.
+(c)~Dispersion relation of phonon polaritons in the monolayer (black) and thin films whose thicknesses are taken to be 1,2,4, and 8 interlayer spacings in hBN.
 For the thin-film, the fundamental mode and the first higher-order mode are plotted.
@@
 
-In Figure ~\ref{fig:2}(a), we present the conductivity of 2D hBN of Eq.~\ref{eq:localsigma}, using parameters from Ref.
-$\cite{sohier2017breakdown}$ calculated from density functional theory within the local density approximation.
+In Figure ~\ref{fig:2}(a), we present the conductivity of 2D hBN of Equation~\ref{eq:localsigma}, using parameters from Ref.~$\citenum{sohier2017breakdown}$ calculated from density functional theory within the local density approximation.
 From the conductivity, the dispersion relation of phonon polaritons on an infinite sheet is given by $q=\frac{2\iu\omega\epsilon_0\epsilon_{\mathrm{env}}}{\sigma(\omega)}$.
 The dispersion, assuming $\tau = \infty$, and $\epsilon_{\mathrm{env}}=1$ is shown in \ref{fig:2}(c) (black line).
-A key figure of merit for applications involving the propagation of phonon polaritons, is the propagation quality factor, defined by $\frac{\text{Re }q}{\text{Im }q} = \frac{\text{Im }\sigma(\omega)}{\text{Re }\sigma(\omega)}$, which is shown in \ref{fig:2}(b).
+A key figure of merit for applications involving the propagation of phonon polaritons, is the propagation quality factor, defined by ${\Re q}/{\Im q} = {\Im\sigma(\omega)}/{\Re\sigma(\omega)}$, which is shown in Figure~\ref{fig:2}(b).
 
 For monolayer hBN, the wavevector grows very rapidly with frequency, due to the extremely low group velocity of 2D LO phonons, which is a remarkable four orders of magnitude slower than the speed of light.
-In particular, at frequencies of \SI{1450}{\per\cm}, the phonon-polariton has a wavelength of about 15 nm, which is much shorter wavelength than any phonon-polariton measured so far, and, similarly shorter than any plasmonic wavelength, even in graphene.
-In fact, this short a wavelength is essentially out of the range of measurability beyond the range of current scattering near-field microscope (SNOM) measurements, which are limited (by tip radii) to wavelengths greater than $\approx\,\SI{60}{\nm}$.
+In particular, at frequencies of \SI{1450}{\per\cm}, the phonon-polariton has a wavelength of about \SI{15}{\nm}, which is much shorter wavelength than any phonon-polariton measured so far, and, similarly shorter than any plasmonic wavelength, even in graphene.
+In fact, this short a wavelength is essentially out of the range of measurability beyond the range of current scattering near-field microscope (SNOM) measurements, which are limited (by tip radii) to wavelengths greater than ${\approx}\,\SI{60}{\nm}$.
 The 2D LO phonon-polariton could in principle be measured by SNOM closer to the TO frequency, where confinement is smaller; unfortunately, as shown in Figure~\ref{fig:2}(b), near the TO frequency, dissipation is far higher (and corresponding propagative quality factors $\Re q/\Im q$ far lower) due to large $\Re\sigma$ (or, equivalently, large $\Im\epsilon_{\mathrm{RPA}}$).
@@
 In the rest of this section, we discuss the relation between the dispersion of an hBN monolayer versus the atomically-thick limit of a thin film of a material with hBN's bulk dielectric function.
-To aid this discussion, in \ref{fig:2}(c), we show the dispersion relation of thin films of `bulk' hBN with film thicknesses of 1, 2, 4, and 8 times the interlayer spacing of bulk hBN, which is roughly 0.33 nm.
-For these plots, we take hBN to be cleaved such that the optic-axis is perpendicular to the plane of the film.
-In such a case, the components of the bulk permittivity perpendicular and parallel to the c-axis ($\epsilon_{\perp}$ and $\epsilon_{||}$ respectively) are given by
+To aid this discussion, in \ref{fig:2}(c), we show the dispersion relation of thin films of `bulk' hBN with film thicknesses of 1, 2, 4, and 8 times the interlayer spacing of bulk hBN, which is roughly \SI{0.33}{\nm}.
+For these plots, we take hBN to be cleaved such that the optical axis is perpendicular to the plane of the film.
+The components of the bulk permittivity perpendicular and parallel to the c-axis ($\epsilon_{\perp}$ and $\epsilon_{\parallel}$, respectively; indexed by $\alpha\in\{\perp,\parallel\}$ below) are then given by
 \begin{equation}
-\epsilon_{\perp,||}(\omega) = \epsilon_{\infty\perp,||}\left(1 + \frac{\omega^2_{\mathrm{LO}\perp,||}-\omega^2_{\mathrm{TO}\perp,||}}{\omega^2-\omega^2_{\mathrm{TO}\perp,||}} \right),
+\epsilon_{\alpha}(\omega) = \epsilon_{\infty,\alpha}\Bigg(1 + \frac{\omega^2_{\mathrm{LO},\alpha}-\omega^2_{\mathrm{TO},\alpha}}{\omega^2-\omega^2_{\mathrm{TO},\alpha}} \Bigg),
 \end{equation}
-with $\epsilon_{\infty\perp,||} = 2.95, 4.87$, $\omega_{\mathrm{TO}\perp,||} = 760,1360$ cm$^{-1}$, and $\omega_{\mathrm{LO}\perp,||} = 830,1614$ cm$^{-1}$ \cite{caldwell2014sub,dai2014tunable}.
+with $\epsilon_{\infty,\perp} = \num{2.95}$, $\omega_{\mathrm{TO},\perp} = \SI{760}{\per\cm}$, and $\omega_{\mathrm{LO},\perp} = \SI{830}{\per\cm}$; and $\epsilon_{\infty,\parallel} = \num{4.87}$, $\omega_{\mathrm{TO},\parallel} = \SI{1360}{\per\cm}$, and $\omega_{\mathrm{LO},\parallel} = \SI{1614}{\per\cm}$\cite{caldwell2014sub,dai2014tunable}.
 Losses are ignored in this discussion altogether. 
-In the range between $\omega_{\mathrm{TO}}$ and $\omega_{\mathrm{LO}}$, the corresponding component of the permittivity is negative, while the other component is positive.
-This hyperbolicity leads to a dispersion for hBN thin films that have multiple branches at a given frequency, as can be seen in \ref{fig:2}(c).
+In the range between $\omega_{\mathrm{TO}, \alpha}$ and $\omega_{\mathrm{LO}, \alpha}$, the corresponding component of the permittivity is negative, while the other component is positive.
+This hyperbolicity leads to a dispersion for hBN thin films that have multiple branches at a given frequency, as can be seen in Figure~\ref{fig:2}(c).
 This trend persists even when the thickness of the bulk is taken down to a single layer, albeit being pushed to high wavevectors.
@@
 This can never happen by extrapolating to a monolayer a dielectric function which has finite LO--TO splitting.
-This `turn-back' point corresponds to a frequency of around $1500$ cm$^{-1}$ and a phonon wavelength of less than 1 nm.
+This `turn-back' point corresponds to a frequency of around \SI{1500}{\per\cm} and a phonon wavelength of less than \SI{1}{\nm}.
 Thus, this deviation is unlikely to be probed in typical optical experiments.
@@
 \begin{figure}[t]
-    \includegraphics[width=80mm]{fig3.pdf}
+    \includegraphics[width=0.95\linewidth]{fig3.pdf}
     \caption{%
@@
 The emitter is concentric with the disk and is at a distance $z=\SI{5}{\nm}$ away from the disk.
-For a disk with a diameter of \SI{20}{\nm}, and assuming a relaxation rate $\tau^{-1} = 5$ cm$^{-1}$, the rate of emission enhancement can be enhanced \num{100} million-fold.
+For a disk with a diameter of \SI{20}{\nm}, and assuming a relaxation rate $\tau^{-1} = \SI{5}{\per\cm}$, the rate of emission enhancement can be enhanced \num{100} million-fold.
 For an emitter with a free-space decay rate of $\SI{1e6}{\per\cm}$ at \SI{7}{\micro\m}, the emitter would experience a decay rate comparable to the frequency of the disk mode, leading to ultra-strong coupling of an external emitter with 2D LO phonons.
-Note that $\tau^{-1} = 5$ cm$^{-1}$ is of the order of the loss rate in bulk hBN.
+Note that $\tau^{-1} = \SI{5}{\per\cm}$ is of the order of the loss rate in bulk hBN.
         \label{fig:3}
@@
 The extreme confinement of electromagnetic fields offered by the 2D LO phonon presents an opportunity for quantum optical applications in which one seeks to couple an external emitter such as an atom, molecule, defect, or artificial atomic system to electromagnetic fields.
-Applications of these couplings are ultra-bright single- or two-photon sources, realizing the strong-coupling regime and the associated phenomenology of Rabi oscillations and polaritons, or resolving spectroscopically `forbidden'' transitions \cite{koppens2011graphene,autore2018boron,rivera2016shrinking,rivera2017making,kurman2018control}.
+Applications of these couplings are ultra-bright single- or two-photon sources, realizing the strong-coupling regime and the associated phenomenology of Rabi oscillations and polaritons, or resolving spectroscopically ``forbidden'' transitions \cite{koppens2011graphene,autore2018boron,rivera2016shrinking,rivera2017making,kurman2018control}.
 
@@
 For simplicity, we consider hBN nanostructured as a disk, which leads to the formation of sharp resonances quantized along the azimuthal and radial directions.
-The disk is taken to have a radius of \SI{20}{cm} and a loss rate $\tau^{-1} = 5$ cm$^{-1}$, which is of the order of the loss rate in bulk hBN.
+The disk is taken to have a radius of \SI{20}{cm} and a loss rate $\tau^{-1} = \SI{5}{\per\cm}$, which is of the order of the loss rate in bulk hBN.
+We parameterize the coupling between the dipole and phonon-polaritons through the rate of spontaneous emission $\Gamma$ of phonon polaritons by the dipole.
+It is related to the dyadic Green function $\mathbf{G}$ of the Maxwell equations for the nanostructure via the expression \cite{novotny2012principles}:
+\begin{equation}
+\frac{\Gamma}{\Gamma_0} = 1 + \frac{6\pi c}{\omega_0}\hat{\mathbf{n}}\cdot\Im\mathbf{G}(\mathbf{r}_0,\mathbf{r}_0,\omega_0)\cdot\hat{\mathbf{n}},
+\end{equation}
+where $\hat{\mathbf{n}}$ is the orientation of the dipole, $\mathbf{r}_0$ is its position, and $\omega_0$ is its frequency.
+The dyadic Green function is computed through a boundary element method (as in Ref.~\citenum{jin2017infrared}) assuming the validity of the electrostatic approximation, which is well-respected due to the extreme sub-wavelength confinement of the electromagnetic field of the phonon polariton.
 In Figure~\ref{fig:3}, we plot the enhancement of the spontaneous emission rate for an external emitter polarized perpendicularly to the plane of the disk and placed \SI{5}{\nm} away from the surface of the disk.
 We find that the rate of spontaneous emission of 2D optical phonons is approximately 8 orders of magnitude larger than the rate of spontaneous emission of photons in the far field.
-For an infrared emitter at a transition wavelength of \SI{7}{\micro\m} with a free-space radiative lifetime of \SI{1}{\micro\s}, \SI{5}{\nm} away from an hBN disk, the coupling rate to 2D optical phonons (about $\SI{1e14}{\per\s}$ or 0.065 eV) would be on the same scale as the optical phonon frequency itself (about $1430$ cm$^{-1}$ or $\SI{2.70e14}{\per\s}$ or 0.18 eV).
+We note that in this approach, the coupling of the dipole to phonon polaritons is manifested through the phonon contribution to the dielectric function of the disk.
+This should be equivalent to an approach that considers the coupling of a bound electron in an emitter to LO phonons in the disk through a (two-dimensional) Fr\"ohlich coupling---\ie a coupling of the atomic electron to the electric potential resulting from the polarization associated with an LO phonon mode \cite{sohier2016two}.
+
+In Figure~\ref{fig:3}, we show that for an infrared emitter at a transition wavelength of \SI{7}{\micro\m} with a free-space radiative lifetime of \SI{1}{\micro\s}, \SI{5}{\nm} away from an hBN disk, the coupling rate to 2D optical phonons (about $\SI{1e14}{\per\s}$ or \SI{65}{\meV}) would be on the same scale as the optical phonon frequency itself (about \SI{1430}{\per\cm} or $\SI{2.70e14}{\per\s}$ or \SI{0.18}{\eV}).
+\comment{tc}{We are using three distinct units for energy, \si{\per\s}, \si{\per\cm}, and \si{\eV}, in various places.
+At most, we should probably use two units (then, I would prefer to get rid of \si{\per\s}.
+Even if we don't do that, we probably don't need to give it in all three units here.
+When writing unitful quantities, please use the \texttt{\textbackslash{}SI\{value\}\{unit\}} macro (\eg \texttt{\textbackslash{}SI\{2\}\{\textbackslash{}eV\}} for \SI{2}{\eV} or \texttt{\textbackslash{}SI\{7\}\{\textbackslash{}per\textbackslash{}cm\}} for \SI{7}{\per\cm}).}
 This coupling rate thus implies coupling between an emitter and the field in the regime of ultra-strong coupling.
-Moreover, the coupling rate for the 20 nm disk (orange), for an emitter with a far field decay rate of $> \SI{3e14}{\per\s}$ would have a sufficient coupling strength to the resonances in orange for its linewidth to span multiple resonances and thus be in a multi-mode ultra-strong coupling regime.
+Moreover, the coupling rate for the \SI{20}{\nm} disk (orange), for an emitter with a far field decay rate of ${>}\,\SI{3e14}{\per\s}$ would have a sufficient coupling strength to the resonances in orange for its linewidth to span multiple resonances and thus be in a multi-mode ultra-strong coupling regime.
 Thus, the extreme confinement of electromagnetic energy associated with LO phonons in two dimensions enables the possibility of realizing ultrastrong coupling of an atom or molecule with optical phonons in a polar material, allowing the potential realization of new coupled states of quantum emitters and phonons such as atom-phonon polariton bound states.
@@
 acknowledges support from the Danish Council for Independent Research (Grant No.\ DFF--6108-00667).
-The authors thank Prof.
-Ido Kaminer (Technion), Siyuan Dai (Columbia), Samuel Moore (Columbia), Jennifer Coulter (Harvard), and Christopher Ciccarino (Harvard) for helpful discussions.
+The authors thank Ido Kaminer, Siyuan Dai, Samuel Moore, Jennifer Coulter, and Christopher Ciccarino for helpful discussions.
 This work was supported by the DOE Photonics at Thermodynamic Limits Energy Frontier Research Center under grant number DE-SC0019140.

### 2019-01-08 paper.tex
more edits and comments

@@
 \usepackage[centering,hmargin=20mm,tmargin=28mm,bmargin=25mm]{geometry}
-\usepackage{multirow}
 \usepackage{newtxtext}
@@
 Does this lack of LO--TO splitting imply the absence of a phonon-polariton in polar monolayers?
-    Here, we derive a first-principles expression for the spatially non-local conductivity of a polar monolayer specified by the wavevector-dependent LO and TO phonon dispersions.
-In the long-wavelength (local) limit, we find a universal form for the conductivity in terms of the TO phonon frequency at the $\Gamma$ point, its lifetime, and the group velocity of the LO phonon.
+    Here, we derive a first-principles expression for the spatially nonlocal conductivity of a polar monolayer specified by the wavevector-dependent LO and TO phonon dispersions.
+In the long-wavelength (local) limit, we find a universal form for the conductivity in terms of the TO phonon frequency at the $\Gamma$ point,
+    \comment{tc}{Why not just say `the LO phonon frequency' rather than `TO' here, since they are degenerate anyway---that way, we would only have to reference LO quantities for the local conductivity (and that case, `and the group velocity of the LO phonon' would be swapped for `and its group velocity').}
+    its lifetime, and the group velocity of the LO phonon.
 Our analysis reveals that the phonon-polariton of the bulk is essentially replaced by the LO phonon of the two-dimensional system.
 We calculate the confinement and propagation losses of the LO phonons, finding that high confinement and reasonable propagation quality factors coincide in regions which are difficult to detect with current near-field optical microscopy techniques.
-    Finally, we study the interaction of external emitters with two-dimensional hBN nanostructures, finding extreme enhancement of spontaneous emission due to coupling with localized 2D LO phonons, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN phonons.
+    Finally, we study the interaction of external emitters with two-dimensional hBN nanostructures, 
+    \comment{tc}{This is the first mention of hBN in the abstract; arguably, it should occur earlier, already at the `We calculate the confinement and propagation losses \ldots' part}
+    finding extreme enhancement of spontaneous emission due to coupling with localized 2D \comment{tc}{`2D' acronym not yet defined.} LO phonons, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN phonons.
+    \comment{tc}{I would suggest to lose this last sentence of the abstract: it doesn't add anything very concrete and it just (undisguised) public-relations work.}
     Using the strong electromagnetic interactions between emitters and 2D LO phonons calculated in this work, it is now possible to design new hybrid states of light, electrons and phonons based on quantum electrodynamical strong coupling.
@@
 Qualitatively, phonon-polaritons share many features with plasmon-polaritons in conductors.
-Recently, it has been shown that phonon-polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra,li2018infrared,ma2018plane}.
-Due to this remarkable confinement and their relatively high lifetimes of around picoseconds, phonon-polaritons open new opportunities for vibrational spectroscopy,\cite{autore2018boron} radiative heat transfer~\cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters~\cite{kumar2015tunable,rivera2017making,kurman2018control}.
-In many cases, extreme confinement of phonon-polaritons is achieved by the use of thin-films, which shrink the in- and out-of-plane wavelength of polaritons monotonically with decreasing film thickness~\cite{dai2014tunable,dubrovkin2018ultra}.
-A monolayer is the ultimate, atomic-scale limit of this effect when the properties of the bulk material can be used in the monolayer limit.
-In polar materials, the transition from 3D to 2D is not so simple, as in a polar monolayer, the LO--TO splitting that gives rise to phonon-polaritons in three-dimensions is absent at the $\Gamma$ point~\cite{sanchez2002vibrational,mele2002electric,serrano2007vibrational,sohier2017breakdown}.
+Recently, it has been shown that phonon-polaritons enable confinement of light to volumes \remove{over} $\inset{\sim}10^6$ times smaller than that of a diffraction-limited photon in free-space~\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra,li2018infrared,ma2018plane}.
+Due to this remarkable confinement and their relatively high lifetimes---around picoseconds---phonon-polaritons open new opportunities for vibrational spectroscopy~\cite{autore2018boron}, radiative heat transfer~\cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters~\cite{kumar2015tunable,rivera2017making,kurman2018control}.
+\swap{In many cases}{Thus far}, extreme confinement of phonon-polaritons \swap{is}{has been} achieved by the use of thin-films, which shrink the in- and out-of-plane wavelength of polaritons monotonically with decreasing film thickness~\cite{dai2014tunable,dubrovkin2018ultra}.
+A monolayer is the ultimate, atomic-scale limit of this effect\remove{ when the properties of the bulk material can be used in the monolayer limit}.
+\swap{In polar materials, the transition from 3D to 2D is not so simple}{The response-transition from three-dimensional (3D) to two-dimensional (2D) polar materials is nontrivial, however}, since in a polar monolayer, the LO--TO splitting that gives rise to phonon-polaritons in \swap{three-dimensions}{3D} is absent at the $\Gamma$ point~\cite{sanchez2002vibrational,mele2002electric,serrano2007vibrational,sohier2017breakdown}.
 This raises a fundamental question about whether strongly confined electromagnetic modes exist in polar monolayers at all.
+\comment{tc}{@NR: Check that 3D \& 2D acronyms are only defined once (above, in my suggestion).}
 
-In this Letter, we develop a theoretical and computational framework with quantitative, concrete examples to enable a new understanding of optical phononics in two-dimensional materials.
-We do so by deriving a universal form for the conductivity of a polar monolayer,  which depends solely on the LO and TO phonon frequencies---and their dispersion with momentum---in the two-dimensional system.
-Using parameters from Ref.~\citenum{sohier2017breakdown} for the canonical two-dimensional polar monolayer---hexagonal boron nitride (hBN)---we present the confinement and propagation losses of the 2D phononic modes, identifying the frequency region where they should be most easily detected.
-Finally, we show that these modes enable extreme light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced by up to eight orders of magnitude by the emitter--LO phonon coupling.
-For an infrared emitter with a sufficiently high free-space radiative decay rate (${>}\, \SI{1e6}{\per\s}$), the enhanced linewidth of the emitter becomes comparable to the spacing between different phonon polaritonic resonances of an hBN nanostructure.
-This suggests the possibility of realizing the multi-mode strong coupling and ultra-strong coupling regimes of quantum electrodynamics in a two-dimensional hBN platform.
+In this Letter, we develop a theoretical and computational framework with quantitative, concrete examples that \swap{enable}{enables} a \swap{new}{first-principles} understanding of \swap{optical phononics}{phonon polaritons} in \swap{two-dimensional}{2D} materials.
+\comment{tc}{I don't think `optical phononics' is a generally accepted identifier for phonon polaritons.
+We use the two somewhat interchangably in this manuscript: maybe settle on just one?
+(Only a single use of `phononics' (this one)---we have three uses of `phononic' elsewhere though: I think only one should be changed.)}
+We do so by deriving a universal form for the conductivity of a polar monolayer,  which depends solely on the LO and TO phonon frequencies---and their dispersion with momentum---in the \swap{two-dimensional}{2D} system.
+Using parameters from Ref.~\citenum{sohier2017breakdown} for the canonical \swap{two-dimensional}{2D} polar monolayer---hexagonal boron nitride (hBN)---we present the confinement and propagation losses of the 2D \swap{phononic}{phonon-polariton }modes, identifying the frequency region where they should be most easily detected.
+Finally, we \swap{show}{find} that these modes enable extreme light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced by up to eight orders of magnitude \swap{by}{through/via} the emitter--LO phonon coupling.
+For an infrared emitter with a sufficiently high free-space radiative decay rate (${\gtrsim}\, \SI{1e6}{\per\s}$), \inset{we find that} the \swap{enhanced}{associated} linewidth of the emitter \swap{becomes}{is} comparable to the spacing between different phonon polaritonic resonances of an hBN nanostructure.
+This suggests the possibility of realizing the multi-mode strong coupling and ultra-strong coupling regimes of quantum electrodynamics in a \swap{two-dimensional}{2D} hBN platform.
 Our results for hBN are particularly relevant due its widespread use in 2D van der Waals heterostructures.
-In addition to providing functionality as a layer which improves the electrical and optical properties of other 2D materials, such as graphene, our results suggest that in these heterostructures, the hBN layers could provide a platform for nanophotonics and quantum optics which operates in the mid-infrared spectral range.
-While we focus on hBN in this manuscript, many of our findings apply to other polar monolayers.
+In addition to providing functionality as a layer which improves the electrical and optical properties of other 2D materials, \swap{such as}{\eg} graphene, our results suggest that in these heterostructures, \remove{the }hBN layers could provide a \inset{mid-infrared} platform for nanophotonics and quantum optics\remove{ which operates in the mid-infrared spectral range}.
+While we focus on hBN in this manuscript, \swap{many of}{the salient features of} our findings apply to other polar monolayers\inset{ as well}.
 
 \section{Optical response of optical phonons in two-dimensions}
-In this section, we develop a theory of electromagnetic response due to optical phonons in two-dimensional systems.
-The key prediction is the conductivity of the monolayer.
-To obtain it, we consider the response of the ions of the monolayer due to a longitudinal electromagnetic potential $\phi$.
-The interaction Hamiltonian in this case is
+In this section, we develop a theory of electromagnetic response due to optical phonons in \swap{two-dimensional}{2D} systems.
+The key \swap{prediction}{response function of interest} is the conductivity of the monolayer.
+To \swap{obtain it}{that end}, we consider the response of the ions of the monolayer due to a longitudinal electromagnetic potential $\phi$.
+For that case, the interaction Hamiltonian is
 \begin{equation}
@@
 Specifically, to first order, the induced polarization is (in repeated-index notation):
+\comment{tc}{Later on, we write things like $\sum_{\kappa}\mathbf{Z}_{\kappa}\boldsymbol{\eta}_\kappa$, so we are not entirely consistent in our application of repeated-index notation.
+Perhaps we should just write vectorial, explicit summation notation here as well (\ie write $\mathbf{P} - \mathbf{P}^0 = \sum_\kappa(\boldsymbol{\nabla}_{\mathbf{u}_\kappa}\mathbf{P})\mathbf{u}_\kappa \equiv \frac{1}{\Omega}\sum_\kappa\mathbf{Z}_{\kappa}\mathbf{u}_{\kappa}$)}
 \begin{equation}
-    P_i - P_i^{0} = \frac{\partial P_i}{\partial u_{\kappa,j}}u_{\kappa,j} \equiv \frac{1}{\Omega}Z_{\kappa,ij}u_{\kappa,j},
+    P_i - P_i^{0} \swap{=}{\simeq} \frac{\partial P_i}{\partial u_{\kappa,j}}u_{\kappa,j} \equiv \frac{1}{\Omega}Z_{\kappa,ij}u_{\kappa,j},
 \end{equation} 
-where $Z_{\kappa,ij} \equiv \Omega\frac{\partial P_i}{\partial u_{\kappa,j}} $ is the Born effective charge tensor of ion $\kappa$ and $\Omega$ is the unit cell area.
-$P_i^{0}$ is the equilibrium polarization in the absence of displacements, which is zero here.
+where \swap{$Z_{\kappa,ij} \equiv \Omega\frac{\partial P_i}{\partial u_{\kappa,j}}$}{$\mathbf{Z}_\kappa \equiv \Omega\boldsymbol{\nabla}_{\mathbf{u}_\kappa}\mathbf{P}$} is the Born effective charge tensor of ion $\kappa$ and $\Omega$ is the unit cell area.
+\swap{$P_i^{0}$}{$\mathbf{P}^0$} is the equilibrium polarization in the absence of displacements, which is zero here.
 With this relation between polarization and ionic displacements, the interaction Hamiltonian in Equation~\eqref{eq:coupling} couples the scalar potential and the ionic displacements.
-We consider the response of the monolayer to a scalar potential of the form $\phi(\mathbf{r}) = \phi(\mathbf{q},\omega)\e^{\iu\mathbf{q}\cdot\mathbf{r}-\iu\omega t}$, where $\mathbf{q}$ is a two-dimensional wavevector in the plane of the monolayer.
+We consider the response of the monolayer to a scalar potential of the form $\phi(\mathbf{r}) = \phi(\mathbf{q},\omega)\e^{\iu\mathbf{q}\cdot\mathbf{r}-\iu\omega t}$, where $\mathbf{q}$ is a \swap{two-dimensional}{2D} wavevector in the plane of the monolayer.
 Such a potential corresponds to a longitudinal electric field  $\mathbf{E}(\mathbf{r}) = \iu\mathbf{q}\phi(\mathbf{r})$.
@@
 \end{equation}
-where $m,n$ are states of the phononic Fock space of the monolayer, $P_{mn,i}(\mathbf{q}) = \sum_{\kappa,j}Z_{\kappa,ij}\langle m | u_{\kappa,j}(\mathbf{q}) | n \rangle$ are matrix elements of the polarization associated with phonon modes (where $u_{\kappa,j}(\mathbf{q})$ is the Fourier transform of the phonon displacement operator), $E_{m}$ ($E_n$) is the energy of state $m$ ($n$), $\beta \equiv 1/k_{\text{\textsc{B}}}T$ is the inverse temperature, and $\mathcal{Z}$ is the grand partition function.
+where $m,n$ are states of the phononic Fock space of the monolayer, \swap{$P_{mn,i}(\mathbf{q}) \swap{=}{\equiv} \sum_{\kappa,j}Z_{\kappa,ij}\langle m | u_{\kappa,j}(\mathbf{q}) | n \rangle$}{$\mathbf{P}_{mn}(\mathbf{q}) \equiv \sum_\kappa \mathbf{Z}_\kappa\langle m|\mathbf{u}_\kappa(\mathbf{q})|n\rangle$} are matrix elements of the polarization associated with phonon modes (where \swap{$u_{\kappa,j}(\mathbf{q})$}{$\mathbf{u}_\kappa(\mathbf{q})$} is the Fourier transform of the phonon displacement operator), $E_{m}$ ($E_n$) is the energy of state $m$ ($n$), $\beta \equiv 1/k_{\text{\textsc{B}}}T$ is the inverse temperature, and $\mathcal{Z}$ is the grand partition function.
 We now evaluate the contribution of optical phonons to the polarization-polarization response in the low-temperature limit $T \ll \hbar\omega_{\mathrm{ph}}/k_{\mathrm{B}}$ with $\omega_{\mathrm{ph}}$ a characteristic optical phonon frequency.
@@
 \begin{equation}
-    \sigma(\mathbf{q},\omega) = -\frac{\iu\omega }{\Omega}\frac{\Big|\hat{\mathbf{q}}\cdot\sum\limits_{\kappa}\mathbf{Z}_{\kappa}\boldsymbol{\eta}_{\kappa} \Big|^2}{\omega^2_{\mathbf{q},\mathrm{L}}-\omega^2-\iu\omega\tau^{-1}}
+    \sigma(\mathbf{q},\omega) = -\frac{\iu\omega }{\Omega}\frac{\Big|\hat{\mathbf{q}}\cdot\sum\limits_{\kappa}\mathbf{Z}_{\kappa}\boldsymbol{\eta}_{\kappa} \Big|^2}{\omega^2_{\mathbf{q},\mathrm{L}}-\omega^2-\iu\omega\tau^{-1}}.
     \label{eq:conductivity_simplified}
@@
 \end{equation}
+\comment{tc}{The tilde-quantity'ed $\tilde{\mathbf{Z}}_\kappa$ is not defined, nor is the $r$-quantity'ed $Z^r_{\kappa,ij}$ later on.
+I suspect they are the same thing, and that it is the $e\tilde{\mathbf{Z}}_\kappa \equiv \mathbf{Z}_\kappa$ from an earlier draft.
+@NR: Need to check and fix.}
   where $V(q)$ is the Coulomb interaction in Fourier space, which in two dimensions, is given by $V(q) = \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}$.
 Given that the zeros of the dielectric function are associated with longitudinal modes, one expects that $\omega_{\mathbf{q}}$ is in fact the frequency of the LO phonon mode.
-This is consistent with the result of Ref.~\citenum{sohier2017breakdown}, in which \relax{{}\ }\citeauthor{sohier2017breakdown} show that in two-dimensional polar materials, the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO--TO splitting and zero LO--TO splitting at the $\Gamma$ point of the Brillouin zone.
+This is consistent with the result of Ref.~\citenum{sohier2017breakdown}, in which \relax{{}\ }\citeauthor{sohier2017breakdown} show that in \swap{two-dimensional}{2D} polar materials, the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO--TO splitting and zero LO--TO splitting at the $\Gamma$ point of the Brillouin zone.
 
@@
   \begin{equation}
-  \sigma(q,\omega) = \frac{-2\iu\epsilon_0\epsilon_{\mathrm{env}}}{q}\frac{\omega^2_{\mathbf{q},\mathrm{LO}}-\omega^2_{\mathrm{TO}}}{\omega^2_{\mathrm{TO}}-\omega^2-\iu\omega\tau^{-1}}.
+  \sigma(\mathbf{q},\omega) = \frac{-2\iu\epsilon_0\epsilon_{\mathrm{env}}}{q}\frac{\omega^2_{\mathbf{q},\mathrm{LO}}-\omega^2_{\mathrm{TO}}}{\omega^2_{\mathrm{TO}}-\omega^2-\iu\omega\tau^{-1}}.
   \end{equation}
@@
 \end{equation}
+\comment{tc}{There is an unhappy situation here: there is a lingering $i$-index, that is not summed out.
+Perhaps this should've been $|\hat{\mathbf{q}}\cdot \sum_\kappa \mathbf{Z}_\kappa\boldsymbol{\eta}_\kappa|^2$?}
 Thus, in the long wavelength limit, we have the following universal parameterization of the conductivity of a polar monolayer:
@@
 \end{equation}
+\comment{tc}{@NR: As we discussed, it would probably be good to explicitly point out that $\sigma(\omega)$ \emph{doesn't} have an explicit $\epsilon_{\mathrm{env}}$ dependence, as it cancels with the dependence in $v_\mathrm{g}$.}
 From this relation, it follows that given the properties of the 2D phonons (from experiments or from \emph{ab initio} calculations), one can immediately specify the conductivity.
@@
 The second comment is that much of what has been discussed here has a strong analogy with the theory of optical response in electron gases in 2D, and particularly the relation between plasmons in 2D and 3D.
-To elaborate on this analogy, To do so, we take Equation~\eqref{eq:zerosofresponse} in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $e^2\big|\sum_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \big|^2 \equiv  \Big(\frac{eZ^r_*}{\sqrt{M_*}}\Big)^2 \equiv \Big(\frac{Q_*}{\sqrt{M_*}}\Big)^2$.
+To elaborate on this analogy, we take Equation~\eqref{eq:zerosofresponse} in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $e^2\big|\sum_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \big|^2 \equiv  \Big(\frac{eZ^r_*}{\sqrt{M_*}}\Big)^2 \equiv \Big(\frac{Q_*}{\sqrt{M_*}}\Big)^2$.
+\comment{tc}{Another lingering $i$-index here.}
 Then, the LO--TO splitting can be written as $\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} = \frac{Q_*^2}{2\epsilon_0\epsilon_{\mathrm{env}} M_*}q$.
-Now we note that the RHS is exactly the squared-frequency $\omega^2_{\mathbf{q}p}$ for a plasma oscillation in a 2D gas of charged particles with charge $Q_*$ and mass $M_*$.
+Now we note that the RHS is exactly the squared-frequency $\omega^2_{\mathbf{q}\mathrm{p}}$ for a plasma oscillation in a 2D gas of charged particles with charge $Q_*$ and mass $M_*$.
+\comment{tc}{You mention the meanings of $Q_*$ and $M_*$: you probably also ought to mention the (of course reasonably obvious) meaning of $Z_*$ (and, probably, it should be $\tilde{Z}_*$).
+Actually, it is not clear to me why we would need both $Q_*$ and $\tilde{Z}_*$ here?}
 This squared frequency can be thought of as the ``LP--TP'' splitting between longitudinal and transverse plasma oscillations.
@@
 \end{equation}
-where $\epsilon_1$ applies for $z>0$ and $\epsilon_2$ applies for $z<0$.
+where $\epsilon_+$ applies for $z>0$ and $\epsilon_-$ applies for $z<0$.
 %\comment{tc}{Probably makes sense to use $\varepsilon_+$ instead of $\varepsilon_1$ and $\varepsilon_-$ for $\varepsilon_2$, to have just a single class of subscripts for this notion.}
 We consider a solution of the form $H(z) = h_\pm\e^{\mp\kappa_\pm z}$ with $\kappa_\pm = \sqrt{q^2-\epsilon_{\pm} k_0^2}$ with $\pm$ corresponding to $\pm z > 0$ respectively.
-The boundary condition on the magnetic field is $h^+-h^- = -K_x = \iu\omega\mathbf{P}_x$ where $\mathbf{K}$ is the surface current density, expressed through the polarization density $\mathbf{P}$ of the monolayer.
-This condition enforces $\frac{h^-}{h^+} =(1+\frac{\epsilon_0}{\epsilon_-} \kappa_-\Pi)^{-1}$.
-Continuity of the electric field in the $x$ direction enforces $\frac{\epsilon_-}{\epsilon_+} = -\frac{\kappa_-}{\kappa_+}\frac{h^-}{h^+}$. 
+The boundary condition on the magnetic field is $h_+ - h_- = -K_x = \iu\omega P_x$ where $\mathbf{K}$ is the surface current density, expressed through the polarization density $\mathbf{P}$ of the monolayer.
+This condition enforces ${h_-}/{h_+} =(1+\frac{\epsilon_0}{\epsilon_-} \kappa_-\Pi)^{-1}$.
+Continuity of the electric field in the $x$ direction enforces ${\epsilon_-}/{\epsilon_+} = -{\kappa_-h_-}/{\kappa_+h_+}$. 
 Combining the two conditions, we obtain the usual dispersion equation for the TM mode of a polarizable 2D monolayer, namely $\epsilon_+/\kappa_+ + \epsilon_-/\kappa_- +\epsilon_0\Pi=0$.
@@
 with $\epsilon_{\mathrm{env}}\equiv (\epsilon_++\epsilon_-)/2$.
-Thus, the condition for `phonon-polaritons' coincides precisely with the condition for longitudinal optical phonons.
+Thus, the condition for `phonon-polaritons' \comment{tc}{Remove quotes} coincides precisely with the condition for longitudinal optical phonons.
 Thus, one of the main results of our manuscript is that despite the lack of LO--TO splitting at the $\Gamma$ point, there nevertheless exists a strongly confined evanescent mode in many respects similar to the phonon-polaritons of thin films.
@@
 In 2D, the plasma frequency at zero wavevector is zero, but the electromagnetic physics is determined by the dispersion of 2D plasmons, which replace the plasmon polariton of bulk and thin films.
-        %\comment{tc}{We already discussed reasonable changes to the figures in-person, so I won't comment on that this time around.}
         \label{fig:1}
@@
     \caption{%
-        \textbf{Properties of phonon polaritons in a monolayer of boron nitride compared to bulk.} (a)~Real and imaginary parts of the conductivity of 2D hexagonal boron nitride in units of $e^2/\hbar$ for different values of the loss-rate.
+        \textbf{Properties of phonon polaritons in a monolayer of \swap{\inset{hexagonal} boron nitride}{hBN} compared to bulk.} (a)~Real and imaginary parts of the conductivity of 2D \swap{hexagonal boron nitride}{hBN} \remove{in units of $e^2/\hbar$} for different values of the loss-rate.
 (b)~Propagation quality factor, which measures the number of wavelengths of propagation of the 2D phonon polariton.
-(c)~Dispersion relation of phonon polaritons in the monolayer (black) and thin films whose thicknesses are taken to be 1,2,4, and 8 interlayer spacings in hBN.
-For the thin-film, the fundamental mode and the first higher-order mode are plotted.
+(c)~Dispersion relation of phonon polaritons in the monolayer (black) and thin films whose thicknesses are taken to be 1, 2, 4, and 8 interlayer spacings in hBN.
+For the thin-film, the fundamental mode \inset{(dash-dotted)} and the first higher-order mode \inset{(dashed)} are plotted.
+\comment{tc}{@NR: Check that the dash-dotted and dashed-labelling is correct.
+From our discussion yesterday, I had the impression that the dashed lines were the ``fundamental'' modes and the dash-dotted the spatially-more-oscillatory ones: if that is the case, then `higher-order' is an unfortunate wording, as it suggests a higher frequency as well.}
         \label{fig:2}
@@
 In Figure ~\ref{fig:2}(a), we present the conductivity of 2D hBN of Equation~\ref{eq:localsigma}, using parameters from Ref.~$\citenum{sohier2017breakdown}$ calculated from density functional theory within the local density approximation.
-From the conductivity, the dispersion relation of phonon polaritons on an infinite sheet is given by $q=\frac{2\iu\omega\epsilon_0\epsilon_{\mathrm{env}}}{\sigma(\omega)}$.
-The dispersion, assuming $\tau = \infty$, and $\epsilon_{\mathrm{env}}=1$ is shown in \ref{fig:2}(c) (black line).
+From the conductivity, the dispersion relation of phonon polaritons on an infinite sheet is given by $q={2\iu\omega\epsilon_0\epsilon_{\mathrm{env}}}/{\sigma(\omega)}$.
+The dispersion, assuming $\tau = \infty$ and $\epsilon_{\mathrm{env}}=1$, is shown in Figure~\ref{fig:2}(c) (black line).
 A key figure of merit for applications involving the propagation of phonon polaritons, is the propagation quality factor, defined by ${\Re q}/{\Im q} = {\Im\sigma(\omega)}/{\Re\sigma(\omega)}$, which is shown in Figure~\ref{fig:2}(b).
@@
 For monolayer hBN, the wavevector grows very rapidly with frequency, due to the extremely low group velocity of 2D LO phonons, which is a remarkable four orders of magnitude slower than the speed of light.
-In particular, at frequencies of \SI{1450}{\per\cm}, the phonon-polariton has a wavelength of about \SI{15}{\nm}, which is much shorter wavelength than any phonon-polariton measured so far, and, similarly shorter than any plasmonic wavelength, even in graphene.
-In fact, this short a wavelength is essentially out of the range of measurability beyond the range of current scattering near-field microscope (SNOM) measurements, which are limited (by tip radii) to wavelengths greater than ${\approx}\,\SI{60}{\nm}$.
-The 2D LO phonon-polariton could in principle be measured by SNOM closer to the TO frequency, where confinement is smaller; unfortunately, as shown in Figure~\ref{fig:2}(b), near the TO frequency, dissipation is far higher (and corresponding propagative quality factors $\Re q/\Im q$ far lower) due to large $\Re\sigma$ (or, equivalently, large $\Im\epsilon_{\mathrm{RPA}}$).
+In particular, at frequencies of \SI{1450}{\per\cm}, the phonon-polariton has a wavelength of about \SI{15}{\nm}, significantly shorter than any phonon-polariton measured so far, and, similarly shorter than any plasmonic wavelength, even in graphene.
+In fact, this short a wavelength is essentially beyond the range of measurability of current scattering near-field microscopy (SNOM) setups, which are limited (by tip radii) to wavelengths greater than ${\approx}\,\SI{60}{\nm}$.
+The 2D LO phonon-polariton could in principle be measured by SNOM closer to the TO frequency, where confinement is smaller; unfortunately, as shown in Figure~\ref{fig:2}(b), near the TO frequency, dissipation is far higher (and corresponding propagation quality factors $\Re q/\Im q$ far lower) due to large $\Re\sigma$ (or, equivalently, large $\Im\epsilon_{\mathrm{RPA}}$).
 
-These considerations imply that access to the lower-loss and higher-confined portions of the dispersion relation of phonon-polaritons, in the absence of a sharp tip, requires a near-field probe such as a free electron probe, as used in electron energy loss spectroscopy (EELS), where slow electrons can be used to probe plasmon wavelengths of just a few nanometers in monolayer metals, as well as the nonlocal bulk plasmon dispersion in metals~\cite{nagao2001dispersion,de2010optical,diaconescu2007low}.
+These considerations imply that access to the lower-loss and higher-confined portions of the dispersion relation of phonon-polaritons, in the absence of a sharper tip, requires a near-field probe \swap{that probes larger momenta.
+Free electrons exemplify such high-momentum probes}{such as a free electron probe,} as used in electron energy loss spectroscopy (EELS), where slow electrons can be used to probe plasmon wavelengths of just a few nanometers in monolayer metals, as well as the nonlocal bulk plasmon dispersion in metals~\cite{nagao2001dispersion,de2010optical,diaconescu2007low}.
 EELS has been recently employed to measure phonon-polaritons in ultrathin films of hBN~\cite{govyadinov2017probing}.
@@
 In the rest of this section, we discuss the relation between the dispersion of an hBN monolayer versus the atomically-thick limit of a thin film of a material with hBN's bulk dielectric function.
-To aid this discussion, in \ref{fig:2}(c), we show the dispersion relation of thin films of `bulk' hBN with film thicknesses of 1, 2, 4, and 8 times the interlayer spacing of bulk hBN, which is roughly \SI{0.33}{\nm}.
+To aid this discussion, in Figure~\ref{fig:2}(c), we show the dispersion relation of thin films of `bulk' hBN with film thicknesses of 1, 2, 4, and 8 times the interlayer spacing of bulk hBN, which is roughly \SI{0.33}{\nm}.
 For these plots, we take hBN to be cleaved such that the optical axis is perpendicular to the plane of the film.
@@
 Note that $\tau^{-1} = \SI{5}{\per\cm}$ is of the order of the loss rate in bulk hBN.
+\comment{tc}{@NR: Citation needed.}
         \label{fig:3}
@@
 The disk is taken to have a radius of \SI{20}{cm} and a loss rate $\tau^{-1} = \SI{5}{\per\cm}$, which is of the order of the loss rate in bulk hBN.
+\comment{tc}{@NR: Citation needed.}
 We parameterize the coupling between the dipole and phonon-polaritons through the rate of spontaneous emission $\Gamma$ of phonon polaritons by the dipole.
@@
 \begin{equation}
-\frac{\Gamma}{\Gamma_0} = 1 + \frac{6\pi c}{\omega_0}\hat{\mathbf{n}}\cdot\Im\mathbf{G}(\mathbf{r}_0,\mathbf{r}_0,\omega_0)\cdot\hat{\mathbf{n}},
+	\frac{\Gamma}{\Gamma_0} 
+	= 
+	1 + \frac{6\pi c}{\omega_0}\hat{\mathbf{n}}\cdot\Im\mathbf{G}(\mathbf{r}_0,\mathbf{r}_0,\omega_0)\cdot\hat{\mathbf{n}},
 \end{equation}
 where $\hat{\mathbf{n}}$ is the orientation of the dipole, $\mathbf{r}_0$ is its position, and $\omega_0$ is its frequency.
+\comment{tc}{@NR: It could probably be fine to just drop the zero-subscript in $\mathbf{r}$ and $\omega$.}
 The dyadic Green function is computed through a boundary element method (as in Ref.~\citenum{jin2017infrared}) assuming the validity of the electrostatic approximation, which is well-respected due to the extreme sub-wavelength confinement of the electromagnetic field of the phonon polariton.
@@
 We note that in this approach, the coupling of the dipole to phonon polaritons is manifested through the phonon contribution to the dielectric function of the disk.
-This should be equivalent to an approach that considers the coupling of a bound electron in an emitter to LO phonons in the disk through a (two-dimensional) Fr\"ohlich coupling---\ie a coupling of the atomic electron to the electric potential resulting from the polarization associated with an LO phonon mode \cite{sohier2016two}.
+This should be equivalent to an approach that considers the coupling of a bound electron in an emitter to LO phonons in the disk through a \swap{(two-dimensional)}{2D} Fr\"ohlich coupling---\ie a coupling of the atomic electron to the electric potential resulting from the polarization associated with an LO phonon mode \cite{sohier2016two}.
 
@@
 This coupling rate thus implies coupling between an emitter and the field in the regime of ultra-strong coupling.
-Moreover, the coupling rate for the \SI{20}{\nm} disk (orange), for an emitter with a far field decay rate of ${>}\,\SI{3e14}{\per\s}$ would have a sufficient coupling strength to the resonances in orange for its linewidth to span multiple resonances and thus be in a multi-mode ultra-strong coupling regime.
-Thus, the extreme confinement of electromagnetic energy associated with LO phonons in two dimensions enables the possibility of realizing ultrastrong coupling of an atom or molecule with optical phonons in a polar material, allowing the potential realization of new coupled states of quantum emitters and phonons such as atom-phonon polariton bound states.
+Moreover, the coupling rate for the \SI{20}{\nm} disk (purple), for an emitter with a far field decay rate of ${\gtrsim}\,\SI{3e14}{\per\s}$ would have a sufficient coupling strength to the \inset{distinct, radially-quantized} resonances\swap{ in purple} for its linewidth to span multiple resonances and thus be in a multi-mode ultra-strong coupling regime.
+Thus, the extreme confinement of electromagnetic energy associated with LO phonons in two dimensions enables the possibility of realizing ultra-strong coupling of an atom or molecule with optical phonons in a polar material, allowing the potential realization of new coupled states of quantum emitters and phonons such as atom--phonon polariton bound states.
 
@@
 The highly confined optical phonon waves in polar monolayers may also provide interesting new opportunities in near-field radiative heat transfer, in which it has been long known that thin-film surface phonon-polaritons play a critical role.
-An important avenue of future study would be the \textit{ab initio} calculation of lifetimes of 2D LO phonons associated with three-phonon processes and electron-phonon interactions.
+An important avenue of future study would be the \emph{ab initio} calculation of lifetimes of 2D LO phonons associated with three-phonon processes and electron-phonon interactions.
 In further work, it would be of great interest to study the effects of isotopic purification and cryogenic temperatures on reducing the decay rate of these 2D LO phonons.
@@
 \section{Acknowledgments}
-N.
-R.
-recognizes the support of the DOE Computational Science Graduate Fellowship (CSGF) fellowship no.
-DE-FG02-97ER25308.
-T.
-C.
-acknowledges support from the Danish Council for Independent Research (Grant No.\ DFF--6108-00667).
 The authors thank Ido Kaminer, Siyuan Dai, Samuel Moore, Jennifer Coulter, and Christopher Ciccarino for helpful discussions.
+N.\,{}R. recognizes the support of the DOE Computational Science Graduate Fellowship (CSGF) fellowship no.\ DE-FG02-97ER25308. 
+T.\,{}C. acknowledges support from the Danish Council for Independent Research (Grant No.\ DFF--6108-00667). 
 This work was supported by the DOE Photonics at Thermodynamic Limits Energy Frontier Research Center under grant number DE-SC0019140.

### 2019-01-09 paper.tex
minor text things

@@
 
-In Figure ~\ref{fig:2}(a), we present the conductivity of 2D hBN of Equation~\ref{eq:localsigma}, using parameters from Ref.~$\citenum{sohier2017breakdown}$ calculated from density functional theory within the local density approximation.
+In Figure ~\ref{fig:2}(a), we present the conductivity of 2D hBN of Equation~\ref{eq:localsigma}, using parameters from Ref.~\citenum{sohier2017breakdown} calculated from density functional theory within the local density approximation.
+\comment{tc}{It is unfortunate that we never state the actual values used for $v_{\mathrm{g}}$ and $\omega_{\mathrm{TO},\mathrm{LO}}(\mathbf{q}=0)$: without them, one cannot reproduce the figures (as I just realized in trying to do so).}
 From the conductivity, the dispersion relation of phonon polaritons on an infinite sheet is given by $q={2\iu\omega\epsilon_0\epsilon_{\mathrm{env}}}/{\sigma(\omega)}$.
@@
 \begin{figure}[t]
-    \includegraphics[width=0.95\linewidth]{fig3.pdf}
+    \includegraphics[width=1\linewidth]{fig3.pdf}
     \caption{%
         \textbf{Extreme spontaneous emission enhancement due to 2D LO phonons in nanostructured geometries.} Plotted is the enhancement of the spontaneous emission rate for an emitter.
-The emitter is concentric with the disk and is at a distance $z=\SI{5}{\nm}$ away from the disk.
-For a disk with a diameter of \SI{20}{\nm}, and assuming a relaxation rate $\tau^{-1} = \SI{5}{\per\cm}$, the rate of emission enhancement can be enhanced \num{100} million-fold.
+The emitter is \swap{concentric with the disk and is at a distance $z=\SI{5}{\nm}$ away from the disk}{$z=\SI{5}{\nm}$ above the disk's center and is oriented normal to its plane ($\hat{\mathbf{n}} = \hat{\mathbf{z}}$)}.
+For a disk with a diameter of \SI{20}{\nm}, and assuming a relaxation rate $\tau^{-1} = \SI{0.5}{\meV}$, the rate of emission enhancement can be enhanced \num{100} million-fold.
 For an emitter with a free-space decay rate of $\SI{1e6}{\per\cm}$ at \SI{7}{\micro\m}, the emitter would experience a decay rate comparable to the frequency of the disk mode, leading to ultra-strong coupling of an external emitter with 2D LO phonons.
-Note that $\tau^{-1} = \SI{5}{\per\cm}$ is of the order of the loss rate in bulk hBN.
+Note that $\tau^{-1} = \SI{0.5}{\meV}$ is of the order of the loss rate in bulk hBN.
 \comment{tc}{@NR: Citation needed.}
+        \comment{tc}{@NR: As we discussed, it would be very pertinent to add the infinite-sheet Purcell factor here as well (at $z=\SI{5}{\nm}$)---that is a very useful reference in assessing the impact of nanostructuring.}
         \label{fig:3}
@@
 For simplicity, we consider hBN nanostructured as a disk, which leads to the formation of sharp resonances quantized along the azimuthal and radial directions.
-The disk is taken to have a radius of \SI{20}{cm} and a loss rate $\tau^{-1} = \SI{5}{\per\cm}$, which is of the order of the loss rate in bulk hBN.
+The disk is taken to have a radius of \SI{20}{\nm}
+\comment{tc}{@NR: We have four different radii ($\SIlist{10;20;30;40}{\nm}$), not just one}
+and a loss rate $\tau^{-1} = \SI{0.5}{\meV}$, which is of the order of the loss rate in bulk hBN. 
 \comment{tc}{@NR: Citation needed.}
@@
 In Figure~\ref{fig:3}, we show that for an infrared emitter at a transition wavelength of \SI{7}{\micro\m} with a free-space radiative lifetime of \SI{1}{\micro\s}, \SI{5}{\nm} away from an hBN disk, the coupling rate to 2D optical phonons (about $\SI{1e14}{\per\s}$ or \SI{65}{\meV}) would be on the same scale as the optical phonon frequency itself (about \SI{1430}{\per\cm} or $\SI{2.70e14}{\per\s}$ or \SI{0.18}{\eV}).
-\comment{tc}{We are using three distinct units for energy, \si{\per\s}, \si{\per\cm}, and \si{\eV}, in various places.
-At most, we should probably use two units (then, I would prefer to get rid of \si{\per\s}.
-Even if we don't do that, we probably don't need to give it in all three units here.
+\comment{tc}{We are using four (!) distinct units for energy, \si{\per\s}, \si{\per\cm}, \si{\micro\metre}, and \si{\eV}, in various places.
+At most, we should probably use two units (then, I would prefer to get rid of \si{\per\s} and \si{\micro\metre}.
+Even if we don't do that, we probably don't need to give it in three units here.
 When writing unitful quantities, please use the \texttt{\textbackslash{}SI\{value\}\{unit\}} macro (\eg \texttt{\textbackslash{}SI\{2\}\{\textbackslash{}eV\}} for \SI{2}{\eV} or \texttt{\textbackslash{}SI\{7\}\{\textbackslash{}per\textbackslash{}cm\}} for \SI{7}{\per\cm}).}

### 2019-01-10 paper.tex
Few more edits and comments; feel free to take it from here Nick

@@
 \begin{abstract}
-    Extreme confinement of electromagnetic energy by phonon-polaritons holds the promise of strong and new forms of control over the dynamics of matter. 
-    To bring such control to the atomic-scale limit, it is important to consider phonon-polaritons in two dimensional systems.
+    Extreme confinement of electromagnetic energy by phonon polaritons holds the promise of strong and new forms of control over the dynamics of matter. 
+    To bring such control to the atomic-scale limit, it is important to consider phonon polaritons in two dimensional systems.
 Recent studies have pointed out that in two-dimensional systems, splitting between longitudinal and transverse optical (LO and TO) phonons is absent at the $\Gamma$ point.
-Does this lack of LO--TO splitting imply the absence of a phonon-polariton in polar monolayers?
+    Does this lack of LO--TO splitting imply the absence of a phonon polariton in polar monolayers? 
     Here, we derive a first-principles expression for the spatially nonlocal conductivity of a polar monolayer specified by the wavevector-dependent LO and TO phonon dispersions.
@@
     its lifetime, and the group velocity of the LO phonon.
-Our analysis reveals that the phonon-polariton of the bulk is essentially replaced by the LO phonon of the two-dimensional system.
+Our analysis reveals that the phonon polariton of the bulk is essentially replaced by the LO phonon of the two-dimensional system.
 We calculate the confinement and propagation losses of the LO phonons, finding that high confinement and reasonable propagation quality factors coincide in regions which are difficult to detect with current near-field optical microscopy techniques.
@@
 Phonon-polaritons, hybrid quasiparticles of photons and optical phonons supported in polar materials, hold tremendous promise for deeply sub-diffraction control of electromagnetic fields at mid-infrared and terahertz frequencies.
-Qualitatively, phonon-polaritons share many features with plasmon-polaritons in conductors.
-Recently, it has been shown that phonon-polaritons enable confinement of light to volumes \remove{over} $\inset{\sim}10^6$ times smaller than that of a diffraction-limited photon in free-space~\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra,li2018infrared,ma2018plane}.
-Due to this remarkable confinement and their relatively high lifetimes---around picoseconds---phonon-polaritons open new opportunities for vibrational spectroscopy~\cite{autore2018boron}, radiative heat transfer~\cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters~\cite{kumar2015tunable,rivera2017making,kurman2018control}.
-\swap{In many cases}{Thus far}, extreme confinement of phonon-polaritons \swap{is}{has been} achieved by the use of thin-films, which shrink the in- and out-of-plane wavelength of polaritons monotonically with decreasing film thickness~\cite{dai2014tunable,dubrovkin2018ultra}.
+Qualitatively, phonon polaritons share many features with plasmon polaritons in conductors.
+Recently, it has been shown that phonon polaritons enable confinement of light to volumes \remove{over} $\inset{\sim}10^6$ times smaller than that of a diffraction-limited photon in free-space~\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra,li2018infrared,ma2018plane}.
+Due to this remarkable confinement and their relatively high lifetimes---around picoseconds---phonon polaritons open new opportunities for vibrational spectroscopy~\cite{autore2018boron}, radiative heat transfer~\cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters~\cite{kumar2015tunable,rivera2017making,kurman2018control}.
+\swap{In many cases}{Thus far}, extreme confinement of phonon polaritons \swap{is}{has been} achieved by the use of thin-films, which shrink the in- and out-of-plane wavelength of polaritons monotonically with decreasing film thickness~\cite{dai2014tunable,dubrovkin2018ultra}.
 A monolayer is the ultimate, atomic-scale limit of this effect\remove{ when the properties of the bulk material can be used in the monolayer limit}.
-\swap{In polar materials, the transition from 3D to 2D is not so simple}{The response-transition from three-dimensional (3D) to two-dimensional (2D) polar materials is nontrivial, however}, since in a polar monolayer, the LO--TO splitting that gives rise to phonon-polaritons in \swap{three-dimensions}{3D} is absent at the $\Gamma$ point~\cite{sanchez2002vibrational,mele2002electric,serrano2007vibrational,sohier2017breakdown}.
+\swap{In polar materials, the transition from 3D to 2D is not so simple}{The response-transition from three-dimensional (3D) to two-dimensional (2D) polar materials is nontrivial, however}, since in a polar monolayer, the LO--TO splitting that gives rise to phonon polaritons in \swap{three-dimensions}{3D} is absent at the $\Gamma$ point~\cite{sanchez2002vibrational,mele2002electric,serrano2007vibrational,sohier2017breakdown}.
 This raises a fundamental question about whether strongly confined electromagnetic modes exist in polar monolayers at all.
@@
 We do so by deriving a universal form for the conductivity of a polar monolayer,  which depends solely on the LO and TO phonon frequencies---and their dispersion with momentum---in the \swap{two-dimensional}{2D} system.
-Using parameters from Ref.~\citenum{sohier2017breakdown} for the canonical \swap{two-dimensional}{2D} polar monolayer---hexagonal boron nitride (hBN)---we present the confinement and propagation losses of the 2D \swap{phononic}{phonon-polariton }modes, identifying the frequency region where they should be most easily detected.
+Using parameters from Ref.~\citenum{sohier2017breakdown} for the canonical \swap{two-dimensional}{2D} polar monolayer---hexagonal boron nitride (hBN)---we present the confinement and propagation losses of the 2D \swap{phononic}{phonon polariton }modes, identifying the frequency region where they should be most easily detected.
 Finally, we \swap{show}{find} that these modes enable extreme light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced by up to eight orders of magnitude \swap{by}{through/via} the emitter--LO phonon coupling.
@@
 
-Before moving on to analyze the electrodynamics of 2D phonon-polaritons, we make two comments on lack of LO--TO splitting in 2D polar materials.
+Before moving on to analyze the electrodynamics of 2D phonon polaritons, we make two comments on lack of LO--TO splitting in 2D polar materials.
 The first is that this situation is in stark contrast to the situation of polar materials in 3D, which have a finite LO--TO splitting at the $\Gamma$ point.
@@
 Combining the two conditions, we obtain the usual dispersion equation for the TM mode of a polarizable 2D monolayer, namely $\epsilon_+/\kappa_+ + \epsilon_-/\kappa_- +\epsilon_0\Pi=0$.
-Given the deeply subwavelength nature of 2D phonon-polaritons, \ie since $q\gg\omega/c$ such that $\kappa_{\pm}\simeq q$, the dispersion equation can be reduced to its quasistatic limit without consequential loss of accuracy:
+Given the deeply subwavelength nature of 2D phonon polaritons, \ie since $q\gg\omega/c$ such that $\kappa_{\pm}\simeq q$, the dispersion equation can be reduced to its quasistatic limit without consequential loss of accuracy:
 \begin{equation}
@@
 with $\epsilon_{\mathrm{env}}\equiv (\epsilon_++\epsilon_-)/2$.
-Thus, the condition for `phonon-polaritons' \comment{tc}{Remove quotes} coincides precisely with the condition for longitudinal optical phonons.
-Thus, one of the main results of our manuscript is that despite the lack of LO--TO splitting at the $\Gamma$ point, there nevertheless exists a strongly confined evanescent mode in many respects similar to the phonon-polaritons of thin films.
+Thus, the condition for `phonon polaritons' \comment{tc}{Remove quotes} coincides precisely with the condition for longitudinal optical phonons.
+Thus, one of the main results of our manuscript is that despite the lack of LO--TO splitting at the $\Gamma$ point, there nevertheless exists a strongly confined evanescent mode in many respects similar to the phonon polaritons of thin films.
 We now analyze the dispersion relation of phonon polaritons in a specific material, hexagonal boron nitride, in more detail, showing the possibility of highly confined electromagnetic modes with a large local density of states.
@@
 In 3D, there is a finite LO--TO splitting at zero wavevector, while in 2D there is none.
-Despite this, the 2D LO phonon plays the role of the phonon-polariton in 3D and thin films.
+Despite this, the 2D LO phonon plays the role of the phonon polariton in 3D and thin films.
 (c) Analogous physics appears in electron gases in 3D and 2D, where the 3D plasma frequency is similar to the 3D LO--TO splitting.
@@
 For monolayer hBN, the wavevector grows very rapidly with frequency, due to the extremely low group velocity of 2D LO phonons, which is a remarkable four orders of magnitude slower than the speed of light.
-In particular, at frequencies of \SI{1450}{\per\cm}, the phonon-polariton has a wavelength of about \SI{15}{\nm}, significantly shorter than any phonon-polariton measured so far, and, similarly shorter than any plasmonic wavelength, even in graphene.
+In particular, at frequencies of \SI{1450}{\per\cm}, the phonon polariton has a wavelength of about \SI{15}{\nm}, significantly shorter than any phonon polariton measured so far, and, similarly shorter than any plasmonic wavelength, even in graphene.
 In fact, this short a wavelength is essentially beyond the range of measurability of current scattering near-field microscopy (SNOM) setups, which are limited (by tip radii) to wavelengths greater than ${\approx}\,\SI{60}{\nm}$.
-The 2D LO phonon-polariton could in principle be measured by SNOM closer to the TO frequency, where confinement is smaller; unfortunately, as shown in Figure~\ref{fig:2}(b), near the TO frequency, dissipation is far higher (and corresponding propagation quality factors $\Re q/\Im q$ far lower) due to large $\Re\sigma$ (or, equivalently, large $\Im\epsilon_{\mathrm{RPA}}$).
+The 2D LO phonon polariton could in principle be measured by SNOM closer to the TO frequency, where confinement is smaller; unfortunately, as shown in Figure~\ref{fig:2}(b), near the TO frequency, dissipation is far higher (and corresponding propagation quality factors $\Re q/\Im q$ far lower) due to large $\Re\sigma$ (or, equivalently, large $\Im\epsilon_{\mathrm{RPA}}$).
 
-These considerations imply that access to the lower-loss and higher-confined portions of the dispersion relation of phonon-polaritons, in the absence of a sharper tip, requires a near-field probe \swap{that probes larger momenta.
+These considerations imply that access to the lower-loss and higher-confined portions of the dispersion relation of phonon polaritons, in the absence of a sharper tip, requires a near-field probe \swap{that probes larger momenta.
 Free electrons exemplify such high-momentum probes}{such as a free electron probe,} as used in electron energy loss spectroscopy (EELS), where slow electrons can be used to probe plasmon wavelengths of just a few nanometers in monolayer metals, as well as the nonlocal bulk plasmon dispersion in metals~\cite{nagao2001dispersion,de2010optical,diaconescu2007low}.
-EELS has been recently employed to measure phonon-polaritons in ultrathin films of hBN~\cite{govyadinov2017probing}.
+EELS has been recently employed to measure phonon polaritons in ultrathin films of hBN~\cite{govyadinov2017probing}.
 Another interesting class of near-field probes, with relevance to fundamental physics and quantum optics applications, is a quantum emitter such as an atom, molecule, or quantum well.
-Recently, it was demonstrated using nanostructures of bulk hBN that the interaction of vibrational emitters with phonon-polaritons is on the border of the strong coupling regime~\cite{autore2018boron}.
+Recently, it was demonstrated using nanostructures of bulk hBN that the interaction of vibrational emitters with phonon polaritons is on the border of the strong coupling regime~\cite{autore2018boron}.
 
@@
 \begin{figure}[t]
-    \includegraphics[width=1\linewidth]{fig3.pdf}
+    \includegraphics[width=.925\linewidth]{fig3.pdf}
     \caption{%
@@
 \comment{tc}{@NR: Citation needed.}
-        \comment{tc}{@NR: As we discussed, it would be very pertinent to add the infinite-sheet Purcell factor here as well (at $z=\SI{5}{\nm}$)---that is a very useful reference in assessing the impact of nanostructuring.}
+        \comment{tc}{@NR: Caption should be updated to reflect addition to figure; 1. potential maps (in the $z=0$ plane) and 2. sheet limit.}
         \label{fig:3}
@@
 
-In Figure~\ref{fig:3}, we consider the coupling of a dipole emitter to localized phonon-polaritons of nanostructured monolayer hBN.
+In Figure~\ref{fig:3}, we consider the coupling of a dipole emitter to localized phonon polaritons of nanostructured monolayer hBN.
 For simplicity, we consider hBN nanostructured as a disk, which leads to the formation of sharp resonances quantized along the azimuthal and radial directions.
@@
 \comment{tc}{@NR: Citation needed.}
-We parameterize the coupling between the dipole and phonon-polaritons through the rate of spontaneous emission $\Gamma$ of phonon polaritons by the dipole.
+We parameterize the coupling between the dipole and phonon polaritons through the rate of spontaneous emission $\Gamma$ of phonon polaritons by the dipole.
 It is related to the dyadic Green function $\mathbf{G}$ of the Maxwell equations for the nanostructure via the expression \cite{novotny2012principles}:
@@
 \end{equation}
+\comment{tc}{@NR: $\Gamma_0$ has not been defined.}
 where $\hat{\mathbf{n}}$ is the orientation of the dipole, $\mathbf{r}_0$ is its position, and $\omega_0$ is its frequency.
 \comment{tc}{@NR: It could probably be fine to just drop the zero-subscript in $\mathbf{r}$ and $\omega$.}
-The dyadic Green function is computed through a boundary element method (as in Ref.~\citenum{jin2017infrared}) assuming the validity of the electrostatic approximation, which is well-respected due to the extreme sub-wavelength confinement of the electromagnetic field of the phonon polariton.
-In Figure~\ref{fig:3}, we plot the enhancement of the spontaneous emission rate for an external emitter polarized perpendicularly to the plane of the disk and placed \SI{5}{\nm} away from the surface of the disk.
+The dyadic Green function is computed using a \inset{quasistatic} boundary element method (as in Ref.~\citenum{jin2017infrared})\remove{, within the electrostatic approximation, which is well-respected due to the extreme sub-wavelength confinement of the electromagnetic field of the phonon polariton}.
+In Figure~\ref{fig:3}, we plot the enhancement of the spontaneous emission rate $\Gamma/\Gamma_0$ for an external emitter polarized perpendicularly to the plane of the disk and placed \SI{5}{\nm} away from the surface of the disk.
+\comment{tc}{Maybe good to add that these choices lead to coupling with axisymmetric $l=0$ modes only.}
 We find that the rate of spontaneous emission of 2D optical phonons is approximately 8 orders of magnitude larger than the rate of spontaneous emission of photons in the far field.
 We note that in this approach, the coupling of the dipole to phonon polaritons is manifested through the phonon contribution to the dielectric function of the disk.
-This should be equivalent to an approach that considers the coupling of a bound electron in an emitter to LO phonons in the disk through a \swap{(two-dimensional)}{2D} Fr\"ohlich coupling---\ie a coupling of the atomic electron to the electric potential resulting from the polarization associated with an LO phonon mode \cite{sohier2016two}.
+This should be equivalent to an approach that considers the coupling of a bound electron in an emitter to LO phonons in the disk through a \swap{(two-dimensional)}{2D} Fr\"ohlich coupling---\ie a coupling of the atomic electron to the electric potential resulting from the polarization associated with an LO phonon mode~\cite{sohier2016two}.
 
-In Figure~\ref{fig:3}, we show that for an infrared emitter at a transition wavelength of \SI{7}{\micro\m} with a free-space radiative lifetime of \SI{1}{\micro\s}, \SI{5}{\nm} away from an hBN disk, the coupling rate to 2D optical phonons (about $\SI{1e14}{\per\s}$ or \SI{65}{\meV}) would be on the same scale as the optical phonon frequency itself (about \SI{1430}{\per\cm} or $\SI{2.70e14}{\per\s}$ or \SI{0.18}{\eV}).
+In Figure~\ref{fig:3}, we show that for an infrared emitter at a transition wavelength of \SI{7}{\micro\m} with a free-space radiative lifetime of \SI{1}{\micro\s}, \SI{5}{\nm} away from an hBN disk, the coupling rate to 2D optical phonons (about \SI{1e14}{\per\s} or \SI{65}{\meV}) would be on the same scale as the optical phonon frequency itself (about \SI{1430}{\per\cm} or \SI{2.70e14}{\per\s} or \SI{0.18}{\eV}).
 \comment{tc}{We are using four (!) distinct units for energy, \si{\per\s}, \si{\per\cm}, \si{\micro\metre}, and \si{\eV}, in various places.
@@
 When writing unitful quantities, please use the \texttt{\textbackslash{}SI\{value\}\{unit\}} macro (\eg \texttt{\textbackslash{}SI\{2\}\{\textbackslash{}eV\}} for \SI{2}{\eV} or \texttt{\textbackslash{}SI\{7\}\{\textbackslash{}per\textbackslash{}cm\}} for \SI{7}{\per\cm}).}
-This coupling rate thus implies coupling between an emitter and the field in the regime of ultra-strong coupling.
+This rate thus implies coupling between an emitter and the field in the regime of ultra-strong coupling.
 Moreover, the coupling rate for the \SI{20}{\nm} disk (purple), for an emitter with a far field decay rate of ${\gtrsim}\,\SI{3e14}{\per\s}$ would have a sufficient coupling strength to the \inset{distinct, radially-quantized} resonances\swap{ in purple} for its linewidth to span multiple resonances and thus be in a multi-mode ultra-strong coupling regime.
@@
 The ability to probe low-loss and highly confined electromagnetic modes associated with 2D LO phonons in polar materials provides a new platform for nanophotonics in the mid- and far-infrared spectral range.
-The identification of the phonon-polariton of bulk and thin-film geometries with the 2D LO phonon made in this Letter would extend the rich phenomenology of optical phonons to nanophotonic applications.
+The identification of the phonon polariton of bulk and thin-film geometries with the 2D LO phonon made in this Letter would extend the rich phenomenology of optical phonons to nanophotonic applications.
 This work also points the way to useful new approaches to study LO phonons, arising from the fact that 2D LO phonons, unlike their 3D counterparts, have their electromagnetic energy extend a considerable distance from the material boundary.
 Due to the strong electromagnetic interactions between emitters and 2D LO phonons shown in this Letter, it is now possible to design interesting new hybrid states of matter and phonons based on quantum electrodynamical strong coupling.
-The highly confined optical phonon waves in polar monolayers may also provide interesting new opportunities in near-field radiative heat transfer, in which it has been long known that thin-film surface phonon-polaritons play a critical role.
+The highly confined optical phonon waves in polar monolayers may also provide interesting new opportunities in near-field radiative heat transfer, in which it has been long known that thin-film surface phonon polaritons play a critical role.
 An important avenue of future study would be the \emph{ab initio} calculation of lifetimes of 2D LO phonons associated with three-phonon processes and electron-phonon interactions.

### 2019-01-12 paper.tex
all comments addressed, doing one final spot check

@@
     Here, we derive a first-principles expression for the spatially nonlocal conductivity of a polar monolayer specified by the wavevector-dependent LO and TO phonon dispersions.
-In the long-wavelength (local) limit, we find a universal form for the conductivity in terms of the TO phonon frequency at the $\Gamma$ point,
-    \comment{tc}{Why not just say `the LO phonon frequency' rather than `TO' here, since they are degenerate anyway---that way, we would only have to reference LO quantities for the local conductivity (and that case, `and the group velocity of the LO phonon' would be swapped for `and its group velocity').}
+In the long-wavelength (local) limit, we find a universal form for the conductivity in terms of the LO phonon frequency at the $\Gamma$ point,
+   % \comment{tc}{Why not just say `the LO phonon frequency' rather than `TO' here, since they are degenerate anyway---that way, we would only have to reference LO quantities for the local conductivity (and that case, `and the group velocity of the LO phonon' would be swapped for `and its group velocity').}
     its lifetime, and the group velocity of the LO phonon.
 Our analysis reveals that the phonon polariton of the bulk is essentially replaced by the LO phonon of the two-dimensional system.
-We calculate the confinement and propagation losses of the LO phonons, finding that high confinement and reasonable propagation quality factors coincide in regions which are difficult to detect with current near-field optical microscopy techniques.
+For the specific example of hexagonal boron nitride, we estimate the confinement and propagation losses of the LO phonons, finding that high confinement and reasonable propagation quality factors coincide in regions which are difficult to detect with current near-field optical microscopy techniques.
     Finally, we study the interaction of external emitters with two-dimensional hBN nanostructures, 
-    \comment{tc}{This is the first mention of hBN in the abstract; arguably, it should occur earlier, already at the `We calculate the confinement and propagation losses \ldots' part}
-    finding extreme enhancement of spontaneous emission due to coupling with localized 2D \comment{tc}{`2D' acronym not yet defined.} LO phonons, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN phonons.
-    \comment{tc}{I would suggest to lose this last sentence of the abstract: it doesn't add anything very concrete and it just (undisguised) public-relations work.}
-    Using the strong electromagnetic interactions between emitters and 2D LO phonons calculated in this work, it is now possible to design new hybrid states of light, electrons and phonons based on quantum electrodynamical strong coupling.
+  %  \comment{tc}{This is the first mention of hBN in the abstract; arguably, it should occur earlier, already at the `We calculate the confinement and propagation losses \ldots' part}
+    finding extreme enhancement of spontaneous emission due to coupling with localized two-dimensional (2D) LO phonons, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN phonons.
+This may lead to the design of new hybrid states of light, electrons and phonons based on strong coupling.
+   % \comment{tc}{I would suggest to lose this last sentence of the abstract: it doesn't add anything very concrete and it just (undisguised) public-relations work.}
+%    Using the strong electromagnetic interactions between emitters and 2D LO phonons calculated in this work, it is now possible to design new hybrid states of light, electrons and phonons based on quantum electrodynamical strong coupling.
 \end{abstract}
@@
 
-\comment{tc}{@NR: There's a somewhat inconsistent use of quotations: \ie `X' vs.\ ``X'' throughout the paper.
+%\comment{tc}{@NR: There's a somewhat inconsistent use of quotations: \ie `X' vs.\ ``X'' throughout the paper.
 In most cases, double quotations are probably correct.
@@
 Qualitatively, phonon polaritons share many features with plasmon polaritons in conductors.
-Recently, it has been shown that phonon polaritons enable confinement of light to volumes \remove{over} $\inset{\sim}10^6$ times smaller than that of a diffraction-limited photon in free-space~\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra,li2018infrared,ma2018plane}.
-Due to this remarkable confinement and their relatively high lifetimes---around picoseconds---phonon polaritons open new opportunities for vibrational spectroscopy~\cite{autore2018boron}, radiative heat transfer~\cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters~\cite{kumar2015tunable,rivera2017making,kurman2018control}.
-\swap{In many cases}{Thus far}, extreme confinement of phonon polaritons \swap{is}{has been} achieved by the use of thin-films, which shrink the in- and out-of-plane wavelength of polaritons monotonically with decreasing film thickness~\cite{dai2014tunable,dubrovkin2018ultra}.
-A monolayer is the ultimate, atomic-scale limit of this effect\remove{ when the properties of the bulk material can be used in the monolayer limit}.
-\swap{In polar materials, the transition from 3D to 2D is not so simple}{The response-transition from three-dimensional (3D) to two-dimensional (2D) polar materials is nontrivial, however}, since in a polar monolayer, the LO--TO splitting that gives rise to phonon polaritons in \swap{three-dimensions}{3D} is absent at the $\Gamma$ point~\cite{sanchez2002vibrational,mele2002electric,serrano2007vibrational,sohier2017breakdown}.
+Recently, it has been shown that phonon polaritons enable confinement of light to volumes  $\sim10^6$ times smaller than that of a diffraction-limited photon in free-space~\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra,li2018infrared,ma2018plane}.
+Due to this remarkable confinement and their relatively high lifetimes---around picoseconds---phonon polaritons open new opportunities for vibrational spectroscopy~\cite{autore2018boron}, radiative heat transfer~\cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters~\cite{kumar2015tunable,rivera2017making,kurman2018control,flick2018strong}.
+Thus far, extreme confinement of phonon polaritons has been achieved by the use of thin-films, which shrink the in- and out-of-plane wavelength of polaritons monotonically with decreasing film thickness~\cite{dai2014tunable,dubrovkin2018ultra}.
+A monolayer is the ultimatelimit of this effect.
+The response-transition from three-dimensional (3D) to two-dimensional (2D) polar materials is nontrivial, however, since in a polar monolayer, the LO--TO splitting that gives rise to phonon polaritons in 3D is absent at the $\Gamma$ point~\cite{sanchez2002vibrational,mele2002electric,serrano2007vibrational,sohier2017breakdown}.
 This raises a fundamental question about whether strongly confined electromagnetic modes exist in polar monolayers at all.
-\comment{tc}{@NR: Check that 3D \& 2D acronyms are only defined once (above, in my suggestion).}
 
-In this Letter, we develop a theoretical and computational framework with quantitative, concrete examples that \swap{enable}{enables} a \swap{new}{first-principles} understanding of \swap{optical phononics}{phonon polaritons} in \swap{two-dimensional}{2D} materials.
-\comment{tc}{I don't think `optical phononics' is a generally accepted identifier for phonon polaritons.
+We now develop a theoretical and computational framework with quantitative, concrete examples that enables a first-principles understanding of phonon polaritons in 2D materials. 
+%\comment{tc}{I don't think `optical phononics' is a generally accepted identifier for phonon polaritons.
 We use the two somewhat interchangably in this manuscript: maybe settle on just one?
 (Only a single use of `phononics' (this one)---we have three uses of `phononic' elsewhere though: I think only one should be changed.)}
-We do so by deriving a universal form for the conductivity of a polar monolayer,  which depends solely on the LO and TO phonon frequencies---and their dispersion with momentum---in the \swap{two-dimensional}{2D} system.
-Using parameters from Ref.~\citenum{sohier2017breakdown} for the canonical \swap{two-dimensional}{2D} polar monolayer---hexagonal boron nitride (hBN)---we present the confinement and propagation losses of the 2D \swap{phononic}{phonon polariton }modes, identifying the frequency region where they should be most easily detected.
-Finally, we \swap{show}{find} that these modes enable extreme light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced by up to eight orders of magnitude \swap{by}{through/via} the emitter--LO phonon coupling.
-For an infrared emitter with a sufficiently high free-space radiative decay rate (${\gtrsim}\, \SI{1e6}{\per\s}$), \inset{we find that} the \swap{enhanced}{associated} linewidth of the emitter \swap{becomes}{is} comparable to the spacing between different phonon polaritonic resonances of an hBN nanostructure.
-This suggests the possibility of realizing the multi-mode strong coupling and ultra-strong coupling regimes of quantum electrodynamics in a \swap{two-dimensional}{2D} hBN platform.
+We do so by deriving a universal form for the conductivity of a polar monolayer,  which depends solely on the LO and TO phonon frequencies---and their dispersion with momentum---in the 2D system.
+Using parameters from Ref.~\citenum{sohier2017breakdown} for the canonical 2D polar monolayer---hexagonal boron nitride (hBN)---we present the confinement and propagation losses of the 2D phonon polariton modes, identifying the frequency region where they should be most easily detected.
+Finally, we find that these modes enable extreme light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced by up to eight orders of magnitude through the emitter--LO phonon coupling.
+For an infrared emitter with a sufficiently high free-space radiative decay rate (${\gtrsim}\, \SI{1e6}{\per\s}$), we find that the associated linewidth of the emitter is comparable to the spacing between different phonon polaritonic resonances of an hBN nanostructure.
+This suggests the possibility of realizing the multi-mode strong coupling and ultra-strong coupling regimes of quantum electrodynamics in a 2D hBN platform.
 Our results for hBN are particularly relevant due its widespread use in 2D van der Waals heterostructures.
-In addition to providing functionality as a layer which improves the electrical and optical properties of other 2D materials, \swap{such as}{\eg} graphene, our results suggest that in these heterostructures, \remove{the }hBN layers could provide a \inset{mid-infrared} platform for nanophotonics and quantum optics\remove{ which operates in the mid-infrared spectral range}.
-While we focus on hBN in this manuscript, \swap{many of}{the salient features of} our findings apply to other polar monolayers\inset{ as well}.
+In addition to providing functionality as a layer which improves the electrical and optical properties of other 2D materials, \eg graphene, our results suggest that in these heterostructures, hBN layers could provide a mid-infrared platform for nanophotonics and quantum optics.
+While we focus on hBN in this manuscript, the salient features of our findings apply to other polar monolayers as well.
 
 \section{Optical response of optical phonons in two-dimensions}
-In this section, we develop a theory of electromagnetic response due to optical phonons in \swap{two-dimensional}{2D} systems.
-The key \swap{prediction}{response function of interest} is the conductivity of the monolayer.
-To \swap{obtain it}{that end}, we consider the response of the ions of the monolayer due to a longitudinal electromagnetic potential $\phi$.
+In this section, we develop a theory of electromagnetic response due to optical phonons in 2D systems.
+The key response function of interest is the conductivity of the monolayer.
+To that end, we consider the response of the ions of the monolayer due to a longitudinal electromagnetic potential $\phi$.
 For that case, the interaction Hamiltonian is
@@
 Within linear response theory, the induced polarization density can be straightforwardly evaluated from the displacement amplitude $\mathbf{u}_{\kappa}$ of every atom $\kappa$ within the unit cell.
-Specifically, to first order, the induced polarization is (in repeated-index notation):
-\comment{tc}{Later on, we write things like $\sum_{\kappa}\mathbf{Z}_{\kappa}\boldsymbol{\eta}_\kappa$, so we are not entirely consistent in our application of repeated-index notation.
+Specifically, to first order, the induced polarization is %(in repeated-index notation):
+%\comment{tc}{Later on, we write things like $\sum_{\kappa}\mathbf{Z}_{\kappa}\boldsymbol{\eta}_\kappa$, so we are not entirely consistent in our application of repeated-index notation.
 Perhaps we should just write vectorial, explicit summation notation here as well (\ie write $\mathbf{P} - \mathbf{P}^0 = \sum_\kappa(\boldsymbol{\nabla}_{\mathbf{u}_\kappa}\mathbf{P})\mathbf{u}_\kappa \equiv \frac{1}{\Omega}\sum_\kappa\mathbf{Z}_{\kappa}\mathbf{u}_{\kappa}$)}
 \begin{equation}
-    P_i - P_i^{0} \swap{=}{\simeq} \frac{\partial P_i}{\partial u_{\kappa,j}}u_{\kappa,j} \equiv \frac{1}{\Omega}Z_{\kappa,ij}u_{\kappa,j},
+%P_i - P_i^{0} \swap{=}{\simeq} \frac{\partial P_i}{\partial u_{\kappa,j}}u_{\kappa,j} \equiv \frac{1}{\Omega}Z_{\kappa,ij}u_{\kappa,j},
+\mathbf{P} - \mathbf{P}^0 = \sum_\kappa(\boldsymbol{\nabla}_{\mathbf{u}_\kappa}\mathbf{P})\mathbf{u}_\kappa \equiv \frac{1}{\Omega}\sum_\kappa\mathbf{Z}_{\kappa}\mathbf{u}_{\kappa},
 \end{equation} 
-where \swap{$Z_{\kappa,ij} \equiv \Omega\frac{\partial P_i}{\partial u_{\kappa,j}}$}{$\mathbf{Z}_\kappa \equiv \Omega\boldsymbol{\nabla}_{\mathbf{u}_\kappa}\mathbf{P}$} is the Born effective charge tensor of ion $\kappa$ and $\Omega$ is the unit cell area.
-\swap{$P_i^{0}$}{$\mathbf{P}^0$} is the equilibrium polarization in the absence of displacements, which is zero here.
+where $\mathbf{Z}_\kappa \equiv \Omega\boldsymbol{\nabla}_{\mathbf{u}_\kappa}\mathbf{P}$ is the Born effective charge tensor of ion $\kappa$ and $\Omega$ is the unit cell area.
+$\mathbf{P}^0$ is the equilibrium polarization in the absence of displacements, which is zero here.
 With this relation between polarization and ionic displacements, the interaction Hamiltonian in Equation~\eqref{eq:coupling} couples the scalar potential and the ionic displacements.
-We consider the response of the monolayer to a scalar potential of the form $\phi(\mathbf{r}) = \phi(\mathbf{q},\omega)\e^{\iu\mathbf{q}\cdot\mathbf{r}-\iu\omega t}$, where $\mathbf{q}$ is a \swap{two-dimensional}{2D} wavevector in the plane of the monolayer.
+We consider the response of the monolayer to a scalar potential of the form $\phi(\mathbf{r}) = \phi(\mathbf{q},\omega)\e^{\iu\mathbf{q}\cdot\mathbf{r}-\iu\omega t}$, where $\mathbf{q}$ is a 2D wavevector in the plane of the monolayer.
 Such a potential corresponds to a longitudinal electric field  $\mathbf{E}(\mathbf{r}) = \iu\mathbf{q}\phi(\mathbf{r})$.
@@
 \end{equation}
-where $m,n$ are states of the phononic Fock space of the monolayer, \swap{$P_{mn,i}(\mathbf{q}) \swap{=}{\equiv} \sum_{\kappa,j}Z_{\kappa,ij}\langle m | u_{\kappa,j}(\mathbf{q}) | n \rangle$}{$\mathbf{P}_{mn}(\mathbf{q}) \equiv \sum_\kappa \mathbf{Z}_\kappa\langle m|\mathbf{u}_\kappa(\mathbf{q})|n\rangle$} are matrix elements of the polarization associated with phonon modes (where \swap{$u_{\kappa,j}(\mathbf{q})$}{$\mathbf{u}_\kappa(\mathbf{q})$} is the Fourier transform of the phonon displacement operator), $E_{m}$ ($E_n$) is the energy of state $m$ ($n$), $\beta \equiv 1/k_{\text{\textsc{B}}}T$ is the inverse temperature, and $\mathcal{Z}$ is the grand partition function.
+where $m,n$ are states of the phononic Fock space of the monolayer, $\mathbf{P}_{mn}(\mathbf{q}) \equiv \sum_\kappa \mathbf{Z}_\kappa\langle m|\mathbf{u}_\kappa(\mathbf{q})|n\rangle$ are matrix elements of the polarization associated with phonon modes (where $\mathbf{u}_\kappa(\mathbf{q})$ is the Fourier transform of the phonon displacement operator), $E_{m}$ ($E_n$) is the energy of state $m$ ($n$), $\beta \equiv 1/k_{\text{\textsc{B}}}T$ is the inverse temperature, and $\mathcal{Z}$ is the grand partition function.
 We now evaluate the contribution of optical phonons to the polarization-polarization response in the low-temperature limit $T \ll \hbar\omega_{\mathrm{ph}}/k_{\mathrm{B}}$ with $\omega_{\mathrm{ph}}$ a characteristic optical phonon frequency.
-Taking a material with long-wavelength isotropy, such as hBN, we only have to consider the $qq$-component in the response tensor, where $qq$ denotes a pair of directions parallel to the wavevector.
+Considering the long-wavelength (small wavevector) limit, and taking a material with long-wavelength isotropy, such as hBN, we only have to consider the $qq$-component in the response tensor, where $qq$ denotes a pair of directions parallel to the wavevector.
 Denoting $\sigma_{qq}$ as simply $\sigma$, we find that the conductivity is given by \cite{rivera2018ab}:
@@
 \end{equation}
-In this expression, $\omega_\mathrm{L}$, the LO phonon frequency prior to LO--TO splitting, has been re-named as $\omega_{\mathrm{TO}}$, the transverse optical phonon frequency, because in the absence of LO--TO splitting, they are degenerate.
+%In this expression, $\omega_\mathrm{L}$, the LO phonon frequency prior to LO--TO splitting, has been re-named as $\omega_{\mathrm{TO}}$, the transverse optical phonon frequency, because in the absence of LO--TO splitting, they are degenerate.
 This is done in order to make the form of the final results more closely resemble their 3D counterparts.
-Additionally, we have defined scaled eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}\equiv \frac{\hat{\mathbf{e}}_{\kappa\mathbf{q},\mathrm{L}}}{\sqrt{M_{\kappa}}}$, where $\hat{\mathbf{e}}_{\kappa\mathbf{q},\mathrm{L}}$ is the unit-normalized polarization vector of atom $\kappa$ in the unit cell oscillating according to a longitudinal phonon of wavevector $\mathbf{q}$ and $M_{\kappa}$ is the mass of atom $\kappa$.
+We have defined scaled eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}\equiv \frac{\hat{\mathbf{e}}_{\kappa\mathbf{q},\mathrm{L}}}{\sqrt{M_{\kappa}}}$, where $\hat{\mathbf{e}}_{\kappa\mathbf{q},\mathrm{L}}$ is the unit-normalized polarization vector of atom $\kappa$ in the unit cell oscillating according to a longitudinal phonon of wavevector $\mathbf{q}$ and $M_{\kappa}$ is the mass of atom $\kappa$.
 
@@
 Here, $\epsilon_{\mathrm{env}}$ is the average permittivity of the bulk above and below the monolayer, and is added to take into account the polarization arising from these bulk materials.
+Note that we have neglected any intrinsic high-frequency screening in the monolayer itself, which is only relevant for wavevectors comparable to the inverse layer spacing between monolayers.
+When considering non-local corrections to the conductivity at these large wavevectors, these must be taken into account \cite{sohier2016two,thygesen2017calculating,sohier2017breakdown}.
 Based on Equation~\eqref{eq:conductivity_simplified}, the zeros $\omega_{\mathbf{q}}$ of the RPA dielectric function satisfy:
@@
 \end{equation}
-\comment{tc}{The tilde-quantity'ed $\tilde{\mathbf{Z}}_\kappa$ is not defined, nor is the $r$-quantity'ed $Z^r_{\kappa,ij}$ later on.
+%\comment{tc}{The tilde-quantity'ed $\tilde{\mathbf{Z}}_\kappa$ is not defined, nor is the $r$-quantity'ed $Z^r_{\kappa,ij}$ later on.
 I suspect they are the same thing, and that it is the $e\tilde{\mathbf{Z}}_\kappa \equiv \mathbf{Z}_\kappa$ from an earlier draft.
 @NR: Need to check and fix.}
-  where $V(q)$ is the Coulomb interaction in Fourier space, which in two dimensions, is given by $V(q) = \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}$.
+  where $e\tilde{\mathbf{Z}}_\kappa \equiv \mathbf{Z}_\kappa$ and $V(q)$ is the Coulomb interaction in Fourier space, which in two dimensions, is given by $V(q) = \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}$.
 Given that the zeros of the dielectric function are associated with longitudinal modes, one expects that $\omega_{\mathbf{q}}$ is in fact the frequency of the LO phonon mode.
-This is consistent with the result of Ref.~\citenum{sohier2017breakdown}, in which \relax{{}\ }\citeauthor{sohier2017breakdown} show that in \swap{two-dimensional}{2D} polar materials, the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO--TO splitting and zero LO--TO splitting at the $\Gamma$ point of the Brillouin zone.
+This is consistent with the result of Ref.~\citenum{sohier2017breakdown}, in which \relax{{}\ }\citeauthor{sohier2017breakdown} show that in 2D polar materials, the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO--TO splitting and zero LO--TO splitting at the $\Gamma$ point of the Brillouin zone.
 
@@
   \begin{equation}
-  \sigma(\mathbf{q},\omega) = \frac{-2\iu\epsilon_0\epsilon_{\mathrm{env}}}{q}\frac{\omega^2_{\mathbf{q},\mathrm{LO}}-\omega^2_{\mathrm{TO}}}{\omega^2_{\mathrm{TO}}-\omega^2-\iu\omega\tau^{-1}}.
+  \sigma(\mathbf{q},\omega) = \frac{-2\iu\epsilon_0\epsilon_{\mathrm{env}}\omega}{q}\frac{\omega^2_{\mathbf{q},\mathrm{LO}}-\omega^2_{\mathrm{TO}}}{\omega^2_{\mathrm{TO}}-\omega^2-\iu\omega\tau^{-1}}.
   \end{equation}
+ In this expression, $\omega_L$, the LO phonon frequency prior to LO-TO splitting, has been re-named as $\omega_{\mathrm{TO}}$ , the transverse optical phonon frequency, because in the absence of LO–TO splitting, they are degenerate.
+This is done in order to make the form of the final results more closely resemble their 3D counterparts, in which the dielectric function is expressed in terms of the TO frequency.
 For small $q$,  the Born charges are (to lowest-order) constant, and so the LO phonon dispersion takes the form $\omega_{\mathbf{q},\mathrm{LO}}=\sqrt{\omega^2_{\mathrm{TO}}+2v_{\mathrm{g}}\omega_{\mathrm{TO}}q} \simeq \omega_{\mathrm{TO}} + v_{\mathrm{g}}q$, where $v_{\mathrm{g}}$, the LO phonon group velocity, is defined from microscopic parameters through the relation
  \begin{equation}
-    v_{\mathrm{g}} = \frac{e^2 }{4\epsilon_0 \epsilon_{\mathrm{env}}\omega_{\Gamma, \mathrm{TO}}\Omega}\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2. 
+    v_{\mathrm{g}} = \frac{e^2 }{4\epsilon_0 \epsilon_{\mathrm{env}}\omega_{\Gamma, \mathrm{TO}}\Omega}\Big|\hat{\mathbf{q}}\cdot \sum_\kappa \tilde{\mathbf{Z}}_\kappa\boldsymbol{\eta}_\kappa\Big|^2. 
 \end{equation}
-\comment{tc}{There is an unhappy situation here: there is a lingering $i$-index, that is not summed out.
+%\comment{tc}{There is an unhappy situation here: there is a lingering $i$-index, that is not summed out.
 Perhaps this should've been $|\hat{\mathbf{q}}\cdot \sum_\kappa \mathbf{Z}_\kappa\boldsymbol{\eta}_\kappa|^2$?}
@@
 \begin{equation}
-    \sigma(\omega) =  \frac{-4\iu\epsilon_0\epsilon_{\mathrm{env}}\omega_{\mathrm{TO}}v_{\mathrm{g}}}{\omega^2_{\mathrm{TO}}-\omega^2-\iu\omega\tau^{-1}}.
+    \sigma(\omega) =  \frac{-4\iu\epsilon_0\epsilon_{\mathrm{env}}\omega\omega_{\mathrm{TO}}v_{\mathrm{g}}}{\omega^2_{\mathrm{TO}}-\omega^2-\iu\omega\tau^{-1}}.
     \label{eq:localsigma}
 \end{equation}
-\comment{tc}{@NR: As we discussed, it would probably be good to explicitly point out that $\sigma(\omega)$ \emph{doesn't} have an explicit $\epsilon_{\mathrm{env}}$ dependence, as it cancels with the dependence in $v_\mathrm{g}$.}
+%\comment{tc}{@NR: As we discussed, it would probably be good to explicitly point out that $\sigma(\omega)$ \emph{doesn't} have an explicit $\epsilon_{\mathrm{env}}$ dependence, as it cancels with the dependence in $v_\mathrm{g}$.}
+We note that despite its appearance, $\sigma(\omega)$ does not depend on $\epsilon_{\mathrm{env}}$, as $v_g$ has an opposite dependence on $\epsilon_{\mathrm{env}}$.
 From this relation, it follows that given the properties of the 2D phonons (from experiments or from \emph{ab initio} calculations), one can immediately specify the conductivity.
@@
 In the absence of such LO--TO splitting in 3D, there would be no frequency compatible with the existence of a phonon  polariton.
-On the contrary, we will show that in 2D, despite the absence of LO--TO splitting at the $\Gamma$ point, there persists a strongly confined evanescent electromagnetic mode with a high local density of states which in all respects is similar to a phonon polariton of a thin film, but is in fact the 2D LO phonon of the polar monolayer.
+On the contrary, we will show that in 2D, despite the absence of LO--TO splitting at the $\Gamma$ point, there persists a strongly confined evanescent electromagnetic mode with a high local density of states which in all respects is similar to a phonon polariton of a thin film, but is in fact the 2D LO phonon of the polar monolayer (thus the phrases "phonon polariton" and "2D LO phonon may be used somewhat interchangably as is the case in plasmonics where the terms "plasmon polariton" and "2D plasmon" are often used interchangably).
 
 The second comment is that much of what has been discussed here has a strong analogy with the theory of optical response in electron gases in 2D, and particularly the relation between plasmons in 2D and 3D.
-To elaborate on this analogy, we take Equation~\eqref{eq:zerosofresponse} in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $e^2\big|\sum_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \big|^2 \equiv  \Big(\frac{eZ^r_*}{\sqrt{M_*}}\Big)^2 \equiv \Big(\frac{Q_*}{\sqrt{M_*}}\Big)^2$.
-\comment{tc}{Another lingering $i$-index here.}
+To elaborate on this analogy, we take Equation~\eqref{eq:zerosofresponse} in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $e^2\Big|\hat{\mathbf{q}}\cdot \sum_\kappa \tilde{\mathbf{Z}}_\kappa\boldsymbol{\eta}_\kappa\Big|^2 \equiv \frac{Q_*^2}{M_*}$, with $Q_*$ being an effective charge and $M_*$ being an effective mass.
+%\comment{tc}{Another lingering $i$-index here.}
 Then, the LO--TO splitting can be written as $\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} = \frac{Q_*^2}{2\epsilon_0\epsilon_{\mathrm{env}} M_*}q$.
 Now we note that the RHS is exactly the squared-frequency $\omega^2_{\mathbf{q}\mathrm{p}}$ for a plasma oscillation in a 2D gas of charged particles with charge $Q_*$ and mass $M_*$.
-\comment{tc}{You mention the meanings of $Q_*$ and $M_*$: you probably also ought to mention the (of course reasonably obvious) meaning of $Z_*$ (and, probably, it should be $\tilde{Z}_*$).
+%\comment{tc}{You mention the meanings of $Q_*$ and $M_*$: you probably also ought to mention the (of course reasonably obvious) meaning of $Z_*$ (and, probably, it should be $\tilde{Z}_*$).
 Actually, it is not clear to me why we would need both $Q_*$ and $\tilde{Z}_*$ here?}
@@
 \begin{equation}
-    \epsilon_{\mathrm{env}} + \frac{1}{2}\epsilon_0 q \Pi(q,\omega) \equiv \epsilon_{\mathrm{RPA}}(\omega) = 0,
+    \epsilon_{\mathrm{env}} + \frac{1}{2}\epsilon_0 q \Pi(q,\omega) \equiv \epsilon_{\mathrm{RPA}}(\omega) = 0 \implies q=\frac{2i\omega\epsilon_0\epsilon_{\mathrm{env}}}{\sigma(q,\omega)},
     \label{eq:dispeq}
@@
 with $\epsilon_{\mathrm{env}}\equiv (\epsilon_++\epsilon_-)/2$.
-Thus, the condition for `phonon polaritons' \comment{tc}{Remove quotes} coincides precisely with the condition for longitudinal optical phonons.
-Thus, one of the main results of our manuscript is that despite the lack of LO--TO splitting at the $\Gamma$ point, there nevertheless exists a strongly confined evanescent mode in many respects similar to the phonon polaritons of thin films.
+Thus, the condition for phonon polaritons coincides precisely with the condition for longitudinal optical phonons.
+One of the main results of our manuscript is that despite the lack of LO--TO splitting at the $\Gamma$ point, there nevertheless exists a strongly confined evanescent mode in many respects similar to the phonon polaritons of thin films.
 We now analyze the dispersion relation of phonon polaritons in a specific material, hexagonal boron nitride, in more detail, showing the possibility of highly confined electromagnetic modes with a large local density of states.
@@
     \caption{%
-        \textbf{Properties of phonon polaritons in a monolayer of \swap{\inset{hexagonal} boron nitride}{hBN} compared to bulk.} (a)~Real and imaginary parts of the conductivity of 2D \swap{hexagonal boron nitride}{hBN} \remove{in units of $e^2/\hbar$} for different values of the loss-rate.
+        \textbf{Properties of phonon polaritons in a monolayer of hBN compared to bulk.} (a)~Real and imaginary parts of the conductivity of 2D hBN for different values of the loss-rate.
 (b)~Propagation quality factor, which measures the number of wavelengths of propagation of the 2D phonon polariton.
 (c)~Dispersion relation of phonon polaritons in the monolayer (black) and thin films whose thicknesses are taken to be 1, 2, 4, and 8 interlayer spacings in hBN.
-For the thin-film, the fundamental mode \inset{(dash-dotted)} and the first higher-order mode \inset{(dashed)} are plotted.
-\comment{tc}{@NR: Check that the dash-dotted and dashed-labelling is correct.
+For the thin-film, the fundamental mode (dashed) and the first higher-order mode (dash-dotted) are plotted.
+%\comment{tc}{@NR: Check that the dash-dotted and dashed-labelling is correct.
 From our discussion yesterday, I had the impression that the dashed lines were the ``fundamental'' modes and the dash-dotted the spatially-more-oscillatory ones: if that is the case, then `higher-order' is an unfortunate wording, as it suggests a higher frequency as well.}
@@
 
-\section{Properties of 2D LO phonon modes in monolayer hBN}
+%\section{Properties of 2D phonon polaritons in monolayer hBN}
 
 In Figure ~\ref{fig:2}(a), we present the conductivity of 2D hBN of Equation~\ref{eq:localsigma}, using parameters from Ref.~\citenum{sohier2017breakdown} calculated from density functional theory within the local density approximation.
-\comment{tc}{It is unfortunate that we never state the actual values used for $v_{\mathrm{g}}$ and $\omega_{\mathrm{TO},\mathrm{LO}}(\mathbf{q}=0)$: without them, one cannot reproduce the figures (as I just realized in trying to do so).}
+In this plot $\epsilon_{\mathrm{env}}=1,$ $\omega_{\mathrm{TO}} = $ \SI{1387}{\per\cm} and $v_g = 1.2\times 10^{-4}c$, with $c$ the speed of light in vacuum.
+%\comment{tc}{It is unfortunate that we never state the actual values used for $v_{\mathrm{g}}$ and $\omega_{\mathrm{TO},\mathrm{LO}}(\mathbf{q}=0)$: without them, one cannot reproduce the figures (as I just realized in trying to do so).}
 From the conductivity, the dispersion relation of phonon polaritons on an infinite sheet is given by $q={2\iu\omega\epsilon_0\epsilon_{\mathrm{env}}}/{\sigma(\omega)}$.
@@
 In fact, this short a wavelength is essentially beyond the range of measurability of current scattering near-field microscopy (SNOM) setups, which are limited (by tip radii) to wavelengths greater than ${\approx}\,\SI{60}{\nm}$.
-The 2D LO phonon polariton could in principle be measured by SNOM closer to the TO frequency, where confinement is smaller; unfortunately, as shown in Figure~\ref{fig:2}(b), near the TO frequency, dissipation is far higher (and corresponding propagation quality factors $\Re q/\Im q$ far lower) due to large $\Re\sigma$ (or, equivalently, large $\Im\epsilon_{\mathrm{RPA}}$).
+The 2D phonon polariton could in principle be measured by SNOM closer to the TO frequency, where confinement is smaller; unfortunately, as shown in Figure~\ref{fig:2}(b), near the TO frequency, dissipation is far higher (and corresponding propagation quality factors $\Re q/\Im q$ far lower) due to large $\Re\sigma$ (or, equivalently, large $\Im\epsilon_{\mathrm{RPA}}$).
 
-These considerations imply that access to the lower-loss and higher-confined portions of the dispersion relation of phonon polaritons, in the absence of a sharper tip, requires a near-field probe \swap{that probes larger momenta.
-Free electrons exemplify such high-momentum probes}{such as a free electron probe,} as used in electron energy loss spectroscopy (EELS), where slow electrons can be used to probe plasmon wavelengths of just a few nanometers in monolayer metals, as well as the nonlocal bulk plasmon dispersion in metals~\cite{nagao2001dispersion,de2010optical,diaconescu2007low}.
+These considerations imply that access to the lower-loss and higher-confined portions of the dispersion relation of phonon polaritons, in the absence of a sharper tip, requires a near-field probe such as a free electron probe, as used in electron energy loss spectroscopy (EELS), where slow electrons can be used to probe plasmon wavelengths of just a few nanometers in monolayer metals, as well as the nonlocal bulk plasmon dispersion in metals~\cite{nagao2001dispersion,de2010optical,diaconescu2007low}.
 EELS has been recently employed to measure phonon polaritons in ultrathin films of hBN~\cite{govyadinov2017probing}.
@@
 In the rest of this section, we discuss the relation between the dispersion of an hBN monolayer versus the atomically-thick limit of a thin film of a material with hBN's bulk dielectric function.
-To aid this discussion, in Figure~\ref{fig:2}(c), we show the dispersion relation of thin films of `bulk' hBN with film thicknesses of 1, 2, 4, and 8 times the interlayer spacing of bulk hBN, which is roughly \SI{0.33}{\nm}.
+To aid this discussion, in Figure~\ref{fig:2}(c), we show the dispersion relation of thin films of bulk hBN with film thicknesses of 1, 2, 4, and 8 times the interlayer spacing of bulk hBN, which is roughly \SI{0.33}{\nm}.
 For these plots, we take hBN to be cleaved such that the optical axis is perpendicular to the plane of the film.
@@
 This can never happen by extrapolating to a monolayer a dielectric function which has finite LO--TO splitting.
-This `turn-back' point corresponds to a frequency of around \SI{1500}{\per\cm} and a phonon wavelength of less than \SI{1}{\nm}.
+This turn-back point corresponds to a frequency of around \SI{1500}{\per\cm} and a phonon wavelength of less than \SI{1}{\nm}.
 Thus, this deviation is unlikely to be probed in typical optical experiments.
 However it means that for the monolayer, the range of accessible phonon polaritonic frequencies for optical applications is reduced compared to thick films.
-Finally, while we have focused on anomalies between the atomically-thin limit of bulk and a true monolayer, a comparison of the monolayer with the fundamental phonon polariton mode of the `one-atom-thick thin film' suggests that we have a reasonable qualitative understanding of the monolayer dispersion from this limit.
+Finally, while we have focused on anomalies between the atomically-thin limit of bulk and a true monolayer, a comparison of the monolayer with the fundamental phonon polariton mode of the one-atom-thick thin film suggests that we have a reasonable qualitative understanding of the monolayer dispersion from this limit.
 
@@
         \textbf{Extreme spontaneous emission enhancement due to 2D LO phonons in nanostructured geometries.} Plotted is the enhancement of the spontaneous emission rate for an emitter.
-The emitter is \swap{concentric with the disk and is at a distance $z=\SI{5}{\nm}$ away from the disk}{$z=\SI{5}{\nm}$ above the disk's center and is oriented normal to its plane ($\hat{\mathbf{n}} = \hat{\mathbf{z}}$)}.
+The emitter is $z=\SI{5}{\nm}$ above the disk's center and is oriented normal to its plane ($\hat{\mathbf{n}} = \hat{\mathbf{z}}$).
 For a disk with a diameter of \SI{20}{\nm}, and assuming a relaxation rate $\tau^{-1} = \SI{0.5}{\meV}$, the rate of emission enhancement can be enhanced \num{100} million-fold.
 For an emitter with a free-space decay rate of $\SI{1e6}{\per\cm}$ at \SI{7}{\micro\m}, the emitter would experience a decay rate comparable to the frequency of the disk mode, leading to ultra-strong coupling of an external emitter with 2D LO phonons.
+For reference, we compare the spontaneous emission enhancement in a nanodisk geometry to that of a disk of infinite radius (i.e., a flat sheet), showing clear enhancement relative to the flat sheet due to concentration of local density of states around phonon polariton resonances.
 Note that $\tau^{-1} = \SI{0.5}{\meV}$ is of the order of the loss rate in bulk hBN.
-\comment{tc}{@NR: Citation needed.}
-        \comment{tc}{@NR: Caption should be updated to reflect addition to figure; 1. potential maps (in the $z=0$ plane) and 2. sheet limit.}
+Also shown in the figure are maps of the electric potential on the surface of the disk for modes corresponding to selected phonon polariton resonances in the plot.
+%\comment{tc}{@NR: Citation needed.}
+       % $$\comment{tc}{@NR: Caption should be updated to reflect addition to figure; 1. potential maps (in the $z=0$ plane) and 2. sheet limit.}
         \label{fig:3}
@@
 For simplicity, we consider hBN nanostructured as a disk, which leads to the formation of sharp resonances quantized along the azimuthal and radial directions.
-The disk is taken to have a radius of \SI{20}{\nm}
-\comment{tc}{@NR: We have four different radii ($\SIlist{10;20;30;40}{\nm}$), not just one}
-and a loss rate $\tau^{-1} = \SI{0.5}{\meV}$, which is of the order of the loss rate in bulk hBN. 
-\comment{tc}{@NR: Citation needed.}
-We parameterize the coupling between the dipole and phonon polaritons through the rate of spontaneous emission $\Gamma$ of phonon polaritons by the dipole.
+The disk is taken to have a radii varying from \SI{10}{\nm} to \SI{40}{nm}
+%\comment{tc}{@NR: We have four different radii ($\SIlist{10;20;30;40}{\nm}$), not just one}
+and a loss rate $\tau^{-1} = \SI{0.5}{\meV}$, which is of the order of the loss rate in bulk hBN \cite{caldwell2014sub, dai2014tunable}.
+We also show (dashed line) the results for a disk of infinite radius, i.e., a flat sheet of monolayer hBN.
+We parameterize the coupling between the dipole and phonon polaritons through the rate of spontaneous emission $\Gamma$ of phonon polaritons by the dipole, normalized to the rate of spontaneous emission in free space $\Gamma_0$.
 It is related to the dyadic Green function $\mathbf{G}$ of the Maxwell equations for the nanostructure via the expression \cite{novotny2012principles}:
@@
 	\frac{\Gamma}{\Gamma_0} 
-	= 
-	1 + \frac{6\pi c}{\omega_0}\hat{\mathbf{n}}\cdot\Im\mathbf{G}(\mathbf{r}_0,\mathbf{r}_0,\omega_0)\cdot\hat{\mathbf{n}},
+	=  \frac{6\pi c}{\omega_0}\hat{\mathbf{n}}\cdot\Im\mathbf{G}(\mathbf{r},\mathbf{r},\omega)\cdot\hat{\mathbf{n}},
 \end{equation}
-\comment{tc}{@NR: $\Gamma_0$ has not been defined.}
-where $\hat{\mathbf{n}}$ is the orientation of the dipole, $\mathbf{r}_0$ is its position, and $\omega_0$ is its frequency.
-\comment{tc}{@NR: It could probably be fine to just drop the zero-subscript in $\mathbf{r}$ and $\omega$.}
-The dyadic Green function is computed using a \inset{quasistatic} boundary element method (as in Ref.~\citenum{jin2017infrared})\remove{, within the electrostatic approximation, which is well-respected due to the extreme sub-wavelength confinement of the electromagnetic field of the phonon polariton}.
+%\comment{tc}{@NR: $\Gamma_0$ has not been defined.}
+where $\hat{\mathbf{n}}$ is the orientation of the dipole, $\mathbf{r}$ is its position, and $\omega$ is its frequency.
+%\comment{tc}{@NR: It could probably be fine to just drop the zero-subscript in $\mathbf{r}$ and $\omega$.}
+The dyadic Green function is computed using a quasi-electrostatic boundary element method (as in Ref.~\citenum{jin2017infrared}).
 In Figure~\ref{fig:3}, we plot the enhancement of the spontaneous emission rate $\Gamma/\Gamma_0$ for an external emitter polarized perpendicularly to the plane of the disk and placed \SI{5}{\nm} away from the surface of the disk.
-\comment{tc}{Maybe good to add that these choices lead to coupling with axisymmetric $l=0$ modes only.}
-We find that the rate of spontaneous emission of 2D optical phonons is approximately 8 orders of magnitude larger than the rate of spontaneous emission of photons in the far field.
+Due to the orientation and position of the dipole , which maintains the axial symmetry of the disk, the emitter only couples to axially symmetric ($\l = 0$)  modes with zero orbital angular momentum.
+We find that the rate of spontaneous emission of 2D optical phonons is approximately 8 orders of magnitude larger than the rate of spontaneous emission of photons in the far field at frequencies corresponding to resonant modes of the hBN disk.
+Such enhancement is much larger than the enhancement presented by an unstructured, infinite sheet at the same frequency, due to the concentration of local density of states around the resonances.
+Nevertheless, the average spontaneous emission enhancement, defined by the integral of the enhancement over frequencies, is comparable to that of the flat sheet, in keeping with sum rules for spontaneous emission enhancement \cite{scheel2008macroscopic}.
 We note that in this approach, the coupling of the dipole to phonon polaritons is manifested through the phonon contribution to the dielectric function of the disk.
-This should be equivalent to an approach that considers the coupling of a bound electron in an emitter to LO phonons in the disk through a \swap{(two-dimensional)}{2D} Fr\"ohlich coupling---\ie a coupling of the atomic electron to the electric potential resulting from the polarization associated with an LO phonon mode~\cite{sohier2016two}.
+This should be equivalent to an approach that considers the coupling of a bound electron in an emitter to LO phonons in the disk through a 2D Fr\"ohlich coupling---\ie a coupling of the atomic electron to the electric potential resulting from the polarization associated with an LO phonon mode~\cite{sohier2016two}.
 
-In Figure~\ref{fig:3}, we show that for an infrared emitter at a transition wavelength of \SI{7}{\micro\m} with a free-space radiative lifetime of \SI{1}{\micro\s}, \SI{5}{\nm} away from an hBN disk, the coupling rate to 2D optical phonons (about \SI{1e14}{\per\s} or \SI{65}{\meV}) would be on the same scale as the optical phonon frequency itself (about \SI{1430}{\per\cm} or \SI{2.70e14}{\per\s} or \SI{0.18}{\eV}).
-\comment{tc}{We are using four (!) distinct units for energy, \si{\per\s}, \si{\per\cm}, \si{\micro\metre}, and \si{\eV}, in various places.
-At most, we should probably use two units (then, I would prefer to get rid of \si{\per\s} and \si{\micro\metre}.
-Even if we don't do that, we probably don't need to give it in three units here.
-When writing unitful quantities, please use the \texttt{\textbackslash{}SI\{value\}\{unit\}} macro (\eg \texttt{\textbackslash{}SI\{2\}\{\textbackslash{}eV\}} for \SI{2}{\eV} or \texttt{\textbackslash{}SI\{7\}\{\textbackslash{}per\textbackslash{}cm\}} for \SI{7}{\per\cm}).}
+In Figure~\ref{fig:3}, we show that for an infrared emitter at a transition wavelength of \SI{7}{\micro\m} with a free-space radiative lifetime of \SI{1}{\micro\s}, \SI{5}{\nm} away from an hBN disk, the coupling rate to 2D optical phonons (about \SI{65}{\meV}) would be on the same scale as the optical phonon frequency itself (about \SI{180}{\meV}).
+
 This rate thus implies coupling between an emitter and the field in the regime of ultra-strong coupling.
-Moreover, the coupling rate for the \SI{20}{\nm} disk (purple), for an emitter with a far field decay rate of ${\gtrsim}\,\SI{3e14}{\per\s}$ would have a sufficient coupling strength to the \inset{distinct, radially-quantized} resonances\swap{ in purple} for its linewidth to span multiple resonances and thus be in a multi-mode ultra-strong coupling regime.
+Moreover, the coupling rate for the \SI{20}{\nm} disk (purple), for an emitter with a far field decay rate of ${\gtrsim}\,\SI{3e14}{\per\s}$ would have a sufficient coupling strength to the distinct, radially-quantized resonances in purple for its linewidth to span multiple resonances and thus be in a multi-mode ultra-strong coupling regime.
 Thus, the extreme confinement of electromagnetic energy associated with LO phonons in two dimensions enables the possibility of realizing ultra-strong coupling of an atom or molecule with optical phonons in a polar material, allowing the potential realization of new coupled states of quantum emitters and phonons such as atom--phonon polariton bound states.

### 2019-01-12 paper.tex
miniscule edits -- all yours pri

@@
     finding extreme enhancement of spontaneous emission due to coupling with localized two-dimensional (2D) LO phonons, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN phonons.
-This may lead to the design of new hybrid states of light, electrons and phonons based on strong coupling.
+This may lead to the design of new hybrid states of electrons and phonons based on strong coupling.
    % \comment{tc}{I would suggest to lose this last sentence of the abstract: it doesn't add anything very concrete and it just (undisguised) public-relations work.}
@@
 Due to this remarkable confinement and their relatively high lifetimes---around picoseconds---phonon polaritons open new opportunities for vibrational spectroscopy~\cite{autore2018boron}, radiative heat transfer~\cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters~\cite{kumar2015tunable,rivera2017making,kurman2018control,flick2018strong}.
-Thus far, extreme confinement of phonon polaritons has been achieved by the use of thin-films, which shrink the in- and out-of-plane wavelength of polaritons monotonically with decreasing film thickness~\cite{dai2014tunable,dubrovkin2018ultra}.
-A monolayer is the ultimatelimit of this effect.
-The response-transition from three-dimensional (3D) to two-dimensional (2D) polar materials is nontrivial, however, since in a polar monolayer, the LO--TO splitting that gives rise to phonon polaritons in 3D is absent at the $\Gamma$ point~\cite{sanchez2002vibrational,mele2002electric,serrano2007vibrational,sohier2017breakdown}.
-This raises a fundamental question about whether strongly confined electromagnetic modes exist in polar monolayers at all.
+Thus far, extreme confinement of phonon polaritons has been achieved by the use of thin-films (or nanostructuring), which shrink the in- and out-of-plane wavelength of polaritons monotonically with decreasing film thickness~\cite{dai2014tunable,dubrovkin2018ultra}.
+A monolayer is the ultimate limit of this effect.
+Concerning the optical response, the transition from three-dimensional (3D) to two-dimensional (2D) polar materials is nontrivial, however, since in a polar monolayer, the LO--TO splitting that gives rise to phonon polaritons in 3D is absent at the $\Gamma$ point~\cite{sanchez2002vibrational,mele2002electric,serrano2007vibrational,sohier2017breakdown}.
+This raises a fundamental question about the nature of  electromagnetic modes in polar monolayers.
 
-We now develop a theoretical and computational framework with quantitative, concrete examples that enables a first-principles understanding of phonon polaritons in 2D materials. 
+We now develop a framework with quantitative, concrete examples that enables a first-principles understanding of phonon polaritons in 2D materials. 
 %\comment{tc}{I don't think `optical phononics' is a generally accepted identifier for phonon polaritons.
@@
 The key response function of interest is the conductivity of the monolayer.
-To that end, we consider the response of the ions of the monolayer due to a longitudinal electromagnetic potential $\phi$.
+To that end, we consider the response of the ions of the monolayer due to an electric potential $\phi$.
 For that case, the interaction Hamiltonian is
@@
 \end{equation} 
-with $\rho$ the induced charge density and $\mathbf{P}$ the total polarization density associated with the ionic motion. 
+with $\rho$ the induced charge density and $\mathbf{P}$ the polarization density associated with the ionic motion. 
 Within linear response theory, the induced polarization density can be straightforwardly evaluated from the displacement amplitude $\mathbf{u}_{\kappa}$ of every atom $\kappa$ within the unit cell.
@@
 With this relation between polarization and ionic displacements, the interaction Hamiltonian in Equation~\eqref{eq:coupling} couples the scalar potential and the ionic displacements.
-We consider the response of the monolayer to a scalar potential of the form $\phi(\mathbf{r}) = \phi(\mathbf{q},\omega)\e^{\iu\mathbf{q}\cdot\mathbf{r}-\iu\omega t}$, where $\mathbf{q}$ is a 2D wavevector in the plane of the monolayer.
+We consider the response of the monolayer to a potential of the form $\phi(\mathbf{r}) = \phi(\mathbf{q},\omega)\e^{\iu\mathbf{q}\cdot\mathbf{r}-\iu\omega t}$, where $\mathbf{q}$ is a 2D wavevector in the plane of the monolayer.
 Such a potential corresponds to a longitudinal electric field  $\mathbf{E}(\mathbf{r}) = \iu\mathbf{q}\phi(\mathbf{r})$.
@@
   \end{equation}
- In this expression, $\omega_L$, the LO phonon frequency prior to LO-TO splitting, has been re-named as $\omega_{\mathrm{TO}}$ , the transverse optical phonon frequency, because in the absence of LO–TO splitting, they are degenerate.
+ In this expression, $\omega_L$, the LO phonon frequency prior to LO-TO splitting, has been re-named as $\omega_{\mathrm{TO}}$ , the transverse optical phonon frequency, because in the absence of LO--TO splitting, they are degenerate.
 This is done in order to make the form of the final results more closely resemble their 3D counterparts, in which the dielectric function is expressed in terms of the TO frequency.

### 2019-01-13 paper.tex
Final edits. Comments removed.

@@
 In the long-wavelength (local) limit, we find a universal form for the conductivity in terms of the LO phonon frequency at the $\Gamma$ point,
-   % \comment{tc}{Why not just say `the LO phonon frequency' rather than `TO' here, since they are degenerate anyway---that way, we would only have to reference LO quantities for the local conductivity (and that case, `and the group velocity of the LO phonon' would be swapped for `and its group velocity').}
     its lifetime, and the group velocity of the LO phonon.
@@
     Finally, we study the interaction of external emitters with two-dimensional hBN nanostructures, 
-  %  \comment{tc}{This is the first mention of hBN in the abstract; arguably, it should occur earlier, already at the `We calculate the confinement and propagation losses \ldots' part}
     finding extreme enhancement of spontaneous emission due to coupling with localized two-dimensional (2D) LO phonons, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN phonons.
 This may lead to the design of new hybrid states of electrons and phonons based on strong coupling.
-   % \comment{tc}{I would suggest to lose this last sentence of the abstract: it doesn't add anything very concrete and it just (undisguised) public-relations work.}
-%    Using the strong electromagnetic interactions between emitters and 2D LO phonons calculated in this work, it is now possible to design new hybrid states of light, electrons and phonons based on quantum electrodynamical strong coupling.
+
 \end{abstract}
@@
 \maketitle
-
-%\comment{tc}{@NR: There's a somewhat inconsistent use of quotations: \ie `X' vs.\ ``X'' throughout the paper.
-In most cases, double quotations are probably correct.
-Generally though, I would favor removing them in most cases, as they often read mostly as implied uncertainty or vagueness.}
 
@@
 We now develop a framework with quantitative, concrete examples that enables a first-principles understanding of phonon polaritons in 2D materials. 
-%\comment{tc}{I don't think `optical phononics' is a generally accepted identifier for phonon polaritons.
-We use the two somewhat interchangably in this manuscript: maybe settle on just one?
-(Only a single use of `phononics' (this one)---we have three uses of `phononic' elsewhere though: I think only one should be changed.)}
 We do so by deriving a universal form for the conductivity of a polar monolayer,  which depends solely on the LO and TO phonon frequencies---and their dispersion with momentum---in the 2D system.
@@
 Within linear response theory, the induced polarization density can be straightforwardly evaluated from the displacement amplitude $\mathbf{u}_{\kappa}$ of every atom $\kappa$ within the unit cell.
-Specifically, to first order, the induced polarization is %(in repeated-index notation):
-%\comment{tc}{Later on, we write things like $\sum_{\kappa}\mathbf{Z}_{\kappa}\boldsymbol{\eta}_\kappa$, so we are not entirely consistent in our application of repeated-index notation.
-Perhaps we should just write vectorial, explicit summation notation here as well (\ie write $\mathbf{P} - \mathbf{P}^0 = \sum_\kappa(\boldsymbol{\nabla}_{\mathbf{u}_\kappa}\mathbf{P})\mathbf{u}_\kappa \equiv \frac{1}{\Omega}\sum_\kappa\mathbf{Z}_{\kappa}\mathbf{u}_{\kappa}$)}
+Specifically, to first order, the induced polarization is
 \begin{equation}
-%P_i - P_i^{0} \swap{=}{\simeq} \frac{\partial P_i}{\partial u_{\kappa,j}}u_{\kappa,j} \equiv \frac{1}{\Omega}Z_{\kappa,ij}u_{\kappa,j},
 \mathbf{P} - \mathbf{P}^0 = \sum_\kappa(\boldsymbol{\nabla}_{\mathbf{u}_\kappa}\mathbf{P})\mathbf{u}_\kappa \equiv \frac{1}{\Omega}\sum_\kappa\mathbf{Z}_{\kappa}\mathbf{u}_{\kappa},
@@
 \end{equation} 
-%In this expression, $\omega_\mathrm{L}$, the LO phonon frequency prior to LO--TO splitting, has been re-named as $\omega_{\mathrm{TO}}$, the transverse optical phonon frequency, because in the absence of LO--TO splitting, they are degenerate.
-This is done in order to make the form of the final results more closely resemble their 3D counterparts.
 We have defined scaled eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}\equiv \frac{\hat{\mathbf{e}}_{\kappa\mathbf{q},\mathrm{L}}}{\sqrt{M_{\kappa}}}$, where $\hat{\mathbf{e}}_{\kappa\mathbf{q},\mathrm{L}}$ is the unit-normalized polarization vector of atom $\kappa$ in the unit cell oscillating according to a longitudinal phonon of wavevector $\mathbf{q}$ and $M_{\kappa}$ is the mass of atom $\kappa$.
@@
 \end{equation}
-%\comment{tc}{The tilde-quantity'ed $\tilde{\mathbf{Z}}_\kappa$ is not defined, nor is the $r$-quantity'ed $Z^r_{\kappa,ij}$ later on.
-I suspect they are the same thing, and that it is the $e\tilde{\mathbf{Z}}_\kappa \equiv \mathbf{Z}_\kappa$ from an earlier draft.
-@NR: Need to check and fix.}
   where $e\tilde{\mathbf{Z}}_\kappa \equiv \mathbf{Z}_\kappa$ and $V(q)$ is the Coulomb interaction in Fourier space, which in two dimensions, is given by $V(q) = \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}$.
@@
 \end{equation}
-%\comment{tc}{There is an unhappy situation here: there is a lingering $i$-index, that is not summed out.
-Perhaps this should've been $|\hat{\mathbf{q}}\cdot \sum_\kappa \mathbf{Z}_\kappa\boldsymbol{\eta}_\kappa|^2$?}
 Thus, in the long wavelength limit, we have the following universal parameterization of the conductivity of a polar monolayer:
@@
 \end{equation}
-%\comment{tc}{@NR: As we discussed, it would probably be good to explicitly point out that $\sigma(\omega)$ \emph{doesn't} have an explicit $\epsilon_{\mathrm{env}}$ dependence, as it cancels with the dependence in $v_\mathrm{g}$.}
+
 We note that despite its appearance, $\sigma(\omega)$ does not depend on $\epsilon_{\mathrm{env}}$, as $v_g$ has an opposite dependence on $\epsilon_{\mathrm{env}}$.
@@
 In the absence of such LO--TO splitting in 3D, there would be no frequency compatible with the existence of a phonon  polariton.
-On the contrary, we will show that in 2D, despite the absence of LO--TO splitting at the $\Gamma$ point, there persists a strongly confined evanescent electromagnetic mode with a high local density of states which in all respects is similar to a phonon polariton of a thin film, but is in fact the 2D LO phonon of the polar monolayer (thus the phrases "phonon polariton" and "2D LO phonon may be used somewhat interchangably as is the case in plasmonics where the terms "plasmon polariton" and "2D plasmon" are often used interchangably).
+On the contrary, we will show that in 2D, despite the absence of LO--TO splitting at the $\Gamma$ point, there persists a strongly confined evanescent electromagnetic mode with a high local density of states which in all respects is similar to a phonon polariton of a thin film, but is in fact the 2D LO phonon of the polar monolayer (thus the phrases "phonon polariton" and "2D LO phonon may be used somewhat interchangeably as is the case in plasmonics where the terms "plasmon polariton" and "2D plasmon" are often used interchangeably).
 
@@
 To elaborate on this analogy, we take Equation~\eqref{eq:zerosofresponse} in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $e^2\Big|\hat{\mathbf{q}}\cdot \sum_\kappa \tilde{\mathbf{Z}}_\kappa\boldsymbol{\eta}_\kappa\Big|^2 \equiv \frac{Q_*^2}{M_*}$, with $Q_*$ being an effective charge and $M_*$ being an effective mass.
-%\comment{tc}{Another lingering $i$-index here.}
+
 Then, the LO--TO splitting can be written as $\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} = \frac{Q_*^2}{2\epsilon_0\epsilon_{\mathrm{env}} M_*}q$.
 Now we note that the RHS is exactly the squared-frequency $\omega^2_{\mathbf{q}\mathrm{p}}$ for a plasma oscillation in a 2D gas of charged particles with charge $Q_*$ and mass $M_*$.
-%\comment{tc}{You mention the meanings of $Q_*$ and $M_*$: you probably also ought to mention the (of course reasonably obvious) meaning of $Z_*$ (and, probably, it should be $\tilde{Z}_*$).
-Actually, it is not clear to me why we would need both $Q_*$ and $\tilde{Z}_*$ here?}
+
 This squared frequency can be thought of as the ``LP--TP'' splitting between longitudinal and transverse plasma oscillations.
@@
 For the thin-film, the fundamental mode (dashed) and the first higher-order mode (dash-dotted) are plotted.
-%\comment{tc}{@NR: Check that the dash-dotted and dashed-labelling is correct.
-From our discussion yesterday, I had the impression that the dashed lines were the ``fundamental'' modes and the dash-dotted the spatially-more-oscillatory ones: if that is the case, then `higher-order' is an unfortunate wording, as it suggests a higher frequency as well.}
         \label{fig:2}
@@
 
-%\section{Properties of 2D phonon polaritons in monolayer hBN}
-
 In Figure ~\ref{fig:2}(a), we present the conductivity of 2D hBN of Equation~\ref{eq:localsigma}, using parameters from Ref.~\citenum{sohier2017breakdown} calculated from density functional theory within the local density approximation.
 In this plot $\epsilon_{\mathrm{env}}=1,$ $\omega_{\mathrm{TO}} = $ \SI{1387}{\per\cm} and $v_g = 1.2\times 10^{-4}c$, with $c$ the speed of light in vacuum.
-%\comment{tc}{It is unfortunate that we never state the actual values used for $v_{\mathrm{g}}$ and $\omega_{\mathrm{TO},\mathrm{LO}}(\mathbf{q}=0)$: without them, one cannot reproduce the figures (as I just realized in trying to do so).}
+
 From the conductivity, the dispersion relation of phonon polaritons on an infinite sheet is given by $q={2\iu\omega\epsilon_0\epsilon_{\mathrm{env}}}/{\sigma(\omega)}$.
@@
 Also shown in the figure are maps of the electric potential on the surface of the disk for modes corresponding to selected phonon polariton resonances in the plot.
-%\comment{tc}{@NR: Citation needed.}
-       % $$\comment{tc}{@NR: Caption should be updated to reflect addition to figure; 1. potential maps (in the $z=0$ plane) and 2. sheet limit.}
         \label{fig:3}
@@
 The disk is taken to have a radii varying from \SI{10}{\nm} to \SI{40}{nm}
-%\comment{tc}{@NR: We have four different radii ($\SIlist{10;20;30;40}{\nm}$), not just one}
 and a loss rate $\tau^{-1} = \SI{0.5}{\meV}$, which is of the order of the loss rate in bulk hBN \cite{caldwell2014sub, dai2014tunable}.
@@
 \end{equation}
-%\comment{tc}{@NR: $\Gamma_0$ has not been defined.}
 where $\hat{\mathbf{n}}$ is the orientation of the dipole, $\mathbf{r}$ is its position, and $\omega$ is its frequency.
-%\comment{tc}{@NR: It could probably be fine to just drop the zero-subscript in $\mathbf{r}$ and $\omega$.}
+
 The dyadic Green function is computed using a quasi-electrostatic boundary element method (as in Ref.~\citenum{jin2017infrared}).

### 2019-01-13 paper.tex
final edits -- ready

@@
     Extreme confinement of electromagnetic energy by phonon polaritons holds the promise of strong and new forms of control over the dynamics of matter. 
-    To bring such control to the atomic-scale limit, it is important to consider phonon polaritons in two dimensional systems.
-Recent studies have pointed out that in two-dimensional systems, splitting between longitudinal and transverse optical (LO and TO) phonons is absent at the $\Gamma$ point.
+    To bring such control to the atomic-scale limit, it is important to consider phonon polaritons in two-dimensional (2D) systems.
+Recent studies have pointed out that in 2D, splitting between longitudinal and transverse optical (LO and TO) phonons is absent at the $\Gamma$ point, even for polar materials.
     Does this lack of LO--TO splitting imply the absence of a phonon polariton in polar monolayers? 
-    Here, we derive a first-principles expression for the spatially nonlocal conductivity of a polar monolayer specified by the wavevector-dependent LO and TO phonon dispersions.
+    Here, we derive a first-principles expression for the conductivity of a polar monolayer specified by the wavevector-dependent LO and TO phonon dispersions.
 In the long-wavelength (local) limit, we find a universal form for the conductivity in terms of the LO phonon frequency at the $\Gamma$ point,
     its lifetime, and the group velocity of the LO phonon.
-Our analysis reveals that the phonon polariton of the bulk is essentially replaced by the LO phonon of the two-dimensional system.
-For the specific example of hexagonal boron nitride, we estimate the confinement and propagation losses of the LO phonons, finding that high confinement and reasonable propagation quality factors coincide in regions which are difficult to detect with current near-field optical microscopy techniques.
-    Finally, we study the interaction of external emitters with two-dimensional hBN nanostructures, 
-    finding extreme enhancement of spontaneous emission due to coupling with localized two-dimensional (2D) LO phonons, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN phonons.
+Our analysis reveals that the phonon polariton of 2D is simply the LO phonon of the 2D system.
+For the specific example of hexagonal boron nitride (hBN), we estimate the confinement and propagation losses of the LO phonons, finding that high confinement and reasonable propagation quality factors coincide in regions which may be difficult to detect with current near-field optical microscopy techniques.
+Finally, we study the interaction of external emitters with two-dimensional hBN nanostructures, finding extreme enhancement of spontaneous emission due to coupling with localized 2D LO phonons, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN phonons.
 This may lead to the design of new hybrid states of electrons and phonons based on strong coupling.
@@
 % --- INTRODUCTION ---
-Phonon-polaritons, hybrid quasiparticles of photons and optical phonons supported in polar materials, hold tremendous promise for deeply sub-diffraction control of electromagnetic fields at mid-infrared and terahertz frequencies.
+Phonon-polaritons, hybrid quasiparticles of photons and optical phonons supported in polar materials, hold promise for nanoscale control of electromagnetic fields at mid-infrared and terahertz frequencies.
 Qualitatively, phonon polaritons share many features with plasmon polaritons in conductors.
@@
 Due to this remarkable confinement and their relatively high lifetimes---around picoseconds---phonon polaritons open new opportunities for vibrational spectroscopy~\cite{autore2018boron}, radiative heat transfer~\cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters~\cite{kumar2015tunable,rivera2017making,kurman2018control,flick2018strong}.
-Thus far, extreme confinement of phonon polaritons has been achieved by the use of thin-films (or nanostructuring), which shrink the in- and out-of-plane wavelength of polaritons monotonically with decreasing film thickness~\cite{dai2014tunable,dubrovkin2018ultra}.
+Thus far, extreme confinement of phonon polaritons has been achieved by the use of thin-films (or nanostructuring), which shrink the in- and out-of-plane wavelength of polaritons with decreasing feature size (such as the film thickness)~\cite{dai2014tunable,dubrovkin2018ultra}.
 A monolayer is the ultimate limit of this effect.
@@
 \end{equation} 
-with $\rho$ the induced charge density and $\mathbf{P}$ the polarization density associated with the ionic motion. 
-Within linear response theory, the induced polarization density can be straightforwardly evaluated from the displacement amplitude $\mathbf{u}_{\kappa}$ of every atom $\kappa$ within the unit cell.
-Specifically, to first order, the induced polarization is
+with $\rho$ the induced charge density and $\mathbf{P}$ the induced polarization density associated with the ionic motion.
+Note that boldfaced quantities refer to vectors or tensors as appropriate.
+Within linear response theory, the polarization density can be straightforwardly evaluated from the displacement  $\mathbf{u}_{\kappa}$ of every atom $\kappa$ within the unit cell.
+Specifically, to first order, the polarization density is
 \begin{equation}
-\mathbf{P} - \mathbf{P}^0 = \sum_\kappa(\boldsymbol{\nabla}_{\mathbf{u}_\kappa}\mathbf{P})\mathbf{u}_\kappa \equiv \frac{1}{\Omega}\sum_\kappa\mathbf{Z}_{\kappa}\mathbf{u}_{\kappa},
+\mathbf{P} - \mathbf{P}^0 = \sum_\kappa(\mathbf{u}_\kappa\cdot\boldsymbol{\nabla}_{\mathbf{u}_\kappa})\mathbf{P} \equiv \frac{1}{\Omega}\sum_\kappa\mathbf{Z}_{\kappa}\mathbf{u}_{\kappa},
 \end{equation} 
@@
 Within the RPA, these Coulombic interactions are accounted for by taking the induced polarization $\mathbf{P}(\mathbf{q},\omega)$ to be proportional \emph{total} electric field, $\mathbf{E}_{\mathrm{tot}}(\mathbf{q},\omega)$, defined to be the sum of the externally applied electric field and the electric field created by the induced polarization.
-The polarization and total-field are connected by the polarization-polarization response (tensor) function  $\boldsymbol{\Pi}(\mathbf{q},\omega)$ via
+The polarization and total field are connected by the polarization-polarization response (tensor) function  $\boldsymbol{\Pi}(\mathbf{q},\omega)$ via
 \begin{equation}
@@
 \end{equation}
-where boldfaced quantities refer to vectors or tensors as appropriate.
 The polarization-polarization response function is related to the conductivity via the relation $\boldsymbol{\sigma}(\mathbf{q},\omega) = -\iu\omega\epsilon_0\boldsymbol{\Pi}(\mathbf{q},\omega)$.
@@
 \end{equation}
-where $m,n$ are states of the phononic Fock space of the monolayer, $\mathbf{P}_{mn}(\mathbf{q}) \equiv \sum_\kappa \mathbf{Z}_\kappa\langle m|\mathbf{u}_\kappa(\mathbf{q})|n\rangle$ are matrix elements of the polarization associated with phonon modes (where $\mathbf{u}_\kappa(\mathbf{q})$ is the Fourier transform of the phonon displacement operator), $E_{m}$ ($E_n$) is the energy of state $m$ ($n$), $\beta \equiv 1/k_{\text{\textsc{B}}}T$ is the inverse temperature, and $\mathcal{Z}$ is the grand partition function.
+where $m,n$ are eigenstates in the phononic Fock space of the monolayer, $\mathbf{P}_{mn}(\mathbf{q}) \equiv \sum_\kappa \mathbf{Z}_\kappa\langle m|\mathbf{u}_\kappa(\mathbf{q})|n\rangle$ are matrix elements of the polarization associated with phonon modes (where $\mathbf{u}_\kappa(\mathbf{q})$ is the Fourier transform of the phonon displacement operator), $E_{m}$ ($E_n$) is the energy of state $m$ ($n$), $\beta \equiv 1/k_{\text{\textsc{B}}}T$ is the inverse temperature, and $\mathcal{Z}$ is the grand partition function.
 We now evaluate the contribution of optical phonons to the polarization-polarization response in the low-temperature limit $T \ll \hbar\omega_{\mathrm{ph}}/k_{\mathrm{B}}$ with $\omega_{\mathrm{ph}}$ a characteristic optical phonon frequency.
@@
 We have also phenomenologically included the phonon dissipation rate $\tau^{-1}$, consistently with  a relaxation-time prescription.
-The frequency $\omega_{\mathbf{q},\mathrm{L}}$ in the denominator, as in the case of bulk phonons, is the frequency of the longitudinal phonon of wavevector $\mathbf{q}$ prior to considering LO--TO splitting \cite{BornHuang:1954}.
+The frequency $\omega_{\mathbf{q},\mathrm{L}}$ in the denominator, as in the case of bulk phonons, is the frequency of the longitudinal phonon of wavevector $\mathbf{q}$ prior to considering LO--TO splitting \cite{BornHuang:1954} (and near the $\Gamma$ point is approximately equal to the TO phonon frequency).
 This is consistent with the fact that LO--TO splitting is a collective effect arising from Coulomb interactions and the fact that the equation above represents a single-particle susceptibility.
@@
 \begin{equation}
-    \omega^2_{\mathbf{q}} - \omega^2_{\mathrm{TO}} = V(q)\frac{1}{\Omega}q^2\Big|\hat{\mathbf{q}}\cdot\sum\limits_{\kappa}\tilde{\mathbf{Z}}_{\kappa}\boldsymbol{\eta}_{\kappa}\Big|^2,
+    \omega^2_{\mathbf{q}} - \omega^2_{\mathrm{TO}} = \frac{V(q)}{e^2}\frac{1}{\Omega}q^2\Big|\hat{\mathbf{q}}\cdot\sum\limits_{\kappa}\mathbf{Z}_{\kappa}\boldsymbol{\eta}_{\kappa}\Big|^2,
     \label{eq:zerosofresponse}
 \end{equation}
-  where $e\tilde{\mathbf{Z}}_\kappa \equiv \mathbf{Z}_\kappa$ and $V(q)$ is the Coulomb interaction in Fourier space, which in two dimensions, is given by $V(q) = \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}$.
+  where $V(q)$ is the Coulomb interaction in Fourier space, which in two dimensions, is given by $V(q) = \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}$.
 Given that the zeros of the dielectric function are associated with longitudinal modes, one expects that $\omega_{\mathbf{q}}$ is in fact the frequency of the LO phonon mode.
@@
 
-Given these results, we now re-express the conductivity explicitly in terms of the 2D phonon dispersion, and derive a universal form for the conductivity in the local ($q\rightarrow 0$) limit specified in terms of three parameters: the TO phonon frequency, the group velocity of the LO phonon at the $\Gamma$ point, and the damping rate.
+Given these results, we now re-express the conductivity explicitly in terms of the 2D phonon dispersion, and derive a universal form for the conductivity in the local ($q\rightarrow 0$) limit specified in terms of three parameters: the LO phonon frequency at the $\Gamma$ point (i.e., $\omega_{\mathrm{TO}}$), the group velocity of the LO phonon at the $\Gamma$ point, and the damping rate.
 From Equation~\eqref{eq:zerosofresponse}, we can immediately write the conductivity as
@@
   \end{equation}
- In this expression, $\omega_L$, the LO phonon frequency prior to LO-TO splitting, has been re-named as $\omega_{\mathrm{TO}}$ , the transverse optical phonon frequency, because in the absence of LO--TO splitting, they are degenerate.
-This is done in order to make the form of the final results more closely resemble their 3D counterparts, in which the dielectric function is expressed in terms of the TO frequency.
+ In this expression, $\omega_\mathrm{L}$, the LO phonon frequency prior to LO-TO splitting, has been re-named as $\omega_{\mathrm{TO}}$ , the transverse optical phonon frequency, because in the absence of LO--TO splitting, they are degenerate.
+The RPA zeros $\omega_{\mathbf{q}$ have also been renamed as $\omega_{\mathbf{q},\mathrm{LO}}$.
+This is done in order to make the form of the final results more closely resemble their 3D counterparts, in which the dielectric function is expressed in terms of the TO frequency (see for example Eq.
+(\ref{eq:3deps})).
+
  For small $q$,  the Born charges are (to lowest-order) constant, and so the LO phonon dispersion takes the form $\omega_{\mathbf{q},\mathrm{LO}}=\sqrt{\omega^2_{\mathrm{TO}}+2v_{\mathrm{g}}\omega_{\mathrm{TO}}q} \simeq \omega_{\mathrm{TO}} + v_{\mathrm{g}}q$, where $v_{\mathrm{g}}$, the LO phonon group velocity, is defined from microscopic parameters through the relation
  \begin{equation}
-    v_{\mathrm{g}} = \frac{e^2 }{4\epsilon_0 \epsilon_{\mathrm{env}}\omega_{\Gamma, \mathrm{TO}}\Omega}\Big|\hat{\mathbf{q}}\cdot \sum_\kappa \tilde{\mathbf{Z}}_\kappa\boldsymbol{\eta}_\kappa\Big|^2. 
+    v_{\mathrm{g}} = \frac{\Big|\hat{\mathbf{q}}\cdot \sum_\kappa \mathbf{Z}_\kappa\boldsymbol{\eta}_\kappa\Big|^2 }{4\epsilon_0 \epsilon_{\mathrm{env}}\omega_{\mathrm{TO}}\Omega}. 
 \end{equation}
@@
 The second comment is that much of what has been discussed here has a strong analogy with the theory of optical response in electron gases in 2D, and particularly the relation between plasmons in 2D and 3D.
-To elaborate on this analogy, we take Equation~\eqref{eq:zerosofresponse} in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $e^2\Big|\hat{\mathbf{q}}\cdot \sum_\kappa \tilde{\mathbf{Z}}_\kappa\boldsymbol{\eta}_\kappa\Big|^2 \equiv \frac{Q_*^2}{M_*}$, with $Q_*$ being an effective charge and $M_*$ being an effective mass.
-
+To elaborate on this analogy, we take Equation~\eqref{eq:zerosofresponse} in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $\Big|\hat{\mathbf{q}}\cdot \sum_\kappa \mathbf{Z}_\kappa\boldsymbol{\eta}_\kappa\Big|^2 \equiv \frac{Q_*^2}{M_*}$, with $Q_*$ being an effective charge and $M_*$ being an effective mass.
 Then, the LO--TO splitting can be written as $\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} = \frac{Q_*^2}{2\epsilon_0\epsilon_{\mathrm{env}} M_*}q$.
 Now we note that the RHS is exactly the squared-frequency $\omega^2_{\mathbf{q}\mathrm{p}}$ for a plasma oscillation in a 2D gas of charged particles with charge $Q_*$ and mass $M_*$.
-
-This squared frequency can be thought of as the ``LP--TP'' splitting between longitudinal and transverse plasma oscillations.
+To connect to LO-TO splitting in phonons, this squared frequency $\omega^2_{\mathbf{q}\mathrm{p}}$ can be thought of as the ``LP--TP'' splitting between longitudinal and transverse plasma oscillations.
 Of course, there are no transverse plasma oscillations due to the structure-less nature of the electron, and so ``$\omega_{\mathrm{TP}}"$  should be considered as 0.
@@
 The change in the dispersion of 3D versus 2D plasmons, like 3D versus 2D LO phonons, arises from the change in dimensionality of the Coulomb interaction from 3D to 2D.
-This analogy between phononic and plasmonic behavior as a function of dimension is illustrated schematically in Figures~\ref{fig:1}(b,c).
+This analogy between phononic and plasmonic behavior as a function of dimension is illustrated schematically in Figures~\ref{fig:1}(b,c) in order to help unify the understanding of 2D plasmonics and phononics.
+These considerations should also extend to the one-dimensional case.
 
 \section{Electrodynamics of optical phonons in two-dimensions}
-To relate for example the polarization-polarization response function  to the electromagnetic modes supported by a polar monolayer, we solve Maxwell's equations for an evanescent electromagnetic mode supported by a surface with polarization-polarization response tensor $\mathbf{\Pi}$.
+To relate the conductivity function  to the electromagnetic modes supported by a polar monolayer, we solve Maxwell's equations for an evanescent electromagnetic mode supported by a surface with conductivity $\sigma$.
 We consider the monolayer to be sandwiched by a superstrate of permittivity $\epsilon_+$ and a substrate of permittivity $\epsilon_-$.
@@
 In a monolayer geometry with translation invariance and in-plane isotropy, the solutions of Maxwell's equations can be decomposed into transverse magnetic (TM) and transverse electric (TE) parts, where the magnetic or electric field respectively is transverse to the in-plane wavevector of the mode.
-In practice, it is the TM mode which is associated with highly confined electromagnetic waves that are of use in nanophotonics.
-Given the isotropy, we may suppress indices from the response tensor.
-We may also consider without loss of generality a TM mode with wavenumber $q$ along the $x$-direction in the monolayer and magnetic field $H(z)\e^{\iu qx-\iu\omega t}$ along the $y$-direction of the monolayer.
+In practice, it is the TM mode which is associated with highly confined electromagnetic waves.
+We consider without loss of generality a TM mode with wavenumber $q$ along the $x$-direction in the monolayer and magnetic field $H(z)\e^{\iu qx-\iu\omega t}$ along the $y$-direction of the monolayer.
 The direction transverse to the monolayer is denoted as $z$.
@@
 %\comment{tc}{Probably makes sense to use $\varepsilon_+$ instead of $\varepsilon_1$ and $\varepsilon_-$ for $\varepsilon_2$, to have just a single class of subscripts for this notion.}
-We consider a solution of the form $H(z) = h_\pm\e^{\mp\kappa_\pm z}$ with $\kappa_\pm = \sqrt{q^2-\epsilon_{\pm} k_0^2}$ with $\pm$ corresponding to $\pm z > 0$ respectively.
-The boundary condition on the magnetic field is $h_+ - h_- = -K_x = \iu\omega P_x$ where $\mathbf{K}$ is the surface current density, expressed through the polarization density $\mathbf{P}$ of the monolayer.
-This condition enforces ${h_-}/{h_+} =(1+\frac{\epsilon_0}{\epsilon_-} \kappa_-\Pi)^{-1}$.
+We consider a solution of the form $H(z) = h_\pm\e^{\mp\kappa_\pm z}$ with $\kappa_\pm = \sqrt{q^2-\epsilon_{\pm} \frac{\omega^2}{c^2}}$ with $\pm$ corresponding to $\pm z > 0$ respectively.
+The boundary condition on the magnetic field is $h_+ - h_- = -K_x = -\sigma E_x$ where $\mathbf{K}$ is the surface current density, and $\mathbf{E} = -\frac{1}{i\omega\epsilon}\nabla\times\mathbf{H}$ is the electric field.
+This condition enforces $h_+-h_- = \frac{\sigma}{i\omega\epsilon_+}\kappa_+h_+$.
 Continuity of the electric field in the $x$ direction enforces ${\epsilon_-}/{\epsilon_+} = -{\kappa_-h_-}/{\kappa_+h_+}$. 
-Combining the two conditions, we obtain the usual dispersion equation for the TM mode of a polarizable 2D monolayer, namely $\epsilon_+/\kappa_+ + \epsilon_-/\kappa_- +\epsilon_0\Pi=0$.
+Combining the two conditions, we obtain the usual dispersion equation for the TM mode of a polarizable 2D monolayer, namely $\epsilon_+/\kappa_+ + \epsilon_-/\kappa_- = \frac{\sigma}{i\omega}$.
 Given the deeply subwavelength nature of 2D phonon polaritons, \ie since $q\gg\omega/c$ such that $\kappa_{\pm}\simeq q$, the dispersion equation can be reduced to its quasistatic limit without consequential loss of accuracy:
 \begin{equation}
-    \epsilon_{\mathrm{env}} + \frac{1}{2}\epsilon_0 q \Pi(q,\omega) \equiv \epsilon_{\mathrm{RPA}}(\omega) = 0 \implies q=\frac{2i\omega\epsilon_0\epsilon_{\mathrm{env}}}{\sigma(q,\omega)},
+ q=\frac{2i\omega\epsilon_0\epsilon_{\mathrm{env}}}{\sigma},
     \label{eq:dispeq}
@@
 with $\epsilon_{\mathrm{env}}\equiv (\epsilon_++\epsilon_-)/2$.
+This condition, as can be seen from the relation between $\Pi$ and $\sigma$, is equivalent to $\epsilon_{\mathrm{RPA}}=0$.
 Thus, the condition for phonon polaritons coincides precisely with the condition for longitudinal optical phonons.
@@
 In particular, at frequencies of \SI{1450}{\per\cm}, the phonon polariton has a wavelength of about \SI{15}{\nm}, significantly shorter than any phonon polariton measured so far, and, similarly shorter than any plasmonic wavelength, even in graphene.
-In fact, this short a wavelength is essentially beyond the range of measurability of current scattering near-field microscopy (SNOM) setups, which are limited (by tip radii) to wavelengths greater than ${\approx}\,\SI{60}{\nm}$.
+In fact, this short a wavelength well-below that of any polariton in current scattering near-field microscopy (SNOM) measurements.
 The 2D phonon polariton could in principle be measured by SNOM closer to the TO frequency, where confinement is smaller; unfortunately, as shown in Figure~\ref{fig:2}(b), near the TO frequency, dissipation is far higher (and corresponding propagation quality factors $\Re q/\Im q$ far lower) due to large $\Re\sigma$ (or, equivalently, large $\Im\epsilon_{\mathrm{RPA}}$).
@@
 EELS has been recently employed to measure phonon polaritons in ultrathin films of hBN~\cite{govyadinov2017probing}.
-Another interesting class of near-field probes, with relevance to fundamental physics and quantum optics applications, is a quantum emitter such as an atom, molecule, or quantum well.
+Another interesting class of near-field probes, with relevance to fundamental physics and quantum optics applications, is a quantum emitter such as an atom, molecule, or artificial atom such as a quantum dot, quantum well, or vacancy center.
 Recently, it was demonstrated using nanostructures of bulk hBN that the interaction of vibrational emitters with phonon polaritons is on the border of the strong coupling regime~\cite{autore2018boron}.
@@
 \begin{equation}
-	\epsilon_{\alpha}(\omega) = \epsilon_{\infty,\alpha}\Bigg(1 + \frac{\omega^2_{\mathrm{LO},\alpha}-\omega^2_{\mathrm{TO},\alpha}}{\omega^2-\omega^2_{\mathrm{TO},\alpha}} \Bigg),
+	\epsilon_{\alpha}(\omega) = \epsilon_{\infty,\alpha}\Bigg(1 + \frac{\omega^2_{\mathrm{LO},\alpha}-\omega^2_{\mathrm{TO},\alpha}}{\omega^2_{\mathrm{TO},\alpha}-\omega^2} \Bigg),
+	\label{eq:3deps}
 \end{equation}
-with $\epsilon_{\infty,\perp} = \num{2.95}$, $\omega_{\mathrm{TO},\perp} = \SI{760}{\per\cm}$, and $\omega_{\mathrm{LO},\perp} = \SI{830}{\per\cm}$; and $\epsilon_{\infty,\parallel} = \num{4.87}$, $\omega_{\mathrm{TO},\parallel} = \SI{1360}{\per\cm}$, and $\omega_{\mathrm{LO},\parallel} = \SI{1614}{\per\cm}$\cite{caldwell2014sub,dai2014tunable}.
+with $\epsilon_{\infty,\parallel} = \num{2.95}$, $\omega_{\mathrm{TO},\parallel} = \SI{760}{\per\cm}$, and $\omega_{\mathrm{LO},\parallel} = \SI{830}{\per\cm}$; and $\epsilon_{\infty,\perp} = \num{4.87}$, $\omega_{\mathrm{TO},\perp} = \SI{1360}{\per\cm}$, and $\omega_{\mathrm{LO},\perp} = \SI{1614}{\per\cm}$\cite{caldwell2014sub,dai2014tunable}.
 Losses are ignored in this discussion altogether. 
@@
 This turn-back point corresponds to a frequency of around \SI{1500}{\per\cm} and a phonon wavelength of less than \SI{1}{\nm}.
-Thus, this deviation is unlikely to be probed in typical optical experiments.
+Thus, this deviation is challenging to probe in typical optical experiments.
 However it means that for the monolayer, the range of accessible phonon polaritonic frequencies for optical applications is reduced compared to thick films.
@@
 The extreme confinement of electromagnetic fields offered by the 2D LO phonon presents an opportunity for quantum optical applications in which one seeks to couple an external emitter such as an atom, molecule, defect, or artificial atomic system to electromagnetic fields.
-Applications of these couplings are ultra-bright single- or two-photon sources, realizing the strong-coupling regime and the associated phenomenology of Rabi oscillations and polaritons, or resolving spectroscopically ``forbidden'' transitions \cite{koppens2011graphene,autore2018boron,rivera2016shrinking,rivera2017making,kurman2018control}.
+Applications of these couplings are ultra-bright single- or two-photon sources, realizing the strong-coupling regime and the associated phenomenology of Rabi oscillations and polaritons, or resolving spectroscopically ``forbidden'' transitions \cite{koppens2011graphene,autore2018boron,rivera2016shrinking,rivera2017making,kurman2018control} to achieve near-field spectroscopies with momentum and angular momentum resolution not accessible in far-field spectroscopies.
 
@@
 	\frac{\Gamma}{\Gamma_0} 
-	=  \frac{6\pi c}{\omega_0}\hat{\mathbf{n}}\cdot\Im\mathbf{G}(\mathbf{r},\mathbf{r},\omega)\cdot\hat{\mathbf{n}},
+	=  \frac{6\pi c}{\omega}\hat{\mathbf{n}}\cdot\Im\mathbf{G}(\mathbf{r},\mathbf{r},\omega)\cdot\hat{\mathbf{n}},
 \end{equation}
@@
 The dyadic Green function is computed using a quasi-electrostatic boundary element method (as in Ref.~\citenum{jin2017infrared}).
-In Figure~\ref{fig:3}, we plot the enhancement of the spontaneous emission rate $\Gamma/\Gamma_0$ for an external emitter polarized perpendicularly to the plane of the disk and placed \SI{5}{\nm} away from the surface of the disk.
-Due to the orientation and position of the dipole , which maintains the axial symmetry of the disk, the emitter only couples to axially symmetric ($\l = 0$)  modes with zero orbital angular momentum.
+In Figure~\ref{fig:3}, we plot the enhancement of the spontaneous emission rate $\Gamma/\Gamma_0$ for an external emitter polarized perpendicularly to the plane of the disk and placed \SI{5}{\nm} away from the center of the disk.
+Due to the orientation and position of the dipole , which maintains the axial symmetry of the disk, the emitter only couples to axially symmetric ($\ell = 0$)  modes with zero orbital angular momentum.
 We find that the rate of spontaneous emission of 2D optical phonons is approximately 8 orders of magnitude larger than the rate of spontaneous emission of photons in the far field at frequencies corresponding to resonant modes of the hBN disk.
-Such enhancement is much larger than the enhancement presented by an unstructured, infinite sheet at the same frequency, due to the concentration of local density of states around the resonances.
+Such enhancement is much larger than the enhancement presented by an unstructured, infinite sheet at the same frequency, due to the concentration of electromagnetic local density of states around the resonances.
 Nevertheless, the average spontaneous emission enhancement, defined by the integral of the enhancement over frequencies, is comparable to that of the flat sheet, in keeping with sum rules for spontaneous emission enhancement \cite{scheel2008macroscopic}.
@@
 In Figure~\ref{fig:3}, we show that for an infrared emitter at a transition wavelength of \SI{7}{\micro\m} with a free-space radiative lifetime of \SI{1}{\micro\s}, \SI{5}{\nm} away from an hBN disk, the coupling rate to 2D optical phonons (about \SI{65}{\meV}) would be on the same scale as the optical phonon frequency itself (about \SI{180}{\meV}).
-
 This rate thus implies coupling between an emitter and the field in the regime of ultra-strong coupling.
@@
 The ability to probe low-loss and highly confined electromagnetic modes associated with 2D LO phonons in polar materials provides a new platform for nanophotonics in the mid- and far-infrared spectral range.
-The identification of the phonon polariton of bulk and thin-film geometries with the 2D LO phonon made in this Letter would extend the rich phenomenology of optical phonons to nanophotonic applications.
+The identification of the phonon polariton of bulk and thin-film geometries with the 2D LO phonon made in this manuscript would extend the rich phenomenology of optical phonons to nanophotonic applications.
 This work also points the way to useful new approaches to study LO phonons, arising from the fact that 2D LO phonons, unlike their 3D counterparts, have their electromagnetic energy extend a considerable distance from the material boundary.

### 2019-01-13 paper.tex
refs added

@@
 For the specific example of hexagonal boron nitride (hBN), we estimate the confinement and propagation losses of the LO phonons, finding that high confinement and reasonable propagation quality factors coincide in regions which may be difficult to detect with current near-field optical microscopy techniques.
-Finally, we study the interaction of external emitters with two-dimensional hBN nanostructures, finding extreme enhancement of spontaneous emission due to coupling with localized 2D LO phonons, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN phonons.
+Finally, we study the interaction of external emitters with two-dimensional hBN nanostructures, finding extreme enhancement of spontaneous emission due to coupling with localized 2D phonon polaritons, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN phonons.
 This may lead to the design of new hybrid states of electrons and phonons based on strong coupling.
@@
 % --- INTRODUCTION ---
-Phonon-polaritons, hybrid quasiparticles of photons and optical phonons supported in polar materials, hold promise for nanoscale control of electromagnetic fields at mid-infrared and terahertz frequencies.
+Phonon polaritons, hybrid quasiparticles of photons and optical phonons supported in polar materials, hold promise for nanoscale control of electromagnetic fields at mid-infrared and terahertz frequencies.
 Qualitatively, phonon polaritons share many features with plasmon polaritons in conductors.
@@
 Due to this remarkable confinement and their relatively high lifetimes---around picoseconds---phonon polaritons open new opportunities for vibrational spectroscopy~\cite{autore2018boron}, radiative heat transfer~\cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters~\cite{kumar2015tunable,rivera2017making,kurman2018control,flick2018strong}.
+
 Thus far, extreme confinement of phonon polaritons has been achieved by the use of thin-films (or nanostructuring), which shrink the in- and out-of-plane wavelength of polaritons with decreasing feature size (such as the film thickness)~\cite{dai2014tunable,dubrovkin2018ultra}.
-A monolayer is the ultimate limit of this effect.
+A monolayer is the ultimate limit of this effect, making it critical to have a fundamental understanding of the optical response of 2D polar materials \cite{thygesen2017calculating}.
 Concerning the optical response, the transition from three-dimensional (3D) to two-dimensional (2D) polar materials is nontrivial, however, since in a polar monolayer, the LO--TO splitting that gives rise to phonon polaritons in 3D is absent at the $\Gamma$ point~\cite{sanchez2002vibrational,mele2002electric,serrano2007vibrational,sohier2017breakdown}.
@@
     \caption{%
-        \textbf{Extreme spontaneous emission enhancement due to 2D LO phonons in nanostructured geometries.} Plotted is the enhancement of the spontaneous emission rate for an emitter.
-The emitter is $z=\SI{5}{\nm}$ above the disk's center and is oriented normal to its plane ($\hat{\mathbf{n}} = \hat{\mathbf{z}}$).
+        \textbf{Extreme spontaneous emission enhancement due to 2D phonon polaritons in nanostructured geometries.} Plotted is the enhancement of the spontaneous emission rate for an emitter  $z=\SI{5}{\nm}$ above the disk's center and  oriented normal to its plane ($\hat{\mathbf{n}} = \hat{\mathbf{z}}$).
 For a disk with a diameter of \SI{20}{\nm}, and assuming a relaxation rate $\tau^{-1} = \SI{0.5}{\meV}$, the rate of emission enhancement can be enhanced \num{100} million-fold.
-For an emitter with a free-space decay rate of $\SI{1e6}{\per\cm}$ at \SI{7}{\micro\m}, the emitter would experience a decay rate comparable to the frequency of the disk mode, leading to ultra-strong coupling of an external emitter with 2D LO phonons.
+For an emitter with a free-space decay rate of $\SI{1e6}{\per\cm}$ at \SI{7}{\micro\m}, the emitter would experience a decay rate comparable to the frequency of the disk mode, leading to ultra-strong coupling of an external emitter with 2D phonon polaritons.
 For reference, we compare the spontaneous emission enhancement in a nanodisk geometry to that of a disk of infinite radius (i.e., a flat sheet), showing clear enhancement relative to the flat sheet due to concentration of local density of states around phonon polariton resonances.
@@
 
-The extreme confinement of electromagnetic fields offered by the 2D LO phonon presents an opportunity for quantum optical applications in which one seeks to couple an external emitter such as an atom, molecule, defect, or artificial atomic system to electromagnetic fields.
+The extreme confinement of electromagnetic fields offered by the 2D phonon polariton presents an opportunity for quantum optical applications in which one seeks to couple an external emitter such as an atom, molecule, defect, or artificial atomic system to electromagnetic fields.
 Applications of these couplings are ultra-bright single- or two-photon sources, realizing the strong-coupling regime and the associated phenomenology of Rabi oscillations and polaritons, or resolving spectroscopically ``forbidden'' transitions \cite{koppens2011graphene,autore2018boron,rivera2016shrinking,rivera2017making,kurman2018control} to achieve near-field spectroscopies with momentum and angular momentum resolution not accessible in far-field spectroscopies.
@@
 
-The ability to probe low-loss and highly confined electromagnetic modes associated with 2D LO phonons in polar materials provides a new platform for nanophotonics in the mid- and far-infrared spectral range.
+The ability to probe low-loss and highly confined electromagnetic modes associated with optical phonons in 2D polar materials provides a new platform for nanophotonics in the mid- and far-infrared spectral range.
 The identification of the phonon polariton of bulk and thin-film geometries with the 2D LO phonon made in this manuscript would extend the rich phenomenology of optical phonons to nanophotonic applications.
 This work also points the way to useful new approaches to study LO phonons, arising from the fact that 2D LO phonons, unlike their 3D counterparts, have their electromagnetic energy extend a considerable distance from the material boundary.
-Due to the strong electromagnetic interactions between emitters and 2D LO phonons shown in this Letter, it is now possible to design interesting new hybrid states of matter and phonons based on quantum electrodynamical strong coupling.
-The highly confined optical phonon waves in polar monolayers may also provide interesting new opportunities in near-field radiative heat transfer, in which it has been long known that thin-film surface phonon polaritons play a critical role.
-An important avenue of future study would be the \emph{ab initio} calculation of lifetimes of 2D LO phonons associated with three-phonon processes and electron-phonon interactions.
-In further work, it would be of great interest to study the effects of isotopic purification and cryogenic temperatures on reducing the decay rate of these 2D LO phonons.
+Due to the strong electromagnetic interactions between emitters and 2D phonon polaritons shown here, it is now possible to design interesting new hybrid states of matter and phonons based on quantum electrodynamical strong coupling.
+The highly confined phonon polaritons in polar monolayers may also provide interesting new opportunities in near-field radiative heat transfer, in which it has been long known that thin-film surface phonon polaritons play a critical role.
+Additional opportunities come from considering the near- and far-field optical properties of periodically structured layers involving hBN and other materials such as graphene \cite{papadakis2017ultralight}.
+An important avenue of future study would be the \emph{ab initio} calculation of lifetimes of 2D LO phonons associated with three-phonon processes \cite{caldwell2015low},srivastava1990physics} and electron-phonon interactions \cite{sundararaman2014theoretical,sohier2016two},ciccarino2018dynamics}.
+In further work, it would be of great interest to study the effects of isotopic purification and cryogenic temperatures on reducing the decay rate of these 2D LO phonons \cite{giles2017ultra}.
 

### 2019-01-13 paper.tex
unmerged edits.

--- 2019-01-13 paper.tex
@@
 The highly confined optical phonon waves in polar monolayers may also provide interesting new opportunities in near-field radiative heat transfer, in which it has been long known that thin-film surface phonon polaritons play a critical role.
-An important avenue of future study would be the \emph{ab initio} calculation of lifetimes of 2D LO phonons associated with three-phonon processes and electron-phonon interactions.
+An important avenue of future study would be the \emph{ab initio} calculation of lifetimes of 2D LO phonons associated with three-phonon processes and electron-phonon interactions.
 In further work, it would be of great interest to study the effects of isotopic purification and cryogenic temperatures on reducing the decay rate of these 2D LO phonons.

### 2019-01-13 paper.tex
resolved!
minor fix.

--- 2019-01-13 paper.tex
@@
 Additional opportunities come from considering the near- and far-field optical properties of periodically structured layers involving hBN and other materials such as graphene \cite{papadakis2017ultralight}.
-An important avenue of future study would be the \emph{ab initio} calculation of lifetimes of 2D LO phonons associated with three-phonon processes \cite{caldwell2015low},srivastava1990physics} and electron-phonon interactions \cite{sundararaman2014theoretical,sohier2016two},ciccarino2018dynamics}.
+An important avenue of future study would be the \emph{ab initio} calculation of lifetimes of 2D LO phonons associated with three-phonon processes \cite{caldwell2015low ,srivastava1990physics} and electron-phonon interactions \cite{sundararaman2014theoretical, sohier2016two, ciccarino2018dynamics}.
 In further work, it would be of great interest to study the effects of isotopic purification and cryogenic temperatures on reducing the decay rate of these 2D LO phonons \cite{giles2017ultra}.

### 2019-01-13 paper.tex
fix compilation error & quotes & space before cites

--- 2019-01-13 paper.tex
@@
 Thus far, extreme confinement of phonon polaritons has been achieved by the use of thin-films (or nanostructuring), which shrink the in- and out-of-plane wavelength of polaritons with decreasing feature size (such as the film thickness)~\cite{dai2014tunable,dubrovkin2018ultra}.
-A monolayer is the ultimate limit of this effect, making it critical to have a fundamental understanding of the optical response of 2D polar materials \cite{thygesen2017calculating}.
+A monolayer is the ultimate limit of this effect, making it critical to have a fundamental understanding of the optical response of 2D polar materials~\cite{thygesen2017calculating}.
 Concerning the optical response, the transition from three-dimensional (3D) to two-dimensional (2D) polar materials is nontrivial, however, since in a polar monolayer, the LO--TO splitting that gives rise to phonon polaritons in 3D is absent at the $\Gamma$ point~\cite{sanchez2002vibrational,mele2002electric,serrano2007vibrational,sohier2017breakdown}.
@@
 Considering the long-wavelength (small wavevector) limit, and taking a material with long-wavelength isotropy, such as hBN, we only have to consider the $qq$-component in the response tensor, where $qq$ denotes a pair of directions parallel to the wavevector.
-Denoting $\sigma_{qq}$ as simply $\sigma$, we find that the conductivity is given by \cite{rivera2018ab}:
+Denoting $\sigma_{qq}$ as simply $\sigma$, we find that the conductivity is given by~\cite{rivera2018ab}:
 \begin{equation}
@@
 We have also phenomenologically included the phonon dissipation rate $\tau^{-1}$, consistently with  a relaxation-time prescription.
-The frequency $\omega_{\mathbf{q},\mathrm{L}}$ in the denominator, as in the case of bulk phonons, is the frequency of the longitudinal phonon of wavevector $\mathbf{q}$ prior to considering LO--TO splitting \cite{BornHuang:1954} (and near the $\Gamma$ point is approximately equal to the TO phonon frequency).
+The frequency $\omega_{\mathbf{q},\mathrm{L}}$ in the denominator, as in the case of bulk phonons, is the frequency of the longitudinal phonon of wavevector $\mathbf{q}$ prior to considering LO--TO splitting~\cite{BornHuang:1954} (and near the $\Gamma$ point is approximately equal to the TO phonon frequency).
 This is consistent with the fact that LO--TO splitting is a collective effect arising from Coulomb interactions and the fact that the equation above represents a single-particle susceptibility.
@@
 While the conductivity is the main electromagnetic quantity of interest for electrodynamics applications, we briefly state the form of the (2D) permittivity, as its zeros immediately yield the longitudinal modes of the system, which are the LO phonons.
-The permittivity within the RPA, denoted $\epsilon_{\textrm{RPA}}$ is related to the polarization-polarization response function via \cite{jablan2009plasmonics} $\epsilon_{\textrm{RPA}} = \epsilon_{\mathrm{env}} + \frac{1}{2}q\Pi(q,\omega)$ and the conductivity via $\epsilon_{\textrm{RPA}} = \epsilon_{\mathrm{env}} + \frac{\iu q\sigma(q,\omega)}{2\omega\epsilon_0}$.
+The permittivity within the RPA, denoted $\epsilon_{\textrm{RPA}}$ is related to the polarization-polarization response function via~\cite{jablan2009plasmonics} $\epsilon_{\textrm{RPA}} = \epsilon_{\mathrm{env}} + \frac{1}{2}q\Pi(q,\omega)$ and the conductivity via $\epsilon_{\textrm{RPA}} = \epsilon_{\mathrm{env}} + \frac{\iu q\sigma(q,\omega)}{2\omega\epsilon_0}$.
 Here, $\epsilon_{\mathrm{env}}$ is the average permittivity of the bulk above and below the monolayer, and is added to take into account the polarization arising from these bulk materials.
 Note that we have neglected any intrinsic high-frequency screening in the monolayer itself, which is only relevant for wavevectors comparable to the inverse layer spacing between monolayers.
-When considering non-local corrections to the conductivity at these large wavevectors, these must be taken into account \cite{sohier2016two,thygesen2017calculating,sohier2017breakdown}.
+When considering non-local corrections to the conductivity at these large wavevectors, these must be taken into account~\cite{sohier2016two,thygesen2017calculating,sohier2017breakdown}.
 Based on Equation~\eqref{eq:conductivity_simplified}, the zeros $\omega_{\mathbf{q}}$ of the RPA dielectric function satisfy:
@@
  In this expression, $\omega_\mathrm{L}$, the LO phonon frequency prior to LO-TO splitting, has been re-named as $\omega_{\mathrm{TO}}$ , the transverse optical phonon frequency, because in the absence of LO--TO splitting, they are degenerate.
-The RPA zeros $\omega_{\mathbf{q}$ have also been renamed as $\omega_{\mathbf{q},\mathrm{LO}}$.
+The RPA zeros $\omega_{\mathbf{q}}$ have also been renamed as $\omega_{\mathbf{q},\mathrm{LO}}$.
 This is done in order to make the form of the final results more closely resemble their 3D counterparts, in which the dielectric function is expressed in terms of the TO frequency (see for example Eq.
@@
 In the absence of such LO--TO splitting in 3D, there would be no frequency compatible with the existence of a phonon  polariton.
-On the contrary, we will show that in 2D, despite the absence of LO--TO splitting at the $\Gamma$ point, there persists a strongly confined evanescent electromagnetic mode with a high local density of states which in all respects is similar to a phonon polariton of a thin film, but is in fact the 2D LO phonon of the polar monolayer (thus the phrases "phonon polariton" and "2D LO phonon may be used somewhat interchangeably as is the case in plasmonics where the terms "plasmon polariton" and "2D plasmon" are often used interchangeably).
+On the contrary, we will show that in 2D, despite the absence of LO--TO splitting at the $\Gamma$ point, there persists a strongly confined evanescent electromagnetic mode with a high local density of states which in all respects is similar to a phonon polariton of a thin film, but is in fact the 2D LO phonon of the polar monolayer (thus the phrases ``phonon polariton'' and ``2D LO phonon'' may be used somewhat interchangeably as is the case in plasmonics where the terms ``plasmon polariton'' and ``2D plasmon'' are often used interchangeably).
 
@@
 To connect to LO-TO splitting in phonons, this squared frequency $\omega^2_{\mathbf{q}\mathrm{p}}$ can be thought of as the ``LP--TP'' splitting between longitudinal and transverse plasma oscillations.
-Of course, there are no transverse plasma oscillations due to the structure-less nature of the electron, and so ``$\omega_{\mathrm{TP}}"$  should be considered as 0.
+Of course, there are no transverse plasma oscillations due to the structure-less nature of the electron, and so ``$\omega_{\mathrm{TP}}$''  should be considered equal to zero.
 In the three-dimensional plasmon case, ``$\omega_{\mathrm{TP}} = 0$'', but the difference between the squared longitudinal and transverse plasma oscillation frequencies at zero-wavevector is non-zero and given by $\omega_{\mathrm{p}}^2$.
@@
 The extreme confinement of electromagnetic fields offered by the 2D phonon polariton presents an opportunity for quantum optical applications in which one seeks to couple an external emitter such as an atom, molecule, defect, or artificial atomic system to electromagnetic fields.
-Applications of these couplings are ultra-bright single- or two-photon sources, realizing the strong-coupling regime and the associated phenomenology of Rabi oscillations and polaritons, or resolving spectroscopically ``forbidden'' transitions \cite{koppens2011graphene,autore2018boron,rivera2016shrinking,rivera2017making,kurman2018control} to achieve near-field spectroscopies with momentum and angular momentum resolution not accessible in far-field spectroscopies.
+Applications of these couplings are ultra-bright single- or two-photon sources, realizing the strong-coupling regime and the associated phenomenology of Rabi oscillations and polaritons, or resolving spectroscopically ``forbidden'' transitions~\cite{koppens2011graphene,autore2018boron,rivera2016shrinking,rivera2017making,kurman2018control} to achieve near-field spectroscopies with momentum and angular momentum resolution not accessible in far-field spectroscopies.
 
@@
 The disk is taken to have a radii varying from \SI{10}{\nm} to \SI{40}{nm}
-and a loss rate $\tau^{-1} = \SI{0.5}{\meV}$, which is of the order of the loss rate in bulk hBN \cite{caldwell2014sub, dai2014tunable}.
+and a loss rate $\tau^{-1} = \SI{0.5}{\meV}$, which is of the order of the loss rate in bulk hBN~\cite{caldwell2014sub, dai2014tunable}.
 We also show (dashed line) the results for a disk of infinite radius, i.e., a flat sheet of monolayer hBN.
 We parameterize the coupling between the dipole and phonon polaritons through the rate of spontaneous emission $\Gamma$ of phonon polaritons by the dipole, normalized to the rate of spontaneous emission in free space $\Gamma_0$.
-It is related to the dyadic Green function $\mathbf{G}$ of the Maxwell equations for the nanostructure via the expression \cite{novotny2012principles}:
+It is related to the dyadic Green function $\mathbf{G}$ of the Maxwell equations for the nanostructure via the expression~\cite{novotny2012principles}:
 \begin{equation}
@@
 Such enhancement is much larger than the enhancement presented by an unstructured, infinite sheet at the same frequency, due to the concentration of electromagnetic local density of states around the resonances.
-Nevertheless, the average spontaneous emission enhancement, defined by the integral of the enhancement over frequencies, is comparable to that of the flat sheet, in keeping with sum rules for spontaneous emission enhancement \cite{scheel2008macroscopic}.
+Nevertheless, the average spontaneous emission enhancement, defined by the integral of the enhancement over frequencies, is comparable to that of the flat sheet, in keeping with sum rules for spontaneous emission enhancement~\cite{scheel2008macroscopic}.
 We note that in this approach, the coupling of the dipole to phonon polaritons is manifested through the phonon contribution to the dielectric function of the disk.
@@
 The highly confined phonon polaritons in polar monolayers may also provide interesting new opportunities in near-field radiative heat transfer, in which it has been long known that thin-film surface phonon polaritons play a critical role.
-Additional opportunities come from considering the near- and far-field optical properties of periodically structured layers involving hBN and other materials such as graphene \cite{papadakis2017ultralight}.
-An important avenue of future study would be the \emph{ab initio} calculation of lifetimes of 2D LO phonons associated with three-phonon processes \cite{caldwell2015low},srivastava1990physics} and electron-phonon interactions \cite{sundararaman2014theoretical,sohier2016two},ciccarino2018dynamics}.
-In further work, it would be of great interest to study the effects of isotopic purification and cryogenic temperatures on reducing the decay rate of these 2D LO phonons \cite{giles2017ultra}.
+Additional opportunities come from considering the near- and far-field optical properties of periodically structured layers involving hBN and other materials such as graphene~\cite{papadakis2017ultralight}.
+An important avenue of future study would be the \emph{ab initio} calculation of lifetimes of 2D LO phonons associated with three-phonon processes~\cite{caldwell2015low ,srivastava1990physics} and electron-phonon interactions~\cite{sundararaman2014theoretical, sohier2016two, ciccarino2018dynamics}.
+In further work, it would be of great interest to study the effects of isotopic purification and cryogenic temperatures on reducing the decay rate of these 2D LO phonons~\cite{giles2017ultra}.
 

### 2019-01-13 paper.tex
fix v_g in .tex and unicode in .bib

@@
 
-We note that despite its appearance, $\sigma(\omega)$ does not depend on $\epsilon_{\mathrm{env}}$, as $v_g$ has an opposite dependence on $\epsilon_{\mathrm{env}}$.
+We note that despite its appearance, $\sigma(\omega)$ does not depend on $\epsilon_{\mathrm{env}}$, as $v_{\mathrm{g}}$ has an opposite dependence on $\epsilon_{\mathrm{env}}$.
 From this relation, it follows that given the properties of the 2D phonons (from experiments or from \emph{ab initio} calculations), one can immediately specify the conductivity.
@@
 In Figure ~\ref{fig:2}(a), we present the conductivity of 2D hBN of Equation~\ref{eq:localsigma}, using parameters from Ref.~\citenum{sohier2017breakdown} calculated from density functional theory within the local density approximation.
-In this plot $\epsilon_{\mathrm{env}}=1,$ $\omega_{\mathrm{TO}} = $ \SI{1387}{\per\cm} and $v_g = 1.2\times 10^{-4}c$, with $c$ the speed of light in vacuum.
+In this plot $\epsilon_{\mathrm{env}}=1,$ $\omega_{\mathrm{TO}} = $ \SI{1387}{\per\cm} and $v_{\mathrm{g}} = 1.2\times 10^{-4}c$, with $c$ the speed of light in vacuum.
 

### 2019-01-13 arXiv/paper.tex
fix citation-unicode-bug & prepare for arXiv submission

--- 2019-01-13 paper.tex
@@
 Additional opportunities come from considering the near- and far-field optical properties of periodically structured layers involving hBN and other materials such as graphene~\cite{papadakis2017ultralight}.
-An important avenue of future study would be the \emph{ab initio} calculation of lifetimes of 2D LO phonons associated with three-phonon processes~\cite{caldwell2015low ,srivastava1990physics} and electron-phonon interactions~\cite{sundararaman2014theoretical, sohier2016two, ciccarino2018dynamics}.
+An important avenue of future study would be the \emph{ab initio} calculation of lifetimes of 2D LO phonons associated with three-phonon processes~\cite{caldwell2015low,srivastava1990physics} and electron-phonon interactions~\cite{sundararaman2014theoretical, sohier2016two, ciccarino2018dynamics}.
 In further work, it would be of great interest to study the effects of isotopic purification and cryogenic temperatures on reducing the decay rate of these 2D LO phonons~\cite{giles2017ultra}.

### 2019-01-18 arXiv/paper.tex
epsilon-scale adjustments

@@
 We now develop a framework with quantitative, concrete examples that enables a first-principles understanding of phonon polaritons in 2D materials. 
-We do so by deriving a universal form for the conductivity of a polar monolayer,  which depends solely on the LO and TO phonon frequencies---and their dispersion with momentum---in the 2D system.
+We do so by deriving a universal form for the conductivity of a polar monolayer,  which depends solely on the LO and TO phonon frequencies---and their dispersion with momentum---in the 2D system (Equation (\ref{eq:localsigma})).
 Using parameters from Ref.~\citenum{sohier2017breakdown} for the canonical 2D polar monolayer---hexagonal boron nitride (hBN)---we present the confinement and propagation losses of the 2D phonon polariton modes, identifying the frequency region where they should be most easily detected.
@@
 
-\section{Optical response of optical phonons in two-dimensions}
+\section{Optical response of  2D optical phonons}
 In this section, we develop a theory of electromagnetic response due to optical phonons in 2D systems.
@@
 
-\section{Electrodynamics of optical phonons in two-dimensions}
+\section{Electrodynamics of 2D phonon polaritons}
 To relate the conductivity function  to the electromagnetic modes supported by a polar monolayer, we solve Maxwell's equations for an evanescent electromagnetic mode supported by a surface with conductivity $\sigma$.
@@
 For a disk with a diameter of \SI{20}{\nm}, and assuming a relaxation rate $\tau^{-1} = \SI{0.5}{\meV}$, the rate of emission enhancement can be enhanced \num{100} million-fold.
-For an emitter with a free-space decay rate of $\SI{1e6}{\per\cm}$ at \SI{7}{\micro\m}, the emitter would experience a decay rate comparable to the frequency of the disk mode, leading to ultra-strong coupling of an external emitter with 2D phonon polaritons.
+For an emitter with a free-space decay rate of $\SI{1e6}{\per\s}$ at \SI{7}{\micro\m}, the emitter would experience a decay rate comparable to the frequency of the disk mode, leading to ultra-strong coupling of an external emitter with 2D phonon polaritons.
 For reference, we compare the spontaneous emission enhancement in a nanodisk geometry to that of a disk of infinite radius (i.e., a flat sheet), showing clear enhancement relative to the flat sheet due to concentration of local density of states around phonon polariton resonances.
@@
 
-\section{Strong light-matter interactions enabled by 2D optical phonons}
+\section{Strong light-matter interactions enabled by 2D phonon polaritons}
 
@@
 This rate thus implies coupling between an emitter and the field in the regime of ultra-strong coupling.
-Moreover, the coupling rate for the \SI{20}{\nm} disk (purple), for an emitter with a far field decay rate of ${\gtrsim}\,\SI{3e14}{\per\s}$ would have a sufficient coupling strength to the distinct, radially-quantized resonances in purple for its linewidth to span multiple resonances and thus be in a multi-mode ultra-strong coupling regime.
+Moreover, the coupling rate for the \SI{20}{\nm} disk (purple), for an emitter with a far field decay rate of ${\gtrsim}\,\SI{3e6}{\per\s}$ would have a sufficient coupling strength to the distinct, radially-quantized resonances in purple for its linewidth to span multiple resonances and thus be in a multi-mode ultra-strong coupling regime.
 Thus, the extreme confinement of electromagnetic energy associated with LO phonons in two dimensions enables the possibility of realizing ultra-strong coupling of an atom or molecule with optical phonons in a polar material, allowing the potential realization of new coupled states of quantum emitters and phonons such as atom--phonon polariton bound states.

### 2019-02-04 arXiv/paper.tex
more explicit referencing in intro

@@
 A monolayer is the ultimate limit of this effect, making it critical to have a fundamental understanding of the optical response of 2D polar materials~\cite{thygesen2017calculating}.
-Concerning the optical response, the transition from three-dimensional (3D) to two-dimensional (2D) polar materials is nontrivial, however, since in a polar monolayer, the LO--TO splitting that gives rise to phonon polaritons in 3D is absent at the $\Gamma$ point~\cite{sanchez2002vibrational,mele2002electric,serrano2007vibrational,sohier2017breakdown}.
+Concerning the optical response, the transition from three-dimensional (3D) to two-dimensional (2D) polar materials is nontrivial, however, since in a polar monolayer, the LO--TO splitting that gives rise to phonon polaritons in 3D is absent at the $\Gamma$ point as has been realized and discussed in early works by Wirtz \emph{et al.} \cite{wirtz2003ab}, Sanchez-Portal \emph{et al.} \cite{sanchez2002vibrational}, Serrano \emph{et al.} \cite{serrano2007vibrational}, and also recent works such as Sohier \emph{et al.} \cite{sohier2017breakdown}.
 This raises a fundamental question about the nature of  electromagnetic modes in polar monolayers.

### 2019-03-09 paper.tex
minor fixes to LDOS section

@@
   \end{equation}
- In this expression, $\omega_\mathrm{L}$, the LO phonon frequency prior to LO-TO splitting, has been re-named as $\omega_{\mathrm{TO}}$ , the transverse optical phonon frequency, because in the absence of LO--TO splitting, they are degenerate.
+ In this expression, $\omega_\mathrm{L}$, the LO phonon frequency prior to LO-TO splitting, has been re-named as $\omega_{\mathrm{TO}}$, the transverse optical phonon frequency, because in the absence of LO--TO splitting, they are degenerate.
 The RPA zeros $\omega_{\mathbf{q}}$ have also been renamed as $\omega_{\mathbf{q},\mathrm{LO}}$.
@@
 \end{equation}
-where $\hat{\mathbf{n}}$ is the orientation of the dipole, $\mathbf{r}$ is its position, and $\omega$ is its frequency.
+where $\hat{\mathbf{n}}$ is the orientation of the dipole, $\mathbf{r}$ its position, and $\omega$ its frequency.
 
-The dyadic Green function is computed using a quasi-electrostatic boundary element method (as in Ref.~\citenum{jin2017infrared}).
+The dyadic Green function is computed using a quasistatic boundary element method (as in Ref.~\citenum{jin2017infrared}).
 In Figure~\ref{fig:3}, we plot the enhancement of the spontaneous emission rate $\Gamma/\Gamma_0$ for an external emitter polarized perpendicularly to the plane of the disk and placed \SI{5}{\nm} away from the center of the disk.
-Due to the orientation and position of the dipole , which maintains the axial symmetry of the disk, the emitter only couples to axially symmetric ($\ell = 0$)  modes with zero orbital angular momentum.
+Due to the orientation and position of the dipole, which maintains the axial symmetry of the disk, the emitter only couples to axially symmetric ($\ell = 0$)  modes with zero orbital angular momentum.
 We find that the rate of spontaneous emission of 2D optical phonons is approximately 8 orders of magnitude larger than the rate of spontaneous emission of photons in the far field at frequencies corresponding to resonant modes of the hBN disk.
@@
 Nevertheless, the average spontaneous emission enhancement, defined by the integral of the enhancement over frequencies, is comparable to that of the flat sheet, in keeping with sum rules for spontaneous emission enhancement~\cite{scheel2008macroscopic}.
-We note that in this approach, the coupling of the dipole to phonon polaritons is manifested through the phonon contribution to the dielectric function of the disk.
+We note that in this approach, the coupling of the dipole to phonon polaritons is manifested through the phonon contribution to the conductivity of the disk.
 This should be equivalent to an approach that considers the coupling of a bound electron in an emitter to LO phonons in the disk through a 2D Fr\"ohlich coupling---\ie a coupling of the atomic electron to the electric potential resulting from the polarization associated with an LO phonon mode~\cite{sohier2016two}.

### 2019-03-18 paper.tex
edits from reviewers

@@
 
-\newcommand{\HarvardSEAS}{John A. Paulson School of Engineering and Applied Sciences, Harvard University, Cambridge, MA, USA}
-\newcommand{\MITPhy}{Department of Physics, Massachusetts Institute of Technology, Cambridge, MA, USA}
+\newcommand{\HarvardSEAS}{John A. Paulson School of Engineering and Applied Sciences, Harvard University, Cambridge, MA, 02139, USA}
+\newcommand{\MITPhy}{Department of Physics, Massachusetts Institute of Technology, Cambridge, MA, 02139, USA}
 
@@
 
-Before moving on to analyze the electrodynamics of 2D phonon polaritons, we make two comments on lack of LO--TO splitting in 2D polar materials.
+Before moving on to analyze the electrodynamics of 2D phonon polaritons, we make three comments on lack of LO--TO splitting in 2D polar materials.
 The first is that this situation is in stark contrast to the situation of polar materials in 3D, which have a finite LO--TO splitting at the $\Gamma$ point.
@@
 These considerations should also extend to the one-dimensional case.
+
+The third comment is that the considerations of this section can be extended to few-layer systems, such as hBN bilayers, which may still be considered 2D for polariton wavelengths large compared to the bilayer thickness.
+The considerations may also be extended to 1D polar materials, such as BN nanotubes.
+For bilayers, as was discussed in Ref.~ \cite{sohier2017breakdown}, one may describe the LO-TO splitting in terms of Eq.
+~\ref{eq:zerosofresponse}, with the 2D Coulomb interaction, but with eigendisplacements appropriate to the few layer system.
+For a small number of layers $N$, this should enhance the LO-TO splitting by a factor of $N$.
+For 1D systems, beyond using the eigendisplacements appropriate to the 1D system, one must also use the 1D Coulomb interaction, cutoff by a length $R$ which for a nanotube, would be its radius.
+In that case, the 1D Coulomb interaction is $\frac{e^2K_0(qr)}{2\pi\epsilon_0}$, where $K_0$ is a modified Bessel function of the second kind and of order zero.
+Thus, the $q$-dependence of the LO-TO splitting is $q^2K_0(qR)$.
+This is in close analogy with the relation for the plasma frequency of 1D plasmons, as expected from the considerations of the previous paragraph.
+As regards the optical properties, the optical conductivity is still governed by Eq.~\ref{eq:2dsusceptibility}.
+
+%\textcolor{blue}{The third comment is that the considerations of this section can be extended to few-layer systems, such as hBN bilayers, which may still be considered 2D for polariton wavelengths large compared to the bilayer thickness.
+The considerations may also be extended to 1D polar materials, such as BN nanotubes.
+For bilayers, as was discussed in Ref.~ \cite{sohier2017breakdown}, one may describe the LO-TO splitting in terms of Eq.
+~\ref{eq:zerosofresponse}, with the 2D Coulomb interaction, but with eigendisplacements appropriate to the few layer system.
+For a small number of layers $N$, this should enhance the LO-TO splitting by a factor of $N$.
+For 1D systems, beyond using the eigendisplacements appropriate to the 1D system, one must also use the 1D Coulomb interaction, cutoff by a length $R$ which for a nanotube, would be its radius.
+In that case, the 1D Coulomb interaction is $\frac{e^2K_0(qr)}{2\pi\epsilon_0}$, where $K_0$ is a modified Bessel function of the second kind and of order zero.
+Thus, the $q$-dependence of the LO-TO splitting is $q^2K_0(qR)$.
+This is in close analogy with the relation for the plasma frequency of 1D plasmons, as expected from the considerations of the previous paragraph.
+As regards the optical properties, the optical conductivity is still governed by Eq.~\ref{eq:2dsusceptibility}.}
 
@@
 Given that the phonon polariton of 2D is the LO phonon, there can be only one branch of the dispersion.
+The bulk dielectric function of Eq.~\ref{eq:3deps} fails to capture this lack of higher-order modes, and thus hyperbolicity.
+For a bilayer however, there is a higher-order LO mode, associated with out-of-phase oscillation between the two layers, and thus one expects to see a higher-order mode.
+However, it is expected challenging to observe such a mode, as can be anticipated from the blue higher-order mode in Fig.~\ref{fig:2}(c), because even for a polariton wavelength of 12 nm, which is very short, the polariton frequency is only a few cm$^{-1}$ above the TO frequency, within the damping linewidth of the TO frequency.
+%\textcolor{blue}{The bulk dielectric function of Eq.~\ref{eq:3deps} fails to capture this lack of higher-order modes, and thus hyperbolicity.
+For a bilayer however, there is a higher-order LO mode, associated with out-of-phase oscillation between the two layers, and thus one expects to see a higher-order mode.
+However, it is expected challenging to observe such a mode, as can be anticipated from the blue higher-order mode in Fig.~\ref{fig:2}(c), because even for a polariton wavelength of 12 nm, which is very short, the polariton frequency is only a few cm$^{-1}$ above the TO frequency, within the damping linewidth of the TO frequency.}
+
 Another anomaly between 2D and 3D is that the dispersion relation of the 2D LO phonon is non-monotonic, and the frequency starts to decrease for sufficiently large wavevector.
@@
 Finally, while we have focused on anomalies between the atomically-thin limit of bulk and a true monolayer, a comparison of the monolayer with the fundamental phonon polariton mode of the one-atom-thick thin film suggests that we have a reasonable qualitative understanding of the monolayer dispersion from this limit.
+This qualitative similarity is to be somewhat expected, as in many cases, a 2D layer can be considered as an very thin film of bulk, provided that the microscopic properties of the bulk and monolayer are the same.
+In fact, if we take an thin film of thickness $t$ of hBN of dielectric function given by Eq.~\ref{eq:3deps}, and focus only on the in-plane response, then there is a mathematical condition for the thin film to exactly mirror the bulk.
+It is:
+\begin{equation}
+(\omega^{2 (\mathrm{bulk})}_{\mathrm{LO}} - \omega^{2 (\mathrm{bulk})}_{\mathrm{TO}})t = 4\omega^{(\mathrm{mono})}_{\mathrm{TO}}v^{(\mathrm{mono})}_g,
+\end{equation}
+where "bulk" denotes a property of the bulk dielectric function and "mono" denotes that of the monolayer conductivity.
+
+%\textcolor{blue}{This qualitative similarity is to be somewhat expected, as in many cases, a 2D layer can be considered as an very thin film of bulk, provided that the microscopic properties of the bulk and monolayer are the same.
+In fact, if we take an thin film of thickness $t$ of hBN of dielectric function given by Eq.~\ref{eq:3deps}, and focus only on the in-plane response, then there is a mathematical condition for the thin film to exactly mirror the bulk.
+It is:
+%\begin{equation}
+%(\omega^{2 (\mathrm{bulk})}_{\mathrm{LO}} - \omega^{2 (\mathrm{bulk})}_{\mathrm{TO}})t = 4\omega^{(\mathrm{mono})}_{\mathrm{TO}}v^{(\mathrm{mono})}_g,
+%\end{equation}
+%where "bulk" denotes a property of the bulk dielectric function and "mono" denotes that of the monolayer conductivity.}
 
@@
 Applications of these couplings are ultra-bright single- or two-photon sources, realizing the strong-coupling regime and the associated phenomenology of Rabi oscillations and polaritons, or resolving spectroscopically ``forbidden'' transitions~\cite{koppens2011graphene,autore2018boron,rivera2016shrinking,rivera2017making,kurman2018control} to achieve near-field spectroscopies with momentum and angular momentum resolution not accessible in far-field spectroscopies.
+One expects that a monolayer offers the largest spontaneous emission enhancement compared to any multilayer system for a sufficiently nearby emitter.
+This assertion is shown in the Supplementary Materials.
+
+%\textcolor{blue}{One expects that a monolayer offers the largest spontaneous emission enhancement compared to any multilayer system for a sufficiently nearby emitter.
+This assertion is shown in the Supplementary Materials.}
 

### 2019-03-18 paper.tex
just fixing some latex stuff; serious look later in the afternoon

@@
   \end{equation}
- In this expression, $\omega_\mathrm{L}$, the LO phonon frequency prior to LO-TO splitting, has been re-named as $\omega_{\mathrm{TO}}$, the transverse optical phonon frequency, because in the absence of LO--TO splitting, they are degenerate.
+ In this expression, $\omega_\mathrm{L}$, the LO phonon frequency prior to LO--TO splitting, has been re-named as $\omega_{\mathrm{TO}}$, the transverse optical phonon frequency, because in the absence of LO--TO splitting, they are degenerate.
 The RPA zeros $\omega_{\mathbf{q}}$ have also been renamed as $\omega_{\mathbf{q},\mathrm{LO}}$.
@@
 Now we note that the RHS is exactly the squared-frequency $\omega^2_{\mathbf{q}\mathrm{p}}$ for a plasma oscillation in a 2D gas of charged particles with charge $Q_*$ and mass $M_*$.
-To connect to LO-TO splitting in phonons, this squared frequency $\omega^2_{\mathbf{q}\mathrm{p}}$ can be thought of as the ``LP--TP'' splitting between longitudinal and transverse plasma oscillations.
+To connect to LO--TO splitting in phonons, this squared frequency $\omega^2_{\mathbf{q}\mathrm{p}}$ can be thought of as the ``LP--TP'' splitting between longitudinal and transverse plasma oscillations.
 Of course, there are no transverse plasma oscillations due to the structure-less nature of the electron, and so ``$\omega_{\mathrm{TP}}$''  should be considered equal to zero.
@@
 The considerations may also be extended to 1D polar materials, such as BN nanotubes.
-For bilayers, as was discussed in Ref.~ \cite{sohier2017breakdown}, one may describe the LO-TO splitting in terms of Eq.
-~\ref{eq:zerosofresponse}, with the 2D Coulomb interaction, but with eigendisplacements appropriate to the few layer system.
-For a small number of layers $N$, this should enhance the LO-TO splitting by a factor of $N$.
+For bilayers, as was discussed in Ref.~\citenum{sohier2017breakdown}, one may describe the LO--TO splitting in terms of Eq.~\eqref{eq:zerosofresponse}, with the 2D Coulomb interaction, but with eigendisplacements appropriate to the few layer system.
+For a small number of layers $N$, this should enhance the LO--TO splitting by a factor of $N$.
 For 1D systems, beyond using the eigendisplacements appropriate to the 1D system, one must also use the 1D Coulomb interaction, cutoff by a length $R$ which for a nanotube, would be its radius.
-In that case, the 1D Coulomb interaction is $\frac{e^2K_0(qr)}{2\pi\epsilon_0}$, where $K_0$ is a modified Bessel function of the second kind and of order zero.
-Thus, the $q$-dependence of the LO-TO splitting is $q^2K_0(qR)$.
+In that case, the 1D Coulomb interaction is $e^2K_0(qr)/2\pi\epsilon_0$, where $K_0$ is the (zeroth order) modified Bessel function of the second kind.
+Thus, the $q$-dependence of the LO--TO splitting is $q^2K_0(qR)$.
 This is in close analogy with the relation for the plasma frequency of 1D plasmons, as expected from the considerations of the previous paragraph.
-As regards the optical properties, the optical conductivity is still governed by Eq.~\ref{eq:2dsusceptibility}.
+As regards the optical properties, the optical conductivity is still governed by Eq.~\eqref{eq:2dsusceptibility}.
 
@@
 The considerations may also be extended to 1D polar materials, such as BN nanotubes.
-For bilayers, as was discussed in Ref.~ \cite{sohier2017breakdown}, one may describe the LO-TO splitting in terms of Eq.
+For bilayers, as was discussed in Ref.~ \cite{sohier2017breakdown}, one may describe the LO--TO splitting in terms of Eq.
 ~\ref{eq:zerosofresponse}, with the 2D Coulomb interaction, but with eigendisplacements appropriate to the few layer system.
-For a small number of layers $N$, this should enhance the LO-TO splitting by a factor of $N$.
+For a small number of layers $N$, this should enhance the LO--TO splitting by a factor of $N$.
 For 1D systems, beyond using the eigendisplacements appropriate to the 1D system, one must also use the 1D Coulomb interaction, cutoff by a length $R$ which for a nanotube, would be its radius.
 In that case, the 1D Coulomb interaction is $\frac{e^2K_0(qr)}{2\pi\epsilon_0}$, where $K_0$ is a modified Bessel function of the second kind and of order zero.
-Thus, the $q$-dependence of the LO-TO splitting is $q^2K_0(qR)$.
+Thus, the $q$-dependence of the LO--TO splitting is $q^2K_0(qR)$.
 This is in close analogy with the relation for the plasma frequency of 1D plasmons, as expected from the considerations of the previous paragraph.
@@
 Given that the phonon polariton of 2D is the LO phonon, there can be only one branch of the dispersion.
-The bulk dielectric function of Eq.~\ref{eq:3deps} fails to capture this lack of higher-order modes, and thus hyperbolicity.
+The bulk dielectric function of Eq.~\eqref{eq:3deps} fails to capture this lack of higher-order modes, and thus hyperbolicity.
 For a bilayer however, there is a higher-order LO mode, associated with out-of-phase oscillation between the two layers, and thus one expects to see a higher-order mode.
-However, it is expected challenging to observe such a mode, as can be anticipated from the blue higher-order mode in Fig.~\ref{fig:2}(c), because even for a polariton wavelength of 12 nm, which is very short, the polariton frequency is only a few cm$^{-1}$ above the TO frequency, within the damping linewidth of the TO frequency.
+However, it is expected challenging to observe such a mode, as can be anticipated from the blue higher-order mode in Fig.~\ref{fig:2}(c), because even for a polariton wavelength of \SI{12}{\nm}, which is very short, the polariton frequency is only a few \si{\per\cm} above the TO frequency, within the damping linewidth of the TO frequency.
 %\textcolor{blue}{The bulk dielectric function of Eq.~\ref{eq:3deps} fails to capture this lack of higher-order modes, and thus hyperbolicity.
@@
 This qualitative similarity is to be somewhat expected, as in many cases, a 2D layer can be considered as an very thin film of bulk, provided that the microscopic properties of the bulk and monolayer are the same.
-In fact, if we take an thin film of thickness $t$ of hBN of dielectric function given by Eq.~\ref{eq:3deps}, and focus only on the in-plane response, then there is a mathematical condition for the thin film to exactly mirror the bulk.
+In fact, if we take an thin film of thickness $t$ of hBN of dielectric function given by Eq.~\eqref{eq:3deps}, and focus only on the in-plane response, then there is a mathematical condition for the thin film to exactly mirror the bulk.
 It is:
@@
 \end{equation}
-where "bulk" denotes a property of the bulk dielectric function and "mono" denotes that of the monolayer conductivity.
+where `bulk' denotes a property of the bulk dielectric function and `mono' denotes that of the monolayer conductivity.
 

### 2019-03-18 paper.tex
discussion of 2.5D scenario

@@
 For a small number of layers $N$, this should enhance the LO--TO splitting by a factor of $N$.
+To go beyond an effectively 2D description of a monolayer, and make the transition from monolayer to bulk, we may make use of the "2.5D" Coulomb interaction \cite{thygesen2017calculating} in Eq.~\ref{eq:zerosofresponse}.
+In that case, for a film of thickness $d$, the interaction takes the form $V(q) = \frac{e^2}{\epsilon_0 dq^2}\left(1-\frac{2}{qd}e^{-\frac{qd}{2}}\sinh\left(\frac{qd}{2}\right) \right)$.
+This form interpolates between the 2D and 3D interactions in the limits $qd \rightarrow 0$ and $qd \rightarrow \infinity$.
+For sufficiently large thickness and finite-q, the 2.5D Coulomb interaction scales as $\frac{1}{dq^2}$.
+Noting the $q^2$ factor in Eq.~\ref{eq:zerosofresponse}, as well as the $N$-dependence of the quantity involving eigendisplacements, one finds an LO-TO splitting independent of thickness, as expected for bulk samples.
+
 For 1D systems, beyond using the eigendisplacements appropriate to the 1D system, one must also use the 1D Coulomb interaction, cutoff by a length $R$ which for a nanotube, would be its radius.
@@
 For a small number of layers $N$, this should enhance the LO--TO splitting by a factor of $N$.
-For 1D systems, beyond using the eigendisplacements appropriate to the 1D system, one must also use the 1D Coulomb interaction, cutoff by a length $R$ which for a nanotube, would be its radius.
+To go beyond an effectively 2D description of a monolayer, and make the transition from monolayer to bulk, we may make use of the "2.5D" Coulomb interaction \cite{thygesen2017calculating} in Eq.~\ref{eq:zerosofresponse}.
+In that case, for a film of thickness $d$, the interaction takes the form $V(q) = \frac{e^2}{\epsilon_0 dq^2}\left(1-\frac{2}{qd}e^{-\frac{qd}{2}}\sinh\left(\frac{qd}{2}\right) \right)$.
+This form interpolates between the 2D and 3D interactions in the limits $qd \rightarrow 0$ and $qd \rightarrow \infinity$.
+For sufficiently large thickness and finite-q, the 2.5D Coulomb interaction scales as $\frac{1}{dq^2}$.
+Noting the $q^2$ factor in Eq.~\ref{eq:zerosofresponse}, as well as the $N$-dependence of the quantity involving eigendisplacements, one finds an LO-TO splitting independent of thickness, as expected for bulk samples.}
+
+% \textcolor{blue}{For 1D systems, beyond using the eigendisplacements appropriate to the 1D system, one must also use the 1D Coulomb interaction, cutoff by a length $R$ which for a nanotube, would be its radius.
 In that case, the 1D Coulomb interaction is $\frac{e^2K_0(qr)}{2\pi\epsilon_0}$, where $K_0$ is a modified Bessel function of the second kind and of order zero.

### 2019-03-18 paper.tex
minor obvious fixes

@@
 For a small number of layers $N$, this should enhance the LO--TO splitting by a factor of $N$.
-To go beyond an effectively 2D description of a monolayer, and make the transition from monolayer to bulk, we may make use of the "2.5D" Coulomb interaction \cite{thygesen2017calculating} in Eq.~\ref{eq:zerosofresponse}.
-In that case, for a film of thickness $d$, the interaction takes the form $V(q) = \frac{e^2}{\epsilon_0 dq^2}\left(1-\frac{2}{qd}e^{-\frac{qd}{2}}\sinh\left(\frac{qd}{2}\right) \right)$.
-This form interpolates between the 2D and 3D interactions in the limits $qd \rightarrow 0$ and $qd \rightarrow \infinity$.
-For sufficiently large thickness and finite-q, the 2.5D Coulomb interaction scales as $\frac{1}{dq^2}$.
-Noting the $q^2$ factor in Eq.~\ref{eq:zerosofresponse}, as well as the $N$-dependence of the quantity involving eigendisplacements, one finds an LO-TO splitting independent of thickness, as expected for bulk samples.
+To go beyond an effectively 2D description of a monolayer, and make the transition from monolayer to bulk, we may make use of the quasi-2.5D Coulomb interaction~\cite{thygesen2017calculating} in Eq.~\eqref{eq:zerosofresponse}.
+In that case, for a film of effective thickness $d$, the interaction is approximately $V_{\text{Q2.5D}}(q) = \frac{e^2}{\epsilon_0 dq^2}\big(1-\frac{2}{qd}e^{-\frac{qd}{2}}\sinh\frac{qd}{2} \big)$.
+This form interpolates between the 2D and 3D interactions in the limits $qd \rightarrow 0$ and $qd \rightarrow \infty$.
+For sufficiently large thickness and finite-$q$, the 2.5D Coulomb interaction scales as $d^{-1}q^{-2}$.
+Noting the $q^2$ factor in Eq.~\eqref{eq:zerosofresponse}, as well as the $N$-dependence of the quantity involving eigendisplacements, one finds an LO--TO splitting independent of thickness, as expected for bulk samples.
 
@@
 For a bilayer however, there is a higher-order LO mode, associated with out-of-phase oscillation between the two layers, and thus one expects to see a higher-order mode.
-However, it is expected challenging to observe such a mode, as can be anticipated from the blue higher-order mode in Fig.~\ref{fig:2}(c), because even for a polariton wavelength of \SI{12}{\nm}, which is very short, the polariton frequency is only a few \si{\per\cm} above the TO frequency, within the damping linewidth of the TO frequency.
+However, it is expected challenging to observe such a mode, as can be anticipated from the blue higher-order mode in Figure~\ref{fig:2}(c), because even for a polariton wavelength of \SI{12}{\nm}, which is very short, the polariton frequency is only a few \si{\per\cm} above the TO frequency, within the damping linewidth of the TO frequency.
 %\textcolor{blue}{The bulk dielectric function of Eq.~\ref{eq:3deps} fails to capture this lack of higher-order modes, and thus hyperbolicity.
 For a bilayer however, there is a higher-order LO mode, associated with out-of-phase oscillation between the two layers, and thus one expects to see a higher-order mode.
-However, it is expected challenging to observe such a mode, as can be anticipated from the blue higher-order mode in Fig.~\ref{fig:2}(c), because even for a polariton wavelength of 12 nm, which is very short, the polariton frequency is only a few cm$^{-1}$ above the TO frequency, within the damping linewidth of the TO frequency.}
+However, it is expected challenging to observe such a mode, as can be anticipated from the blue higher-order mode in Figure~\ref{fig:2}(c), because even for a polariton wavelength of 12 nm, which is very short, the polariton frequency is only a few cm$^{-1}$ above the TO frequency, within the damping linewidth of the TO frequency.}
 
@@
 Finally, while we have focused on anomalies between the atomically-thin limit of bulk and a true monolayer, a comparison of the monolayer with the fundamental phonon polariton mode of the one-atom-thick thin film suggests that we have a reasonable qualitative understanding of the monolayer dispersion from this limit.
-This qualitative similarity is to be somewhat expected, as in many cases, a 2D layer can be considered as an very thin film of bulk, provided that the microscopic properties of the bulk and monolayer are the same.
-In fact, if we take an thin film of thickness $t$ of hBN of dielectric function given by Eq.~\eqref{eq:3deps}, and focus only on the in-plane response, then there is a mathematical condition for the thin film to exactly mirror the bulk.
-It is:
+This qualitative similarity is to be somewhat expected, as in many cases, a 2D layer can be considered as a very thin film of bulk, provided that the microscopic properties of the bulk and monolayer are the same.
+In fact, if we take a thin film of thickness $t$ of hBN of dielectric function given by Eq.~\eqref{eq:3deps}, and focus only on the in-plane response, then there is a mathematical condition for the thin film to exactly mirror the bulk, namely
 \begin{equation}

### 2019-03-18 paper.tex
Some fixes to/revisions of main text

--- 2019-03-18 paper.tex
@@
 \newboolean{togglechanges} 
+\newboolean{toggleadditioncolor}
 
@@
 \setboolean{togglechanges}{false} 
+% toggle to true to see comments (otherwise hidden)
+\setboolean{toggleadditioncolor}{true}
 
@@
     {{}#1}  % TC-only version
-    {\textcolor{green!50!black}{{}#1}}}
-\makeatletter
+    {{}{\color{green!50!black}#1}}}
 \newcommand{\citeremind}[1]{%
@@
 	}\xspace}
-\makeatother
+\newcommand{\addition}[1]{\ifbool{toggleadditioncolor}
+	{\textcolor{blue!65!black}{{}#1}}  % colored version
+	{#1}}
 \newcommand{\FIXME}[1]{\textcolor{red!70!black}{[#1]}}
@@
 
+\addition{%
 The third comment is that the considerations of this section can be extended to few-layer systems, such as hBN bilayers, which may still be considered 2D for polariton wavelengths large compared to the bilayer thickness.
@@
 For bilayers, as was discussed in Ref.~\citenum{sohier2017breakdown}, one may describe the LO--TO splitting in terms of Eq.~\eqref{eq:zerosofresponse}, with the 2D Coulomb interaction, but with eigendisplacements appropriate to the few layer system.
-For a small number of layers $N$, this should enhance the LO--TO splitting by a factor of $N$.
+For a small number of layers $N$, \FIXME{Mention scaling of $|\hat{\mathbf{q}}\cdot\ldots|^2 = NQ_*^2/M_*$} this should enhance the LO--TO splitting by a factor of $N$. 
+To go beyond an effectively 2D description of a monolayer, and make the transition from monolayer to bulk, we may make use of the quasi-2.5D Coulomb interaction~\cite{thygesen2017calculating} in Eq.~\eqref{eq:zerosofresponse}.
+In that case, for a film of effective thickness $t$, the interaction is approximately $V_{\text{Q2.5D}}(q) = \frac{e^2}{\epsilon_0 tq^2}\big(1-\frac{2}{qt}e^{-\frac{qt}{2}}\sinh\frac{qt}{2} \big)$. 
+This form interpolates between the 2D and 3D interactions in the limits $qt \rightarrow 0$ and $qt \rightarrow \infty$. 
+For sufficiently large thickness and finite-$q$, the 2.5D Coulomb interaction scales as $t^{-1}q^{-2}$. 
+Noting the $q^2$ factor in Eq.~\eqref{eq:zerosofresponse}, as well as the $N$-dependence of the quantity involving eigendisplacements, one finds an LO--TO splitting independent of thickness, as expected for bulk samples.
+}
+
+\addition{%
 For 1D systems, beyond using the eigendisplacements appropriate to the 1D system, one must also use the 1D Coulomb interaction, cutoff by a length $R$ which for a nanotube, would be its radius. 
@@
 As regards the optical properties, the optical conductivity is still governed by Eq.~\eqref{eq:2dsusceptibility}.
-
-%\textcolor{blue}{The third comment is that the considerations of this section can be extended to few-layer systems, such as hBN bilayers, which may still be considered 2D for polariton wavelengths large compared to the bilayer thickness.
-The considerations may also be extended to 1D polar materials, such as BN nanotubes.
-For bilayers, as was discussed in Ref.~ \cite{sohier2017breakdown}, one may describe the LO--TO splitting in terms of Eq.
-~\ref{eq:zerosofresponse}, with the 2D Coulomb interaction, but with eigendisplacements appropriate to the few layer system.
-For a small number of layers $N$, this should enhance the LO--TO splitting by a factor of $N$.
-For 1D systems, beyond using the eigendisplacements appropriate to the 1D system, one must also use the 1D Coulomb interaction, cutoff by a length $R$ which for a nanotube, would be its radius.
-In that case, the 1D Coulomb interaction is $\frac{e^2K_0(qr)}{2\pi\epsilon_0}$, where $K_0$ is a modified Bessel function of the second kind and of order zero.
-Thus, the $q$-dependence of the LO--TO splitting is $q^2K_0(qR)$.
-This is in close analogy with the relation for the plasma frequency of 1D plasmons, as expected from the considerations of the previous paragraph.
-As regards the optical properties, the optical conductivity is still governed by Eq.~\ref{eq:2dsusceptibility}.}
+}
 
@@
 where $\epsilon_+$ applies for $z>0$ and $\epsilon_-$ applies for $z<0$.
-%\comment{tc}{Probably makes sense to use $\varepsilon_+$ instead of $\varepsilon_1$ and $\varepsilon_-$ for $\varepsilon_2$, to have just a single class of subscripts for this notion.}
 We consider a solution of the form $H(z) = h_\pm\e^{\mp\kappa_\pm z}$ with $\kappa_\pm = \sqrt{q^2-\epsilon_{\pm} \frac{\omega^2}{c^2}}$ with $\pm$ corresponding to $\pm z > 0$ respectively.
@@
 This hyperbolicity leads to a dispersion for hBN thin films that have multiple branches at a given frequency, as can be seen in Figure~\ref{fig:2}(c).
-This trend persists even when the thickness of the bulk is taken down to a single layer, albeit being pushed to high wavevectors.
-This is in contrast with the true monolayer, in which there is only one LO phonon mode.
+This trend persists even when the thickness of the bulk is taken down to a single layer, albeit pushed to high wavevectors.
+This is in contrast with the true monolayer, where there is only one LO phonon mode.
 Given that the phonon polariton of 2D is the LO phonon, there can be only one branch of the dispersion.
-The bulk dielectric function of Eq.~\eqref{eq:3deps} fails to capture this lack of higher-order modes, and thus hyperbolicity.
+\addition{%
+The bulk dielectric function of Eq.~\eqref{eq:3deps} fails to capture this removal of higher-order modes (\ie hyperbolicity).
+	\comment{tc}{But, na\"ively, I guess this is to be expected, since \eqref{eq:3deps} is explicitly hyperbolic?}
 For a bilayer however, there is a higher-order LO mode, associated with out-of-phase oscillation between the two layers, and thus one expects to see a higher-order mode. 
-However, it is expected challenging to observe such a mode, as can be anticipated from the blue higher-order mode in Fig.~\ref{fig:2}(c), because even for a polariton wavelength of \SI{12}{\nm}, which is very short, the polariton frequency is only a few \si{\per\cm} above the TO frequency, within the damping linewidth of the TO frequency.
-%\textcolor{blue}{The bulk dielectric function of Eq.~\ref{eq:3deps} fails to capture this lack of higher-order modes, and thus hyperbolicity.
-For a bilayer however, there is a higher-order LO mode, associated with out-of-phase oscillation between the two layers, and thus one expects to see a higher-order mode.
-However, it is expected challenging to observe such a mode, as can be anticipated from the blue higher-order mode in Fig.~\ref{fig:2}(c), because even for a polariton wavelength of 12 nm, which is very short, the polariton frequency is only a few cm$^{-1}$ above the TO frequency, within the damping linewidth of the TO frequency.}
+However, it is expected challenging to observe such a mode, as can be anticipated from the blue higher-order mode in Figure~\ref{fig:2}(c), because even for a polariton wavelength of \SI{12}{\nm}, which is very short, the polariton frequency is only a few \si{\per\cm} above the TO frequency, within the damping linewidth of the TO frequency.
+}
 
+	\comment{tc}{Isn't this a bit strange to discuss, given that our results do not really show this (since we are sticking to the local case? This is only a feature in \citet{sohier2016two} results?}
 Another anomaly between 2D and 3D is that the dispersion relation of the 2D LO phonon is non-monotonic, and the frequency starts to decrease for sufficiently large wavevector.
@@
 Finally, while we have focused on anomalies between the atomically-thin limit of bulk and a true monolayer, a comparison of the monolayer with the fundamental phonon polariton mode of the one-atom-thick thin film suggests that we have a reasonable qualitative understanding of the monolayer dispersion from this limit.
-This qualitative similarity is to be somewhat expected, as in many cases, a 2D layer can be considered as an very thin film of bulk, provided that the microscopic properties of the bulk and monolayer are the same.
-In fact, if we take an thin film of thickness $t$ of hBN of dielectric function given by Eq.~\eqref{eq:3deps}, and focus only on the in-plane response, then there is a mathematical condition for the thin film to exactly mirror the bulk.
-It is:
+%
+\addition{%
+This qualitative similarity is to be somewhat expected, as a 2D layer can be considered as a very thin film of bulk material, provided that the microscopic properties of the bulk and monolayer do not deviate substantially.
+The differences between the $t = t_{\text{2D}} = \SI{0.33}{\nm}$ bulky thin-film and monolayer that nevertheless do exist, reflect such microscopic deviations.
+Occasionally, \eg for compatibility with standard numerical tools, it is useful to introduce a fictitious, effective bulk material whose fundamental thin-film mode \emph{exactly} matches the monolayer's.
+The dielectric function of such a ``bulkified'' monolayer is thickness-dependent and anisotropic: for a film in the $xy$ plane, it is $\boldsymbol{\epsilon}_{\text{Q2D}} = (\hat{\mathbf{x}}\hat{\mathbf{x}} + \hat{\mathbf{y}}\hat{\mathbf{y}})\epsilon_{\text{Q2D}, \parallel} + \hat{\mathbf{z}}\hat{\mathbf{z}}\epsilon_{\text{Q2D}, \perp}$ with $\epsilon_{\text{Q2D}, \perp} = 1$ and
 \begin{equation}
-(\omega^{2 (\mathrm{bulk})}_{\mathrm{LO}} - \omega^{2 (\mathrm{bulk})}_{\mathrm{TO}})t = 4\omega^{(\mathrm{mono})}_{\mathrm{TO}}v^{(\mathrm{mono})}_g,
+	\epsilon_{\text{Q2D}, \parallel}(\omega) = 1 + \frac{1}{t}\frac{4\epsilon_{\mathrm{env}}\omega_{\mathrm{TO}}v_\mathrm{g}}{\omega^2-\omega_{\mathrm{TO}}^2+\iu\omega\tau^{-1}}.
 \end{equation}
-where `bulk' denotes a property of the bulk dielectric function and `mono' denotes that of the monolayer conductivity.
-
-%\textcolor{blue}{This qualitative similarity is to be somewhat expected, as in many cases, a 2D layer can be considered as an very thin film of bulk, provided that the microscopic properties of the bulk and monolayer are the same.
-In fact, if we take an thin film of thickness $t$ of hBN of dielectric function given by Eq.~\ref{eq:3deps}, and focus only on the in-plane response, then there is a mathematical condition for the thin film to exactly mirror the bulk.
-It is:
-%\begin{equation}
-%(\omega^{2 (\mathrm{bulk})}_{\mathrm{LO}} - \omega^{2 (\mathrm{bulk})}_{\mathrm{TO}})t = 4\omega^{(\mathrm{mono})}_{\mathrm{TO}}v^{(\mathrm{mono})}_g,
-%\end{equation}
-%where "bulk" denotes a property of the bulk dielectric function and "mono" denotes that of the monolayer conductivity.}
+}
 
@@
 Applications of these couplings are ultra-bright single- or two-photon sources, realizing the strong-coupling regime and the associated phenomenology of Rabi oscillations and polaritons, or resolving spectroscopically ``forbidden'' transitions~\cite{koppens2011graphene,autore2018boron,rivera2016shrinking,rivera2017making,kurman2018control} to achieve near-field spectroscopies with momentum and angular momentum resolution not accessible in far-field spectroscopies.
-One expects that a monolayer offers the largest spontaneous emission enhancement compared to any multilayer system for a sufficiently nearby emitter.
-This assertion is shown in the Supplementary Materials.
-
-%\textcolor{blue}{One expects that a monolayer offers the largest spontaneous emission enhancement compared to any multilayer system for a sufficiently nearby emitter.
-This assertion is shown in the Supplementary Materials.}
 

### 2019-03-18 supp.tex
supp

--- 2019-03-18 paper.tex
@@
-\documentclass[aps,prb,twocolumn,
+\documentclass[aps,prb,twocolumn, final,
 	           groupedaddress,superscriptaddress,
@@
 \newboolean{togglechanges}
-\newboolean{toggleadditioncolor}
 
@@
 \setboolean{togglechanges}{false}
-% toggle to true to see comments (otherwise hidden)
-\setboolean{toggleadditioncolor}{true}
 
@@
     {{}#1}  % TC-only version
-    {{}{\color{green!50!black}#1}}}
+    {\textcolor{green!50!black}{{}#1}}}
+\makeatletter
 \newcommand{\citeremind}[1]{%
@@
 	}\xspace}
-\newcommand{\addition}[1]{\ifbool{toggleadditioncolor}
-	{\textcolor{blue!65!black}{{}#1}}  % colored version
-	{#1}}
+\makeatother
 \newcommand{\FIXME}[1]{\textcolor{red!70!black}{[#1]}}
@@
 % --- TITLE & AUTHORSHIP ---
-\title{Phonon polaritonics in two-dimensional materials}
+\title{Supporting Information for: Phonon polaritonics in two-dimensional materials}
 
@@
 
-\date{\today}
-
 % --- ABSTRACT ---
-\begin{abstract}
-    Extreme confinement of electromagnetic energy by phonon polaritons holds the promise of strong and new forms of control over the dynamics of matter. 
-    To bring such control to the atomic-scale limit, it is important to consider phonon polaritons in two-dimensional (2D) systems.
-Recent studies have pointed out that in 2D, splitting between longitudinal and transverse optical (LO and TO) phonons is absent at the $\Gamma$ point, even for polar materials.
-    Does this lack of LO--TO splitting imply the absence of a phonon polariton in polar monolayers? 
-    Here, we derive a first-principles expression for the conductivity of a polar monolayer specified by the wavevector-dependent LO and TO phonon dispersions.
-In the long-wavelength (local) limit, we find a universal form for the conductivity in terms of the LO phonon frequency at the $\Gamma$ point,
-    its lifetime, and the group velocity of the LO phonon.
-Our analysis reveals that the phonon polariton of 2D is simply the LO phonon of the 2D system.
-For the specific example of hexagonal boron nitride (hBN), we estimate the confinement and propagation losses of the LO phonons, finding that high confinement and reasonable propagation quality factors coincide in regions which may be difficult to detect with current near-field optical microscopy techniques.
-Finally, we study the interaction of external emitters with two-dimensional hBN nanostructures, finding extreme enhancement of spontaneous emission due to coupling with localized 2D phonon polaritons, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN phonons.
-This may lead to the design of new hybrid states of electrons and phonons based on strong coupling.
-
-\end{abstract}
 
@@
 
-% --- INTRODUCTION ---
-Phonon polaritons, hybrid quasiparticles of photons and optical phonons supported in polar materials, hold promise for nanoscale control of electromagnetic fields at mid-infrared and terahertz frequencies.
-Qualitatively, phonon polaritons share many features with plasmon polaritons in conductors.
-Recently, it has been shown that phonon polaritons enable confinement of light to volumes  $\sim10^6$ times smaller than that of a diffraction-limited photon in free-space~\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra,li2018infrared,ma2018plane}.
-Due to this remarkable confinement and their relatively high lifetimes---around picoseconds---phonon polaritons open new opportunities for vibrational spectroscopy~\cite{autore2018boron}, radiative heat transfer~\cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters~\cite{kumar2015tunable,rivera2017making,kurman2018control,flick2018strong}.
-
-Thus far, extreme confinement of phonon polaritons has been achieved by the use of thin-films (or nanostructuring), which shrink the in- and out-of-plane wavelength of polaritons with decreasing feature size (such as the film thickness)~\cite{dai2014tunable,dubrovkin2018ultra}.
-A monolayer is the ultimate limit of this effect, making it critical to have a fundamental understanding of the optical response of 2D polar materials~\cite{thygesen2017calculating}.
-Concerning the optical response, the transition from three-dimensional (3D) to two-dimensional (2D) polar materials is nontrivial, however, since in a polar monolayer, the LO--TO splitting that gives rise to phonon polaritons in 3D is absent at the $\Gamma$ point~\cite{sanchez2002vibrational,mele2002electric,serrano2007vibrational,sohier2017breakdown}.
-This raises a fundamental question about the nature of  electromagnetic modes in polar monolayers.
-
-We now develop a framework with quantitative, concrete examples that enables a first-principles understanding of phonon polaritons in 2D materials. 
-We do so by deriving a universal form for the conductivity of a polar monolayer,  which depends solely on the LO and TO phonon frequencies---and their dispersion with momentum---in the 2D system.
-Using parameters from Ref.~\citenum{sohier2017breakdown} for the canonical 2D polar monolayer---hexagonal boron nitride (hBN)---we present the confinement and propagation losses of the 2D phonon polariton modes, identifying the frequency region where they should be most easily detected.
-Finally, we find that these modes enable extreme light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced by up to eight orders of magnitude through the emitter--LO phonon coupling.
-For an infrared emitter with a sufficiently high free-space radiative decay rate (${\gtrsim}\, \SI{1e6}{\per\s}$), we find that the associated linewidth of the emitter is comparable to the spacing between different phonon polaritonic resonances of an hBN nanostructure.
-This suggests the possibility of realizing the multi-mode strong coupling and ultra-strong coupling regimes of quantum electrodynamics in a 2D hBN platform.
-Our results for hBN are particularly relevant due its widespread use in 2D van der Waals heterostructures.
-In addition to providing functionality as a layer which improves the electrical and optical properties of other 2D materials, \eg graphene, our results suggest that in these heterostructures, hBN layers could provide a mid-infrared platform for nanophotonics and quantum optics.
-While we focus on hBN in this manuscript, the salient features of our findings apply to other polar monolayers as well.
-
-\section{Optical response of optical phonons in two-dimensions}
-In this section, we develop a theory of electromagnetic response due to optical phonons in 2D systems.
-The key response function of interest is the conductivity of the monolayer.
-To that end, we consider the response of the ions of the monolayer due to an electric potential $\phi$.
-For that case, the interaction Hamiltonian is
-\begin{equation}
-    H_{\mathrm{int}} = \int \dd^2x ~\rho \phi = -\int \dd^2x~ (\nabla\cdot\mathbf{P})\phi,
-    \label{eq:coupling}
-\end{equation} 
-with $\rho$ the induced charge density and $\mathbf{P}$ the induced polarization density associated with the ionic motion.
-Note that boldfaced quantities refer to vectors or tensors as appropriate.
-Within linear response theory, the polarization density can be straightforwardly evaluated from the displacement  $\mathbf{u}_{\kappa}$ of every atom $\kappa$ within the unit cell.
-Specifically, to first order, the polarization density is
-\begin{equation}
-\mathbf{P} - \mathbf{P}^0 = \sum_\kappa(\mathbf{u}_\kappa\cdot\boldsymbol{\nabla}_{\mathbf{u}_\kappa})\mathbf{P} \equiv \frac{1}{\Omega}\sum_\kappa\mathbf{Z}_{\kappa}\mathbf{u}_{\kappa},
-\end{equation} 
-where $\mathbf{Z}_\kappa \equiv \Omega\boldsymbol{\nabla}_{\mathbf{u}_\kappa}\mathbf{P}$ is the Born effective charge tensor of ion $\kappa$ and $\Omega$ is the unit cell area.
-$\mathbf{P}^0$ is the equilibrium polarization in the absence of displacements, which is zero here.
-With this relation between polarization and ionic displacements, the interaction Hamiltonian in Equation~\eqref{eq:coupling} couples the scalar potential and the ionic displacements.
-We consider the response of the monolayer to a potential of the form $\phi(\mathbf{r}) = \phi(\mathbf{q},\omega)\e^{\iu\mathbf{q}\cdot\mathbf{r}-\iu\omega t}$, where $\mathbf{q}$ is a 2D wavevector in the plane of the monolayer.
-Such a potential corresponds to a longitudinal electric field  $\mathbf{E}(\mathbf{r}) = \iu\mathbf{q}\phi(\mathbf{r})$.
-
-In what follows, we assume the validity of the random-phase approximation (RPA) in calculating Coulombic interactions between ions in the polar lattice.
-Within the RPA, these Coulombic interactions are accounted for by taking the induced polarization $\mathbf{P}(\mathbf{q},\omega)$ to be proportional \emph{total} electric field, $\mathbf{E}_{\mathrm{tot}}(\mathbf{q},\omega)$, defined to be the sum of the externally applied electric field and the electric field created by the induced polarization.
-The polarization and total field are connected by the polarization-polarization response (tensor) function  $\boldsymbol{\Pi}(\mathbf{q},\omega)$ via
-\begin{equation}
-\mathbf{P}(\mathbf{q},\omega)  = \epsilon_0\boldsymbol{\Pi}(\mathbf{q},\omega)\mathbf{E}_{\mathrm{tot}}(\mathbf{q},\omega),
-\end{equation}
-The polarization-polarization response function is related to the conductivity via the relation $\boldsymbol{\sigma}(\mathbf{q},\omega) = -\iu\omega\epsilon_0\boldsymbol{\Pi}(\mathbf{q},\omega)$.
-From the Kubo formula, it follows that the conductivity is
-\begin{equation}\label{eq:2dsusceptibility}
-    \boldsymbol{\sigma}(\mathbf{q},\omega) =  \frac{-\iu \omega}{ \Omega\mathcal{Z}}\sum\limits_{m,n}\frac{\mathbf{P}_{mn}(\mathbf{q})\otimes\mathbf{P}_{nm}(\mathbf{q})}{\hbar\omega + E_{nm}+\iu 0^+}\Big(\e^{-\beta E_m}-\e^{-\beta E_n} \Big),
-\end{equation}
-where $m,n$ are eigenstates in the phononic Fock space of the monolayer, $\mathbf{P}_{mn}(\mathbf{q}) \equiv \sum_\kappa \mathbf{Z}_\kappa\langle m|\mathbf{u}_\kappa(\mathbf{q})|n\rangle$ are matrix elements of the polarization associated with phonon modes (where $\mathbf{u}_\kappa(\mathbf{q})$ is the Fourier transform of the phonon displacement operator), $E_{m}$ ($E_n$) is the energy of state $m$ ($n$), $\beta \equiv 1/k_{\text{\textsc{B}}}T$ is the inverse temperature, and $\mathcal{Z}$ is the grand partition function.
-We now evaluate the contribution of optical phonons to the polarization-polarization response in the low-temperature limit $T \ll \hbar\omega_{\mathrm{ph}}/k_{\mathrm{B}}$ with $\omega_{\mathrm{ph}}$ a characteristic optical phonon frequency.
-Considering the long-wavelength (small wavevector) limit, and taking a material with long-wavelength isotropy, such as hBN, we only have to consider the $qq$-component in the response tensor, where $qq$ denotes a pair of directions parallel to the wavevector.
-Denoting $\sigma_{qq}$ as simply $\sigma$, we find that the conductivity is given by~\cite{rivera2018ab}:
-\begin{equation}
-    \sigma(\mathbf{q},\omega) = \frac{-\iu\omega}{\hbar\Omega} \frac{2\omega_{\mathbf{q},\mathrm{L}}}{\omega^2_{\mathbf{q},\mathrm{L}}-\omega^2-\iu\omega\tau^{-1}}|\hat{\mathbf{q}}\cdot\langle 1_{\mathbf{q},\mathrm{L}}|\mathbf{P}(\mathbf{q})|0_{\mathbf{q},L}\rangle|^2,
-    \label{eq:polpolresponse}
-\end{equation}
-where L-subscripts denote longitudinal polarization, $|0_{\mathbf{q},\mathrm{L}}\rangle$ ($|1_{\mathbf{q},\mathrm{L}}\rangle$) denotes a state with no (one) longitudinal phonon of wavevector $\mathbf{q}$, and $\hat{\mathbf{q}}$ denotes a unit vector in the direction of $\mathbf{q}$.
-We have also phenomenologically included the phonon dissipation rate $\tau^{-1}$, consistently with  a relaxation-time prescription.
-The frequency $\omega_{\mathbf{q},\mathrm{L}}$ in the denominator, as in the case of bulk phonons, is the frequency of the longitudinal phonon of wavevector $\mathbf{q}$ prior to considering LO--TO splitting~\cite{BornHuang:1954} (and near the $\Gamma$ point is approximately equal to the TO phonon frequency).
-This is consistent with the fact that LO--TO splitting is a collective effect arising from Coulomb interactions and the fact that the equation above represents a single-particle susceptibility.
-Coulomb interactions are accounted for in the random phase approximation, and to include them in the single-particle response amounts to an uncontrolled double-counting.
-
-Next, we express the polarization matrix element in Equation~\eqref{eq:polpolresponse} in terms of the Born effective charges of the monolayer and the phonon displacement eigenvectors.
-Considering the longitudinal phonon contribution to the second-quantized ionic displacement, as in Ref.~\citenum{srivastava1990physics}, we find that the conductivity within the RPA is given by
-\begin{equation}
-    \sigma(\mathbf{q},\omega) = -\frac{\iu\omega }{\Omega}\frac{\Big|\hat{\mathbf{q}}\cdot\sum\limits_{\kappa}\mathbf{Z}_{\kappa}\boldsymbol{\eta}_{\kappa} \Big|^2}{\omega^2_{\mathbf{q},\mathrm{L}}-\omega^2-\iu\omega\tau^{-1}}.
-    \label{eq:conductivity_simplified}
-\end{equation} 
-We have defined scaled eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}\equiv \frac{\hat{\mathbf{e}}_{\kappa\mathbf{q},\mathrm{L}}}{\sqrt{M_{\kappa}}}$, where $\hat{\mathbf{e}}_{\kappa\mathbf{q},\mathrm{L}}$ is the unit-normalized polarization vector of atom $\kappa$ in the unit cell oscillating according to a longitudinal phonon of wavevector $\mathbf{q}$ and $M_{\kappa}$ is the mass of atom $\kappa$.
-
-While the conductivity is the main electromagnetic quantity of interest for electrodynamics applications, we briefly state the form of the (2D) permittivity, as its zeros immediately yield the longitudinal modes of the system, which are the LO phonons.
-The permittivity within the RPA, denoted $\epsilon_{\textrm{RPA}}$ is related to the polarization-polarization response function via~\cite{jablan2009plasmonics} $\epsilon_{\textrm{RPA}} = \epsilon_{\mathrm{env}} + \frac{1}{2}q\Pi(q,\omega)$ and the conductivity via $\epsilon_{\textrm{RPA}} = \epsilon_{\mathrm{env}} + \frac{\iu q\sigma(q,\omega)}{2\omega\epsilon_0}$.
-Here, $\epsilon_{\mathrm{env}}$ is the average permittivity of the bulk above and below the monolayer, and is added to take into account the polarization arising from these bulk materials.
-Note that we have neglected any intrinsic high-frequency screening in the monolayer itself, which is only relevant for wavevectors comparable to the inverse layer spacing between monolayers.
-When considering non-local corrections to the conductivity at these large wavevectors, these must be taken into account~\cite{sohier2016two,thygesen2017calculating,sohier2017breakdown}.
-Based on Equation~\eqref{eq:conductivity_simplified}, the zeros $\omega_{\mathbf{q}}$ of the RPA dielectric function satisfy:
-\begin{equation}
-    \omega^2_{\mathbf{q}} - \omega^2_{\mathrm{TO}} = \frac{V(q)}{e^2}\frac{1}{\Omega}q^2\Big|\hat{\mathbf{q}}\cdot\sum\limits_{\kappa}\mathbf{Z}_{\kappa}\boldsymbol{\eta}_{\kappa}\Big|^2,
-    \label{eq:zerosofresponse}
-\end{equation}
-  where $V(q)$ is the Coulomb interaction in Fourier space, which in two dimensions, is given by $V(q) = \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}$.
-Given that the zeros of the dielectric function are associated with longitudinal modes, one expects that $\omega_{\mathbf{q}}$ is in fact the frequency of the LO phonon mode.
-This is consistent with the result of Ref.~\citenum{sohier2017breakdown}, in which \relax{{}\ }\citeauthor{sohier2017breakdown} show that in 2D polar materials, the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO--TO splitting and zero LO--TO splitting at the $\Gamma$ point of the Brillouin zone.
-
-Given these results, we now re-express the conductivity explicitly in terms of the 2D phonon dispersion, and derive a universal form for the conductivity in the local ($q\rightarrow 0$) limit specified in terms of three parameters: the LO phonon frequency at the $\Gamma$ point (i.e., $\omega_{\mathrm{TO}}$), the group velocity of the LO phonon at the $\Gamma$ point, and the damping rate.
-From Equation~\eqref{eq:zerosofresponse}, we can immediately write the conductivity as
-  \begin{equation}
-  \sigma(\mathbf{q},\omega) = \frac{-2\iu\epsilon_0\epsilon_{\mathrm{env}}\omega}{q}\frac{\omega^2_{\mathbf{q},\mathrm{LO}}-\omega^2_{\mathrm{TO}}}{\omega^2_{\mathrm{TO}}-\omega^2-\iu\omega\tau^{-1}}.
-  \end{equation}
- In this expression, $\omega_\mathrm{L}$, the LO phonon frequency prior to LO--TO splitting, has been re-named as $\omega_{\mathrm{TO}}$, the transverse optical phonon frequency, because in the absence of LO--TO splitting, they are degenerate.
-The RPA zeros $\omega_{\mathbf{q}}$ have also been renamed as $\omega_{\mathbf{q},\mathrm{LO}}$.
-This is done in order to make the form of the final results more closely resemble their 3D counterparts, in which the dielectric function is expressed in terms of the TO frequency (see for example Eq.
-(\ref{eq:3deps})).
-
- For small $q$,  the Born charges are (to lowest-order) constant, and so the LO phonon dispersion takes the form $\omega_{\mathbf{q},\mathrm{LO}}=\sqrt{\omega^2_{\mathrm{TO}}+2v_{\mathrm{g}}\omega_{\mathrm{TO}}q} \simeq \omega_{\mathrm{TO}} + v_{\mathrm{g}}q$, where $v_{\mathrm{g}}$, the LO phonon group velocity, is defined from microscopic parameters through the relation
- \begin{equation}
-    v_{\mathrm{g}} = \frac{\Big|\hat{\mathbf{q}}\cdot \sum_\kappa \mathbf{Z}_\kappa\boldsymbol{\eta}_\kappa\Big|^2 }{4\epsilon_0 \epsilon_{\mathrm{env}}\omega_{\mathrm{TO}}\Omega}. 
-\end{equation}
-Thus, in the long wavelength limit, we have the following universal parameterization of the conductivity of a polar monolayer:
-\begin{equation}
-    \sigma(\omega) =  \frac{-4\iu\epsilon_0\epsilon_{\mathrm{env}}\omega\omega_{\mathrm{TO}}v_{\mathrm{g}}}{\omega^2_{\mathrm{TO}}-\omega^2-\iu\omega\tau^{-1}}.
-    \label{eq:localsigma}
-\end{equation}
-
-We note that despite its appearance, $\sigma(\omega)$ does not depend on $\epsilon_{\mathrm{env}}$, as $v_{\mathrm{g}}$ has an opposite dependence on $\epsilon_{\mathrm{env}}$.
-From this relation, it follows that given the properties of the 2D phonons (from experiments or from \emph{ab initio} calculations), one can immediately specify the conductivity.
-Alternatively, from optical measurements (including far-field measurements) which allow one to extract the conductivity, it becomes possible to extract the group velocity of 2D LO phonons and thus the small-wavevector dispersion of those phonons.
-
-Before moving on to analyze the electrodynamics of 2D phonon polaritons, we make three comments on lack of LO--TO splitting in 2D polar materials.
-The first is that this situation is in stark contrast to the situation of polar materials in 3D, which have a finite LO--TO splitting at the $\Gamma$ point.
-In the absence of such LO--TO splitting in 3D, there would be no frequency compatible with the existence of a phonon  polariton.
-On the contrary, we will show that in 2D, despite the absence of LO--TO splitting at the $\Gamma$ point, there persists a strongly confined evanescent electromagnetic mode with a high local density of states which in all respects is similar to a phonon polariton of a thin film, but is in fact the 2D LO phonon of the polar monolayer (thus the phrases ``phonon polariton'' and ``2D LO phonon'' may be used somewhat interchangeably as is the case in plasmonics where the terms ``plasmon polariton'' and ``2D plasmon'' are often used interchangeably).
-
-The second comment is that much of what has been discussed here has a strong analogy with the theory of optical response in electron gases in 2D, and particularly the relation between plasmons in 2D and 3D.
-To elaborate on this analogy, we take Equation~\eqref{eq:zerosofresponse} in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $\Big|\hat{\mathbf{q}}\cdot \sum_\kappa \mathbf{Z}_\kappa\boldsymbol{\eta}_\kappa\Big|^2 \equiv \frac{Q_*^2}{M_*}$, with $Q_*$ being an effective charge and $M_*$ being an effective mass.
-Then, the LO--TO splitting can be written as $\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} = \frac{Q_*^2}{2\epsilon_0\epsilon_{\mathrm{env}} M_*}q$.
-Now we note that the RHS is exactly the squared-frequency $\omega^2_{\mathbf{q}\mathrm{p}}$ for a plasma oscillation in a 2D gas of charged particles with charge $Q_*$ and mass $M_*$.
-To connect to LO--TO splitting in phonons, this squared frequency $\omega^2_{\mathbf{q}\mathrm{p}}$ can be thought of as the ``LP--TP'' splitting between longitudinal and transverse plasma oscillations.
-Of course, there are no transverse plasma oscillations due to the structure-less nature of the electron, and so ``$\omega_{\mathrm{TP}}$''  should be considered equal to zero.
-In the three-dimensional plasmon case, ``$\omega_{\mathrm{TP}} = 0$'', but the difference between the squared longitudinal and transverse plasma oscillation frequencies at zero-wavevector is non-zero and given by $\omega_{\mathrm{p}}^2$.
-In other words, the plasma frequency in electron gases is analogous to the LO--TO splitting in polar materials.
-The change in the dispersion of 3D versus 2D plasmons, like 3D versus 2D LO phonons, arises from the change in dimensionality of the Coulomb interaction from 3D to 2D.
-This analogy between phononic and plasmonic behavior as a function of dimension is illustrated schematically in Figures~\ref{fig:1}(b,c) in order to help unify the understanding of 2D plasmonics and phononics.
-These considerations should also extend to the one-dimensional case.
-
-\addition{%
-The third comment is that the considerations of this section can be extended to few-layer systems, such as hBN bilayers, which may still be considered 2D for polariton wavelengths large compared to the bilayer thickness.
-The considerations may also be extended to 1D polar materials, such as BN nanotubes.
-For bilayers, as was discussed in Ref.~\citenum{sohier2017breakdown}, one may describe the LO--TO splitting in terms of Eq.~\eqref{eq:zerosofresponse}, with the 2D Coulomb interaction, but with eigendisplacements appropriate to the few layer system.
-For a small number of layers $N$, \FIXME{Mention scaling of $|\hat{\mathbf{q}}\cdot\ldots|^2 = NQ_*^2/M_*$} this should enhance the LO--TO splitting by a factor of $N$. 
-To go beyond an effectively 2D description of a monolayer, and make the transition from monolayer to bulk, we may make use of the quasi-2.5D Coulomb interaction~\cite{thygesen2017calculating} in Eq.~\eqref{eq:zerosofresponse}.
-In that case, for a film of effective thickness $t$, the interaction is approximately $V_{\text{Q2.5D}}(q) = \frac{e^2}{\epsilon_0 tq^2}\big(1-\frac{2}{qt}e^{-\frac{qt}{2}}\sinh\frac{qt}{2} \big)$. 
-This form interpolates between the 2D and 3D interactions in the limits $qt \rightarrow 0$ and $qt \rightarrow \infty$. 
-For sufficiently large thickness and finite-$q$, the 2.5D Coulomb interaction scales as $t^{-1}q^{-2}$. 
-Noting the $q^2$ factor in Eq.~\eqref{eq:zerosofresponse}, as well as the $N$-dependence of the quantity involving eigendisplacements, one finds an LO--TO splitting independent of thickness, as expected for bulk samples.
+\begin{figure*}[h]
+    \includegraphics[width=0.6\linewidth]{enhancement.png}
+    \caption{%
+        \textbf{Spontaneous emission enhancement in hBN thin films and monolayer.} Spontaneous emission enhancement for a dipole emitter 5 nm away from the surface of thin films of bulk hBN as a function of emission frequency for film thicknesses of 100 nm, 10 nm, 1 nm, and 0.33 nm, as well as monolayer hBN with dispersion relation and conductivity given in the main text.
+The hBN in all cases is surrounded by air.
+The TO frequency  taken for the bulk is 1387 cm$^{-1}$.
+        \label{fig:s1}
         }
-
-\addition{%
-For 1D systems, beyond using the eigendisplacements appropriate to the 1D system, one must also use the 1D Coulomb interaction, cutoff by a length $R$ which for a nanotube, would be its radius. 
-In that case, the 1D Coulomb interaction is $e^2K_0(qr)/2\pi\epsilon_0$, where $K_0$ is the (zeroth order) modified Bessel function of the second kind. 
-Thus, the $q$-dependence of the LO--TO splitting is $q^2K_0(qR)$.
-This is in close analogy with the relation for the plasma frequency of 1D plasmons, as expected from the considerations of the previous paragraph.
-As regards the optical properties, the optical conductivity is still governed by Eq.~\eqref{eq:2dsusceptibility}.
-}
-
-\section{Electrodynamics of optical phonons in two-dimensions}
-To relate the conductivity function  to the electromagnetic modes supported by a polar monolayer, we solve Maxwell's equations for an evanescent electromagnetic mode supported by a surface with conductivity $\sigma$.
-We consider the monolayer to be sandwiched by a superstrate of permittivity $\epsilon_+$ and a substrate of permittivity $\epsilon_-$.
-To strip the analysis to its bare essentials, we consider optical phonon response with in-plane isotropy in the long-wavelength limit arising from in-plane LO oscillations.
-A relevant example of a system where these conditions are satisfied is in a hexagonal boron nitride monolayer (see Figure~\ref{fig:1}(a) for schematic atomic structure).
-In a monolayer geometry with translation invariance and in-plane isotropy, the solutions of Maxwell's equations can be decomposed into transverse magnetic (TM) and transverse electric (TE) parts, where the magnetic or electric field respectively is transverse to the in-plane wavevector of the mode.
-In practice, it is the TM mode which is associated with highly confined electromagnetic waves.
-We consider without loss of generality a TM mode with wavenumber $q$ along the $x$-direction in the monolayer and magnetic field $H(z)\e^{\iu qx-\iu\omega t}$ along the $y$-direction of the monolayer.
-The direction transverse to the monolayer is denoted as $z$.
-With these definitions in place, the Maxwell equation satisfied by the magnetic field is
-\begin{equation}\label{eq:2dmaxwell}
-    \bigg(-\frac{\dd^2}{\dd{}z^2}+q^2-\epsilon_{\pm}\frac{\omega^2}{c^2} \bigg)H(z) = 0,
-\end{equation}
-where $\epsilon_+$ applies for $z>0$ and $\epsilon_-$ applies for $z<0$.
-We consider a solution of the form $H(z) = h_\pm\e^{\mp\kappa_\pm z}$ with $\kappa_\pm = \sqrt{q^2-\epsilon_{\pm} \frac{\omega^2}{c^2}}$ with $\pm$ corresponding to $\pm z > 0$ respectively.
-The boundary condition on the magnetic field is $h_+ - h_- = -K_x = -\sigma E_x$ where $\mathbf{K}$ is the surface current density, and $\mathbf{E} = -\frac{1}{i\omega\epsilon}\nabla\times\mathbf{H}$ is the electric field.
-This condition enforces $h_+-h_- = \frac{\sigma}{i\omega\epsilon_+}\kappa_+h_+$.
-Continuity of the electric field in the $x$ direction enforces ${\epsilon_-}/{\epsilon_+} = -{\kappa_-h_-}/{\kappa_+h_+}$. 
-Combining the two conditions, we obtain the usual dispersion equation for the TM mode of a polarizable 2D monolayer, namely $\epsilon_+/\kappa_+ + \epsilon_-/\kappa_- = \frac{\sigma}{i\omega}$.
-Given the deeply subwavelength nature of 2D phonon polaritons, \ie since $q\gg\omega/c$ such that $\kappa_{\pm}\simeq q$, the dispersion equation can be reduced to its quasistatic limit without consequential loss of accuracy:
-\begin{equation}
- q=\frac{2i\omega\epsilon_0\epsilon_{\mathrm{env}}}{\sigma},
-    \label{eq:dispeq}
-\end{equation}
-with $\epsilon_{\mathrm{env}}\equiv (\epsilon_++\epsilon_-)/2$.
-This condition, as can be seen from the relation between $\Pi$ and $\sigma$, is equivalent to $\epsilon_{\mathrm{RPA}}=0$.
-Thus, the condition for phonon polaritons coincides precisely with the condition for longitudinal optical phonons.
-One of the main results of our manuscript is that despite the lack of LO--TO splitting at the $\Gamma$ point, there nevertheless exists a strongly confined evanescent mode in many respects similar to the phonon polaritons of thin films.
-We now analyze the dispersion relation of phonon polaritons in a specific material, hexagonal boron nitride, in more detail, showing the possibility of highly confined electromagnetic modes with a large local density of states.
-
-\begin{figure}[t]
-    \includegraphics[width=0.95\linewidth]{fig1.pdf}
-    \caption{%
-        \textbf{LO phonons as the basic electromagnetic waves of a polar monolayer.} (a) Schematic structure of a polar monolayer such as hexagonal boron nitride.
-(b) Properties of LO and TO phonons in 3D and 2D.
-In 3D, there is a finite LO--TO splitting at zero wavevector, while in 2D there is none.
-Despite this, the 2D LO phonon plays the role of the phonon polariton in 3D and thin films.
-(c) Analogous physics appears in electron gases in 3D and 2D, where the 3D plasma frequency is similar to the 3D LO--TO splitting.
-In 2D, the plasma frequency at zero wavevector is zero, but the electromagnetic physics is determined by the dispersion of 2D plasmons, which replace the plasmon polariton of bulk and thin films.
-        \label{fig:1}
-        }
-\end{figure}~
-
-\begin{figure*}[t]
-    \includegraphics[width=.7\linewidth]{figure2hBN.pdf}
-    \caption{%
-        \textbf{Properties of phonon polaritons in a monolayer of hBN compared to bulk.} (a)~Real and imaginary parts of the conductivity of 2D hBN for different values of the loss-rate.
-(b)~Propagation quality factor, which measures the number of wavelengths of propagation of the 2D phonon polariton.
-(c)~Dispersion relation of phonon polaritons in the monolayer (black) and thin films whose thicknesses are taken to be 1, 2, 4, and 8 interlayer spacings in hBN.
-For the thin-film, the fundamental mode (dashed) and the first higher-order mode (dash-dotted) are plotted.
-        \label{fig:2}
-        }
-\end{figure*}
-
-In Figure ~\ref{fig:2}(a), we present the conductivity of 2D hBN of Equation~\ref{eq:localsigma}, using parameters from Ref.~\citenum{sohier2017breakdown} calculated from density functional theory within the local density approximation.
-In this plot $\epsilon_{\mathrm{env}}=1,$ $\omega_{\mathrm{TO}} = $ \SI{1387}{\per\cm} and $v_{\mathrm{g}} = 1.2\times 10^{-4}c$, with $c$ the speed of light in vacuum.
-
-From the conductivity, the dispersion relation of phonon polaritons on an infinite sheet is given by $q={2\iu\omega\epsilon_0\epsilon_{\mathrm{env}}}/{\sigma(\omega)}$.
-The dispersion, assuming $\tau = \infty$ and $\epsilon_{\mathrm{env}}=1$, is shown in Figure~\ref{fig:2}(c) (black line).
-A key figure of merit for applications involving the propagation of phonon polaritons, is the propagation quality factor, defined by ${\Re q}/{\Im q} = {\Im\sigma(\omega)}/{\Re\sigma(\omega)}$, which is shown in Figure~\ref{fig:2}(b).
-
-For monolayer hBN, the wavevector grows very rapidly with frequency, due to the extremely low group velocity of 2D LO phonons, which is a remarkable four orders of magnitude slower than the speed of light.
-In particular, at frequencies of \SI{1450}{\per\cm}, the phonon polariton has a wavelength of about \SI{15}{\nm}, significantly shorter than any phonon polariton measured so far, and, similarly shorter than any plasmonic wavelength, even in graphene.
-In fact, this short a wavelength well-below that of any polariton in current scattering near-field microscopy (SNOM) measurements.
-The 2D phonon polariton could in principle be measured by SNOM closer to the TO frequency, where confinement is smaller; unfortunately, as shown in Figure~\ref{fig:2}(b), near the TO frequency, dissipation is far higher (and corresponding propagation quality factors $\Re q/\Im q$ far lower) due to large $\Re\sigma$ (or, equivalently, large $\Im\epsilon_{\mathrm{RPA}}$).
-
-These considerations imply that access to the lower-loss and higher-confined portions of the dispersion relation of phonon polaritons, in the absence of a sharper tip, requires a near-field probe such as a free electron probe, as used in electron energy loss spectroscopy (EELS), where slow electrons can be used to probe plasmon wavelengths of just a few nanometers in monolayer metals, as well as the nonlocal bulk plasmon dispersion in metals~\cite{nagao2001dispersion,de2010optical,diaconescu2007low}.
-EELS has been recently employed to measure phonon polaritons in ultrathin films of hBN~\cite{govyadinov2017probing}.
-Another interesting class of near-field probes, with relevance to fundamental physics and quantum optics applications, is a quantum emitter such as an atom, molecule, or artificial atom such as a quantum dot, quantum well, or vacancy center.
-Recently, it was demonstrated using nanostructures of bulk hBN that the interaction of vibrational emitters with phonon polaritons is on the border of the strong coupling regime~\cite{autore2018boron}.
-
-In the rest of this section, we discuss the relation between the dispersion of an hBN monolayer versus the atomically-thick limit of a thin film of a material with hBN's bulk dielectric function.
-To aid this discussion, in Figure~\ref{fig:2}(c), we show the dispersion relation of thin films of bulk hBN with film thicknesses of 1, 2, 4, and 8 times the interlayer spacing of bulk hBN, which is roughly \SI{0.33}{\nm}.
-For these plots, we take hBN to be cleaved such that the optical axis is perpendicular to the plane of the film.
-The components of the bulk permittivity perpendicular and parallel to the c-axis ($\epsilon_{\perp}$ and $\epsilon_{\parallel}$, respectively; indexed by $\alpha\in\{\perp,\parallel\}$ below) are then given by
-\begin{equation}
-	\epsilon_{\alpha}(\omega) = \epsilon_{\infty,\alpha}\Bigg(1 + \frac{\omega^2_{\mathrm{LO},\alpha}-\omega^2_{\mathrm{TO},\alpha}}{\omega^2_{\mathrm{TO},\alpha}-\omega^2} \Bigg),
-	\label{eq:3deps}
-\end{equation}
-with $\epsilon_{\infty,\parallel} = \num{2.95}$, $\omega_{\mathrm{TO},\parallel} = \SI{760}{\per\cm}$, and $\omega_{\mathrm{LO},\parallel} = \SI{830}{\per\cm}$; and $\epsilon_{\infty,\perp} = \num{4.87}$, $\omega_{\mathrm{TO},\perp} = \SI{1360}{\per\cm}$, and $\omega_{\mathrm{LO},\perp} = \SI{1614}{\per\cm}$\cite{caldwell2014sub,dai2014tunable}.
-Losses are ignored in this discussion altogether. 
-In the range between $\omega_{\mathrm{TO}, \alpha}$ and $\omega_{\mathrm{LO}, \alpha}$, the corresponding component of the permittivity is negative, while the other component is positive.
-This hyperbolicity leads to a dispersion for hBN thin films that have multiple branches at a given frequency, as can be seen in Figure~\ref{fig:2}(c).
-This trend persists even when the thickness of the bulk is taken down to a single layer, albeit pushed to high wavevectors.
-This is in contrast with the true monolayer, where there is only one LO phonon mode.
-Given that the phonon polariton of 2D is the LO phonon, there can be only one branch of the dispersion.
-\addition{%
-The bulk dielectric function of Eq.~\eqref{eq:3deps} fails to capture this removal of higher-order modes (\ie hyperbolicity).
-	\comment{tc}{But, na\"ively, I guess this is to be expected, since \eqref{eq:3deps} is explicitly hyperbolic?}
-For a bilayer however, there is a higher-order LO mode, associated with out-of-phase oscillation between the two layers, and thus one expects to see a higher-order mode. 
-However, it is expected challenging to observe such a mode, as can be anticipated from the blue higher-order mode in Figure~\ref{fig:2}(c), because even for a polariton wavelength of \SI{12}{\nm}, which is very short, the polariton frequency is only a few \si{\per\cm} above the TO frequency, within the damping linewidth of the TO frequency.
-}
-
-	\comment{tc}{Isn't this a bit strange to discuss, given that our results do not really show this (since we are sticking to the local case? This is only a feature in \citet{sohier2016two} results?}
-Another anomaly between 2D and 3D is that the dispersion relation of the 2D LO phonon is non-monotonic, and the frequency starts to decrease for sufficiently large wavevector.
-This can never happen by extrapolating to a monolayer a dielectric function which has finite LO--TO splitting.
-This turn-back point corresponds to a frequency of around \SI{1500}{\per\cm} and a phonon wavelength of less than \SI{1}{\nm}.
-Thus, this deviation is challenging to probe in typical optical experiments.
-However it means that for the monolayer, the range of accessible phonon polaritonic frequencies for optical applications is reduced compared to thick films.
-Finally, while we have focused on anomalies between the atomically-thin limit of bulk and a true monolayer, a comparison of the monolayer with the fundamental phonon polariton mode of the one-atom-thick thin film suggests that we have a reasonable qualitative understanding of the monolayer dispersion from this limit.
-%
-\addition{%
-This qualitative similarity is to be somewhat expected, as a 2D layer can be considered as a very thin film of bulk material, provided that the microscopic properties of the bulk and monolayer do not deviate substantially.
-The differences between the $t = t_{\text{2D}} = \SI{0.33}{\nm}$ bulky thin-film and monolayer that nevertheless do exist, reflect such microscopic deviations.
-Occasionally, \eg for compatibility with standard numerical tools, it is useful to introduce a fictitious, effective bulk material whose fundamental thin-film mode \emph{exactly} matches the monolayer's.
-The dielectric function of such a ``bulkified'' monolayer is thickness-dependent and anisotropic: for a film in the $xy$ plane, it is $\boldsymbol{\epsilon}_{\text{Q2D}} = (\hat{\mathbf{x}}\hat{\mathbf{x}} + \hat{\mathbf{y}}\hat{\mathbf{y}})\epsilon_{\text{Q2D}, \parallel} + \hat{\mathbf{z}}\hat{\mathbf{z}}\epsilon_{\text{Q2D}, \perp}$ with $\epsilon_{\text{Q2D}, \perp} = 1$ and
-\begin{equation}
-	\epsilon_{\text{Q2D}, \parallel}(\omega) = 1 + \frac{1}{t}\frac{4\epsilon_{\mathrm{env}}\omega_{\mathrm{TO}}v_\mathrm{g}}{\omega^2-\omega_{\mathrm{TO}}^2+\iu\omega\tau^{-1}}.
-\end{equation}
-}
-
-\begin{figure}[t]
-    \includegraphics[width=.925\linewidth]{fig3.pdf}
-    \caption{%
-        \textbf{Extreme spontaneous emission enhancement due to 2D phonon polaritons in nanostructured geometries.} Plotted is the enhancement of the spontaneous emission rate for an emitter  $z=\SI{5}{\nm}$ above the disk's center and  oriented normal to its plane ($\hat{\mathbf{n}} = \hat{\mathbf{z}}$).
-For a disk with a diameter of \SI{20}{\nm}, and assuming a relaxation rate $\tau^{-1} = \SI{0.5}{\meV}$, the rate of emission enhancement can be enhanced \num{100} million-fold.
-For an emitter with a free-space decay rate of $\SI{1e6}{\per\cm}$ at \SI{7}{\micro\m}, the emitter would experience a decay rate comparable to the frequency of the disk mode, leading to ultra-strong coupling of an external emitter with 2D phonon polaritons.
-For reference, we compare the spontaneous emission enhancement in a nanodisk geometry to that of a disk of infinite radius (i.e., a flat sheet), showing clear enhancement relative to the flat sheet due to concentration of local density of states around phonon polariton resonances.
-Note that $\tau^{-1} = \SI{0.5}{\meV}$ is of the order of the loss rate in bulk hBN.
-Also shown in the figure are maps of the electric potential on the surface of the disk for modes corresponding to selected phonon polariton resonances in the plot.
-        \label{fig:3}
-        }
-\end{figure}
-
-\section{Strong light-matter interactions enabled by 2D optical phonons}
-
-The extreme confinement of electromagnetic fields offered by the 2D phonon polariton presents an opportunity for quantum optical applications in which one seeks to couple an external emitter such as an atom, molecule, defect, or artificial atomic system to electromagnetic fields.
-Applications of these couplings are ultra-bright single- or two-photon sources, realizing the strong-coupling regime and the associated phenomenology of Rabi oscillations and polaritons, or resolving spectroscopically ``forbidden'' transitions~\cite{koppens2011graphene,autore2018boron,rivera2016shrinking,rivera2017making,kurman2018control} to achieve near-field spectroscopies with momentum and angular momentum resolution not accessible in far-field spectroscopies.
-
-In Figure~\ref{fig:3}, we consider the coupling of a dipole emitter to localized phonon polaritons of nanostructured monolayer hBN.
-For simplicity, we consider hBN nanostructured as a disk, which leads to the formation of sharp resonances quantized along the azimuthal and radial directions.
-The disk is taken to have a radii varying from \SI{10}{\nm} to \SI{40}{nm}
-and a loss rate $\tau^{-1} = \SI{0.5}{\meV}$, which is of the order of the loss rate in bulk hBN~\cite{caldwell2014sub, dai2014tunable}.
-We also show (dashed line) the results for a disk of infinite radius, i.e., a flat sheet of monolayer hBN.
-We parameterize the coupling between the dipole and phonon polaritons through the rate of spontaneous emission $\Gamma$ of phonon polaritons by the dipole, normalized to the rate of spontaneous emission in free space $\Gamma_0$.
-It is related to the dyadic Green function $\mathbf{G}$ of the Maxwell equations for the nanostructure via the expression~\cite{novotny2012principles}:
-\begin{equation}
-	\frac{\Gamma}{\Gamma_0} 
-	=  \frac{6\pi c}{\omega}\hat{\mathbf{n}}\cdot\Im\mathbf{G}(\mathbf{r},\mathbf{r},\omega)\cdot\hat{\mathbf{n}},
-\end{equation}
-where $\hat{\mathbf{n}}$ is the orientation of the dipole, $\mathbf{r}$ its position, and $\omega$ its frequency.
-
-The dyadic Green function is computed using a quasistatic boundary element method (as in Ref.~\citenum{jin2017infrared}).
-In Figure~\ref{fig:3}, we plot the enhancement of the spontaneous emission rate $\Gamma/\Gamma_0$ for an external emitter polarized perpendicularly to the plane of the disk and placed \SI{5}{\nm} away from the center of the disk.
-Due to the orientation and position of the dipole, which maintains the axial symmetry of the disk, the emitter only couples to axially symmetric ($\ell = 0$)  modes with zero orbital angular momentum.
-We find that the rate of spontaneous emission of 2D optical phonons is approximately 8 orders of magnitude larger than the rate of spontaneous emission of photons in the far field at frequencies corresponding to resonant modes of the hBN disk.
-Such enhancement is much larger than the enhancement presented by an unstructured, infinite sheet at the same frequency, due to the concentration of electromagnetic local density of states around the resonances.
-Nevertheless, the average spontaneous emission enhancement, defined by the integral of the enhancement over frequencies, is comparable to that of the flat sheet, in keeping with sum rules for spontaneous emission enhancement~\cite{scheel2008macroscopic}.
-We note that in this approach, the coupling of the dipole to phonon polaritons is manifested through the phonon contribution to the conductivity of the disk.
-This should be equivalent to an approach that considers the coupling of a bound electron in an emitter to LO phonons in the disk through a 2D Fr\"ohlich coupling---\ie a coupling of the atomic electron to the electric potential resulting from the polarization associated with an LO phonon mode~\cite{sohier2016two}.
-
-In Figure~\ref{fig:3}, we show that for an infrared emitter at a transition wavelength of \SI{7}{\micro\m} with a free-space radiative lifetime of \SI{1}{\micro\s}, \SI{5}{\nm} away from an hBN disk, the coupling rate to 2D optical phonons (about \SI{65}{\meV}) would be on the same scale as the optical phonon frequency itself (about \SI{180}{\meV}).
-This rate thus implies coupling between an emitter and the field in the regime of ultra-strong coupling.
-Moreover, the coupling rate for the \SI{20}{\nm} disk (purple), for an emitter with a far field decay rate of ${\gtrsim}\,\SI{3e14}{\per\s}$ would have a sufficient coupling strength to the distinct, radially-quantized resonances in purple for its linewidth to span multiple resonances and thus be in a multi-mode ultra-strong coupling regime.
-Thus, the extreme confinement of electromagnetic energy associated with LO phonons in two dimensions enables the possibility of realizing ultra-strong coupling of an atom or molecule with optical phonons in a polar material, allowing the potential realization of new coupled states of quantum emitters and phonons such as atom--phonon polariton bound states.
-
-The ability to probe low-loss and highly confined electromagnetic modes associated with optical phonons in 2D polar materials provides a new platform for nanophotonics in the mid- and far-infrared spectral range.
-The identification of the phonon polariton of bulk and thin-film geometries with the 2D LO phonon made in this manuscript would extend the rich phenomenology of optical phonons to nanophotonic applications.
-This work also points the way to useful new approaches to study LO phonons, arising from the fact that 2D LO phonons, unlike their 3D counterparts, have their electromagnetic energy extend a considerable distance from the material boundary.
-Due to the strong electromagnetic interactions between emitters and 2D phonon polaritons shown here, it is now possible to design interesting new hybrid states of matter and phonons based on quantum electrodynamical strong coupling.
-The highly confined phonon polaritons in polar monolayers may also provide interesting new opportunities in near-field radiative heat transfer, in which it has been long known that thin-film surface phonon polaritons play a critical role.
-Additional opportunities come from considering the near- and far-field optical properties of periodically structured layers involving hBN and other materials such as graphene~\cite{papadakis2017ultralight}.
-An important avenue of future study would be the \emph{ab initio} calculation of lifetimes of 2D LO phonons associated with three-phonon processes~\cite{caldwell2015low,srivastava1990physics} and electron-phonon interactions~\cite{sundararaman2014theoretical, sohier2016two, ciccarino2018dynamics}.
-In further work, it would be of great interest to study the effects of isotopic purification and cryogenic temperatures on reducing the decay rate of these 2D LO phonons~\cite{giles2017ultra}.
-
-\section{Acknowledgments}
-The authors thank Ido Kaminer, Siyuan Dai, Samuel Moore, Jennifer Coulter, and Christopher Ciccarino for helpful discussions.
-N.\,{}R. recognizes the support of the DOE Computational Science Graduate Fellowship (CSGF) fellowship no.\ DE-FG02-97ER25308. 
-T.\,{}C. acknowledges support from the Danish Council for Independent Research (Grant No.\ DFF--6108-00667). 
-This work was supported by the DOE Photonics at Thermodynamic Limits Energy Frontier Research Center under grant number DE-SC0019140.
+\end{figure*}~
 

### 2019-03-19 paper.tex
did TODOs / removed discussion of nonlocal anomaly

@@
 The considerations may also be extended to 1D polar materials, such as BN nanotubes.
-For bilayers, as was discussed in Ref.~\citenum{sohier2017breakdown}, one may describe the LO--TO splitting in terms of Eq.~\eqref{eq:zerosofresponse}, with the 2D Coulomb interaction, but with eigendisplacements appropriate to the few layer system.
-For a small number of layers $N$, \FIXME{Mention scaling of $|\hat{\mathbf{q}}\cdot\ldots|^2 = NQ_*^2/M_*$} this should enhance the LO--TO splitting by a factor of $N$. 
+One may describe the LO--TO splitting in terms of Eq.~\eqref{eq:zerosofresponse}, with the 2D Coulomb interaction, but with eigendisplacements appropriate to the few layer system.
+For a small number of layers $N$, the oscillator strength scales as $N$ \cite{sohier2017breakdown}.
+This should enhance the LO--TO splitting by a factor of $N$ and adjust the observed phonon polariton frequencies accordingly.
 To go beyond an effectively 2D description of a monolayer, and make the transition from monolayer to bulk, we may make use of the quasi-2.5D Coulomb interaction~\cite{thygesen2017calculating} in Eq.~\eqref{eq:zerosofresponse}.
-In that case, for a film of effective thickness $t$, the interaction is approximately $V_{\text{Q2.5D}}(q) = \frac{e^2}{\epsilon_0 tq^2}\big(1-\frac{2}{qt}e^{-\frac{qt}{2}}\sinh\frac{qt}{2} \big)$. 
-This form interpolates between the 2D and 3D interactions in the limits $qt \rightarrow 0$ and $qt \rightarrow \infty$. 
+In that case, for a film of effective thickness $t$, the interaction is approximately $V_{\text{Q2.5D}}(q) = \frac{e^2}{\epsilon_0\epsilon_{\infty} tq^2}\big(1-\frac{2}{qt}e^{-\frac{qt}{2}}\sinh\frac{qt}{2} \big)$.
+This form interpolates between the 2D and 3D interactions in the limits $qt \rightarrow 0$ and $qt \rightarrow \infty$, assuming that $\epsilon_{\infty}$ also interpolates between the monolayer value, which is 1, and the bulk value, which is $\epsilon_{\infty}$.
 For sufficiently large thickness and finite-$q$, the 2.5D Coulomb interaction scales as $t^{-1}q^{-2}$.
@@
 In that case, the 1D Coulomb interaction is $e^2K_0(qr)/2\pi\epsilon_0$, where $K_0$ is the (zeroth order) modified Bessel function of the second kind. 
-Thus, the $q$-dependence of the LO--TO splitting is $q^2K_0(qR)$.
-This is in close analogy with the relation for the plasma frequency of 1D plasmons, as expected from the considerations of the previous paragraph.
+Thus, the $q$-dependence of the LO--TO splitting is $q^2K_0(qR)$. This is in close analogy with the relation for the plasma frequency of 1D plasmons, as expected from the discussion of Figure 1. 
 As regards the optical properties, the optical conductivity is still governed by Eq.~\eqref{eq:2dsusceptibility}.
@@
 
-	\comment{tc}{Isn't this a bit strange to discuss, given that our results do not really show this (since we are sticking to the local case? This is only a feature in \citet{sohier2016two} results?}
-Another anomaly between 2D and 3D is that the dispersion relation of the 2D LO phonon is non-monotonic, and the frequency starts to decrease for sufficiently large wavevector.
+%\comment{tc}{Isn't this a bit strange to discuss, given that our results do not really show this (since we are sticking to the local case? This is only a feature in \citet{sohier2016two} results?}
+%Another anomaly between 2D and 3D is that the dispersion relation of the 2D LO phonon is non-monotonic, and the frequency starts to decrease for sufficiently large wavevector.
 This can never happen by extrapolating to a monolayer a dielectric function which has finite LO--TO splitting.
@@
 However it means that for the monolayer, the range of accessible phonon polaritonic frequencies for optical applications is reduced compared to thick films.
-Finally, while we have focused on anomalies between the atomically-thin limit of bulk and a true monolayer, a comparison of the monolayer with the fundamental phonon polariton mode of the one-atom-thick thin film suggests that we have a reasonable qualitative understanding of the monolayer dispersion from this limit.
+
+While we have focused on anomalies between the atomically-thin limit of bulk and a true monolayer, a comparison of the monolayer with the fundamental phonon polariton mode of the one-atom-thick thin film suggests that we have a reasonable qualitative understanding of the monolayer dispersion from the one-atom-thick thin film.
 %

### 2019-03-19 paper.tex
removing a comment

@@
 The bulk dielectric function of Eq.~\eqref{eq:3deps} fails to capture this removal of higher-order modes (\ie hyperbolicity).
-	\comment{tc}{But, na\"ively, I guess this is to be expected, since \eqref{eq:3deps} is explicitly hyperbolic?}
 For a bilayer however, there is a higher-order LO mode, associated with out-of-phase oscillation between the two layers, and thus one expects to see a higher-order mode. 

### 2019-03-19 supp.tex
Fix up SI

@@
-\documentclass[aps,prb,twocolumn, final,
+\documentclass[aps,rmp,onecolumn,
 			   groupedaddress,superscriptaddress,
-               amsfonts,amssymb,amsmath,floatfix,
-	           citeautoscript]{revtex4-1}
+			   amsfonts,amssymb,amsmath,
+			   citeautoscript,bibnotes,
+			   a4paper]{revtex4}
 
-% --- PACKAGES ---
+%-----PACKAGES-----
+\usepackage[pdftex]{hyperref}
+\hypersetup{colorlinks,
+	linkcolor={blue!75!black!80!yellow},
+	citecolor={blue!75!black!80!yellow},
+	urlcolor={blue!75!black!80!yellow},
+	pdfstartview=FitH}
+\usepackage[top=1in, bottom=1.5in, left=1.25in, right=1.25in]{geometry} %Change margins
 \usepackage{graphicx}
-\usepackage[centering,hmargin=20mm,tmargin=28mm,bmargin=25mm]{geometry}
-\usepackage{newtxtext}
-\usepackage[cmintegrals]{newtxmath}
+\usepackage{xspace}
+\newcommand\bmmax{0} % the bm package takes up a lot of alphabets internally, making it impossible to load more font packages (like sansmath); reduce the number of preloaded alphabets to avoid this
+\usepackage{bm}
+\usepackage{braket}
+\usepackage{xcolor}
+\usepackage[UKenglish]{babel}
 
-%-----SI UNITS-----
 \usepackage{siunitx}
@@
 	list-final-separator = {,\ \linebreak[0]\text{and}\ },
-	retain-unity-mantissa = false,
+	detect-all = true}
+\DeclareSIUnit[number-unit-product = ]\percent{\char`\%} % remove space before percentage sign
+
+%-----APPPEND 'S' TO REFERENCES OF ALL KINDS-----
+\renewcommand{\theequation}{S\arabic{equation}} %Equations
+\renewcommand{\thefigure}{S\arabic{figure}} %Figures
+\renewcommand{\bibnumfmt}[1]{[S#1]} %Citations
+\renewcommand{\citenumfont}[1]{S#1}
+\renewcommand{\thetable}{Supplementary Table \arabic{table}}%
+\renewcommand*\descriptionlabel[1]{\hspace\labelsep\normalfont\itshape #1\ \hfil} %Change appearance of description lists
+\renewcommand{\thesection}{S\arabic{section}}
+\renewcommand{\thesubsection}{\Alph{subsection}}
+\renewcommand{\thesubsubsection}{\roman{subsubsection}}
+
+%-----LOADING RELEVANT FONTS [Times Roman]-----
+\usepackage{txfonts}  %Times Roman fonts
+\usepackage{txfontsb} %Addition for txfonts, including old style numerals and greek
+
+%-----REDEFINE THE RMP STYLE OF AUTHORS AND AFFILIATIONS-----
+\makeatletter
+\def\frontmatter@authorformat{%
+	\preprintsty@sw{\vskip0.5pc\relax}{}%
+	\@tempskipa\@flushglue
+	\@flushglue\z@ plus50\p@\relax
+	\raggedright\advance\leftskip.25in\relax
+	\@flushglue\@tempskipa
+	\parskip\z@skip
+}%
+%
+\def\frontmatter@affiliationfont{% Helvetica 9/10.2
+	\small\slshape\selectfont\baselineskip10.5\p@\relax
+	\@tempskipa\@flushglue
+	\@flushglue\z@ plus50\p@\relax
+	\raggedright\advance\leftskip.25in\relax
+	\@flushglue\@tempskipa
 }
-
-%--- REFERENCING ---
-\usepackage{xcolor}
-\usepackage{hyperref}
-\hypersetup{colorlinks,
-	linkcolor={blue!75!black!80!yellow},
-	citecolor={blue!75!black!80!yellow},
-	urlcolor={blue!75!black!80!yellow}
-}
-
-%--- CAPTIONS (sans serif) ---
-\makeatletter
-\renewcommand\@make@capt@title[2]{%
-	\@ifx@empty\float@link{\@firstofone}{\expandafter\href\expandafter{\float@link}}%
-	\sffamily{\textbf{#1}}\@caption@fignum@sep#2
+%	
+\def\paragraph{%
+	\@startsection
+	{paragraph}%
+	{4}%
+	{\parindent}%
+	{\z@}%
+	{-1em}%
+	{\normalfont\small\itshape\sansmath\textsf}%%Change so that paragraphs are also in sans serif font
 }%
-\renewcommand\figurename{Figure}
+\renewcommand*\email[1][]{\begingroup\sanitize@url\@email{#1}} %Remove the 'Electronic adress: ' from RMP style
 \makeatother
 
-% --- SPACING ---
-\usepackage{microtype} % Kerning and symbol-stretching
-\usepackage{xspace}
+\setcitestyle{numbers,square,sort&compress} %We don't want the rmp citation style, but standard style square bracket, numbers style instead
 
+%----- Changes to caption -----
+% These hacks are specific to revtex4 (revtex4-1 solution doesn't work)
+\usepackage[EULERGREEK]{sansmath}
+\makeatletter 
+% set figure caption title to bold sans serif... 
+\renewcommand*{\fnum@figure}{{\sffamily\bfseries{}\figurename~\thefigure{}.}}
+% set figure caption contents to sans serif
+\long\def\@makecaption#1#2{%
+	\vskip\abovecaptionskip
+	\vbox{%
+		\flushing%
+		\sisetup{math-sf=\textsf}%
+		\small\sffamily\sansmath%
+		\noindent
+		#1\nobreak\hskip.5em plus.2em\ignorespaces#2\par
+	}%
+	\vskip\belowcaptionskip
+}%
+\makeatother
+
+%----- Spacing and binary relations/operations -----
 \thickmuskip=5mu plus 2mu minus 1mu  %binary relations (default, 5mu plus 5mu)
 \medmuskip=4mu plus 2mu minus 2mu    %binary operations (default, 4mu plus 2mu minus 4mu)
+\frenchspacing %Ensure that revTeX does not do "double spaces" after punctuation
 
-\frenchspacing %Ensure that revTeX does not do "double spaces" after punctuation
+%-----COMMANDS-----
+\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
+\let\endchangemargin=\endlist 
+\newcommand{\appropto}{\mathrel{\vcenter{
+			\offinterlineskip\halign{\hfil$##$\cr
+				\propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
+
+% hyphenation rules
+\hyphenation{}
 
@@
 
+% ---- Author affiliations -----
 \newcommand{\HarvardSEAS}{John A. Paulson School of Engineering and Applied Sciences, Harvard University, Cambridge, MA, 02139, USA}
 \newcommand{\MITPhy}{Department of Physics, Massachusetts Institute of Technology, Cambridge, MA, 02139, USA}
-
-% ----- COMMENTS AND META-ANNOTATIONS -----
-\usepackage{textcomp} % for \textrightarrow
-\usepackage{xifthen}
-\usepackage{etoolbox}
-\newboolean{togglecomments}
-\newboolean{togglechanges}
-
-% toggle to true to see comments (otherwise hidden)
-\setboolean{togglecomments}{true}  
-% toggle to false to see mixed versions (otherwise edits are shown exclusively)
-\setboolean{togglechanges}{false}
-
-\newcommand{\comment}[2]{%
-    \ifbool{togglecomments}%
-    {\textcolor{blue!70!black}{\small\textsf{%
-    \textsuperscript{\textsc{\textsf{\MakeLowercase{#1}}}}%
-    [#2]}}} % if true, show comments
-    {}}     % if false, do nothing
-\newcommand{\swap}[2]{\ifbool{togglechanges}
-    {#2}  % TC-only version
-    {\textcolor{red!70!black}{[#1]}\text{\textrightarrow{}}\textcolor{green!50!black}{[#2]}}}
-\newcommand{\remove}[1]{\ifbool{togglechanges}
-    {}    % TC-only version
-    {\textcolor{red!70!black}{#1}}}
-\newcommand{\inset}[1]{\ifbool{togglechanges}
-    {{}#1}  % TC-only version
-    {\textcolor{green!50!black}{{}#1}}}
-\makeatletter
-\newcommand{\citeremind}[1]{%
-	\unskip%
-    \textcolor{blue!75!black!80!yellow}{${}^\blacksquare$%
-	\ifthenelse{\isempty{#1}}{}{\textsuperscript{\tiny\textsf{#1}}}%
-	}\xspace}
-\makeatother
-\newcommand{\FIXME}[1]{\textcolor{red!70!black}{[#1]}}
-\newcommand{\optional}[1]{\textcolor{yellow!70!gray}{#1}}
 
@@
 % --- TITLE & AUTHORSHIP ---
-\title{Supporting Information for: Phonon polaritonics in two-dimensional materials}
+\title{SUPPORTING INFORMATION\\Phonon polaritonics in two-dimensional materials}
 
@@
 
+\vspace*{4ex}
 \begin{figure*}[h]
-    \includegraphics[width=0.6\linewidth]{enhancement.png}
+    \includegraphics[width=0.725\linewidth]{enhancement.png}
     \caption{%
-        \textbf{Spontaneous emission enhancement in hBN thin films and monolayer.} Spontaneous emission enhancement for a dipole emitter 5 nm away from the surface of thin films of bulk hBN as a function of emission frequency for film thicknesses of 100 nm, 10 nm, 1 nm, and 0.33 nm, as well as monolayer hBN with dispersion relation and conductivity given in the main text.
-The hBN in all cases is surrounded by air.
-The TO frequency  taken for the bulk is 1387 cm$^{-1}$.
+        \textbf{Spontaneous emission enhancement in hBN thin films and monolayer.} 
+        Spontaneous emission enhancement for a dipole emitter \SI{5}{\nm} away from the surface of thin films of bulk hBN as a function of emission frequency for film thicknesses of \SIlist{100; 10; 1; 0.33}{\nm}, as well as monolayer hBN (dispersion relation and conductivity given in the main text).
+        In all cases, the hBN is assumed surrounded by air. 
+        The bulk TO frequency is \SI{1387}{\per\cm}.
         \label{fig:s1}
         }
-\end{figure*}~
+\end{figure*}
 
-\bibliographystyle{apsrev4-1}
-\bibliography{references}
+\mbox{}
+\vfill
+
+\mbox{}
+
+\renewcommand\refname{} % we don't have any references this time around, so suppress printing of 'References' section head
+%\bibliographystyle{apsrev4-1}
+%\bibliography{references}
 

### 2019-03-19 paper.tex
Some updates to intro (micro+macro sentences) & some revisions of the 2D->3D and 2D->1D discussion

@@
 \newcommand{\super}[1]{\ensuremath{^{\textrm{#1}}}} %Upright multi-character superscript
+
+\newcommand{\appropto}{\mathrel{\vcenter{
+			\offinterlineskip\halign{\hfil$##$\cr
+				\propto\cr\noalign{\kern.5pt}\sim\cr\noalign{\kern-1pt}}}}}
 
@@
     Does this lack of LO--TO splitting imply the absence of a phonon polariton in polar monolayers? 
-    Here, we derive a first-principles expression for the conductivity of a polar monolayer specified by the wavevector-dependent LO and TO phonon dispersions.
+	\addition{To answer this, we connect the microscopic phonon properties with the macroscopic electromagnetic response.}
+	Specifically, we derive a first-principles expression for the conductivity of a polar monolayer specified by the wavevector-dependent LO and TO phonon dispersions.
 In the long-wavelength (local) limit, we find a universal form for the conductivity in terms of the LO phonon frequency at the $\Gamma$ point,
@@
 
-We now develop a framework with quantitative, concrete examples that enables a first-principles understanding of phonon polaritons in 2D materials. 
-We do so by deriving a universal form for the conductivity of a polar monolayer,  which depends solely on the LO and TO phonon frequencies---and their dispersion with momentum---in the 2D system.
+\addition{%
+To resolve this question, a connection between the microscopic phonon properties and the macroscopic electromagnetic response is required.
+}
+To that end, we develop a first-principles framework for phonon polaritons in polar 2D materials and illustrate it with quantitative, concrete examples.
+We derive a universal form for the conductivity of a polar monolayer,  which depends solely on the LO and TO phonon frequencies---and their dispersion with momentum---in the 2D system.
 Using parameters from Ref.~\citenum{sohier2017breakdown} for the canonical 2D polar monolayer---hexagonal boron nitride (hBN)---we present the confinement and propagation losses of the 2D phonon polariton modes, identifying the frequency region where they should be most easily detected.
@@
 The second comment is that much of what has been discussed here has a strong analogy with the theory of optical response in electron gases in 2D, and particularly the relation between plasmons in 2D and 3D.
-To elaborate on this analogy, we take Equation~\eqref{eq:zerosofresponse} in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $\Big|\hat{\mathbf{q}}\cdot \sum_\kappa \mathbf{Z}_\kappa\boldsymbol{\eta}_\kappa\Big|^2 \equiv \frac{Q_*^2}{M_*}$, with $Q_*$ being an effective charge and $M_*$ being an effective mass.
+To elaborate on this analogy, we take Equation~\eqref{eq:zerosofresponse} in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $\big|\hat{\mathbf{q}}\cdot \sum_\kappa \mathbf{Z}_\kappa\boldsymbol{\eta}_\kappa\big|^2 \equiv Q_*^2/M_*$, with $Q_*$ being an effective charge and $M_*$ being an effective mass.
 Then, the LO--TO splitting can be written as $\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} = \frac{Q_*^2}{2\epsilon_0\epsilon_{\mathrm{env}} M_*}q$.
@@
 \addition{%
-The third comment is that the considerations of this section can be extended to few-layer systems, such as hBN bilayers, which may still be considered 2D for polariton wavelengths large compared to the bilayer thickness.
-The considerations may also be extended to 1D polar materials, such as BN nanotubes.
-One may describe the LO--TO splitting in terms of Eq.~\eqref{eq:zerosofresponse}, with the 2D Coulomb interaction, but with eigendisplacements appropriate to the few layer system.
-For a small number of layers $N$, the oscillator strength scales as $N$ \cite{sohier2017breakdown}.
-This should enhance the LO--TO splitting by a factor of $N$ and adjust the observed phonon polariton frequencies accordingly.
-To go beyond an effectively 2D description of a monolayer, and make the transition from monolayer to bulk, we may make use of the quasi-2.5D Coulomb interaction~\cite{thygesen2017calculating} in Eq.~\eqref{eq:zerosofresponse}.
-In that case, for a film of effective thickness $t$, the interaction is approximately $V_{\text{Q2.5D}}(q) = \frac{e^2}{\epsilon_0\epsilon_{\infty} tq^2}\big(1-\frac{2}{qt}e^{-\frac{qt}{2}}\sinh\frac{qt}{2} \big)$.
-This form interpolates between the 2D and 3D interactions in the limits $qt \rightarrow 0$ and $qt \rightarrow \infty$, assuming that $\epsilon_{\infty}$ also interpolates between the monolayer value, which is 1, and the bulk value, which is $\epsilon_{\infty}$.
-For sufficiently large thickness and finite-$q$, the 2.5D Coulomb interaction scales as $t^{-1}q^{-2}$.
-Noting the $q^2$ factor in Eq.~\eqref{eq:zerosofresponse}, as well as the $N$-dependence of the quantity involving eigendisplacements, one finds an LO--TO splitting independent of thickness, as expected for bulk samples.
+The third comment is that the considerations of this section can be extended to few-layer systems, such as hBN bilayers, which remain two-dimensional from an electromagnetic perspective for polariton wavelengths large compared to the bilayer thickness.
+Concretely, the LO--TO splitting in such quasi-2D systems is still given by Eq.~\eqref{eq:zerosofresponse}, provided the appropriate eigendisplacements of the few-layer system are employed, and provided suitable modifications of the Coulomb interaction are made.
+For $N$ weakly coupled layers, the sum of Born effective charges $\big|\hat{\mathbf{q}}\cdot \sum_\kappa \mathbf{Z}_\kappa\boldsymbol{\eta}_\kappa\big|^2$, which is essentially an oscillator strength, is approximately $NQ_*^2/M_*$, \ie enhanced $N$-fold relative to the monolayer or, equivalently, proportional to the thickness $t$~\cite{sohier2017breakdown}.
+For few layers, \ie for $qt\ll 1$ where the Coulomb interaction remains essentially 2D, this simple geometric effect enhances the LO--TO splitting by a factor $N$.
+At larger $qt$, the Coulomb interaction must eventually approach its 3D limit; to study this transition from monolayer to bulk explicitly, a quasi-2.5D Coulomb interaction~\cite{thygesen2017calculating} can be employed in Eq.~\eqref{eq:zerosofresponse}.
+Specifically, for a film of thickness $t$, the interaction is approximately $V_{\text{Q2.5D}}(q) = \frac{e^2}{\epsilon_0\epsilon_{\infty} q^2t}\big(1-\frac{2}{qt}e^{-\frac{qt}{2}}\sinh\frac{qt}{2} \big)$, with the static electronic screening $\epsilon_{\infty}$ interpolating between its monolayer value, 1, and its bulk value, $\epsilon_{\infty}$.
+The $qt \rightarrow 0$ and $qt \rightarrow \infty$ asymptotics of this interaction are the 2D and 3D interactions---$e^2/2\epsilon_0q$ and $e^2/\epsilon_0\epsilon_\infty q^2t$, respectively---which it interpolates between.
+In the large-thickness or -momentum limit, $qt\gg 1$, the ${\appropto}\,t$ scaling of the oscillator strength, the ${\appropto}\,t^{-1}q^{-2}$ scaling of the Coulomb interaction, and the ${\propto}\,q^2$ factor in Eq.~\eqref{eq:zerosofresponse} cancel, producing a thickness- and momentum-independent LO--TO splitting, as expected in the bulk.
 }
@@
 \addition{%
-For 1D systems, beyond using the eigendisplacements appropriate to the 1D system, one must also use the 1D Coulomb interaction, cutoff by a length $R$ which for a nanotube, would be its radius. 
-In that case, the 1D Coulomb interaction is $e^2K_0(qr)/2\pi\epsilon_0$, where $K_0$ is the (zeroth order) modified Bessel function of the second kind. 
-Thus, the $q$-dependence of the LO--TO splitting is $q^2K_0(qR)$. This is in close analogy with the relation for the plasma frequency of 1D plasmons, as expected from the discussion of Figure 1. 
-As regards the optical properties, the optical conductivity is still governed by Eq.~\eqref{eq:2dsusceptibility}.
+Analogously, our considerations can be extended to 1D polar materials, such as BN nanotubes, by employing the 1D Coulomb interaction (and, naturally, the 1D-specific eigendisplacements).
+For a 1D system of transverse extent $R$ (\eg, the radius of a nanotube), the 1D Coulomb interaction is $e^2K_0(qr)/2\pi\epsilon_0$, where $K_0$ is the (zeroth order) modified Bessel function of the second kind.
+Thus, the $q$-dependence of the (squared) LO--TO splitting is $q^2K_0(qR)\simeq - q^2 \log q R$. 
+This closely mirrors the situation in 1D plasmonics, as expected from the discussion of Figure 1. 
+%As regards the optical properties, the optical conductivity is still governed by Eq.~\eqref{eq:2dsusceptibility}.
 }
@@
     \caption{%
-        \textbf{Extreme spontaneous emission enhancement due to 2D phonon polaritons in nanostructured geometries.} Plotted is the enhancement of the spontaneous emission rate for an emitter  $z=\SI{5}{\nm}$ above the disk's center and  oriented normal to its plane ($\hat{\mathbf{n}} = \hat{\mathbf{z}}$).
+        \textbf{Extreme spontaneous emission enhancement due to 2D phonon polaritons in nanostructured geometries.} 
+        Plotted is the enhancement of the spontaneous emission rate for an emitter  $z=\SI{5}{\nm}$ above the disk's center and  oriented normal to its plane ($\hat{\mathbf{n}} = \hat{\mathbf{z}}$).
 For a disk with a diameter of \SI{20}{\nm}, and assuming a relaxation rate $\tau^{-1} = \SI{0.5}{\meV}$, the rate of emission enhancement can be enhanced \num{100} million-fold.

### 2019-03-19 paper.tex
stuff before change to achemso format

@@
 In the absence of such LO--TO splitting in 3D, there would be no frequency compatible with the existence of a phonon  polariton.
-On the contrary, we will show that in 2D, despite the absence of LO--TO splitting at the $\Gamma$ point, there persists a strongly confined evanescent electromagnetic mode with a high local density of states which in all respects is similar to a phonon polariton of a thin film, but is in fact the 2D LO phonon of the polar monolayer (thus the phrases ``phonon polariton'' and ``2D LO phonon'' may be used somewhat interchangeably as is the case in plasmonics where the terms ``plasmon polariton'' and ``2D plasmon'' are often used interchangeably).
+On the contrary, we will show that in 2D, despite the absence of LO--TO splitting at the $\Gamma$ point, there persists a strongly confined evanescent electromagnetic mode with a high local density of states which in all respects is similar to a phonon polariton of a thin film, but is in fact the 2D LO phonon of the polar monolayer (thus the phrases `phonon polariton' and `2D LO phonon' may be used somewhat interchangeably as is the case in plasmonics where the terms `plasmon polariton' and `2D plasmon' are often used interchangeably).
 
@@
 Analogously, our considerations can be extended to 1D polar materials, such as BN nanotubes, by employing the 1D Coulomb interaction (and, naturally, the 1D-specific eigendisplacements).
-For a 1D system of transverse extent $R$ (\eg, the radius of a nanotube), the 1D Coulomb interaction is $e^2K_0(qr)/2\pi\epsilon_0$, where $K_0$ is the (zeroth order) modified Bessel function of the second kind.
-Thus, the $q$-dependence of the (squared) LO--TO splitting is $q^2K_0(qR)\simeq - q^2 \log q R$. 
-This closely mirrors the situation in 1D plasmonics, as expected from the discussion of Figure 1. 
+For a 1D system of transverse extent $R$ (\eg, the radius of a nanotube), the 1D Coulomb interaction is $e^2K_0(qR)/2\pi\epsilon_0$, where $K_0$ is the (zeroth order) modified Bessel function of the second kind.
+Thus, the $q$-dependence of the (squared) LO--TO splitting in 1D is $q^2K_0(qR)\simeq - q^2 \log q R$. 
+This parallels the situation in 1D plasmonics, as expected from the discussion of Figure 1. 
 %As regards the optical properties, the optical conductivity is still governed by Eq.~\eqref{eq:2dsusceptibility}.
@@
 This qualitative similarity is to be somewhat expected, as a 2D layer can be considered as a very thin film of bulk material, provided that the microscopic properties of the bulk and monolayer do not deviate substantially.
-The differences between the $t = t_{\text{2D}} = \SI{0.33}{\nm}$ bulky thin-film and monolayer that nevertheless do exist, reflect such microscopic deviations.
+The differences between the $t = t_{\text{2D}} = \SI{0.33}{\nm}$ bulk thin-film and monolayer that nevertheless do exist, reflect such microscopic deviations.
 Occasionally, \eg for compatibility with standard numerical tools, it is useful to introduce a fictitious, effective bulk material whose fundamental thin-film mode \emph{exactly} matches the monolayer's.
@@
 \begin{equation}
-	\epsilon_{\text{Q2D}, \parallel}(\omega) = 1 + \frac{1}{t}\frac{4\epsilon_{\mathrm{env}}\omega_{\mathrm{TO}}v_\mathrm{g}}{\omega^2-\omega_{\mathrm{TO}}^2+\iu\omega\tau^{-1}}.
+	\epsilon_{\text{Q2D}, \parallel}(\omega) 
+	=
+	 1 + \frac{\iu\sigma}{\varepsilon_0\omega t} 
+ 	=
+	 1 + \frac{1}{t}\frac{4\epsilon_{\mathrm{env}}\omega_{\mathrm{TO}}v_\mathrm{g}}{\omega^2-\omega_{\mathrm{TO}}^2+\iu\omega\tau^{-1}}.
 \end{equation}

### 2019-03-19 paper.tex
Change referencing style to achemso
Last bits

@@
 \end{equation} 
-We have defined scaled eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}\equiv \frac{\hat{\mathbf{e}}_{\kappa\mathbf{q},\mathrm{L}}}{\sqrt{M_{\kappa}}}$, where $\hat{\mathbf{e}}_{\kappa\mathbf{q},\mathrm{L}}$ is the unit-normalized polarization vector of atom $\kappa$ in the unit cell oscillating according to a longitudinal phonon of wavevector $\mathbf{q}$ and $M_{\kappa}$ is the mass of atom $\kappa$.
+We have defined scaled eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}\equiv {\hat{\mathbf{e}}_{\kappa\mathbf{q},\mathrm{L}}}/{\sqrt{M_{\kappa}}}$, where $\hat{\mathbf{e}}_{\kappa\mathbf{q},\mathrm{L}}$ is the unit-normalized polarization vector of atom $\kappa$ in the unit cell oscillating according to a longitudinal phonon of wavevector $\mathbf{q}$ and $M_{\kappa}$ is the mass of atom $\kappa$.
 
 While the conductivity is the main electromagnetic quantity of interest for electrodynamics applications, we briefly state the form of the (2D) permittivity, as its zeros immediately yield the longitudinal modes of the system, which are the LO phonons.
-The permittivity within the RPA, denoted $\epsilon_{\textrm{RPA}}$ is related to the polarization-polarization response function via~\cite{jablan2009plasmonics} $\epsilon_{\textrm{RPA}} = \epsilon_{\mathrm{env}} + \frac{1}{2}q\Pi(q,\omega)$ and the conductivity via $\epsilon_{\textrm{RPA}} = \epsilon_{\mathrm{env}} + \frac{\iu q\sigma(q,\omega)}{2\omega\epsilon_0}$.
+The permittivity within the RPA, denoted $\epsilon_{\textrm{RPA}}$ is related to the polarization-polarization response function via~\cite{jablan2009plasmonics} $\epsilon_{\textrm{RPA}} = \epsilon_{\mathrm{env}} + \frac{1}{2}q\Pi(q,\omega)$ and the conductivity via $\epsilon_{\textrm{RPA}} = \epsilon_{\mathrm{env}} + {\iu q\sigma(q,\omega)}/{2\epsilon_0\omega}$.
 Here, $\epsilon_{\mathrm{env}}$ is the average permittivity of the bulk above and below the monolayer, and is added to take into account the polarization arising from these bulk materials.
@@
 \end{equation}
-  where $V(q)$ is the Coulomb interaction in Fourier space, which in two dimensions, is given by $V(q) = \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}$.
+  where $V(q)$ is the Coulomb interaction in Fourier space, which in two dimensions, is given by $V(q) = {e^2}/{2\epsilon_0\epsilon_{\mathrm{env}}q}$.
 Given that the zeros of the dielectric function are associated with longitudinal modes, one expects that $\omega_{\mathbf{q}}$ is in fact the frequency of the LO phonon mode.
@@
 The bulk dielectric function of Eq.~\eqref{eq:3deps} fails to capture this removal of higher-order modes (\ie hyperbolicity).
-For a bilayer however, there is a higher-order LO mode, associated with out-of-phase oscillation between the two layers, and thus one expects to see a higher-order mode. 
-However, it is expected challenging to observe such a mode, as can be anticipated from the blue higher-order mode in Figure~\ref{fig:2}(c), because even for a polariton wavelength of \SI{12}{\nm}, which is very short, the polariton frequency is only a few \si{\per\cm} above the TO frequency, within the damping linewidth of the TO frequency.
+For a bilayer, however, there is a higher-order LO mode, associated with out-of-phase oscillation between the two layers.
+Heuristically, as can be anticipated from the blue higher-order mode in Figure~\ref{fig:2}(c), this mode would be challenging to observe, given its extremely high confinement and small spectral separation from the TO mode (a few \si{\per\cm}, comparable to the damping linewidth).
 }
@@
  	=
-	 1 + \frac{1}{t}\frac{4\epsilon_{\mathrm{env}}\omega_{\mathrm{TO}}v_\mathrm{g}}{\omega^2-\omega_{\mathrm{TO}}^2+\iu\omega\tau^{-1}}.
+	 1 + \frac{1}{t}\frac{4\epsilon_{\mathrm{env}}\omega_{\mathrm{TO}}v_\mathrm{g}}{\omega_{\mathrm{TO}}^2-\omega^2-\iu\omega\tau^{-1}}.
 \end{equation}
@@
 
-In Figure~\ref{fig:3}, we consider the coupling of a dipole emitter to localized phonon polaritons of nanostructured monolayer hBN.
+In Figure~\ref{fig:3}, we consider the coupling of a dipole emitter to localized phonon polaritons of nanostructured monolayer hBN \addition{(we also compare extended monolayer and thin-film hBN in the Supporting Information)}.
 For simplicity, we consider hBN nanostructured as a disk, which leads to the formation of sharp resonances quantized along the azimuthal and radial directions.
@@
 
-\bibliographystyle{apsrev4-1}
+\bibliographystyle{achemso}
 \bibliography{references}
